### 2019-02-05 docs/manuscipt.Rnw
adjusted alzheimer data analysis

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

% Add history information for the article if required
%\history{Received August 1, 2010;
%	revised October 1, 2010;
%	accepted for publication November 1, 2010}

% objects
\newcommand{\x}{\mathbf{x}}
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\m}{\mathbf{m}}

% functions and operators
\newcommand{\tr}{^{\text{T}}}
\newcommand{\expit}{\text{expit}}
\newcommand{\argmax}{\text{argmax} \,}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\diag}{\text{diag}}
\newcommand{\trace}{\text{tr}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Adaptive group-regularized logistic elastic net regression}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}

\begin{document}

	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge, United Kingdom \\

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features is often available. Examples of such external information in omics research are: (a) $p$-values from a previous study, (b) a summary of prior information, and (c) omics annotation. The inclusion of this information in the analysis may enhance classification performance and feature selection, but is not straightforward in the standard regression setting. As a solution to this problem, we propose a group-regularized (logistic) elastic net regression method, where each penalty parameter corresponds to a group of features based on the external information. The method, termed \texttt{gren}, makes use of the Bayesian formulation of logistic elastic net regression to estimate both the model and penalty parameters in an approximate empirical-variational Bayes framework. Simulations and an application to a colon cancer microRNA study show that, if the partitioning of the features is informative, classification performance and feature selection are indeed enhanced.}
	\end{abstract}

	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes

	\noindent\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}

	\section{Introduction}
	Prediction from high-dimensional data is a challenge common to many fields of research. Examples of such high-dimensional prediction problems arise in computer vision, stock market prediction and disease diagnosis from omics data. In this paper, we specifically focus on the latter setting. High dimensionality of data introduces several issues in the estimation of prediction models, especially: (a) unidentifiable models, (b) highly correlated predictor variables and (c) non-trivial variable selection.

	Attempts to tackle one or more of these issues have lead to the development of many prediction methods \cite[]{fan_selective_2010}. Here we will focus on the penalization approach. A well-known and popular penalization method is elastic net regression \cite[]{zou_regularization_2005}, with its special cases ridge regression \cite[]{hoerl_ridge_1970} and lasso regression \cite[]{tibshirani_regression_1996}. The elastic net approach takes issues (a)-(c) into account and has yielded many successful extensions. Recent work on elastic net regression (and its special cases) has focussed on increasing prediction accuracy by the inclusion of prior knowledge on the variables \cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.

	In omics research, sources of such prior knowledge on the variables are often available. The information can, for example, come in the form of (a) results on the same molecular features obtained in a previous study (e.g. $p$-values), (b) information from a publicly available database that summarizes the prior information on the molecular features involved (e.g., the Cancer Gene Census \cite[]{futreal_census_2004}), and (c) omics annotation (e.g. the location of a gene on the chromosome). Although this information can rarely be directly included in the statistical analysis, it may still be useful and informative for the study at hand.

	An example of such a study with extra information available is investigated in Section \ref{sec:app}. INTRODUCE DATA. This problem is not straightforward to solve with the standard methods available in literature. At first sight, such an \textit{a priori} grouping of the features suggests the group lasso \cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which penalises and selects features at the group level. There are two limitations to the group lasso: (a) it selects groups as a whole instead of single features and (b) it is not adaptive: the amount of penalization a group receives depends on a `global' penalty parameter common to all groups. Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter. An apparent issue with this differential penalization is the estimation of the penalty parameters. Naive estimation may be done by cross-validation. However, cross-validation requires re-estimation of the model over a grid of penalty parameters. The size of this grid increases exponentially with the number of penalty parameters. Consequently, it quickly becomes computationally infeasible. We therefore propose an efficient alternative: empirical Bayes estimation of the penalty parameters, which corresponds to hyperparameter estimation in the Bayesian prior framework. Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net. Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit. Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.

	In line with the above, our proposed method is (a) subjective: the few remaining tuning parameters may be cross-validated, (b) adaptive: prior information is automatically and subjectively weighed with respect to its informativeness, (c) fast compared to full Bayesian analysis or cross-validation, and (d) requires minimal input by the user: only the data and grouping of the features is required.

	We introduce the Bayesian logistic (generalised) elastic net model in Section \ref{sec:model}. In Section \ref{seq:estimation} we derive a variational approximation to this model and use this novel approximation in the empirical Bayes estimation of multiple, group-specific penalty parameters. In Sections \ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a simulation study and a real data example, respectively. In Section \ref{sec:discussion} we conclude by both discussing some differences and parallels between the proposed method, termed \texttt{gren}, and other methods in the literature.

	\section{Model}\label{sec:model}
	\subsection{Logistic elastic net regression}\label{sec:logitmodel}
	In logistic regression the outcome variables are assumed to be binary or sums of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l \in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix} x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} \( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is the binomial distribution with number of trials $m$ and probability $\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + \exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model reduces to a binary logistic regression model. Throughout the rest of the paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots & \x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.

	In elastic net regression, the penalised likelihood is maximised to yield parameter estimates:
	$$
	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
	$$
	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} y_1 & \cdots & y_n \end{bmatrix} \tr$. The likelihood term is thus complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the parameters, multiplied by respective penalty parameters $\lambda_1$ and $\lambda_2$. We have scaled the penalty parameters by $\frac{1}{2}$, to make them correspond with the Bayesian version in the following. The maximiser of the penalised likelihood corresponds to the posterior mode of a Bayesian elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
	\begin{equation}\label{eq:elasticnetprior}
	\beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta_j| + \lambda_2 \beta_j^2)\].
	\end{equation}
	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the Supplementary Material (SM), along with further details on the elastic net prior). The elastic net combines $L_1$ and $L_2$-norm penalisation such that the model parameters are shrunken towards zero. The $L_1$-norm may set some of the estimates exactly to zero, thus automatically selecting features. The $L_2$-norm ensures that correlated features behave similarly, as one would generally require. We extend the elastic net model to the generalised elastic net to allow for different penalty parameters per group of features.

	\subsection{The generalised elastic net}\label{sec:genen}
	Assume we have a partitioning of the features in $G$ groups, such that each feature belongs to one group. Let $\G(g)$ be the feature index set of group $g$ for $g=1, \dots, G$. To apply differential penalisation we include a group-specific weight $w_g \in \mathbb{R}_{>0}$. The model parameters are now estimated by:
	\begin{align}
	\hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 \label{eq:weightedlik}\\ 
	& = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, \nonumber
	\end{align}
	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific weights may be interpreted as penalty multipliers. Throughout the following we assume that the geometric mean of the multipliers, weighted by their respective group sizes, is one, such that the average shrinkage of the model parameters is determined by the `global' $\lambda_1$ and $\lambda_2$. That is, we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G (\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the $L_1$-norm term to ensure that penalisation on the parameter level is the same for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in (\ref{eq:weightedlik}).

	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as a computationally more convenient scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$. Using this result, we write the generalised elastic net model in its Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i \bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this Bayesian formulation the penalty parameters in $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.

	\section{Estimation}\label{seq:estimation}
	\subsection{Empirical Bayes}\label{seq:empiricalbayes}
	If the penalty parameters are known, estimation of the frequentist elastic net model parameters or finding the posterior of the generalized elastic net model is feasible with small adjustments of the available algorithms \cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers \cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is not straightforward.

	In the frequentist elastic net without group-wise penalisation, two main strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by cross-validation over a two-dimensional grid of values or (ii) re-parametrise the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 \lambda_2 + \lambda_1}$ and $\lambda = 2 \lambda_2 + \lambda_1$, fix the proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global penalty parameter $\lambda$. Strategy (i) is advised by \cite{waldron_optimized_2011}, while (ii) is proposed by \cite{friedman_regularization_2010}. In the generalised elastic net setting, strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, respectively. $K$-fold cross validation over $D$ values then results in $K \cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate. Typically $K$ is set to 5, 10, or to the number of samples $n$, while $D$ is in the order of $100$, so that even for small $G$, the number of models to estimate is prohibitively large.

	In the Bayesian framework, estimation of penalty parameters may be avoided by the addition of a hyperprior to the model hierarchy. The hyperprior takes the uncertainty in the penalty parameters into account by integrating over them. This approach introduces two issues. Firstly, the choice of hyperprior is not straightforward. Many authors suggest a hyperprior from the gamma family of distributions \cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, but the precise parametrisation of this gamma prior is not so obvious. A second issue is the loss of correspondence between the  Bayesian and frequentist elastic net. If the ultimate goal is feature selection, this correspondence may be exploited through the automatic feature selection property of the frequentist elastic net. Endowing the penalty parameters with a hyperprior obstructs their point estimation and, consequently, impedes automatic feature selection. Therefore, to circumvent the problem of hyperprior choice and allow for feature selection by the frequentist elastic net, we propose to estimate the penalty parameters by empirical Bayes.

	Many forms of empirical Bayes exist, the most formal one being maximisation of the marginal likelihood with respect to the hyperparameters. The resulting hyperparameter estimates are then plugged into the prior. The marginal likelihood is often introduced as a measure of model evidence given the observed data and is computed by integrating the product of likelihood and prior with respect to the model parameters. In the case of the elastic net introduced in (\ref{eq:enmodel2}) this results in the following marginal empirical Bayes posterior for $\bbeta$:
	\begin{align}
	p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = \frac{\int_{\btau} \L (\y ; \bbeta) \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \, d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
	\hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
	& = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. \label{eq:mmlestimate}
	\end{align}
	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the elastic net. In the omics setting the integrals are also high-dimensional, in which case numerical and Monte Carlo approximation methods become tedious and computationally expensive. Moreover, Laplace approximation is known to suffer from low accuracy in many high-dimensional settings \cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm is described that estimates the hyperparameters. This EM algorithm iteratively maximises the expected joint log likelihood, such that the sequence:
	\begin{equation}\label{eq:ebEM}
	\blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
	\end{equation}
	converges to a local maximum of the marginal likelihood. The difficulty herein is in the calculation of the expected joint log likelihood. \cite{casella_empirical_2001} suggests to approximate the expectation by its Monte Carlo expectation. Although elegant and simple, this method requires a converged MCMC sample from the posterior for every iteration: a computationally intensive procedure. \cite{roy_selection_2017} introduce generalised importance sampling for the Bayesian elastic net, such that we just need a limited, pre-specified number of MCMC chains. However, this still requires several converged chains, so is not feasible in many high-dimensional omics settings. We propose to tackle this problem by approximating the expectation in (\ref{eq:ebEM}) using variational Bayes.

	\subsection{Variational Bayes}\label{sec:variationalbayes}
	Variational Bayes is a widely used method to approximate Bayesian posteriors. It has successfully been applied in a wide range of applications, including genetic association studies \cite[]{carbonetto_scalable_2012} and gene network reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior is approximated by a tractable form and estimated by optimizing a lower bound on the marginal likelihood of this model (see Section 3 of the SM for the lower bound of the proposed model). For an extensive introduction and concise review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, respectively.

	To simplify the computations of our variational approximation, we follow \cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for $i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	\bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of (\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, calculus of variations gives the optimal distributions $q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} [\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the Kullback-Leibler divergence of the posterior to the approximate distribution \cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation}) renders both the posterior parameter calculations and the expected joint log likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal distributions in our variational Bayes implementation for the model parameters as:
	\begin{equation}\label{eq:varbayesdistr}
	q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)\text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\), 
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian distribution (See SM Section 5 for the derivations). The so-called variational parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		\bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		\bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\ 
		c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		\chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries $\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & \cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and $\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th element of a column vector, respectively. Naive calculation of the variational parameters is computationally expensive. In Section 6 of the SM we show that informed calculation of the parameters results in a significant reduction of computational complexity.

	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
	Variational Bayes was shown to underestimate the posterior variance of the parameters, both numerically and theoretically, in several settings \cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,wang_inadequacy_2005}. This coincides with our experience that the global penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To prevent overestimation we use the parametrisation of the elastic net in \cite{friedman_regularization_2010} as introduced in Section \ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by cross-validation of the regular elastic net model, such that the overall penalisation is determined by cross-validation of only $\lambda$. By combining cross-validation of the global penalty parameter $\lambda$ with empirical Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more robust to underestimation of the variational posterior variances. The remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend to either fix $\alpha$ \textit{a priori}, or compare the results for several choices of $\alpha$. We recommend the latter. Otherwise we note that, in our experience, the choice $\alpha=0.5$ often gives good results.

	For estimating the penalty multipliers, the intractable posterior expectation in (\ref{eq:ebEM}) is approximated by the variational posterior:
	$$
	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \] = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
	$$
	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j \in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate of the new penalty multipliers is now given by:
	\begin{subequations}\label{eq:mmlupdateequation}
		\begin{align}
		\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
		& \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
		\end{align}
	\end{subequations}
	Although the solution to (\ref{eq:mmlupdateequation}) is not available in closed form, this convex problem is easily solved by a numerical optimisation routine. The full procedure is summarized in Section 8 of the SM.

	\subsection{Feature selection}
	Feature selection is often desirable in high-dimensional prediction problems and omics data problems are no exception. A selection of biomarkers may lead to a large cost reduction by supporting targeted assays. Bayesian feature selection is often done by inspection of posterior credible intervals. Similarly as in frequentist hypothesis testing, we may select a feature if zero is not contained in the credible interval. However, the Bayesian lasso's credible intervals (a special case of the elastic net) are known to suffer from low frequentist coverage in sparse settings \cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in omics research, so selection by credible intervals is generally inconsistent. We therefore propose to select features in the frequentist paradigm.

	As shortly touched upon in Section \ref{seq:empiricalbayes}, frequentist feature selection is trivial after estimation of the penalty multipliers. We simply plug the estimated penalty parameters into some frequentist elastic net algorithm that allows for differential penalization. In our own package \texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} \cite[]{friedman_regularization_2010}. In the frequentist elastic net, feature selection is then done automatically. Furthermore, to select a specific number of features, we simply adjust the global $\lambda$ until we select the desired number of features.

	\section{Simulations}\label{sec:simulations}
	\subsection{Setup}\label{sec:setup}
	To assess variable selection and predictive performance we conduct a simulation study in which we compare \texttt{gren} to the regular elastic net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, group mimimax concave penalty (\texttt{grMCP}) \cite[]{breheny_penalized_2009}, and the group exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to \texttt{gren} in the sense that it estimates group-specific penalty multipliers. The two main differences with \texttt{gren} are (i) the absence of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} and not adaptive on the group level. In contrast to the original group lasso, all of them are able to select single features, instead of just on the group level.

	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, we fix the proportion of $L_1$-norm penalty to three different values: $\alpha \in \{0.05, 0.5, 0.95 \}$. The first setting closely resembles the ridge setting, where $\alpha=0$, while the third setting is similar to the lasso with $\alpha=1$. For the methods that do automatic variable selection (all methods except the regular ridge and \texttt{GRridge}), we estimate the models for a range of model sizes.

	Predictive performance is measured by area under the receiver operator curve (AUC) and Brier skill score (BSS). AUC gives the area under the curve of sensitivity versus $1-$ specificity and equals the probability that the classifier ranks a randomly chosen case higher than a randomly chosen non-case. The BSS is a normalised version of the Brier score: $\text{BSS} = 1 - \frac{\sum_{i=1}^n (y_i - \hat{y}_i)^2}{\sum_{i=1}^n (y_i - \bar{y})^2}$,
	where $\hat{y}_i=\expit(\x_i \tr \hat{\bbeta})$ is the prediction for observation $i$ and $\bar{y}=n^{-1} \sum_{i=1}^n y_i$. The BSS measures the fraction of Brier score improvement compared to a non-informative model. Quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960} and parameter estimation by mean squared error (MSE). Both are presented in SM Section 10.

	We simulated data according to five different scenarios: 
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations within groups of features, 
		\item differential signal between the groups and heavy-tailed distributed model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 10 of the SM. In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform. In the second scenario, we expect \texttt{gren} and \textt{GRridge} to overfit due to the large number of hyperparameters to estimate. The rest of the methods are conjectured to perform better, due to a lack of overfitting. In scenario three, we expect all methods to perform badly due to the high correlation between the features. In scenario three

	In all scenarios we created a training set of $n=100$ samples to estimate the models and a test set of $n_{\text{test}}=1000$ samples on which we compute the performance measures. The features $\x_i$, $i=1, \dots, n$, are sampled from a $p$-dimensional Gaussian distribution centred at zero, where we set $p=1000$. We let the covariance matrix of the features vary over the scenarios. Likewise for the model parameters $\beta_j$, $j=1, \dots, p$. We simulate the outcome data $y_i$, $i=1, \dots, n$, from a binary logistic model: $y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence of random variation in the simulations, we repeated every scenario 100 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.

	 The results of this simulation are presented in Figures \ref{fig:sim1_performance1_set1}-\ref{fig:sim1_performance1_set3}.
	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set1.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 1.}
		\label{fig:sim1_performance1_set1}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set2.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 1.}
		\label{fig:sim1_performance1_set2}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set3.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 1.}
		\label{fig:sim1_performance1_set3}
	\end{figure}

	The results of this simulation 1 are presented in Figures \ref{fig:sim4_performance1_set1}-\ref{fig:sim4_performance1_set3}.
	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set1.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 4.}
		\label{fig:sim4_performance1_set1}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set2.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 4.}
		\label{fig:sim4_performance1_set2}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set3.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 4.}
		\label{fig:sim4_performance1_set3}
	\end{figure}

	\subsection{Results}\label{sec:results}

	\section{Application}\label{sec:app}

	\section{Discussion}\label{sec:discussion}
	In a taxonomy of Bayesian methods, the proposed method may be considered a local shrinkage model, as opposed to the global-local shrinkage priors that \cite{polson_local_2012,bernardo_shrink_2011} discuss. They characterise certain desirable properties of these global-local shrinkage priors in high dimensions, which, for example, the horseshoe possesses \cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}. In our case, global shrinkage would imply adding another hyperprior for the global $\lambda_1$ and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters. We argue however, that if the groups are informative, the empirical Bayes estimation of the (semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than full Bayes shrinkage of the global penalty parameters, because the latter does not use any known structure to model the variability in the hyperparameters. Nonetheless, an interesting direction of future research is the extension of the group-regularized elastic net to a group-regularized horseshoe model, since the horseshoe has been shown to handle sparsity well and render better coverage of credibility intervals than lasso-type priors \cite[]{van_der_pas_horseshoe_2014}.

	Although our method can be considered weakly adaptive, it is different from the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net \cite[]{zou_adaptive_2009} in the sense that it adapts to external information rather than to the data. It also differs in the scale of adaptation: in the adaptive lasso and elastic net the adaptive weights are feature specific, while in our case they are estimated on the group level, rendering the adaptation more robust. As can be seen from the simulations and real data example in Sections \ref{sec:simulations} and \ref{sec:app}, adaptation to external data may be beneficial for prediction and feature selection. We believe that this is due to the `borrowing of information' effect: estimates that are believed to be similar are shrunken in the same way, yielding overall, better estimates.

	Another obvious comparison is to the group lasso \cite[]{meier_group_2008,yuan_model_2006}. Although it is similar in the sense that it shrinks on the group level, the group lasso is built upon an entirely different philosophy: it sets whole groups of features to zero. The intended application of such a group-wise penalty is to small interpretable groups of features, like, for example, dummies of a categorical variable. Another difference between \texttt{gren} and the group lasso is the number of penalty parameters. \texttt{gren} estimates one parameter per group, while the group lasso estimates one overall penalty parameter; it is thereby less flexible in differential shrinkage of the parameters. In addition, in the simulations and data applications of \cite{van_de_wiel_better_2016}, group lasso prediction performed inferior to \texttt{GRridge} prediction.

	A possible weak point of the proposed method is the double EM loop. The double loop increases the chance of ending up in a local optimum. In the application discussed above we investigated the occurrence of multiple optima, but never encountered them. This does not guarantee that local optima do not occur, but it provides some evidence that local optima are not ubiquitous. Local optima can often be avoided by an informed choice of starting values. In our experience, reasonable starting values are obtained by running a group-regularized ridge regression and using the estimated penalty multipliers as starting values for \texttt{gren}. In addition, we note that even a local optimum might give `good enough' predictions in many practical settings.

	\section*{Software}
	The method is available as an \texttt{R} package from \url{https://CRAN.R-project.org/package=gren}.

	\section*{Supplementary Material}
	Supplementary Material is available online from \url{http://pub.math.leidenuniv.nl/~munchmm/}.

	\section*{Acknowledgements}
	\textit{Conflict of Interest}: None declared.

	\section*{Funding}
	This research has received funding from the European Research Council under ERC Grant Agreement 320637.

	\bibliographystyle{author_short3.bst}
	\bibliography{refs}

\end{document}

### 2019-02-05 docs/manuscript_biostatistics.Rnw
adjusted alzheimer data analysis

--- 2019-02-05 docs/manuscipt.Rnw
@@
-\documentclass[a4paper,hidelinks]{article}
-\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}
+\documentclass[oupdraft]{bio}
+% \usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
+\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable}
 
@@
 
-% settings
-\graphicspath{{../figs/}}
-\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
-\setlength{\textwidth}{15cm}
+\begin{document}
+
+	% Title of paper
 	\title{Adaptive group-regularized logistic elastic net regression}
-\date{}
-\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}
 
-\begin{document}
+	% List of authors, with corresponding author marked by asterisk
+	\author{MAGNUS M. M\"UNCH$^{\ast 1,2}$, CAREL F.W. PEETERS$^{1}$, AAD W. VAN DER VAART$^{2}$, \\ MARK A. VAN DE WIEL$^{1,3}$\\[4pt]
+		% Author addresses
+		\textit{1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB Amsterdam, The Netherlands\\
+			2. Mathematical Institute, Leiden University, Leiden, The Netherlands\\
+			3. Department of Mathematics, VU University, Amsterdam, The Netherlands}
+		\\[2pt]
+		% E-mail address for correspondence
+		{m.munch@vumc.nl}}
+
+	% Running headers of paper:
+	\markboth%
+	% First field is the short list of authors
+	{M. M. M\"unch and others}
+	% Second field is the short title of the paper
+	{Adaptive group-regularized logistic elastic net regression}
 
@@
 
-	\noindent
-	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
-	Amsterdam, The Netherlands\\
-	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
-	3. MRC Biostatistics Unit, Cambridge, United Kingdom \\
+	% Add a footnote for the corresponding author if one has been
+	% identified in the author list
+	\footnotetext{To whom correspondence should be addressed.}
 
@@
 Simulations and an application to a colon cancer microRNA study show that, if the partitioning of the features is informative, classification performance and feature selection are indeed enhanced.}
+		{Empirical Bayes; High-dimensional data; Prediction; Variational Bayes}
 	\end{abstract}
-
-	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes
-
-	\noindent\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}
 
@@
 In this paper, we specifically focus on the latter setting.
-High dimensionality of data introduces several issues in the estimation of prediction models, especially: (a) unidentifiable models, (b) highly correlated predictor variables and (c) non-trivial variable selection.
+High dimensionality of data introduces several issues in the estimation of prediction models, especially: (a) unidentifiable models, (b) highly correlated predictor variables and (c) non-trivial selection of variables.
 
@@
 The elastic net approach takes issues (a)-(c) into account and has yielded many successful extensions.
-Recent work on elastic net regression (and its special cases) has focussed on increasing prediction accuracy by the inclusion of prior knowledge on the variables \cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.
+Recent work on elastic net regression (and its special cases) has focussed on increasing prediction accuracy by the inclusion of prior knowledge on the variables \cite[]{van_de_wiel_better_2016,tai_incorporating_2007}.
 
@@
 The information can, for example, come in the form of (a) results on the same molecular features obtained in a previous study (e.g.
-$p$-values), (b) information from a publicly available database that summarizes the prior information on the molecular features involved (e.g., the Cancer Gene Census \cite[]{futreal_census_2004}), and (c) omics annotation (e.g.
-the location of a gene on the chromosome).
+$p$-values), (b) information from a publicly available database that summarizes the prior information on the molecular features involved (e.g., the Cancer Gene Census \cite[]{futreal_census_2004}), (c) omics annotation (e.g.
+the location of a gene on the chromosome) and (d) response-independent summary statistics, such as sample standard deviation of the features.
 Although this information can rarely be directly included in the statistical analysis, it may still be useful and informative for the study at hand.
 
-	An example of such a study with extra information available is investigated in Section \ref{sec:app}.
-INTRODUCE DATA.
-This problem is not straightforward to solve with the standard methods available in literature.
-At first sight, such an \textit{a priori} grouping of the features suggests the group lasso \cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which penalises and selects features at the group level.
-There are two limitations to the group lasso: (a) it selects groups as a whole instead of single features and (b) it is not adaptive: the amount of penalization a group receives depends on a `global' penalty parameter common to all groups.
-Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
-
-	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter.
+	A natural way of including these information types is through differential penalization: that is, each group of variables receives its own penalty parameter.
 An apparent issue with this differential penalization is the estimation of the penalty parameters.
@@
 Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net.
-Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit.
-Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.
-
-	In line with the above, our proposed method is (a) subjective: the few remaining tuning parameters may be cross-validated, (b) adaptive: prior information is automatically and subjectively weighed with respect to its informativeness, (c) fast compared to full Bayesian analysis or cross-validation, and (d) requires minimal input by the user: only the data and grouping of the features is required.
 
@@
 In Section \ref{seq:estimation} we derive a variational approximation to this model and use this novel approximation in the empirical Bayes estimation of multiple, group-specific penalty parameters.
-In Sections \ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a simulation study and a real data example, respectively.
+In Section \ref{sec:extensions} we introduce some extensions of the method, including, among others, an extension to unpenalized covariates.
+In Sections \ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a simulation study and a microRNA colon cancer data example, respectively.
 In Section \ref{sec:discussion} we conclude by both discussing some differences and parallels between the proposed method, termed \texttt{gren}, and other methods in the literature.
@@
 Although elegant and simple, this method requires a converged MCMC sample from the posterior for every iteration: a computationally intensive procedure.
-\cite{roy_selection_2017} introduce generalised importance sampling for the Bayesian elastic net, such that we just need a limited, pre-specified number of MCMC chains.
-However, this still requires several converged chains, so is not feasible in many high-dimensional omics settings.
 We propose to tackle this problem by approximating the expectation in (\ref{eq:ebEM}) using variational Bayes.
@@
 
+	\section{Extensions}\label{sec:extensions}
+	\subsection{Unpenalized covariates}
+	Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix $\X$ with a column of ones.
+Penalization of such an intercept parameter is not desirable \cite[]{Hastie_Elements_2009}.
+Additionally, it is often desirable to include unpenalized covariates in the model.
+Common examples of such  covariates in clinical research are patient characteristics, such as age, BMI, and sex.
+In the following we assume the intercept to be included in the unpenalized covariates.
+
+	To include unpenalized covariates, we divide the model matrix into two parts $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$ unpenalized variables and $\X_r$ are the $r$ penalized variables.
+Let $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with $u$ zero columns and $u$ zero rows.
+Then we have for the current estimate of the covariance matrix $\bSigma$:
+	\begin{align}\label{eq:unpencov}
+	\bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
+	& = 
+	\begin{bmatrix} 
+	\X_u \tr \bOmega \X_u & \X_u \tr \bOmega \X_r \\
+	\X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
+	\end{bmatrix}^{-1}.
+	\end{align}
+	With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension $p \times p$.
+Inversion of this matrix is done efficiently by applying the Woodbury identity.
+
+	\subsection{Monotonicity of the penalty parameters}
+	Enforcing monotonicity of the penalty multipliers is desirable in some settings.
+An example of such a setting is if we have $p$-values from a previous, related study available.
+A $p$-values-based partitioning of the variables may, \textit{a priori}, be expected to render penalty multipliers that increase monotonically with $p$-value.
+That is, larger $p$-values may be expected to yield at least as large penalty multipliers as variables with smaller $p$-values.
+We propose to include this \textit{a priori} assumption by requiring the penalty multipliers to increase monotonically with $p$-value, thereby also stabilizing their estimates.
+
+	A natural way of enforcing monotonicity is to extend the constraint in (\ref{eq:mmlupdateequation}) with $\lambda'_1 \leq \dots \leq \lambda'_G$.
+The problem is still convex and may be numerically solved.
+From experience, however, we note that in combination with this constrained optimisation, the EM algorithm described in Section~\ref{seq:empiricalbayes} often converges to a local optimum close to the initialisation.
+We therefore enforce monotonicity through a \textit{post hoc} isotonic regression on the penalty multipliers after every optimisation step.
+
 	\subsection{Feature selection}
@@
 
+	\subsection{Multiple partitions}
+	In many cases the features may be partitioned in more than one way.
+For example, we may have both information on annotation and $p$-values from a previous study available.
+A naive way of incorporating multiple partitions is to cross-tabulate the partitions and create a separate group for every combination.
+This poses two problems: (i) The number of penalty parameters increases exponentially with the number of partitions and (ii) some of these combinations may contain only few features, so that the estimation procedure becomes unstable.
+We propose to stabilise the procedure and keep the number of parameters to estimate manageable by modelling the penalty parameters multiplicatively.
+
+	We describe our implementation here for two partitions of the features.
+To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively.
+Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively.
+In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively.
+The generalised frequentist elastic net estimator and corresponding conditional prior are now:
+	\begin{subequations}\label{eq:multpart}
+		\begin{align}
+		\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
+		\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
+		\end{align}
+	\end{subequations}
+	After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
+	\begin{subequations}\label{eq:estmultpart}
+		\begin{alignat}{2}
+		\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
+		\end{alignat}
+	\end{subequations}
+	with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$.
+The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine.
+This method naturally generalises to more than two partitions of the data.
+
+	In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers in a multiplicative way, mainly for convenience.
+First, there are computational reasons: The separation of the square root in the $L_1$-norm penalty term into two terms facilitates the numerical estimation.
+Additionally, multiplicative modelling of the partitions allows for more flexibility in the estimates.
+To see this consider the obvious alternative: additive partition-specific multipliers.
+Since the penalty multipliers are strictly positive, the constraint on the geometric mean makes one large additive penalty estimate difficult to compensate with a smaller estimate, thereby impairing flexibility of the model.
+Large multiplicative penalty multipliers are easier to compensate for by smaller penalty multipliers, rendering this the less restrictive option for modelling the partitions.
+
 	\section{Simulations}\label{sec:simulations}
 	\subsection{Setup}\label{sec:setup}
-	To assess variable selection and predictive performance we conduct a simulation study in which we compare \texttt{gren} to the regular elastic net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, group mimimax concave penalty (\texttt{grMCP}) \cite[]{breheny_penalized_2009}, and the group exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}.
+	To assess variable selection and predictive performance we conduct a simulation study in which we compare \texttt{gren} to the regular elastic net and ridge models, and \texttt{GRridge} \cite[]{van_de_wiel_better_2016}.
 \texttt{GRridge} is similar to \texttt{gren} in the sense that it estimates group-specific penalty multipliers.
 The two main differences with \texttt{gren} are (i) the absence of an $L_1$-norm penalty and (ii) the estimation procedure.
-The other methods are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} and not adaptive on the group level.
-In contrast to the original group lasso, all of them are able to select single features, instead of just on the group level.
+For the regular elastic net and \texttt{gren}, we fix the proportion of $L_1$-norm penalty to three different values: $\alpha \in \{0.05, 0.5, 0.95 \}$.
+The first setting closely resembles the ridge setting, where $\alpha=0$, the third one is similar to the lasso with $\alpha=1$.
+For the methods that do automatic variable selection (the regular elastic net and \texttt{gren}), we estimate the models for a range of model sizes.
+Likewise for \texttt{GRridge}, where we use the \textit{post hoc} variable selection method described by \cite{novianti_better_2017}.
 
-	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, we fix the proportion of $L_1$-norm penalty to three different values: $\alpha \in \{0.05, 0.5, 0.95 \}$.
-The first setting closely resembles the ridge setting, where $\alpha=0$, while the third setting is similar to the lasso with $\alpha=1$.
-For the methods that do automatic variable selection (all methods except the regular ridge and \texttt{GRridge}), we estimate the models for a range of model sizes.
+	The quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960}.
+This kappa is measures similarity between two sets of categorical values.
+One set contains the feature selection indicators: $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
+We calculate Cohen's $\kappa$ as:
+	\begin{align*}
+	\kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}\sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
+	\end{align*}
+	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and expected correctly identified features, respectively.
+A positive or negative kappa indicate better and worse feature selection than expected by chance, respectively.
+In addition, we evaluate estimation accuracy by mean squared error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
 
@@
 The BSS measures the fraction of Brier score improvement compared to a non-informative model.
-Quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960} and parameter estimation by mean squared error (MSE).
-Both are presented in SM Section 10.
 
-	We simulated data according to five different scenarios: 
-	\begin{enumerate}
-		\item differential signal between the groups and uniformly distributed model parameters,
-		\item a large number of small groups of features,
-		\item no differential signal between the groups, but strong correlations within groups of features, 
-		\item differential signal between the groups and heavy-tailed distributed model parameters, and
-		\item a very sparse setting with no signal in some of the groups.
-	\end{enumerate}
-	Full descriptions of the scenarios are given in Section 10 of the SM.
-In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform.
-In the second scenario, we expect \texttt{gren} and \textt{GRridge} to overfit due to the large number of hyperparameters to estimate.
-The rest of the methods are conjectured to perform better, due to a lack of overfitting.
-In scenario three, we expect all methods to perform badly due to the high correlation between the features.
-In scenario three
+	We created a training set of $n=100$ samples to estimate the models and a test set of $n_{\text{test}}=1000$ samples on which we compute the performance measures.
+The features $\x_i$, $i=1, \dots, n$, are sampled from a $p$-dimensional Gaussian distribution centred at zero, where we set $p=1000$.
+We introduce correlation between the features through a block diagonal covariance matrix for the $\x_i$.
+The blocks are of sizes $25 \times 25$, with off-diagonals set to $\rho=0.7$ and diagonals to $\sigma^2 = 1$.
 
-	In all scenarios we created a training set of $n=100$ samples to estimate the models and a test set of $n_{\text{test}}=1000$ samples on which we compute the performance measures.
-The features $\x_i$, $i=1, \dots, n$, are sampled from a $p$-dimensional Gaussian distribution centred at zero, where we set $p=1000$.
-We let the covariance matrix of the features vary over the scenarios.
-Likewise for the model parameters $\beta_j$, $j=1, \dots, p$.
-We simulate the outcome data $y_i$, $i=1, \dots, n$, from a binary logistic model: $y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$.
-To mitigate the influence of random variation in the simulations, we repeated every scenario 100 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.
+	The model parameters $\beta_j$, $j=1, \dots, p$, are simulated in four groups of 250 with differing signal strengths and are created in two steps.
+First, we draw 250 parameters from the elastic net distribution, parametrised as in \cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$ in all groups, and different penalty multipliers, $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$.
+The second step consists of setting the (in absolute value) smallest 125 parameters in a group to zero.
+This results in a total of 500 zero and 500 non-zero parameters, evenly distributed over the four groups.
 
-	 The results of this simulation are presented in Figures \ref{fig:sim1_performance1_set1}-\ref{fig:sim1_performance1_set3}.
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set1.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set1}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set2.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set2}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set3.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set3}
-	\end{figure}
-
-	The results of this simulation 1 are presented in Figures \ref{fig:sim4_performance1_set1}-\ref{fig:sim4_performance1_set3}.
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set1.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set1}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set2.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set2}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set3.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set3}
-	\end{figure}
+	Finally, we simulate the outcome data $y_i$, $i=1, \dots, n$, from a binary logistic model: $y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$.
+To mitigate the influence of random variation in the simulations, we repeated every simulation 50 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.
 
 	\subsection{Results}\label{sec:results}
+	For all models, the estimated penalty parameters follow the expected pattern (Figure \ref{fig:sim1_penalties}); roughly, the penalty multipliers increase with decreasing signal strength.
+\texttt{GRridge}'s pattern of penalty multipliers is more pronounced than that of \texttt{gren}.
+However, both methods show difficulty in distinguishing between the two lower signal groups.
+In addition, we note that the \texttt{GRridge} penalty multipliers estimates show a lot more variation than \texttt{gren} in the three lower signal groups.
 
-	\section{Application}\label{sec:app}
+	For all performance measures, the group-regularized methods outperform the non group-regularized ones (Figure \ref{fig:sim1_performace}).
+Cohen's kappa shows that \texttt{gren} with $\alpha=0.05$ generally outperforms the other methods in terms of variable selection.
+\texttt{gren} with $\alpha=0.05$ is, however, less able to reconstruct the coefficients, evident from the generally higher MSE.
+In addition, SM Section 9 shows that \texttt{gren} strongly improves feature selection stability, compared to the regular elastic net.
+
+	For predictive performance, competitiveness of \texttt{gren} depends on the range of model sizes.
+For the smaller models, the \texttt{gren} models with $\alpha \in \{ 0.5, 0.95\}$ tend to outperform the other methods, while for the larger models, \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ tend to perform better.
+All elastic net methods seem to deteriorate in performance after a certain model size, because of the selection of `noisy' features after (most of) the predictive features have been selected.
+The more lasso-like models (with $\alpha \in \{0.5, 0.95\}$) prefer smaller models compared to \texttt{GRridge} and the elastic net with $\alpha=0.05$, a tendency already noted in the original elastic net paper \cite[]{zou_regularization_2005}.
+In all, when \texttt{gren} does not outperform the other methods, it's performance is on a par.
+
+	\section{Application to microRNAs in colorectal cancer}\label{sec:app}
+	\subsection{Partitioning based on differential expression}\label{sec:mirna}
+	The example data set is from a deep sequencing analysis on microRNA \cite[]{neerincx_combination_2018}.
+The study was done in 88 treated colorectal cancer patients, with the aim of classifying treatment response, coded as either non-progressive/remission (70 patients) or progressive (18 patients).
+After pre-processing and normalisation, 2114 microRNAs remained.
+In addition to the 2114 microRNAs we incorporated 4 unpenalized clinical covariates into the analysis: prior use of adjuvant therapy (binary), the type of systemic treatment regimen (ternary), age, and primary tumor differentiation (binary).
+
+	In a preliminary experiment on different subjects, the microRNA expression levels of metastatic colorectal tumour tissue were compared to normal non-colorectal tissue and primary colorectal tumour tissue was compared to primary colorectal tumour tissue \cite[]{neerincx_mir_2015}.
+This yielded 221 microRNAs that were differentially expressed in both comparisons ($\text{FDR} \leq 0.05$), versus 1893 not differentially expressed microRNAs.
+We expect that incorporation of this partitioning enhances therapy response classification, because tumor-specific miRNAs might be more relevant than non-specific ones.
+In addition, we divided the differentially expressed microRNAs even further into 127 highly differentially expressed microRNAs ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs ($0.001 < \text{FDR} \leq 0.05$).
+We will refer to this second partitioning as the three-group setting, as opposed to the first, the two-group setting.
+
+	We compared \texttt{gren} to ridge and elastic net regression, and \texttt{GRridge}.
+For the elastic net methods (including \texttt{gren}) we set $\alpha \in \{ 0.05, 0.5, 0.95\}$.
+The estimated penalty multipliers are according to expectation: in all group-regularized methods, the 221 differentially expressed microRNAs receive the smallest penalty (Figure \ref{fig:col_bar}a).
+In the three-group setting the pattern is again as expected (Figure \ref{fig:col_bar}b): the highly expressed group, receives the lowest penalty, followed by the medium expressed group, while the non-expressed group receives the strongest penalty.
+\texttt{GRridge} is not able to distinguish between the medium and non-differentially expressed groups of microRNAs.
+
+	Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}.
+All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS.
+In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC.
+This is especially true for the larger models.
+\texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models.
+A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly.
+This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation.
+In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used.
+Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting.
+Note that models with few features are often desirable for clinical implementation.
+
+	\subsection{Random groups}
+	Considering that the features may be partitioned into many groups, with one parameter per group, we have to be aware of overfitting risks.
+We investigated this using the data introduced in Section \ref{sec:mirna}, randomly dividing the features into three groups.
+We fixed the group sizes to the group sizes used in the three-group setting in Section \ref{sec:mirna}.
+Under this random partitioning of the features, we expect all penalty multipliers to be estimated as one if overfitting does not occur.
+
+	We compared the estimated multipliers to the estimates by \texttt{GRridge}.
+Since these results depend on one specific randomisation of the groups, we repeated the procedure 100 times and present the results in Figure \ref{fig:random}.
+From this figure we see that the estimates for \texttt{gren} are close to one.
+The estimates by \texttt{GRridge} show much more variation.
+Additionally, \texttt{GRridge} gives slightly biased estimates: the median penalty parameter estimates are 1.16, 1, and 0.85 for the three groups.
+In contrast, the median estimates of \texttt{gren} are 1, 1, and 0.99.
+We repeated the simulation with ten evenly sized groups and present the results in SM Section 9.
 
@@
 	\section*{Software}
-	The method is available as an \texttt{R} package from \url{https://CRAN.R-project.org/package=gren}.
+	The method is available as an \texttt{R} package from \url{https://github.com/magnusmunch/gren/}.
 
 	\section*{Supplementary Material}
-	Supplementary Material is available online from \url{http://pub.math.leidenuniv.nl/~munchmm/}.
+	Supplementary Material is available online from
+	\url{http://biostatistics.oxfordjournals.org}.
 
@@
 
-	\bibliographystyle{author_short3.bst}
+	\bibliographystyle{biorefs}
 	\bibliography{refs}
 
+	\begin{figure}[!p]
+		\centering
+		\includegraphics[width=0.8\linewidth]{../figs/grEBEN_sim1_res1_penalties.png}
+		\caption{Estimated penalty multipliers of the group-regularized methods based on simulated data.}
+		\label{fig:sim1_penalties}
+	\end{figure}
+
+	\begin{figure}[!p]
+		\centering
+		\includegraphics[width=1\linewidth]{../figs/grEBEN_sim1_res1_performance.png}
+		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for (group-regularized) ridge and (group-regularized) elastic net with $\alpha \in \{ 0.05, 0.5, 0.95\}$ based on simulated data.}
+		\label{fig:sim1_performace}
+	\end{figure}
+
+	\begin{figure}[!p]
+		\centering
+		\includegraphics[width=0.83\linewidth]{../figs/grEBEN_mirseq_Maarten_bar1.png}
+		\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
+		\label{fig:col_bar}
+	\end{figure}
+
+	\begin{figure}[!p]
+		\centering
+		\includegraphics[width=1\linewidth]{../figs/grEBEN_mirseq_Maarten_combined_performance.png}
+		\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
+		\label{fig:col_per}
+	\end{figure}
+
+	\begin{figure}[!p]
+		\centering
+		\includegraphics[width=0.5\linewidth]{../figs/grEBEN_mirseq_Maarten_res6_boxplot.png}
+		\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 3 random groups in colorectal miRNA data.}
+		\label{fig:random}
+	\end{figure}
+
 \end{document}

### 2019-02-05 docs/supplement.Rnw
adjusted alzheimer data analysis

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\card}[1]{\text{card} \left( #1 \right)}
\makeatletter
\newcommand*{\defeq}{\mathrel{\rlap{%
			\raisebox{0.3ex}{$\m@th\cdot$}}%
		\raisebox{-0.3ex}{$\m@th\cdot$}}%
	=}
\makeatother

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0pt}
\title{Adaptive group-regularized logistic elastic net regression Supplementary material}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}

\begin{document}
\maketitle
\noindent
1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
Amsterdam, The Netherlands\\
2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\

\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes

\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}

\section{Introduction}
This document contains supplementary material to the paper `Adaptive group-regularized logistic elastic net regression'. Section \ref{sec:enetprior} contains extra details on the elastic net prior distribution. Section \ref{sec:objectivefunction} gives the objective function that is maximised in the variational Bayes procedure introduced in Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main Document. The derivations of the variational posterior used in the Main Document are shown in Section \ref{sec:variationaldistributions}. Section \ref{sec:computationalcomplexity} shows how we may reduce the computational complexity of our variational Bayes procedure. Section \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected joint log likelihood that is maximised to estimate the penalty multipliers as explained in Section 3 of the Main Document. The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}. We apply our method to a study on microRNAs in colorectal cancer data \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} . Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.

\section{The elastic net prior}\label{sec:enetprior}
\subsection{Density function}
The density function of $\beta$, drawn from an elastic net prior distribution, is given by:
$$
f(\beta) = g(\lambda_1,\lambda_2) \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\].
$$
The normalizing constant here is calculated as:
$$
g(\lambda_1,\lambda_2) =\left\{ \int_{-\infty}^{\infty} \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta \right\}^{-1}= \frac{\sqrt{\lambda_2}}{2} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1},
$$
with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution functions of $x$, respectively.

\subsection{Expectation and variance}
The expectation and variance of $\beta$ are:
\begin{align*}
\E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
\V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta - \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} - \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
\end{align*}

\subsection{Simulation}
Generating samples from the elastic net prior is not straightforward. \cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$: 
\begin{align*}
\beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} \frac{\tau - 1}{\tau} \right), \\
\tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right),
\end{align*}
where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with an efficient truncated gamma sampler available we may sample from the prior using any standard software package that includes a Gaussian sampling routine. More specifically, we require a a sample from the $\mathcal{TG} \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ distribution. The CDF of this distribution is given by:
$$
F_{T}(\tau) =	
\begin{dcases}
0, & \text{if }\tau < 1 \\
1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}{\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
\end{dcases}
$$

Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for $\tau \geq 1$, the quantile function may be written as a function of the probability $p \in (0,1)$:
$$
Q_{T}(p) =	F^{-1}_{T}(p) = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - p) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2,
$$
where $\Phi^{-1}(p)$ is the quantile function of the standard normal distribution, i.e, the probit function. Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now:
\begin{algorithmic}
	\State{Generate $U \sim \mathcal{U} (0,1)$}
	\State{Set $\tau = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - u) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2$}
	\State{Set $\nu^2= \frac{\tau - 1}{\tau \lambda_2}$}
	\State{Generate $\beta \sim \mathcal{N}(0,\nu^2)$}
	\State{\Return{$\beta$}}
\end{algorithmic}

\section{Objective function}\label{sec:objectivefunction}
The lower bound on the marginal likelihood, or evidence lower bound (ELBO), at iteration $t$ is given by:
\begin{subequations}
	\begin{align}
	\text{ELBO} (Q^{(t)}) & = \E_{Q^{(t)}} [\log \L (\y, \bomega, \bbeta, \btau)] - \E_{Q^{(t)}} [\log Q (\bomega, \bbeta, \btau)] \\
	& = \E_{Q^{(t)}} [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \label{eq:elboomega} \\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\bbeta|\btau)] - \E_{Q^{(t)}} [\log q(\bbeta)] \label{eq:elbobeta}\\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\btau)] - \E_{Q^{(t)}} [\log q(\btau)] \label{eq:elbotau}.
	\end{align}
\end{subequations}

We start with the computation of (\ref{eq:elboomega}). To this end we define:
$$
h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}\frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}\exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
$$
Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, in the following, all expectations and variances are with respect to $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):
\begin{subequations} \label{eq:elboomega2}
	\begin{align}
	\E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\}, \\
	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] \E [\omega_i] \\
	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
	\E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n c_i^2 \E [\omega_i] \\
	& \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
	\end{align}
\end{subequations}
where all mathematical operations on vectors and matrices are element wise. After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
\begin{subequations} \label{eq:elboomega3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
	\end{align}
\end{subequations}

Inspection of the updating equations in the Main Document, learns us that the last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that we are left with just (\ref{eq:elboomega4}). After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
\begin{subequations} \label{eq:elbobeta2}
	\begin{align}
	\E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - \frac{p}{2} \log (2 \pi) \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) (\bmu_j^2 + \bSigma_{jj}) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
	\E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - \frac{p}{2} \log |\bSigma| .
	\end{align}
\end{subequations}
We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):
\begin{subequations} \label{eq:elbobeta3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} [\log q (\bbeta )] \\ 
	& = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + \frac{p}{2} \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
	\end{align}
\end{subequations}

Repeating the same exercise for (\ref{eq:elbotau}) gives:
\begin{subequations} \label{eq:elbotau2}
	\begin{align}
	\E [\log \pi (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{5 p }{2} \log 2 - \frac{p}{2} \log \pi - \frac{p}{2} - \frac{p \lambda_1^2}{8 \lambda_2} \\
	& \,\,\,\,\, - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) - \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} - \frac{1}{2} \sum_{j=1}^p \E [\log ( \psi_j + 1)], \\
	\E [\log q (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{3p}{2} \log 2 -\frac{p}{2} \log \pi -\frac{p}{2} - \frac{\lambda_1}{2 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} \\
	& \,\,\,\,\, + \frac{1}{4} \sum_{j=1}^p \log \chi_j - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j].
	\end{align}
\end{subequations}
Again, we combine the terms in (\ref{eq:elbotau2}) to get:
\begin{subequations} \label{eq:elbotau3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\btau)] - \E_{Q^{(t)}} [\log q (\btau)] 
	\\ & = p \log 2 - \frac{p \lambda_1^2}{8 \lambda_2} - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)} \\
	& \,\,\,\,\, + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log ( \psi_j + 1)].
	\end{align}
\end{subequations}

Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
\begin{align*}
\text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2} \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
& + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
& + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
\end{align*}
Here, proportionality is with respect to the variational and penalty parameters.

\section{The P\'{o}lya-Gamma parametrisation}\label{sec:polyagamma}
In Section 3 of the Main Document latent P\'{o}lya-Gamma distributed variables are introduced. Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma variables are introduced into the logistic regression model under a Gaussian prior. Here, we explain how introduction of these latent variables simplifies the variational Bayes calculations.

In \cite{polson_bayesian_2013}, the central identity for this parametrisation is given in their Theorem 1:
\begin{equation}\label{eq:intid}
\frac{\exp(\psi)^{a}}{[1 + \exp(\psi)]^{b}} = 2^{-b} \exp(\kappa \psi) \int_0^{\infty} \exp(-\omega \psi^2 /2) \pi(\omega) d \omega,
\end{equation}
where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed. This holds for all $b > 0$ and $a \in \mathbb{R}$. After setting $a=y_i$, $b=m_i$, and $\psi=\x_i \tr \bbeta$, we may consider the right-hand side of (\ref{eq:intid}) as an un-normalised joint density in $(\bbeta, \omega)$. The likelihood contribution of observation $i$ for $\bbeta$ conditional on $\omega_i$ may now be written as:
$$
\L_i(\bbeta | \omega_i) \propto \exp[\kappa_i \x_i \tr \bbeta - \omega_i (\x_i \tr \bbeta)^2/2].
$$

This allows the conditional posterior of $\bbeta$ to be written as:
$$
p(\bbeta | \bomega, \y) \propto \pi(\bbeta) \prod_{i=1}^n \L_i(\bbeta | \omega_i) = \pi(\bbeta) \exp \[-\frac{1}{2}(\bOmega^{-1} \bkappa - \X \bbeta) \tr \bOmega (\bOmega^{-1} \bkappa - \X \bbeta)\],
$$
where $\bOmega = \diag (\omega_i)$, for $i=1, \dots, n$. Consequently, with $\pi(\bbeta)$ a Gaussian prior, $p(\bbeta | \bomega, \y)$ is also in the Gaussian family.

Furthermore, by treating the integrand in (\ref{eq:intid}) as an un-normalized density over $(\psi, \phi)$, we find the distribution of $\omega$ conditional on $\bbeta$,
$$
p(\omega | \psi) = \frac{\exp(-\omega \psi^2/2) \pi(\omega)}{\int_0^{\infty} \exp(-\omega \psi^2/2) \pi(\omega) d\omega},
$$
distributed as $\omega | \psi \sim \mathcal{PG}(b,\psi)$ \cite[]{polson_bayesian_2013}, with expectation $\E[ \omega ] = \frac{b}{2\psi} \tanh (\psi/2)$. We substitute again $b=m_i$ and $\psi=\x_i \tr \bbeta$, and set the prior $\pi(\bbeta)$ to a Gaussian with mean $\m$ and covariance $\Vm$ to arrive at the following conditional posteriors:
\begin{subequations}\label{eq:gibbssampler}
	\begin{align}
	\omega_i | \bbeta & \sim \mathcal{PG}(m_i, \x_i \tr \bbeta) \\
	\bbeta | \y, \bomega & \sim \N(\bmu, \bSigma),
	\end{align}
\end{subequations}
where
\begin{align*}
\bSigma & = (\X \tr \bOmega \X + \Vm^{-1})^{-1}, \\
\bmu & = \Vm (\X \tr \bkappa + \Vm^{-1} \m),
\end{align*}
with $\bkappa = \begin{bmatrix} \kappa_1 \dots \kappa_n \end{bmatrix} \tr$. Equations (\ref{eq:gibbssampler}) constitute the Gibbs sampler used in \cite{polson_bayesian_2013}. However, from the corresponding variational Bayes derivations in Section \ref{sec:variationaldistributions}, it is obvious that the parametrisation above is also useful in the variational Bayes approximation to the posterior distribution.

\section{Variational posterior}\label{sec:variationaldistributions}
The optimal distributions in our variational Bayes implementation are:
\begin{align}
q^*_{\bbeta} (\bbeta) & {\propto} \exp \{ \E_{\bomega, \btau} [\log p (\bbeta, \bomega, \btau, \y)]\} {\propto} \exp \{\log \L (\y ; \bbeta) + \E_{\bomega} [\log \pi (\bomega | \bbeta)] + \E_{\btau} [ \log \pi (\bbeta | \btau)] \}, \label{eq:qbeta} \\
q^*_{\bomega} (\bomega) & {\propto} \exp \{ \E_{\bbeta, \btau} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi (\bomega | \bbeta)] \},\label{eq:qomega} \\
q^*_{\tau} (\btau) & {\propto} \exp \{ \E_{\bbeta, \bomega} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi ( \bbeta | \btau)] + \log \pi (\btau) \}. \label{eq:qtau}
\end{align}

For convenience we work on the log-scale and occasionally do a change of variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three terms on the right-hand side of (\ref{eq:qbeta}):
\begin{align*}
\log \L (\y ; \bbeta) & = \sum_{i=1}^n \log f (y_i ; \bbeta) \propto \sum_{i=1}^n \log \left\{ \frac{\exp(\x_i \tr \bbeta)^{y_i}}{[1 + \exp(\x_i \tr \bbeta)]^{m_i}} \right\} \\
& = \y \tr \X \bbeta - \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)], \\
\E_{\bomega} [\log \pi (\bomega | \bbeta)] & \propto \sum_{i=1}^n \E_{\omega_i} [\log \pi (\omega_i | \bbeta)] \\ & \propto \sum_{i=1}^n \E_{\omega_i} \( \log \left\{ \cosh\( \frac{\x_i \tr \bbeta}{2}\)^{m_i} \exp \[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i \]\right\} \) \\
& \propto \sum_{i=1}^n m_i \log \cosh \( \frac{\x_i \tr \bbeta }{2} \) - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta \\ & \propto \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)] - \frac{1}{2} \m \tr \X \bbeta - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta, \\
\E_{\btau} [ \log \pi (\bbeta | \btau)] & = \sum_{j=1}^p \E_{\tau_j} [ \log \pi (\beta_j | \tau_j)] \propto \frac{\lambda_2}{2} \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} \E \( \frac{\tau_j}{\tau_j - 1} \) \\
& = \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} [1 + \E ( \psi_j^{-1} )] = - \frac{1}{2} \lambda_2 \bbeta \tr \bLambda' (\I + \Z) \bbeta,
\end{align*}
where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega = \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' = \diag (\lambda'_{g(j)})$. 
Since the difficult $\sum_{i=1}^n m_i \log[1 + \exp(\x_i \tr \bbeta)]$ term appears in both the $\log \L (\y ; \bbeta)$ and $\pi (\omega_i | \bbeta)$ part, it cancels, thereby justifying the introduction of the $\omega_i$. Finally, we have: 
$$
\log q^*_{\bbeta} (\bbeta) = \bkappa \tr \X \bbeta - \frac{1}{2} \bbeta \tr (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z) \bbeta,
$$
which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma \X \tr \bkappa$.

Next we derive the variational posterior for $\bomega$ in (\ref{eq:qomega}):
\begin{align*}
\log q^*_{\bomega} (\bomega) & = \E_{\bbeta} [\log \pi (\bomega | \bbeta)] = \sum_{i=1}^n \E_{\bbeta} [\log \pi (\omega_i | \bbeta)] \\ & = \sum_{i=1}^n \E_{\bbeta} \( m_i \log \left\{ \cosh \( \frac{\x_i \tr \bbeta}{2}\) \exp\[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i\] \cdot f(\omega_i | m_i, 0) \right\} \) \\ 
& = \sum_{i=1}^n \E_{\bbeta} \[ m_i \log \cosh \( \frac{\x_i \tr \bbeta}{2}\) \] - \frac{1}{2} \sum_{i=1}^n \omega_i \E[ (\x_i \tr \bbeta)^2] + \sum_{i=1}^n \log f(\omega_i | m_i, 0) \\ 
& \propto \sum_{i=1}^n m_i \log \cosh \left\{ \frac{\sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}}{2}\right\} - \frac{1}{2} \sum_{i=1}^n \omega_i \left\{ [\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \right\} \\ 
& \,\,\, + \sum_{i=1}^n \log f(\omega_i | m_i, 0),
\end{align*}
where proportionality is with respect to $\bomega$. If exponentiated, this is a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.

The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated as:
\begin{align*}
\E_{\bbeta} [\log \pi ( \bbeta | \btau)] & = \sum_{j=1}^p \E_{\bbeta_j} [\log \pi(\beta_j | \tau_j )] \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}, \\
\log \pi (\btau) & = \sum_{j=1}^p \log \pi (\tau_j) \propto - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j. 
\end{align*}
Combining the terms we find the log-kernel of the variational posterior as:
\begin{align*}
\log q^*_{\tau} (\btau) & \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}  - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j \\
& = \frac{1}{2} \sum_{j=1}^p \log [(\tau_j - 1)^{-1}] - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2}\tau_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}\),
\end{align*}
where proportionality is with respect to $\btau$. Again applying the change of variables $\psi_j = \tau_j - 1$, we have:
$$
\log q^*_{\psi} (\psi) \propto \frac{1}{2} \sum_{j=1}^p \log \psi_j^{-1} - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2} \psi_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \psi_j^{-1}\),
$$
which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$ random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + \V(\beta_j)]$.

\section{Computational complexity}\label{sec:computationalcomplexity}
Calculation of $\bSigma$ involves the inversion of a $(p \times p)$-dimensional matrix, which occurs at every iteration of the algorithm. If naively done, every inversion is of computational complexity $\mathcal{O}(p^3)$. With $p$ large, this is a serious computational bottleneck of the algorithm, especially with possibly many iterations. However, inspection of the updating equations in (3.9) in the main document, reveals that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the $c_i$ and $\chi_j$. In Section 2.2 of the main document we show that for the EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, and the diagonal of $\bSigma$. Here we provide details on how to compute these with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the computational cost of the algorithm. If necessary, we may calculate the full $\bSigma$ once after the algorithm has converged.

In the following we slightly change notation and let $\diag(\A)$, with $\A$ a $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$ to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$. We start by calculating the complexity of $\diag(\bSigma)$. To this end, let $\H = \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z$. Now, by the Woodbury identity, we may decompose $\diag(\bSigma)$ as:
\begin{subequations}\label{eq:diagsigma}
	\begin{align}
	\diag(\bSigma) ~& = \diag(\H^{-1}) - \diag[\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1}] \\
	& = \diag(\H^{-1}) - [(\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1}) \circ \H^{-1} \X \tr] \cdot \ones_{n \times 1},
	\end{align}
\end{subequations}
where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is a $p \times n$-dimensional matrix of ones. By recognising that $\H$ is a diagonal matrix, we can post-multiply it by a $p \times n$-dimensional matrix with computational complexity $\mathcal{O}(np)$, which leads to a total complexity of $\mathcal{O}(n^2p)$ for $\diag(\bSigma)$. Next we consider the calculation of $\bmu$. Again using the Woodbury identity, we may write $\bmu$ as:
\begin{align}\label{eq:mu}
\bmu = \bSigma \X \tr \bkappa = [\H^{-1} \X \tr - \H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1} \X \tr] \bkappa,
\end{align}
which is, by a similar argument as before, again an operation of complexity $\mathcal{O}(n^2p)$.

Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$ as:
\begin{align*}
\c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
& = \sqrt{\diag \[\X \bSigma \X \tr\] +  \diag \[\X \bmu \bmu \tr \X \tr\]} \\
& = \sqrt{ [(\X \bSigma) \circ \X ] \cdot \ones_{p \times 1} + (\X \bmu) \circ (\X \bmu)}.
\end{align*}
The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$, Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}(n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.

The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p \end{bmatrix} \tr$ are rewritten as:
\begin{subequations}\label{eq:chicomp}
	\begin{align}
	\bm{\chi} & = \begin{bmatrix} \lambda'_{g(1)} \lambda_2 (\bSigma_{11} + \bmu_1^2) & \cdots & \lambda'_{g(p)} \lambda_2 (\bSigma_{pp} + \bmu_p^2) \end{bmatrix} \tr\\
	& = \lambda_2 \bLambda' \[ \diag (\bSigma ) + \diag (\bmu \bmu \tr )\] \\
	& = \lambda_2 \blambda' \circ \[ \diag (\bSigma) + \bmu \circ \bmu \].
	\end{align}
\end{subequations}
 By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we see that the computation of $\bm{\chi}$ consists of one matrix diagonal of $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two Hadamard products of $\mathcal{O}(p)$. Computation of $\bm{\chi}$ is therefore of complexity of $\mathcal{O}(n^2p)$.

\section{Variational Bayes expected joint log likelihood}\label{sec:expectedloglikelihood}
Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then the variational Bayes expected joint log likelihood is derived as follows:
\begin{align*}
\E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ] &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}] \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}(\beta_j | \tau_j) | \blambda'^{(k)}] \\
& \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\ 
& \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
\end{align*}
with the coefficients $d_g^{(k)}$ calculated as:
\begin{align*}
d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\
& = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} ( \beta_j | \blambda'^{(k)} ) \] \\
& = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
\end{align*}

\section{Full estimation procedure}\label{sec:fullestimationprocedure}
The expectation and maximisation steps introduced in the Main Document result in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
\begin{algorithm}
	\caption{Group-regularized empirical Bayes elastic net}\label{alg:summaryalg}
	\begin{algorithmic}[h]
		\Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
		\Ensure $\lambda, \blambda', \bSigma, \bmu$
		\State{Estimate $\lambda$ by cross-validation of the regular elastic net model}
		\While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| > \epsilon_1$}
		\While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
		\State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ for $j=1, \dots, p$ using (3.9) in the Main Document}
		\EndWhile
		\State{Update $\blambda'$ by (3.10) in the Main Document}
		\EndWhile
	\end{algorithmic}
\end{algorithm}

\section{Extensions}\label{sec:extensions}
\subsection{Unpenalized covariates}
Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix $\X$ with a column of ones. Penalization of such an intercept parameter is not desirable \cite[]{Hastie_Elements_2009}. Additionally, it is often desirable to include unpenalized covariates in the model. Common examples of such  covariates in clinical research are patient characteristics, such as age, BMI, and sex. In the following we assume the intercept to be included in the unpenalized covariates.

To include unpenalized covariates, we divide the model matrix into two parts $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$ unpenalized variables and $\X_r$ are the $r$ penalized variables. Let $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with $u$ zero columns and $u$ zero rows. Then we have for the current estimate of the covariance matrix $\bSigma$:
\begin{align}\label{eq:unpencov}
\bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
& = 
\begin{bmatrix} 
\X_u \tr \bOmega \X_u & \X_u \tr \bOmega \X_r \\
\X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
\end{bmatrix}^{-1}.
\end{align}
With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension $p \times p$. Inversion of this matrix is done efficiently by applying the Woodbury identity.

\subsection{Monotonicity of the penalty parameters}
Enforcing monotonicity of the penalty multipliers is desirable in some settings. An example of such a setting is if we have $p$-values from a previous, related study available. A $p$-values-based partitioning of the variables may, \textit{a priori}, be expected to render penalty multipliers that increase monotonically with $p$-value. That is, larger $p$-values may be expected to yield at least as large penalty multipliers as variables with smaller $p$-values. We propose to include this \textit{a priori} assumption by requiring the penalty multipliers to increase monotonically with $p$-value, thereby also stabilizing their estimates.

A natural way of enforcing monotonicity is to extend the constraint in (10) in the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$. The problem is still convex and may be numerically solved. From experience, however, we note that in combination with this constrained optimisation, the EM algorithm described in Section~3 in the Main Document often converges to a local optimum close to the initialisation. We therefore enforce monotonicity through a \textit{post hoc} isotonic regression on the penalty multipliers after every optimisation step.

\subsection{Multiple partitions}
In many cases the features may be partitioned in more than one way. For example, we may have both information on annotation and $p$-values from a previous study available. A naive way of incorporating multiple partitions is to cross-tabulate the partitions and create a separate group for every combination. This poses two problems: (i) The number of penalty parameters increases exponentially with the number of partitions and (ii) some of these combinations may contain only few features, so that the estimation procedure becomes unstable. We propose to stabilise the procedure and keep the number of parameters to estimate manageable by modelling the penalty parameters multiplicatively.

We describe our implementation here for two partitions of the features. To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively. Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively. In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively. The generalised frequentist elastic net estimator and corresponding conditional prior are now:
\begin{subequations}\label{eq:multpart}
	\begin{align}
	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
	\end{align}
\end{subequations}
After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
\begin{subequations}\label{eq:estmultpart}
	\begin{alignat}{2}
	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
	\end{alignat}
\end{subequations}
with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine. This method naturally generalises to more than two partitions of the data.

In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers in a multiplicative way, mainly for convenience. First, there are computational reasons: The separation of the square root in the $L_1$-norm penalty term into two terms facilitates the numerical estimation. Additionally, multiplicative modelling of the partitions allows for more flexibility in the estimates. To see this consider the obvious alternative: additive partition-specific multipliers. Since the penalty multipliers are strictly positive, the constraint on the geometric mean makes one large additive penalty estimate difficult to compensate with a smaller estimate, thereby impairing flexibility of the model. Large multiplicative penalty multipliers are easier to compensate for by smaller penalty multipliers, rendering this the less restrictive option for modelling the partitions.

\section{Simulations}
The quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960}. This kappa is measures similarity between two sets of categorical values. One set contains the feature selection indicators: $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$. We calculate Cohen's $\kappa$ as:
	\begin{align*}
	\kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}\sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
	\end{align*}
	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and expected correctly identified features, respectively. A positive or negative kappa indicate better and worse feature selection than expected by chance, respectively. In addition, we evaluate estimation accuracy by mean squared error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.

\subsection{Scenario 1}
	In this sparse scenario we assume differential signal between 5 groups of features. The average signal in a group increases by a factor 1.6 with group number:
	\begin{align*}
	\bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & \beta_{50}^1 \end{matrix}}_{50}} & \cdots & \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr, \\
	& \, \\
	\beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots, 5  \text{, } j=1, \dots, 50,
	\end{align*}
	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. The model parameters clearly do not follow the relatively heavy-tailed elastic net distribution. Because of this mismatch between the assumed and true distributions, we expect poor performance by \texttt{gren}. Furthermore, we introduce correlation between the features by having covariance elements $\bSigma_{ij} = 0.5^{|i-j|}$.

	\subsection{Scenario 2}
	The second scenario contains many groups of features and is geared towards the non-adaptive group-regularized methods. These methods are able to incorporate the grouping of the features, but were originally designed for small interpretable groups, such as dummies of a categorical variable. \texttt{gren} on the other hand, relies on large groups of features to accurately estimate group-specific penalty weights. The estimation of many penalty weights in small groups is inefficient and computationally challenging. We therefore expect \texttt{gren} to under-perform compared to the other methods.

	For this scenario we follow the second simulation example in \cite{meier_group_2008}. They set $n=100$, $G=250$, and $p=1000$, such that we have 4 features per group. The first 10 groups are active and those 40 model parameters are set to 0.2. The predictors are simulated from a $p$-dimensional Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.

	\subsection{Scenario 3}
	A third scenario is modelled by considering a situation in which there is no differential signal between the groups, but the correlation within groups is very high. We conjecture that this leads to overestimation of effect within the groups by \texttt{gren} and consequently, performance will suffer.

	To that end we sample the predictors in $G=5$ separate blocks of 200 variables, corresponding to the partitioning. The variances of the features are set to one, while all covariances are 0.7. The 1000 model parameters are simulated in two steps. First, we draw 1000 parameters from the elastic net distribution, parametrised as in \cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$. The second step consists of setting the (in absolute value) smallest 100 parameters in a group to zero. This results in a total of 500 zero and 500 non-zero parameters, evenly distributed over the five groups.

	\subsection{Scenario 4}
	This scenario is similar to Scenario 3, except each groups penalty parameter is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the covariance matrix of the predictors is block diagonal, with block sizes $25 \times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$. We expect \texttt{gren} to pick up this differential signal and outperform the other methods.

	\subsection{Scenario 5}
	In the last scenario we create a sparse true model that contains some differential signal. We impl

\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
\subsection{Partitioning based on differential expression}\label{sec:mirna}
The data set is from a deep sequencing analysis on microRNA \cite[]{neerincx_combination_2018}. The study was done in 88 treated colorectal cancer patients, with the aim of classifying treatment response, coded as either non-progressive/remission (70 patients) or progressive (18 patients). After pre-processing and normalisation, 2114 microRNAs remained. In addition to the 2114 microRNAs we incorporated 4 unpenalized clinical covariates into the analysis: prior use of adjuvant therapy (binary), the type of systemic treatment regimen (ternary), age, and primary tumor differentiation (binary).

In a preliminary experiment on different subjects, the microRNA expression levels of metastatic colorectal tumour tissue were compared to normal non-colorectal tissue and primary colorectal tumour tissue was compared to primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 221 microRNAs that were differentially expressed in both comparisons ($\text{FDR} \leq 0.05$), versus 1893 not differentially expressed microRNAs. We expect that incorporation of this partitioning enhances therapy response classification, because tumor-specific miRNAs might be more relevant than non-specific ones. In addition, we divided the differentially expressed microRNAs even further into 127 highly differentially expressed microRNAs ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs ($0.001 < \text{FDR} \leq 0.05$).  We will refer to this second partitioning as the three-group setting, as opposed to the first, the two-group setting.

We compared \texttt{gren} to ridge and elastic net regression, and \texttt{GRridge}. For the elastic net methods (including \texttt{gren}) we set $\alpha \in \{ 0.05, 0.5, 0.95\}$. The estimated penalty multipliers are according to expectation: in all group-regularized methods, the 221 differentially expressed microRNAs receive the smallest penalty (Figure \ref{fig:col_bar}a). In the three-group setting the pattern is again as expected (Figure \ref{fig:col_bar}b): the highly expressed group, receives the lowest penalty, followed by the medium expressed group, while the non-expressed group receives the strongest penalty. \texttt{GRridge} is not able to distinguish between the medium and non-differentially expressed groups of microRNAs.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
	\label{fig:col_bar}
\end{figure}
Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}. All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS. In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC. This is especially true for the larger models. \texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models. A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly. This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation. In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used. Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting. Note that models with few features are often desirable for clinical implementation.
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
	\label{fig:col_per}
\end{figure}

\subsection{Random groups}
Considering that the features may be partitioned into many groups, with one parameter per group, we have to be aware of overfitting risks. We investigated this using the data introduced in Section \ref{sec:mirna}, randomly dividing the features into three groups. We fixed the group sizes to the group sizes used in the three-group setting in Section \ref{sec:mirna}. Under this random partitioning of the features, we expect all penalty multipliers to be estimated as one if overfitting does not occur.

We compared the estimated multipliers to the estimates by \texttt{GRridge}. Since these results depend on one specific randomisation of the groups, we repeated the procedure 100 times and present the results in Figure \ref{fig:random1}. From this figure we see that the estimates for \texttt{gren} are close to one. The estimates by \texttt{GRridge} show much more variation. Additionally, \texttt{GRridge} gives slightly biased estimates: the median penalty parameter estimates are 1.16, 1, and 0.85 for the three groups. In contrast, the median estimates of \texttt{gren} are 1, 1, and 0.99. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\linewidth]{grEBEN_mirseq_Maarten_res6_boxplot.png}
	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 3 random groups in colorectal miRNA data.}
	\label{fig:random1}
\end{figure}

We repeated the simulation with ten evenly sized groups. Figure \ref{fig:random2} presents the estimated penalty multipliers for 100 random splits of the microRNA features into 10 evenly sized groups for \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$. From the Figure we see that \texttt{GRridge} shows more variation around one then the other models. In addition, the median penalty multipliers for the three \texttt{gren} models are all between 0.99 and 1.02, while the \texttt{GRridge} median estimates range from 1.04 to 1.09. This indicates that \texttt{GRridge} suffers from a slight bias in the estimation, while the \texttt{gren} estimates appear to be unbiased.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.9\linewidth]{grEBEN_mirseq_Maarten_res8_boxplot.png}
	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 10 random groups in colorectal miRNA data.}
	\label{fig:random2}
\end{figure}

\subsection{Stability of selection}
One common criticism on the elastic net, and especially the lasso, is the instability of feature selection. That is, different data sets yield different sets of selected features. To investigate the stability of feature selection by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA data \cite[]{neerincx_combination_2018}, where we stratified by treatment response. For each of these bootstrap samples we estimated the regular elastic net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected 25 features (including the 5 unpenalized covariates). We calculated the size of all ${50}\choose{2}$ intersections of selected features and present the results in Figure \ref{fig:bootstrap}. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.9\linewidth]{gren_mirseq_Maarten_overlap.png}
	\caption{Overlap between stratified bootstrap samples for the elastic net models with (a) $\alpha=0.05$, (b) $\alpha=0.5$, and (c) $\alpha=0.95$.}
	\label{fig:bootstrap}
\end{figure}
From this figure it can be seen that stability of feature selection in \texttt{gren} is higher than in the regular elastic net for all three $\alpha$'s.

\section{Empirical-variational Bayes for the ridge and lasso}\label{sec:evbridgelasso}
\subsection{Ridge regression}
Ridge regression is a special case of the elastic net, that does not include an $L_1$-norm on the parameters. Its estimator is given by:
$$
\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_2}{2} \norm{\bbeta}_2.
$$
The Bayesian version of ridge regression is regression under a Gaussian prior. The generalised version of this prior is:
$$
\bbeta \sim \prod_{g=1}^G \prod_{j \in \G(g)} \N(0,(\lambda_2 \lambda'_g)^{-1}).
$$
Although the posterior is analytically calculated in the linear case, in the logistic case we must approximate it iteratively. The variational distributions in this case are:
\begin{align*}
q^*_{\bbeta} (\bbeta) & \sim \mathcal{N} (\bmu, \bSigma), \\ 
q^*_{\bomega} (\bomega) & \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i). \\
\end{align*}
The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net and are updated, together with $\bSigma$, until convergence:
$$
\bSigma^{t+1} = (\X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda')^{-1}\text{, with } \bOmega^{(t)} = \diag \[ \frac{m_i}{2 c^{(t)}_i} \tanh \( \frac{c^{(t)}_i}{2} \) \].
$$

Updating the penalty parameters is done by solving the following (convex) constraint optimisation problem:
\begin{align*}
\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
& \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
\end{align*}
with the terms $d^{(k)}_g$ given by:
$$
d^{(k)}_g = \sum_{j \in \G(g)} ((\bmu^{(k)}_j)^2 + \bSigma^{(k)}_{jj}).
$$

\subsection{Lasso regression}
In lasso regression, another special case of the elastic net, we have the following estimator:
$$
\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1.
$$
The Bayesian counterpart is regression under a Laplace prior on the model parameters. The generalised Bayesian prior may decomposed as:
\begin{align*}
\bbeta | \btau & \sim \prod_{j=1}^p \mathcal{N} \left(0,\tau_j \right), \\
\btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \text{Exp} \left( \frac{\sqrt{\lambda'_g}\lambda_1^2}{8} \right).
\end{align*}

The variational distributions are similar as in the elastic net, except that we do not re-parametrise the $\tau_j$:
\begin{align*}
q^*_{\bbeta} (\bbeta) & \sim \mathcal{N} (\bmu, \bSigma), \\ 
q^*_{\bomega} (\bomega) & \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i), \\
q^*_{\btau}(\btau) & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{GIG} \(\frac{1}{2}, \phi_g, \chi_j\).
\end{align*}
The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic net, while the $\phi_g$ are given by:
$$
\phi_g = \lambda_{g} \frac{\lambda_1^2}{4}.
$$
The following parameters are updated, together with the unchanged $c_i$ and $\bmu$, until convergence:
\begin{align*}
\bSigma^{(t + 1)} &= ( \X \tr \bOmega^{(t)} \X + \Z^{(t)})^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\(\sqrt{\frac{\phi_g(j)}{\chi_j^{(t)}}}\), \\
\chi_j^{(t + 1)} &= \bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \text{, for } j=1, \dots, p.
\end{align*}

Updating the penalty parameters is done by solving the following (convex) constraint problem:
\begin{align*}
\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_1^2}{8} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
& \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
\end{align*}
with the terms $d^{(k)}_g$ given by:
$$
d^{(k)}_g = \sum_{j \in \G(g)} \( \frac{1}{\phi_g^{(k)}} + \frac{\chi^{(k)}_j}{\phi_g^{(k)}} \).
$$ 
Note that the $\phi_g$ are updated after every penalty multiplier update iteration as well, in contrast to the elastic net.

\bibliographystyle{author_short3.bst} 
\bibliography{refs}

\end{document}

### 2019-02-05 docs/supplement_biostatistics.Rnw
adjusted alzheimer data analysis

--- 2019-02-05 docs/supplement.Rnw
@@
 \documentclass[a4paper,hidelinks]{article}
-\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}
+\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont}
 
@@
 % syntax and readability
+\newcommand{\para}{\bigskip\noindent}
 \renewcommand{\(}{\left(}
@@
 % settings
-\graphicspath{{../figs/}}
+\graphicspath{{/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/}}
 \setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
@@
 2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
-3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\
+3. Department of Mathematics, VU University, Amsterdam, The Netherlands
 
-\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes
+\para
+\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes \\
 
-\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}
+\para
+\textbf{Software available from}: \url{https://github.com/magnusmunch/gren/}\\
 
@@
 The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}.
-We apply our method to a study on microRNAs in colorectal cancer data \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} .
+We give some extra results on the microRNA in colorectal cancer data application in the Main Document in Section \ref{sec:extra} \cite[]{neerincx_combination_2018}.
 Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.
@@
 The density function of $\beta$, drawn from an elastic net prior distribution, is given by:
+
 $$
@@
 $$
+
 The normalizing constant here is calculated as:
+
 $$
@@
 $$
+
 with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution functions of $x$, respectively.
@@
 The expectation and variance of $\beta$ are:
+
 \begin{align*}
@@
 \cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$:
+
 \begin{align*}
@@
 \end{align*}
+
 where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$.
@@
 The CDF of this distribution is given by:
+
 $$
@@
 Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for $\tau \geq 1$, the quantile function may be written as a function of the probability $p \in (0,1)$:
+
 $$
@@
 $$
+
 where $\Phi^{-1}(p)$ is the quantile function of the standard normal distribution, i.e, the probit function.
-Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now:
+Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now: \\
+
 \begin{algorithmic}
@@
 \section{Objective function}\label{sec:objectivefunction}
+
 The lower bound on the marginal likelihood, or evidence lower bound (ELBO), at iteration $t$ is given by:
+
 \begin{subequations}
@@
 We start with the computation of (\ref{eq:elboomega}). To this end we define:
+
 $$
@@
 $$
+
 Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$.
@@
 Next we write out the three terms in (\ref{eq:elboomega}):
+
 \begin{subequations} \label{eq:elboomega2}
@@
 	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
-	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] \E [\omega_i] \\
+	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i \tr] \E [\omega_i] \\
 	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
@@
 \end{subequations}
+
 where all mathematical operations on vectors and matrices are element wise.
 After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
+
 \begin{subequations} \label{eq:elboomega3}
@@
 After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
+
 \begin{subequations} \label{eq:elbobeta2}
@@
 \end{subequations}
+
 We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):
+
 \begin{subequations} \label{eq:elbobeta3}
@@
 Repeating the same exercise for (\ref{eq:elbotau}) gives:
+
 \begin{subequations} \label{eq:elbotau2}
@@
 \end{subequations}
+
 Again, we combine the terms in (\ref{eq:elbotau2}) to get:
+
 \begin{subequations} \label{eq:elbotau3}
@@
 Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
+
 \begin{align*}
@@
 \end{align*}
+
 Here, proportionality is with respect to the variational and penalty parameters.
@@
 
+\para
 In \cite{polson_bayesian_2013}, the central identity for this parametrisation is given in their Theorem 1:
+
 \begin{equation}\label{eq:intid}
@@
 \end{equation}
+
 where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed.
@@
 The likelihood contribution of observation $i$ for $\bbeta$ conditional on $\omega_i$ may now be written as:
+
 $$
@@
 This allows the conditional posterior of $\bbeta$ to be written as:
+
 $$
@@
 $$
+
  where $\bOmega = \diag (\omega_i)$, for $i=1, \dots, n$. Consequently, with $\pi(\bbeta)$ a Gaussian prior, $p(\bbeta | \bomega, \y)$ is also in the Gaussian family.
 
+\para
 Furthermore, by treating the integrand in (\ref{eq:intid}) as an un-normalized density over $(\psi, \phi)$, we find the distribution of $\omega$ conditional on $\bbeta$,
+
 $$
@@
 $$
+
 distributed as $\omega | \psi \sim \mathcal{PG}(b,\psi)$ \cite[]{polson_bayesian_2013}, with expectation $\E[ \omega ] = \frac{b}{2\psi} \tanh (\psi/2)$.
 We substitute again $b=m_i$ and $\psi=\x_i \tr \bbeta$, and set the prior $\pi(\bbeta)$ to a Gaussian with mean $\m$ and covariance $\Vm$ to arrive at the following conditional posteriors:
+
 \begin{subequations}\label{eq:gibbssampler}
@@
 \end{subequations}
+
 where
+
 \begin{align*}
@@
 \end{align*}
+
 with $\bkappa = \begin{bmatrix} \kappa_1 \dots \kappa_n \end{bmatrix} \tr$.
@@
 The optimal distributions in our variational Bayes implementation are:
+
 \begin{align}
@@
 For convenience we work on the log-scale and occasionally do a change of variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three terms on the right-hand side of (\ref{eq:qbeta}):
+
 \begin{align*}
@@
 \end{align*}
+
 where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega = \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' = \diag (\lambda'_{g(j)})$. 
@@
 Finally, we have:
+
 $$
@@
 $$
+
 which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma \X \tr \bkappa$.
 
+\para
 Next we derive the variational posterior for $\bomega$ in (\ref{eq:qomega}):
+
 \begin{align*}
@@
 \end{align*}
+
 where proportionality is with respect to $\bomega$.
-If exponentiated, this is a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.
+If exponentiated, this is a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}$.
 
+\para
 The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated as:
+
 \begin{align*}
@@
 \end{align*}
+
 Combining the terms we find the log-kernel of the variational posterior as:
+
 \begin{align*}
@@
 \end{align*}
+
 where proportionality is with respect to $\btau$. Again applying the change of variables $\psi_j = \tau_j - 1$, we have:
+
 $$
@@
 $$
+
 which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$ random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + \V(\beta_j)]$.
@@
 
+\para
 In the following we slightly change notation and let $\diag(\A)$, with $\A$ a $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$ to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$.
@@
 Now, by the Woodbury identity, we may decompose $\diag(\bSigma)$ as:
+
 \begin{subequations}\label{eq:diagsigma}
@@
 \end{subequations}
+
 where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is a $p \times n$-dimensional matrix of ones.
@@
 Again using the Woodbury identity, we may write $\bmu$ as:
+
 \begin{align}\label{eq:mu}
@@
 \end{align}
+
 which is, by a similar argument as before, again an operation of complexity $\mathcal{O}(n^2p)$.
 
+\para
 Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$ as:
+
 \begin{align*}
@@
 \end{align*}
+
 The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$, Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}(n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.
 
+\para
 The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p \end{bmatrix} \tr$ are rewritten as:
+
 \begin{subequations}\label{eq:chicomp}
@@
 \end{subequations}
+
  By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we see that the computation of $\bm{\chi}$ consists of one matrix diagonal of $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two Hadamard products of $\mathcal{O}(p)$.
@@
 Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then the variational Bayes expected joint log likelihood is derived as follows:
+
 \begin{align*}
@@
 \end{align*}
+
 with the coefficients $d_g^{(k)}$ calculated as:
+
 \begin{align*}
@@
 \section{Full estimation procedure}\label{sec:fullestimationprocedure}
+
 The expectation and maximisation steps introduced in the Main Document result in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
@@
 
-\section{Extensions}\label{sec:extensions}
-\subsection{Unpenalized covariates}
-Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix $\X$ with a column of ones.
-Penalization of such an intercept parameter is not desirable \cite[]{Hastie_Elements_2009}.
-Additionally, it is often desirable to include unpenalized covariates in the model.
-Common examples of such  covariates in clinical research are patient characteristics, such as age, BMI, and sex.
-In the following we assume the intercept to be included in the unpenalized covariates.
+\section{Extra results for microRNA data}\label{sec:extra}
+\subsection{Stability of selection}
 
-To include unpenalized covariates, we divide the model matrix into two parts $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$ unpenalized variables and $\X_r$ are the $r$ penalized variables.
-Let $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with $u$ zero columns and $u$ zero rows.
-Then we have for the current estimate of the covariance matrix $\bSigma$:
-\begin{align}\label{eq:unpencov}
-\bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
-& = 
-\begin{bmatrix} 
-\X_u \tr \bOmega \X_u & \X_u \tr \bOmega \X_r \\
-\X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
-\end{bmatrix}^{-1}.
-\end{align}
-With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension $p \times p$.
-Inversion of this matrix is done efficiently by applying the Woodbury identity.
-
-\subsection{Monotonicity of the penalty parameters}
-Enforcing monotonicity of the penalty multipliers is desirable in some settings.
-An example of such a setting is if we have $p$-values from a previous, related study available.
-A $p$-values-based partitioning of the variables may, \textit{a priori}, be expected to render penalty multipliers that increase monotonically with $p$-value.
-That is, larger $p$-values may be expected to yield at least as large penalty multipliers as variables with smaller $p$-values.
-We propose to include this \textit{a priori} assumption by requiring the penalty multipliers to increase monotonically with $p$-value, thereby also stabilizing their estimates.
-
-A natural way of enforcing monotonicity is to extend the constraint in (10) in the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$.
-The problem is still convex and may be numerically solved.
-From experience, however, we note that in combination with this constrained optimisation, the EM algorithm described in Section~3 in the Main Document often converges to a local optimum close to the initialisation.
-We therefore enforce monotonicity through a \textit{post hoc} isotonic regression on the penalty multipliers after every optimisation step.
-
-\subsection{Multiple partitions}
-In many cases the features may be partitioned in more than one way.
-For example, we may have both information on annotation and $p$-values from a previous study available.
-A naive way of incorporating multiple partitions is to cross-tabulate the partitions and create a separate group for every combination.
-This poses two problems: (i) The number of penalty parameters increases exponentially with the number of partitions and (ii) some of these combinations may contain only few features, so that the estimation procedure becomes unstable.
-We propose to stabilise the procedure and keep the number of parameters to estimate manageable by modelling the penalty parameters multiplicatively.
-
-We describe our implementation here for two partitions of the features.
-To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively.
-Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively.
-In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively.
-The generalised frequentist elastic net estimator and corresponding conditional prior are now:
-\begin{subequations}\label{eq:multpart}
-	\begin{align}
-	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
-	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
-	\end{align}
-\end{subequations}
-After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
-\begin{subequations}\label{eq:estmultpart}
-	\begin{alignat}{2}
-	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
-	\end{alignat}
-\end{subequations}
-with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$.
-The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine.
-This method naturally generalises to more than two partitions of the data.
-
-In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers in a multiplicative way, mainly for convenience.
-First, there are computational reasons: The separation of the square root in the $L_1$-norm penalty term into two terms facilitates the numerical estimation.
-Additionally, multiplicative modelling of the partitions allows for more flexibility in the estimates.
-To see this consider the obvious alternative: additive partition-specific multipliers.
-Since the penalty multipliers are strictly positive, the constraint on the geometric mean makes one large additive penalty estimate difficult to compensate with a smaller estimate, thereby impairing flexibility of the model.
-Large multiplicative penalty multipliers are easier to compensate for by smaller penalty multipliers, rendering this the less restrictive option for modelling the partitions.
-
-\section{Simulations}
-The quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960}.
-This kappa is measures similarity between two sets of categorical values.
-One set contains the feature selection indicators: $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
-We calculate Cohen's $\kappa$ as:
-	\begin{align*}
-	\kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}\sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
-	\end{align*}
-	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and expected correctly identified features, respectively.
-A positive or negative kappa indicate better and worse feature selection than expected by chance, respectively.
-In addition, we evaluate estimation accuracy by mean squared error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
-
-\subsection{Scenario 1}
-	In this sparse scenario we assume differential signal between 5 groups of features. The average signal in a group increases by a factor 1.6 with group number:
-	\begin{align*}
-	\bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & \beta_{50}^1 \end{matrix}}_{50}} & \cdots & \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr, \\
-	& \, \\
-	\beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots, 5  \text{, } j=1, \dots, 50,
-	\end{align*}
-	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$.
-The model parameters clearly do not follow the relatively heavy-tailed elastic net distribution.
-Because of this mismatch between the assumed and true distributions, we expect poor performance by \texttt{gren}.
-Furthermore, we introduce correlation between the features by having covariance elements $\bSigma_{ij} = 0.5^{|i-j|}$.
-
-	\subsection{Scenario 2}
-	The second scenario contains many groups of features and is geared towards the non-adaptive group-regularized methods.
-These methods are able to incorporate the grouping of the features, but were originally designed for small interpretable groups, such as dummies of a categorical variable.
-\texttt{gren} on the other hand, relies on large groups of features to accurately estimate group-specific penalty weights.
-The estimation of many penalty weights in small groups is inefficient and computationally challenging.
-We therefore expect \texttt{gren} to under-perform compared to the other methods.
-
-	For this scenario we follow the second simulation example in \cite{meier_group_2008}.
-They set $n=100$, $G=250$, and $p=1000$, such that we have 4 features per group.
-The first 10 groups are active and those 40 model parameters are set to 0.2.
-The predictors are simulated from a $p$-dimensional Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.
-
-	\subsection{Scenario 3}
-	A third scenario is modelled by considering a situation in which there is no differential signal between the groups, but the correlation within groups is very high.
-We conjecture that this leads to overestimation of effect within the groups by \texttt{gren} and consequently, performance will suffer.
-
-	To that end we sample the predictors in $G=5$ separate blocks of 200 variables, corresponding to the partitioning.
-The variances of the features are set to one, while all covariances are 0.7.
-The 1000 model parameters are simulated in two steps.
-First, we draw 1000 parameters from the elastic net distribution, parametrised as in \cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$.
-The second step consists of setting the (in absolute value) smallest 100 parameters in a group to zero.
-This results in a total of 500 zero and 500 non-zero parameters, evenly distributed over the five groups.
-
-	\subsection{Scenario 4}
-	This scenario is similar to Scenario 3, except each groups penalty parameter is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the covariance matrix of the predictors is block diagonal, with block sizes $25 \times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$.
-We expect \texttt{gren} to pick up this differential signal and outperform the other methods.
-
-	\subsection{Scenario 5}
-	In the last scenario we create a sparse true model that contains some differential signal. We impl
-
-\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
-\subsection{Partitioning based on differential expression}\label{sec:mirna}
-The data set is from a deep sequencing analysis on microRNA \cite[]{neerincx_combination_2018}.
-The study was done in 88 treated colorectal cancer patients, with the aim of classifying treatment response, coded as either non-progressive/remission (70 patients) or progressive (18 patients).
-After pre-processing and normalisation, 2114 microRNAs remained.
-In addition to the 2114 microRNAs we incorporated 4 unpenalized clinical covariates into the analysis: prior use of adjuvant therapy (binary), the type of systemic treatment regimen (ternary), age, and primary tumor differentiation (binary).
-
-In a preliminary experiment on different subjects, the microRNA expression levels of metastatic colorectal tumour tissue were compared to normal non-colorectal tissue and primary colorectal tumour tissue was compared to primary colorectal tumour tissue \cite[]{neerincx_mir_2015}.
-This yielded 221 microRNAs that were differentially expressed in both comparisons ($\text{FDR} \leq 0.05$), versus 1893 not differentially expressed microRNAs.
-We expect that incorporation of this partitioning enhances therapy response classification, because tumor-specific miRNAs might be more relevant than non-specific ones.
-In addition, we divided the differentially expressed microRNAs even further into 127 highly differentially expressed microRNAs ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs ($0.001 < \text{FDR} \leq 0.05$).
-We will refer to this second partitioning as the three-group setting, as opposed to the first, the two-group setting.
-
-We compared \texttt{gren} to ridge and elastic net regression, and \texttt{GRridge}.
-For the elastic net methods (including \texttt{gren}) we set $\alpha \in \{ 0.05, 0.5, 0.95\}$.
-The estimated penalty multipliers are according to expectation: in all group-regularized methods, the 221 differentially expressed microRNAs receive the smallest penalty (Figure \ref{fig:col_bar}a).
-In the three-group setting the pattern is again as expected (Figure \ref{fig:col_bar}b): the highly expressed group, receives the lowest penalty, followed by the medium expressed group, while the non-expressed group receives the strongest penalty.
-\texttt{GRridge} is not able to distinguish between the medium and non-differentially expressed groups of microRNAs.
+One common criticism on the elastic net, and especially the lasso, is the instability of feature selection.
+That is, different data sets yield different sets of selected features.
+To investigate the stability of feature selection by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA data used in Section 6 of the Main Document \cite[]{neerincx_combination_2018}, where we stratified by treatment response.
+For each of these bootstrap samples we estimated the regular elastic net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected 25 features (including the 5 unpenalized covariates).
+We calculated the size of all ${50}\choose{2}$ intersections of selected features and present the results in Figure \ref{fig:bootstrap}.
+From this figure it can be seen that stability of feature selection in \texttt{gren} is higher than in the regular elastic net for all three $\alpha$'s.
 
@@
 	\centering
-	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
-	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
-	\label{fig:col_bar}
-\end{figure}
-Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}.
-All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS.
-In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC.
-This is especially true for the larger models.
-\texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models.
-A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly.
-This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation.
-In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used.
-Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting.
-Note that models with few features are often desirable for clinical implementation.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
-	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
-	\label{fig:col_per}
+	\includegraphics[width=0.9\linewidth]{../figs/gren_mirseq_Maarten_overlap.png}
+	\caption{Overlap between stratified bootstrap samples for the elastic net models with (a) $\alpha=0.05$, (b) $\alpha=0.5$, and (c) $\alpha=0.95$.}
+	\label{fig:bootstrap}
 \end{figure}
 
-\subsection{Random groups}
-Considering that the features may be partitioned into many groups, with one parameter per group, we have to be aware of overfitting risks.
-We investigated this using the data introduced in Section \ref{sec:mirna}, randomly dividing the features into three groups.
-We fixed the group sizes to the group sizes used in the three-group setting in Section \ref{sec:mirna}.
-Under this random partitioning of the features, we expect all penalty multipliers to be estimated as one if overfitting does not occur.
+\subsection{Penalty multipliers under random groups}
 
-We compared the estimated multipliers to the estimates by \texttt{GRridge}.
-Since these results depend on one specific randomisation of the groups, we repeated the procedure 100 times and present the results in Figure \ref{fig:random1}.
-From this figure we see that the estimates for \texttt{gren} are close to one.
-The estimates by \texttt{GRridge} show much more variation.
-Additionally, \texttt{GRridge} gives slightly biased estimates: the median penalty parameter estimates are 1.16, 1, and 0.85 for the three groups.
-In contrast, the median estimates of \texttt{gren} are 1, 1, and 0.99.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.5\linewidth]{grEBEN_mirseq_Maarten_res6_boxplot.png}
-	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 3 random groups in colorectal miRNA data.}
-	\label{fig:random1}
-\end{figure}
-
-We repeated the simulation with ten evenly sized groups.
-Figure \ref{fig:random2} presents the estimated penalty multipliers for 100 random splits of the microRNA features into 10 evenly sized groups for \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+Figure \ref{fig:random} presents the estimated penalty multipliers for 100 random splits of the microRNA features into 10 evenly sized groups for \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+Since we created the groups at random, we expect that the penalty multipliers are estimated as one, to reflect the non-informativeness of the groups.
 From the Figure we see that \texttt{GRridge} shows more variation around one then the other models.
@@
 This indicates that \texttt{GRridge} suffers from a slight bias in the estimation, while the \texttt{gren} estimates appear to be unbiased.
+
 \begin{figure}[!h]
 	\centering
-	\includegraphics[width=0.9\linewidth]{grEBEN_mirseq_Maarten_res8_boxplot.png}
+	\includegraphics[width=0.9\linewidth]{../figs/grEBEN_mirseq_Maarten_res8_boxplot.png}
 	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 10 random groups in colorectal miRNA data.}
-	\label{fig:random2}
+	\label{fig:random}
 \end{figure}
-
-\subsection{Stability of selection}
-One common criticism on the elastic net, and especially the lasso, is the instability of feature selection.
-That is, different data sets yield different sets of selected features.
-To investigate the stability of feature selection by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA data \cite[]{neerincx_combination_2018}, where we stratified by treatment response.
-For each of these bootstrap samples we estimated the regular elastic net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected 25 features (including the 5 unpenalized covariates).
-We calculated the size of all ${50}\choose{2}$ intersections of selected features and present the results in Figure \ref{fig:bootstrap}.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.9\linewidth]{gren_mirseq_Maarten_overlap.png}
-	\caption{Overlap between stratified bootstrap samples for the elastic net models with (a) $\alpha=0.05$, (b) $\alpha=0.5$, and (c) $\alpha=0.95$.}
-	\label{fig:bootstrap}
-\end{figure}
-From this figure it can be seen that stability of feature selection in \texttt{gren} is higher than in the regular elastic net for all three $\alpha$'s.
 
@@
 Ridge regression is a special case of the elastic net, that does not include an $L_1$-norm on the parameters. Its estimator is given by:
+
 $$
@@
 $$
+
 The Bayesian version of ridge regression is regression under a Gaussian prior. The generalised version of this prior is:
+
 $$
@@
 $$
+
 Although the posterior is analytically calculated in the linear case, in the logistic case we must approximate it iteratively. The variational distributions in this case are:
+
 \begin{align*}
@@
 \end{align*}
+
 The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net and are updated, together with $\bSigma$, until convergence:
+
 $$
@@
 Updating the penalty parameters is done by solving the following (convex) constraint optimisation problem:
+
 \begin{align*}
@@
 \end{align*}
+
 with the terms $d^{(k)}_g$ given by:
+
 $$
@@
 In lasso regression, another special case of the elastic net, we have the following estimator:
+
 $$
@@
 $$
+
 The Bayesian counterpart is regression under a Laplace prior on the model parameters. The generalised Bayesian prior may decomposed as:
+
 \begin{align*}
@@
 The variational distributions are similar as in the elastic net, except that we do not re-parametrise the $\tau_j$:
+
 \begin{align*}
@@
 \end{align*}
+
 The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic net, while the $\phi_g$ are given by:
+
 $$
@@
 $$
+
 The following parameters are updated, together with the unchanged $c_i$ and $\bmu$, until convergence:
+
 \begin{align*}
@@
 Updating the penalty parameters is done by solving the following (convex) constraint problem:
+
 \begin{align*}
@@
 $$
+
 Note that the $\phi_g$ are updated after every penalty multiplier update iteration as well, in contrast to the elastic net.
 
-\bibliographystyle{author_short3.bst} 
+\bibliographystyle{biorefs.bst} 
 \bibliography{refs}

### 2019-02-05 code/metabolomics_alzheimer.R

@@ -1,46 +1,161 @@
-path.data <- as.character(ifelse(Sys.info()[1]=="Darwin", "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/data/" , "~/EBEN/data/"))
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-
+### libraries
 library(gren)
-library(pROC)
 library(GRridge)
 library(grpreg)
 library(SGL)
-library(Hmisc)
 library(Biobase)
 
-# colors <- bpy.colors(5)[-c(1, 5)]
+### load data
+load("data/ESetMbolCSFPR2.Rdata")
 
-load(paste(path.data, "ESetMbolCSFPR2.Rdata", sep=""))
+### select the patients with APOE E4 allel and controls without APOE E4
+pheno <- pData(ESetMbolCSFPR2)
+pheno.apoe <- pheno[(pheno$D_diag_name=="Probable AD" & pheno$APOE=="E4YES") |
+                      (pheno$D_diag_name=="Subjectieve klachten" & 
+                         pheno$APOE=="E4NO"), ]
 
-x <- scale(t(exprs(ESetMbolCSFPR2)))
-y <- as.numeric(pData(ESetMbolCSFPR2)$D_diag_name) - 1
-colnames(pData(ESetMbolCSFPR2))
-pData(ESetMbolCSFPR2)$APOE=="E4YES"
 
+metabol <- t(exprs(ESetMbolCSFPR2))
+metabol.apoe <- metabol[(pheno$D_diag_name=="Probable AD" & 
+                           pheno$APOE=="E4YES") |
+                          (pheno$D_diag_name=="Subjectieve klachten" & 
+                             pheno$APOE=="E4NO"), ]
 
-part1 <- replace(fData(ESetMbolCSFPR2)$PlatformCode, 
-                 fData(ESetMbolCSFPR2)$PlatformCode==5, 4)
-part2 <- fData(ESetMbolCSFPR2)$PlatformCode
-part3 <- as.numeric(cut2(fData(ESetMbolCSFPR2)$RSDqc, g=3))
-part4 <- as.numeric(cut2(fData(ESetMbolCSFPR2)$RSDqc, 0.06))
+### create co-data
[…]

### 2019-02-08 docs/manuscript.Rnw
added second simulation

--- 2019-02-05 docs/manuscipt.Rnw
@@
 \documentclass[a4paper,hidelinks]{article}
-\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}
+\usepackage{url,bm,algorithm,algpseudocode,bbm,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,threeparttable, booktabs,dsfont,parskip}
 
@@
 	\noindent
-	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
-	Amsterdam, The Netherlands\\
-	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
-	3. MRC Biostatistics Unit, Cambridge, United Kingdom \\
+	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
+	Amsterdam, The Netherlands \\*
+	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
+	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom
 
@@
 Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.
+We would like to emphasize that our method is different from recent developments in grouped penalized regression such as \cite{dondelinger_joint_nodate}, where the grouping is over the observations as opposed to the features
 
@@
 	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$.
-In this Bayesian formulation the penalty parameters in $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.
+In this Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.
 
@@
 In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform.
-In the second scenario, we expect \texttt{gren} and \textt{GRridge} to overfit due to the large number of hyperparameters to estimate.
+In the second scenario, we expect \texttt{gren} and \texttt{GRridge} to overfit due to the large number of hyperparameters to estimate.
 The rest of the methods are conjectured to perform better, due to a lack of overfitting.
@@
 To mitigate the influence of random variation in the simulations, we repeated every scenario 100 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.
-
-	 The results of this simulation are presented in Figures \ref{fig:sim1_performance1_set1}-\ref{fig:sim1_performance1_set3}.
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set1.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set1}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set2.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set2}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set3.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set3}
-	\end{figure}
-
-	The results of this simulation 1 are presented in Figures \ref{fig:sim4_performance1_set1}-\ref{fig:sim4_performance1_set3}.
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set1.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set1}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set2.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set2}
-	\end{figure}
-
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set3.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set3}
-	\end{figure}
 

### 2019-02-08 code/figures.R

@@ -1,7 +1,33 @@
 ################################## manuscript ##################################
 # ---- figures ----
 # ---- boxplot_metabolomics_alzheimer_res1_auc ---- 
+library(pROC)
+library(Biobase)
+labels1 <- c(unique(as.character(feat$Platform))[-c(4, 5)], "Oxidative Stress")
 res1 <- read.table("results/metabolomics_alzheimer_res1.csv", 
                    stringsAsFactors=FALSE)
-labels1 <- c(unique(as.character(feat$Platform))[-c(4, 5)], "Oxidative Stress")
+load("data/ESetMbolCSFPR2.Rdata")
+pheno <- pData(ESetMbolCSFPR2)
+pheno.apoe <- pheno[(pheno$D_diag_name=="Probable AD" & pheno$APOE=="E4YES") |
+                      (pheno$D_diag_name=="Subjectieve klachten" & 
+                         pheno$APOE=="E4NO"), ]
+alzheim.apoe <- as.numeric(pheno.apoe$D_diag_name) - 1
+
+auc1 <- sapply(res1[grepl("pred", rownames(res1)), ], function(s) {
+  pROC::auc(alzheim.apoe, s)})
+psel1 <- colMeans(res1[grepl("psel", rownames(res1)), ])
+
+plot(psel1, auc1)
+boxplot(res1[grepl("psel", rownames(res1)), ])
+
+
+
+barplot(rbind(fit1.gren1$lambdag$part, fit1.grridge$lambdamults$part), 
+        beside=TRUE)
+abline(h=1, lty=2)
+
+
+
+
+
 

### 2019-02-08 code/functions.R

@@ -1,0 +1,40 @@
+### function to find a specific number of features
+froot <- function(lambda, psel, x, y, part, alpha, type) {
+  if(type=="sgl") {
+    fit <- SGL(list(x=x, y=y), part, type="logit", alpha=alpha, standardize=FALSE,
+               lambdas=lambda, nlam=1, min.frac=0.001)
+    out <- psel - sum(fit$beta!=0)
+  } else if(type=="cmcp") {
+    fit <- grpreg(x, y, part, "cMCP", "binomial", lambda=lambda, alpha=alpha)
+    out <- psel - sum(fit$beta[-1, ]!=0)
+  } else if(type=="gel") {
+    fit <- grpreg(x, y, part, "gel", "binomial", lambda=lambda, alpha=alpha)
+    out <- psel - sum(fit$beta[-1, ]!=0)
+  }
+  return(out)
+}
+
+sel.grpreg <- function(X, y, group=1:ncol(X), 
+                       penalty=c("grLasso", "grMCP", "grSCAD", "gel", "cMCP"), 
+                       family=c("gaussian", "binomial", "poisson"),
+                       nlambda=100, lambda, 
+                       lambda.min={if (nrow(X) > ncol(X)) 1e-4 else .05},
+                       log.lambda = TRUE, alpha=1, eps=1e-4, max.iter=10000, 
+                       dfmax=p, gmax=length(unique(group)), 
+                       gamma=ifelse(penalty == "grSCAD", 4, 3), tau = 1/3, 
+                       group.multiplier, warn=TRUE, returnX = FALSE) {
+  
+  
+  
+}
+sapply(psel, function(csel) {
+  fit.root <- uniroot(froot, range(test.init$lambda), psel=csel, x, y, part3,
+                         alpha=0.05, type="gelasso", maxiter=100);
+  fit <- grpreg(x, y, part3, "gel", "binomial", lambda=test.root$root,
+                       alpha=0.05);
+    return(sum(test.fit$beta[-1, ]!=0))})
+  
+  
+  
+  
+  

### 2019-02-08 code/metabolomics_alzheimer.R

@@ -49,6 +49,7 @@ sds.apoe <- rep(c(1:3), sapply(part.sds.apoe, length))[unlist(part.sds.apoe)]
 alzheim.apoe <- as.numeric(pheno.apoe$D_diag_name) - 1
 metabol.apoe.scaled <- scale(metabol.apoe)
 
+################################### model 1 ####################################
 ### fitting the models
 set.seed(2018)
 y <- alzheim.apoe
@@ -186,340 +187,139 @@ for(k in 1:nfolds) {
 }
 
 
-# 
-# 
-# 
-# path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-#                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-# path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-# path.data <- as.character(ifelse(Sys.info()[1]=="Darwin", "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/data/" , "~/EBEN/data/"))
-# 
-# library(sp)
-# 
-# load(paste(path.res, "gren_metabol_alzheimer_res1.Rdata", sep=""))
-# load(paste(path.data, "ESetMbolCSFPR2.Rdata", sep=""))
-# y <- as.numeric(pData(ESetMbolCSFPR2)$D_diag_name) - 1
-# 
-# auc3.ridge <- pROC::auc(y, results3$pred$ridge)
-# auc3.grridge <- apply(results3$pred$grridge, 2, function(s) {pROC::auc(y, s)})
-# auc3.gren1 <- apply(results3$pred$gren1, 2, function(s) {pROC::auc(y, s)})
-# auc3.gren2 <- apply(results3$pred$gren2, 2, function(s) {pROC::auc(y, s)})
-# auc3.gren3 <- apply(results3$pred$gren3, 2, function(s) {pROC::auc(y, s)})
-# auc3.enet1 <- apply(results3$pred$enet1, 2, function(s) {pROC::auc(y, s)})
-# auc3.enet2 <- apply(results3$pred$enet2, 2, function(s) {pROC::auc(y, s)})
-# auc3.enet3 <- apply(results3$pred$enet3, 2, function(s) {pROC::auc(y, s)})
-# 
-# const <- sum((y - mean(y))^2)
-# briers3.ridge <- 1 - sum((y - results3$pred$ridge)^2)/const
[…]

### 2019-02-11 docs/manuscript.Rnw
commit before pull

@@
 \documentclass[a4paper,hidelinks]{article}
-\usepackage{url,bm,algorithm,algpseudocode,bbm,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,threeparttable, booktabs,dsfont,parskip}
+\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
+amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
+titlesec,multicol,hyperref,verbatim,threeparttable,booktabs,dsfont,parskip}
 
@@
 \date{}
-\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}
+\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
+\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
+Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}
 
@@
 	\noindent
-	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
+	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
+	research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
 	Amsterdam, The Netherlands \\*
@@
 	\begin{abstract}
-		{In high-dimensional data settings, additional information on the features is often available.
-Examples of such external information in omics research are: (a) $p$-values from a previous study, (b) a summary of prior information, and (c) omics annotation.
-The inclusion of this information in the analysis may enhance classification performance and feature selection, but is not straightforward in the standard regression setting.
-As a solution to this problem, we propose a group-regularized (logistic) elastic net regression method, where each penalty parameter corresponds to a group of features based on the external information.
-The method, termed \texttt{gren}, makes use of the Bayesian formulation of logistic elastic net regression to estimate both the model and penalty parameters in an approximate empirical-variational Bayes framework.
-Simulations and an application to a colon cancer microRNA study show that, if the partitioning of the features is informative, classification performance and feature selection are indeed enhanced.}
+		{In high-dimensional data settings, additional information on the features 
+		is often available. Examples of such external information in omics research
+		are: (a) $p$-values from a previous study, (b) a summary of prior 
+		information, and (c) omics annotation. The inclusion of this information in
+		the analysis may enhance classification performance and feature selection,
+		but is not straightforward in the standard regression setting. As a solution
+		to this problem, we propose a group-regularized (logistic) elastic net
+		regression method, where each penalty parameter corresponds to a group of
+		features based on the external information. The method, termed 
+		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
+		regression to estimate both the model and penalty parameters in an 
+		approximate empirical-variational Bayes framework. Simulations and an
+		application to a colon cancer microRNA study show that, if the partitioning
+		of the features is informative, classification performance and feature
+		selection are indeed enhanced.}
 	\end{abstract}
 
-	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes
+	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; 
+	Prediction; Variational Bayes
 
-	\noindent\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}
+	\noindent\textbf{Software available from}: 
+	\url{https://CRAN.R-project.org/package=gren}
 
 	\section{Introduction}
-	Prediction from high-dimensional data is a challenge common to many fields of research.
-Examples of such high-dimensional prediction problems arise in computer vision, stock market prediction and disease diagnosis from omics data.
-In this paper, we specifically focus on the latter setting.
-High dimensionality of data introduces several issues in the estimation of prediction models, especially: (a) unidentifiable models, (b) highly correlated predictor variables and (c) non-trivial variable selection.
+	Prediction from high-dimensional data is a challenge common to many fields of
+	research. Examples of such high-dimensional prediction problems arise in
+	computer vision, stock market prediction and disease diagnosis from omics 
+	data. In this paper, we specifically focus on the latter setting. High
+	dimensionality of data introduces several issues in the estimation of 
+	prediction models, especially: (a) unidentifiable models, (b) highly 
+	correlated predictor variables and (c) non-trivial variable selection.
 
-	Attempts to tackle one or more of these issues have lead to the development of many prediction methods \cite[]{fan_selective_2010}.
-Here we will focus on the penalization approach.
-A well-known and popular penalization method is elastic net regression \cite[]{zou_regularization_2005}, with its special cases ridge regression \cite[]{hoerl_ridge_1970} and lasso regression \cite[]{tibshirani_regression_1996}.
-The elastic net approach takes issues (a)-(c) into account and has yielded many successful extensions.
-Recent work on elastic net regression (and its special cases) has focussed on increasing prediction accuracy by the inclusion of prior knowledge on the variables \cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.
+	Attempts to tackle one or more of these issues have lead to the development of
+	many prediction methods \cite[]{fan_selective_2010}. Here we will focus on the
+	penalization approach. A well-known and popular penalization method is elastic
+	net regression \cite[]{zou_regularization_2005}, with its special cases ridge
+	regression \cite[]{hoerl_ridge_1970} and lasso regression 
+	\cite[]{tibshirani_regression_1996}. The elastic net approach takes issues 
+	(a)-(c) into account and has yielded many successful extensions. Recent work 
+	on elastic net regression (and its special cases) has focussed on increasing
+	prediction accuracy by the inclusion of prior knowledge on the variables
+	\cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.
 
-	In omics research, sources of such prior knowledge on the variables are often available.
-The information can, for example, come in the form of (a) results on the same molecular features obtained in a previous study (e.g.
-$p$-values), (b) information from a publicly available database that summarizes the prior information on the molecular features involved (e.g., the Cancer Gene Census \cite[]{futreal_census_2004}), and (c) omics annotation (e.g.
-the location of a gene on the chromosome).
-Although this information can rarely be directly included in the statistical analysis, it may still be useful and informative for the study at hand.
+	In omics research, sources of such prior knowledge on the variables are often
+	available. The information can, for example, come in the form of (a) results 
+	on the same molecular features obtained in a previous study (e.g. $p$-values),
+	(b) information from a publicly available database that summarizes the prior
+	information on the molecular features involved (e.g., the Cancer Gene Census
+	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. the location of 
+	a gene on the chromosome). Although this information can rarely be directly
+	included in the statistical analysis, it may still be useful and informative
+	for the study at hand.
 
-	An example of such a study with extra information available is investigated in Section \ref{sec:app}.
-INTRODUCE DATA.
-This problem is not straightforward to solve with the standard methods available in literature.
-At first sight, such an \textit{a priori} grouping of the features suggests the group lasso \cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which penalises and selects features at the group level.
-There are two limitations to the group lasso: (a) it selects groups as a whole instead of single features and (b) it is not adaptive: the amount of penalization a group receives depends on a `global' penalty parameter common to all groups.
-Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
+	An example of such a study with extra information available is investigated in
+	Section \ref{sec:app}. INTRODUCE DATA. This problem is not straightforward to
+	solve with the standard methods available in literature. At first sight, such
+	an \textit{a priori} grouping of the features suggests the group lasso
+	\cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which
+	penalises and selects features at the group level. There are two limitations 
+	to the group lasso: (a) it selects groups as a whole instead of single 
+	features and (b) it is not adaptive: the amount of penalization a group
+	receives depends on a `global' penalty parameter common to all groups.
+	Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013}
+	partly deal with (a), but do not address (b).
 
-	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter.
-An apparent issue with this differential penalization is the estimation of the penalty parameters.
-Naive estimation may be done by cross-validation.
-However, cross-validation requires re-estimation of the model over a grid of penalty parameters.
-The size of this grid increases exponentially with the number of penalty parameters.
-Consequently, it quickly becomes computationally infeasible.
-We therefore propose an efficient alternative: empirical Bayes estimation of the penalty parameters, which corresponds to hyperparameter estimation in the Bayesian prior framework.
-Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net.
-Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit.
-Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.
-We would like to emphasize that our method is different from recent developments in grouped penalized regression such as \cite{dondelinger_joint_nodate}, where the grouping is over the observations as opposed to the features
+	A natural way to deal with (b) is through differential penalization: that is,
+	each group of variables receives its own penalty parameter. An apparent issue
+	with this differential penalization is the estimation of the penalty 
+	parameters. Naive estimation may be done by cross-validation. However, 
+	cross-validation requires re-estimation of the model over a grid of penalty
+	parameters. The size of this grid increases exponentially with the number of
+	penalty parameters. Consequently, it quickly becomes computationally 
+	infeasible. We therefore propose an efficient alternative: empirical Bayes
+	estimation of the penalty parameters, which corresponds to hyperparameter
+	estimation in the Bayesian prior framework. Because of the ubiquity of binary
+	outcome data in omics research, we focus on the logistic elastic net. 
+	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
+	spirit. Nevertheless, one clear advantage of our method over VBVS is the
+	possibility of adaptive inclusion of external information on the features to
+	aid in prediction and feature selection. We would like to emphasize that our
+	method is different from recent developments in grouped penalized regression
+	such as \cite{dondelinger_joint_nodate}, where the grouping is over the
+	observations as opposed to the features
 
-	In line with the above, our proposed method is (a) subjective: the few remaining tuning parameters may be cross-validated, (b) adaptive: prior information is automatically and subjectively weighed with respect to its informativeness, (c) fast compared to full Bayesian analysis or cross-validation, and (d) requires minimal input by the user: only the data and grouping of the features is required.
+	In line with the above, our proposed method is (a) subjective: the few
+	remaining tuning parameters may be cross-validated, (b) adaptive: prior
+	information is automatically and subjectively weighed with respect to its
+	informativeness, (c) fast compared to full Bayesian analysis or 
+	cross-validation, and (d) requires minimal input by the user: only the data 
+	and grouping of the features is required.
 
-	We introduce the Bayesian logistic (generalised) elastic net model in Section \ref{sec:model}.
-In Section \ref{seq:estimation} we derive a variational approximation to this model and use this novel approximation in the empirical Bayes estimation of multiple, group-specific penalty parameters.
-In Sections \ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a simulation study and a real data example, respectively.
-In Section \ref{sec:discussion} we conclude by both discussing some differences and parallels between the proposed method, termed \texttt{gren}, and other methods in the literature.
+	We introduce the Bayesian logistic (generalised) elastic net model in Section
+	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
+	approximation to this model and use this novel approximation in the empirical
+	Bayes estimation of multiple, group-specific penalty parameters. In Sections
+	\ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a
+	simulation study and a real data example, respectively. In Section 
+	\ref{sec:discussion} we conclude by both discussing some differences and
+	parallels between the proposed method, termed \texttt{gren}, and other methods
+	in the literature.
 
@@
 	\subsection{Logistic elastic net regression}\label{sec:logitmodel}
-	In logistic regression the outcome variables are assumed to be binary or sums of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l \in \{0,1\}$ for $i=1, \dots, n$).
-The binomial logistic model relates the responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix} x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} \( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is the binomial distribution with number of trials $m$ and probability $\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + \exp(\x\tr_i \bbeta)]$.
-Note that if $m_i=1$ for $i=1, \dots, n$, the model reduces to a binary logistic regression model.
-Throughout the rest of the paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots & \x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.
+	In logistic regression the outcome variables are assumed to be binary or sums
+	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
+	\in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the
+	responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix}
+	x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} 
+	\( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is 
+	the binomial distribution with number of trials $m$ and probability 
+	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
+	\exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model
+	reduces to a binary logistic regression model. Throughout the rest of the 
+	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
+	\x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n
+	x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.
 
-	In elastic net regression, the penalised likelihood is maximised to yield parameter estimates:
+	In elastic net regression, the penalised likelihood is maximised to yield
+	parameter estimates:
 	$$
-	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
+	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+	\frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
 	$$
-	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} y_1 & \cdots & y_n \end{bmatrix} \tr$.
-The likelihood term is thus complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the parameters, multiplied by respective penalty parameters $\lambda_1$ and $\lambda_2$.
-We have scaled the penalty parameters by $\frac{1}{2}$, to make them correspond with the Bayesian version in the following.
-The maximiser of the penalised likelihood corresponds to the posterior mode of a Bayesian elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
+	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} 
+	y_1 & \cdots & y_n \end{bmatrix} \tr$. The likelihood term is thus 
+	complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the
+	parameters, multiplied by respective penalty parameters $\lambda_1$ and
+	$\lambda_2$. We have scaled the penalty parameters by $\frac{1}{2}$, to make
+	them correspond with the Bayesian version in the following. The maximiser of
+	the penalised likelihood corresponds to the posterior mode of a Bayesian
+	elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
 	\begin{equation}\label{eq:elasticnetprior}
-	\beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta_j| + \lambda_2 \beta_j^2)\].
+	  \beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1
+	  |\beta_j| + \lambda_2 \beta_j^2)\].
 	\end{equation}
-	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the Supplementary Material (SM), along with further details on the elastic net prior).
-The elastic net combines $L_1$ and $L_2$-norm penalisation such that the model parameters are shrunken towards zero.
-The $L_1$-norm may set some of the estimates exactly to zero, thus automatically selecting features.
-The $L_2$-norm ensures that correlated features behave similarly, as one would generally require.
-We extend the elastic net model to the generalised elastic net to allow for different penalty parameters per group of features.
+	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is 
+	dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the 
+	Supplementary Material (SM), along with further details on the elastic net
+	prior). The elastic net combines $L_1$ and $L_2$-norm penalisation such that
+	the model parameters are shrunken towards zero. The $L_1$-norm may set some of
+	the estimates exactly to zero, thus automatically selecting features. The 
+	$L_2$-norm ensures that correlated features behave similarly, as one would
+	generally require. We extend the elastic net model to the generalised elastic
+	net to allow for different penalty parameters per group of features.
 
 	\subsection{The generalised elastic net}\label{sec:genen}
-	Assume we have a partitioning of the features in $G$ groups, such that each feature belongs to one group.
-Let $\G(g)$ be the feature index set of group $g$ for $g=1, \dots, G$.
-To apply differential penalisation we include a group-specific weight $w_g \in \mathbb{R}_{>0}$.
-The model parameters are now estimated by:
+	Assume we have a partitioning of the features in $G$ groups, such that each
+	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
+	$g$ for $g=1, \dots, G$. To apply differential penalisation we include a 
+	group-specific weight $w_g \in \mathbb{R}_{>0}$. The model parameters are now
+	estimated by:
 	\begin{align}
-	\hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 \label{eq:weightedlik}\\
-	& = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, \nonumber
+	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
+	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
+	  \label{eq:weightedlik}\\
+	  & = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} 
+	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
+	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
+	  \nonumber
 	\end{align}
-	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific weights may be interpreted as penalty multipliers.
-Throughout the following we assume that the geometric mean of the multipliers, weighted by their respective group sizes, is one, such that the average shrinkage of the model parameters is determined by the `global' $\lambda_1$ and $\lambda_2$.
-That is, we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G (\lambda'_g)^{|\G(g)|}=1$.
-The multiplier appears in square root form in the $L_1$-norm term to ensure that penalisation on the parameter level is the same for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in (\ref{eq:weightedlik}).
+	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific
+	weights may be interpreted as penalty multipliers. Throughout the following we
+	assume that the geometric mean of the multipliers, weighted by their 
+	respective group sizes, is one, such that the average shrinkage of the model
+	parameters is determined by the `global' $\lambda_1$ and $\lambda_2$. That is,
+	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
+	(\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the
+	$L_1$-norm term to ensure that penalisation on the parameter level is the same
+	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
+	(\ref{eq:weightedlik}).
 

### 2019-02-11 code/boxcox_test.R
moved old code, results, and data to separate untracked folders

@@ -1,71 +1,0 @@
-#Suppose truth = negative linear relation
-alpha <- -2; lambda <- 1
-xlin <- 1:1000/1000
-data <- alpha/lambda*(xlin^(lambda)-1) + rnorm(1000,0,1)
-plot(xlin,data)
-
-#estimation
-f <- function(par,x,dat) {sum((dat-(par[1]/par[2])*(x^(par[2])-1))^2)}
-opt <- optim(c(-1,1/2),f,x=xlin,dat=data)
-
-#fit
-trans <- function(par,x){(par[1]/par[2])*(x^(par[2])-1)}
-pars <- opt$par
-pars
-points(xlin,sapply(xlin,trans,par=pars),col="red",type="l")
-
-#Now suppose truth = negative inverse sqrt
-alpha <- -2; lambda <- -1/2
-xlin <- 1:1000/1000
-data <- alpha/lambda*(xlin^(lambda)-1) + rnorm(1000,0,10)
-plot(xlin,data)
-
-#estimation
-opt <- optim(c(-1,1/2),f,x=xlin,dat=data)
-pars <- opt$par
-pars
-points(xlin,sapply(xlin,trans,par=pars),col="red",type="l")
-
-#Now suppose truth = positive 1/5
-alpha <- 10; lambda <- 1/5
-xlin <- 1:1000/1000
-data <- alpha*(xlin^(lambda)-1) + rnorm(1000,0,1)
-plot(xlin,data)
-
-#estimation
-opt <- optim(c(1,1/2),f,x=xlin,dat=data)
-pars <- opt$par
-pars
-points(xlin,sapply(xlin,trans,par=pars),col="red",type="l")
-
-#now suppose truth = positive log (limit case for box-cox, lambda -> 0)
-xlin <- 1:1000/1000
-alpha=1
-data <- alpha*log(xlin) + rnorm(1000,0,1)
-plot(xlin,data)
-
-#estimation
-opt <- optim(c(1,1/2),f,x=xlin,dat=data)
-pars <- opt$par
-pars
-points(xlin,sapply(xlin,trans,par=pars),col="red",type="l")
-#conclusion : still fits rather well
-
-
-xlin <- 1:1000/1000
-alpha <- c(-0.5, 1)
-data <- ((xlin + alpha[2])^alpha[1] - 1)/alpha[1]
-plot(xlin, data, type="l")
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2019-02-11 code/check_downloads.R

@@ -1,51 +1,0 @@
-setwd("/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/code")
-
-start <- as.Date("2018-07-30")
-today <- as.Date("2018-08-16")
-all_days <- seq(start, today, by="day")
-
-year <- as.POSIXlt(all_days)$year + 1900
-urls <- paste0("http://cran-logs.rstudio.com/", year, "/", all_days, ".csv.gz")
-
-# only download the files you don't have:
-missing_days <- setdiff(as.character(all_days), 
-                        tools::file_path_sans_ext(dir("CRANlogs"), TRUE))
-
-for (i in 1:length(missing_days)) {
-  print(paste0(i, "/", length(missing_days)))
-  download.file(urls[i], paste0("CRANlogs/", missing_days[i], ".csv.gz"))
-}
-
-
-file_list <- list.files("CRANlogs", full.names=TRUE)
-
-logs <- list()
-for (file in file_list) {
-  print(paste("Reading", file, "..."))
-  clog <- read.table(file, header=TRUE, sep=",", quote='"', dec=".", 
-                     fill=TRUE, comment.char="", as.is=TRUE)
-  logs[[file]] <- clog[clog$package=="gren" & !is.na(clog$package), ]
-}
-rm(clog)
-
-# rbind together all files
-library(data.table)
-dat <- rbindlist(logs)
-
-# add some keys and define variable types
-dat[, date:=as.Date(date)]
-dat[, package:=factor(package)]
-dat[, country:=factor(country)]
-dat[, weekday:=weekdays(date)]
-dat[, week:=strftime(as.POSIXlt(date),format="%Y-%W")]
-setkey(dat, package, date, week, country)
-
-# save data
-save(dat, file="CRANlogs/CRANlogs.RData")
-
-
-plot(unique(dat$date), sapply(unique(dat$date), function(d) {sum(dat$dat==d)}))
-
-
-# for later analyses: load the saved data.table
-load("CRANlogs/CRANlogs.RData")

### 2019-02-11 code/est_param.cpp

@@ -1,211 +1,0 @@
-#include <math.h>
-#include <RcppArmadillo.h>
-
-// [[Rcpp::depends(RcppArmadillo)]]
-
-using namespace Rcpp;
-using namespace arma;
-
-/* This function estimates new variational parameters*/
-// [[Rcpp::export]]
-List est_param(mat xr, mat xu, vec kappa, vec m, int n, int p, vec ciold, 
-               double phi, vec chiold, double lambda2, vec lambdag, 
-               vec lambdagold, bool intercept, bool unpen, bool posterior, 
-               bool elbo, bool start) {
-  
-  vec lambda2vec = lambda2*lambdag;
-  int u = 0;
-  int r = xr.n_cols;
-  int nvars = xr.n_cols;
-  mat x = xr;
-  if(intercept==true && unpen==false) {
-    u = 1;
-    nvars += 1;
-    vec evec(n,fill::ones);
-    x = join_rows(evec, xr);
-  } else if(unpen==true){
-    u = xu.n_cols;
-    nvars += xu.n_cols;
-    x = join_rows(xu,xr);
-  }
-  
-  // initialize c, chi, dsigma and mu
-  double elboval;
-  vec ci(n);
-  vec chi(r);
-  vec dsigma(nvars);
-  mat sigma(nvars,nvars);
-  vec mu(nvars);
-  
-  // used in all situations
-  mat xrt = xr.t();
-  vec h(r);
-  vec om(n);
-  if(start) {
-    // if we are calculating starting values we need the diagonal weight 
-    // matrix with wi = pi/(1-pi) and h=2*lambda
-    om = ciold/(1 - ciold);
-    h = 2*lambda2vec;
-  } else {
-    om = 0.5*m/ciold%tanh(ciold/2);
-    h = lambda2vec + lambda2vec%sqrt(phi/chiold);
-  }
-  vec hinv = 1/h; 
-  mat C = xr.each_row() % hinv.t();
-  mat Ct = C.t();
-  mat D = C*xrt;
-  mat E(n,n), F(r,n), N(nvars,n);
-  
-  double ldetsig;
-  
-  if(p <= n) {
-    sigma = (x.each_col() % om).t()*x;
-    vec happend = h;
-    happend.insert_rows(0, nvars - r, true);
-    sigma.diag() += happend;
-    if(elbo) {
-      ldetsig = - real(log_det(sigma));
-    }
-    sigma = sigma.i();
-    dsigma = sigma.diag();
[…]

### 2019-02-11 code/grEBEN_gene_expr_leukaemia.R

@@ -1,84 +1,0 @@
-################################### preamble ###################################
-# Golub dataset (from original enet paper)                                     #
-# version: 01                                                                  #
-# author: Magnus M?nch                                                         #
-# created: 31-10-2017                                                          #
-# last edited: 31-10-2017                                                      #
-################################################################################
-
-##################################### notes ####################################
-
-### paths
-path.data <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/data/leukaemia_gene-expression_golub/",
-                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/data/leukaemia_gene-expression_golub/")
-path.code <- ifelse(as.character(Sys.info()[1])=="Darwin", "~/EBEN/code/",
-                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/code/")
-path.res <- ifelse(as.character(Sys.info()[1])=="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-
-### libraries
-library(golubEsets)
-library(glmnet)
-library(pROC)
-
-### reading in data
-data(golubTrain)
-data(golubTest)
-
-### data manipulation
-train.data <- t(golubTrain@exprs)
-test.data <- t(golubTest@exprs)
-train.leuk <- as.numeric(golubTrain@phenoData@pData$ALL.AML=="ALL")
-test.leuk <- as.numeric(golubTest@phenoData@pData$ALL.AML=="ALL")
-
-### standardise data
-train.norm <- apply(train.data, 2, function(x) {(x - mean(x))/sd(x)})
-test.norm <- apply(test.data, 2, function(x) {(x - mean(x))/sd(x)})
-
-### data exploration
-ntrain <- nrow(train.norm)
-ptrain <- ncol(train.norm)
-prop.ALL.train <- mean(train.leuk)
-
-### fitting models
-fit1.ridge <- cv.glmnet(train.norm, train.leuk, family="binomial", alpha=0, 
-                        nfolds=ntrain, grouped=FALSE, standardize=FALSE)
[…]

### 2019-02-11 code/grEBEN_metabol.R

@@ -1,179 +1,0 @@
-##############################  preamble  #############################
-# grEBEN on real data                                                 #
-# version: 01                                                         #
-# author: Magnus Munch                                                #
-# created: 04-04-2017                                                 #
-# last edited: 04-04-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#######################################################################
-
-##########################  data description  #########################
-# 248 samples of cases of Alzheimers and controls with measurements   #
-# on 230 metabolites.                                                 #
-#######################################################################
-
-### paths
-path.data <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/data/" ,"~/EBEN/data/"))
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/code/" ,"~/EBEN/code/"))
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/results/" ,"~/EBEN/results/"))
-path.graph <- "/Users/magnusmunch/Documents/PhD/EBEN/graphs/"
-
-### libraries
-library(glmnet)
-library(penalized)
-library(GRridge)
-library(pROC)
-library(Biobase)
-
-### loading data
-load(paste(path.data, "ESetMbolCSFPR2.Rdata", sep=""))
-
-### loading functions 
-source(paste(path.code, "grVBEM.R", sep=""))
-
-# function to cross-validate elastic net
-cv.en <- function(x, y, intercept=intercept) {
-  fit.pen <- cv.pen(x, y, intercept=intercept)
[…]

### 2019-02-11 code/grEBEN_mirna_putri.R

@@ -1,555 +1,0 @@
-##############################  preamble  #############################
-# Putri's miRNA data                                                  #
-# version: 01                                                         #
-# author: Magnus M?nch                                                #
-# created: 06-09-2017                                                 #
-# last edited: 06-09-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#                                                                     #
-#######################################################################
-
-# paths
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/code/" ,"~/EBEN/code/"))
-path.graph <- "/Users/magnusmunch/Documents/PhD/EBEN/graphs/"
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/results/" ,"~/EBEN/results/"))
-path.data <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/data/" ,"~/EBEN/data/"))
-
-### libraries
-library(GRridge)
-library(edgeR)
-library(pROC)
-library(reshape)
-
-### functions
-# source function for variational Bayes
-source(paste(path.code, "grVBEM.R", sep=""))
-
-# version of grridge that doesn't give errors
-source(paste(path.code, "mygrridge.R", sep=""))
-
-# p-values calculation
-logreguni <- function(dat, resp) {
-  #dat <- mirblst[1,];resp <- respblood
-  glm1 <- glm(resp ~ 1, family="binomial")
-  glm2 <- glm(resp ~ 1 + dat, family="binomial")
-  pval <- anova(glm1, glm2, test="Chisq")$Pr[2]
-  return(pval)  
-}
-
-### setting seed for reproducibility
-set.seed(1001)
-
-### loading data
-load(paste(path.data, "normIsoProstate.Rdata", sep=""))
[…]

### 2019-02-11 code/grEBEN_test.R

@@ -1,3716 +1,0 @@
-##############################  preamble  #############################
-# testing grEBEN                                                      #
-# version: 01                                                         #
-# author: Magnus M?nch                                                #
-# created: 27-10-2017                                                 #
-# last edited: 27-10-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#######################################################################
-
-# paths
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin", 
-                                 "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/code/" ,
-                                 "~/EBEN/code/"))
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin", 
-                                "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/" ,
-                                "~/EBEN/results/"))
-
-### libraries
-library(mvtnorm)
-library(GRridge)
-library(pROC)
-
-### source grEBEN
-source(paste(path.code, "grVBEM.R", sep=""))
-source(paste(path.code, "mygrridge.R", sep=""))
-
-# ## simulations
-# ## simulation 1
-# # create data
-# n <- 200
-# p <- 1000
-# G <- 5
-# pblock <- 20
-# rho <- 0.7
-# sigma <- matrix(rho, ncol=pblock, nrow=pblock); diag(sigma) <- 1
-# lambda <- 0.02
-# alpha <- 0.05
-# lambdag <- exp(seq(-1, 1, length.out=G))
-# m <- rep(1, n)
-# part.greben <- list(groups=rep(1:G, each=p/G))
[…]

### 2019-02-11 code/grVBEM_channel3.R

@@ -1,399 +1,0 @@
-##############################  preamble  #############################
-# grMCEM on real data                                                 #
-# version: 01                                                         #
-# author: Magnus M?nch                                                #
-# created: 21-03-2017                                                 #
-# last edited: 21-03-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-# 21-03-2017: Results are used in presentation for channel meeting    #
-#             2017                                                    #
-#######################################################################
-
-##########################  data description  #########################
-# A deep sequencing analysis on small non-coding ribonucleic acid     #
-# (miRNAseq) was performed on 56 samples (24 women with high-grade    #
-# cervical intraepithelial neoplasia (CIN3) and 32 healthy women) for #
-# the purpose of finding relevant screening markers for cervical      #
-# cancer screening. The next generation sequencing analysis resulted  #
-# in 2,576 transcripts. The data was normalized and pre-processed,    #
-# rendering 772 transcripts. More detail description of the data sets #
-# and the preprocessing step are available on the supplementary       #
-# material of this following publication "Better diagnostic           #
-# signatures from RNAseq data through use of auxiliary co-data",      #
-# Bioinformatics (2017).                                              #
-#######################################################################
-
-### paths
[…]

### 2019-02-11 code/grVBEM_sim1.R

@@ -1,112 +1,0 @@
-##############################  preamble  #############################
-# simulations for grVBEM                                              #
-# version: 01                                                         #
-# author: Magnus Münch                                                #
-# created: 15-03-2017                                                 #
-# last edited: 15-03-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#                                                                     #
-#######################################################################
-
-### paths
-path.rcode <- "~/EBEN/code/"
-path.res <- "~/EBEN/results/"
-
-### libraries
-library(Rcpp)
-library(glmnet)
-library(penalized)
-library(GRridge)
-library(mvtnorm)
-library(grpreg)
-library(pROC)
-
-### source grENVB functions
-source(paste(path.rcode, "grVBEM.R", sep=""))
-
-### the simulation
-set.seed(123)
-settings <- cbind(expand.grid(f=c(1.3, 1.6, 2), q=c(0, 0.7, 0.9)), G=10, 
-                  pg=200, meanbeta=0.01, nblock=20, rho=0.3, sigma2=1, m=1)
-n <- 100
-ntest <- 1000
-reps <- 50
-briermat <- aucmat <- matrix(NA, ncol=5, nrow=reps)
-for(s in 1:nrow(settings)) {
-  
-  # specifying parameters for current setting
-  p <- settings$G[s]*settings$pg[s]
-  groups1 <- rep(1:settings$G[s], each=settings$pg[s])
-  groups2 <- CreatePartition(as.factor(groups1))
-  pblock <- p/settings$nblock[s]
-  ubeta <- rep(rev(sapply(0:(settings$G[s] - 1), function(g) {settings$f[s]^(-g)})), 
-               each=settings$pg[s])
-  beta <- ubeta*settings$meanbeta[s]/mean(ubeta)
-  bsigma <- matrix(settings$rho[s], ncol=pblock, nrow=pblock)
-  diag(bsigma) <- settings$sigma2[s]
-  
-  for(r in 1:reps) {
-    
-    print(paste("rep ", r, sep=""))
-    # creating the data
-    x <- do.call(cbind, replicate(settings$nblock[s], 
[…]

### 2019-02-11 code/grVBEM_sim2.R

@@ -1,257 +1,0 @@
-##############################  preamble  #############################
-# simulations for grVBEM                                              #
-# version: 01                                                         #
-# author: Magnus M?nch                                                #
-# created: 15-03-2017                                                 #
-# last edited: 15-03-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#                                                                     #
-#######################################################################
-
-### paths
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/code/" ,"~/EBEN/code/"))
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/results/" ,"~/EBEN/results/"))
-path.data <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/data/" ,"~/EBEN/data/"))
-path.graph <- "/Users/magnusmunch/Documents/PhD/EBEN/graphs/"
-
-### libraries
-library(Rcpp)
-library(glmnet)
-library(penalized)
-library(GRridge)
-library(mvtnorm)
-# library(grpreg)
-library(pROC)
-library(psych)
-
-### functions
-# source grENVB functions
-source(paste(path.code, "grVBEM.R", sep=""))
-
-# function to cross-validate elastic net
-cv.en <- function(x, y, intercept=intercept) {
-  fit.pen <- cv.pen(x, y, intercept=intercept)
-  fit.en <- glmnet(x, y, family="binomial", alpha=fit.pen$alpha[which.min(fit.pen$cvll)],
-                   lambda=fit.pen$lambda[which.min(fit.pen$cvll)])
-  return(fit.en)
-}
-
-### the simulation
-set.seed(6002)
-settings <- cbind(expand.grid(f=c(1.3, 1.6, 2), q=c(0, 0.7, 0.9)), G=10, 
[…]

### 2019-02-11 code/gren_graper_leukaemia.R

@@ -1,28 +1,0 @@
-library(gren)
-
-# load data from the CLL study as provided in MOFAtools
-data("CLL_data", package="MOFAtools")
-
-# use methylation data, gene expression data and drug responses as predictors
-CLL_data <- CLL_data[1:3]
-CLL_data <- lapply(CLL_data,t)
-ngr <- sapply(CLL_data,ncol)
-CLL_data <- Reduce(cbind, CLL_data)
-
-#only include patient samples profiles in all three omics
-CLL_data <- CLL_data[apply(CLL_data,1, function(p) !any(is.na(p))),]
-dim(CLL_data)
-
-# prepare design matrix and response
-x <- scale(CLL_data[,!grepl("D_002", colnames(CLL_data))])
-probs <- rowMeans(CLL_data[,grepl("D_002", colnames(CLL_data))])
-y <- as.numeric(probs > mean(probs))
-part1 <- rep(1:3, times = ngr-c(5,0,0)) # group annotations to drugs, meth and RNA
-
-fit1.gren1 <- gren(x, y, partitions=list(type=part1), alpha=0.05, psel=TRUE)
-fit1.gren2 <- gren(x, y, partitions=list(type=part1), alpha=0.5, psel=TRUE)
-fit1.gren3 <- gren(x, y, partitions=list(type=part1), alpha=0.95, psel=TRUE)
-
-
-
-

### 2019-02-11 code/gren_sim1.R

@@ -1,736 +1,0 @@
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-
-library(mvtnorm)
-library(gren)
-library(pROC)
-library(GRridge)
-library(grpreg)
-library(SGL)
-library(irr)
-
-################################# scenario 1
-n <- 100
-ntest <- 1000
-p <- 1000
-G <- 5
-rho <- 0.5
-rg <- 50
-beta.mean <- 0.07
-f <- 1.6
-
-Sigma <- diag(p)
-for(i in 1:p) {
-  for(j in 1:p) {
-    Sigma[i, j] <- rho^abs(i - j)
-  }
-}
-b1 <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
-beta <- numeric(p)
-
-part1 <- rep(c(1:G), each=p/G)
-csel <- c(seq(1, 10, 2), seq(15, 50, 5), seq(60, 140, 10), seq(160, 200, 20))
-nreps <- 100
-psel1 <- kappa1 <- list(grridge=matrix(NA, nrow=nreps, ncol=length(csel)),
-                        enet1=matrix(NA, nrow=nreps, ncol=100),
-                        enet2=matrix(NA, nrow=nreps, ncol=100),
-                        enet3=matrix(NA, nrow=nreps, ncol=100),
-                        gren1=matrix(NA, nrow=nreps, ncol=100),
-                        gren2=matrix(NA, nrow=nreps, ncol=100),
-                        gren3=matrix(NA, nrow=nreps, ncol=100),
-                        sglasso1=matrix(NA, nrow=nreps, ncol=100),
-                        sglasso2=matrix(NA, nrow=nreps, ncol=100),
-                        sglasso3=matrix(NA, nrow=nreps, ncol=100),
-                        cmcp1=matrix(NA, nrow=nreps, ncol=100),
-                        cmcp2=matrix(NA, nrow=nreps, ncol=100),
-                        cmcp3=matrix(NA, nrow=nreps, ncol=100),
[…]

### 2019-02-11 code/gren_sim3.R

@@ -1,725 +1,0 @@
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-
-library(mvtnorm)
-library(gren)
-library(pROC)
-library(GRridge)
-library(grpreg)
-library(SGL)
-library(irr)
-
-############################# scenario 3
-n <- 100
-ntest <- 1000
-p <- 1000
-alpha <- 0.5
-lambda <- 100
-G <- 5
-q <- 0.5
-
-rho <- 0.7
-Sigma <- matrix(rho, nrow=p/G, ncol=p/G)
-diag(Sigma) <- 1
-
-part1 <- rep(c(1:G), each=p/G)
-csel <- c(seq(1, 10, 2), seq(15, 50, 5), seq(60, 140, 10), seq(160, 200, 20))
-nreps <- 100
-psel3 <- kappa3 <- list(grridge=vector("list", nreps),
-                        enet1=vector("list", nreps),
-                        enet2=vector("list", nreps),
-                        enet3=vector("list", nreps),
-                        gren1=vector("list", nreps),
-                        gren2=vector("list", nreps),
-                        gren3=vector("list", nreps),
-                        sglasso1=vector("list", nreps),
-                        sglasso2=vector("list", nreps),
-                        sglasso3=vector("list", nreps),
-                        cmcp1=vector("list", nreps),
-                        cmcp2=vector("list", nreps),
-                        cmcp3=vector("list", nreps),
-                        gelasso1=vector("list", nreps),
-                        gelasso2=vector("list", nreps),
-                        gelasso3=vector("list", nreps))
-auc3 <- mse3 <- briers3 <- list(ridge=numeric(nreps),
-                                grridge=vector("list", nreps),
-                                enet1=vector("list", nreps),
[…]

### 2019-02-11 code/predict.gren.R

@@ -1,22 +1,0 @@
-# function to make predictions with the gren object
-predict.gren <- function(object, newx, unpenalized, s=NULL, 
-                         type=c("VB", "freq")) {
-  if(is.null(object$freq) & type=="freq") {
-    stop("object should contain a frequentist model fit")
-  } else if(!is.null(s) & !is.numeric(s)) {
-    stop("s is either NULL or a numeric")
-  } else if(!(type %in% c("VB", "freq"))) {
-    stop("type is either VB or freq")
-  }
-  
-  x <- cbind(unpenalized, newx)
-  if(type=="VB") {
-    if(object$args$intercept) {
-      x <- cbind(1, x)
-    } 
-    prob <- as.numeric(1/(1 + exp(-x %*% as.matrix(object$vb.post$mu))))
-  } else {
-    prob <- predict(object$freq, x, s=s, type="response")
-  }
-  return(prob)
-}

### 2019-02-11 code/qtau.R

@@ -1,12 +1,0 @@
-# quantile function of the truncated gamma with shape=1/2, 
-# scale=8*lambda2/lambda1^2, and domain=(1,infinity)
-qtau <- function(p, lambda1, lambda2, log.p=FALSE) {
-  scale <- 8*lambda2/lambda1^2
-  if(log.p) {
-    aux <- pnorm(-sqrt(2/scale), log.p=log.p)
-    tau <- 0.5*scale*qnorm(exp(aux) - exp(aux + p))^2
-  } else {
-    tau <- 0.5*scale*qnorm((1 - p)*pnorm(-sqrt(2/scale)))^2
-  }
-  return(tau)
-}

### 2019-02-11 code/rcpptest.R

@@ -1,139 +1,0 @@
-##############################  preamble  #############################
-# code to test the Rcpp functions                                     #
-# version: 01                                                         #
-# author: Magnus Münch                                                #
-# created: 25-11-2016                                                 #
-# last edited: 05-12-2016                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#                                                                     #
-#######################################################################
-
-### paths
-path.code <- "C:/Users/Magnus/Documents/phd/ENVB/code/"
-
-### libraries
-library(Rcpp)
-library(penalized)
-library(mvtnorm)
-library(GeneralizedHyperbolic)
-
-### functions
-# c++ version
-sourceCpp(paste(path.code, "ENVB2.cpp", sep=""))
-sourceCpp(paste(path.code, "ENgibbs.cpp", sep=""))
-
-# R version
-est_param2 <- function(x, kappa, m, n, p, ciold, phi, chiold, lambda2, intercept) {
-  
-  h <- (1 + sqrt(phi/chiold))*lambda2
-  om <- (0.5*m/ciold)*tanh(ciold/2)
-  omsq <- sqrt(om)
-  hinv <- 1/h
-  sigma <- matrix(0, ncol=p + 1, nrow=p + 1)
-  xaug <- cbind(1, x)
-  
-  # using the Woodbury identity 
-  if(intercept) {
-    invtrom <- 1/sum(om)
-    Omadj <- diag(om) - invtrom*as.matrix(om) %*% t(as.matrix(om))
-    xhinv <- t(t(x)*hinv)
-    Ainv <- diag(hinv) - t(xhinv) %*% Omadj %*% solve(diag(n) + x %*% t(xhinv) %*% Omadj) %*% xhinv
-    ainvxom <- Ainv %*% t(x) %*% as.matrix(om)
[…]

### 2019-02-11 code/renet.R

@@ -1,14 +1,0 @@
-# sampling function of the elastic net prior distribution
-renet <- function(n, lambda1=1, lambda2=1) {
-  if(lambda1==0) {
-    beta <- rnorm(n, 0, sqrt(1/lambda2))
-  } else if(lambda2==0) {
-    u <- runif(n) - 0.5
-    beta <- -2*sign(u)*log(1 - 2*abs(u))/lambda1
-  } else {
-    tau <- rtau(n, lambda1, lambda2)
-    sigma <- (tau - 1)/(lambda2*tau)
-    beta <- rnorm(n, 0, sqrt(sigma))
-  }
-  return(beta)
-}

### 2019-02-11 code/coef.gren.R
moved old code, results, and data to separate untracked folders

@@ -1,14 +1,0 @@
-# retrieve coefficients from gren estimated model
-coef.gren <- function(object, s, type=c("VB", "freq")) {
-  if(!is.null(s) & !is.numeric(s)) {
-    stop("s is either NULL or a numeric")
-  } else if(!(type %in% c("VB", "freq"))) {
-    stop("type is either VB or freq")
-  }
-  if(type=="VB") {
-    coefs <- list(mu=object$vb.post$mu, sigma=object$vb.post$sigma)
-  } else {
-    coefs <- coef(object$freq, s=s)
-  }
-  return(coefs)
-}

### 2019-02-11 code/data_test.R

@@ -1,76 +1,0 @@
-
-path.data <- as.character("/Users/magnusmunch/Downloads/MTBLS315_20170922_111318/")
-object.bsi <- read.table(paste(path.data, "s_NMFI and BSI diagnosis.txt", sep=""), 
-                       header=TRUE, stringsAsFactors=FALSE, fill=TRUE)
-object.gc <- read.table(paste(path.data, "m_GC_nmfi_and_bsi_diagnosis_v2_maf.tsv", sep=""), 
-                      header=TRUE, stringsAsFactors=FALSE, fill=TRUE)
-object.lc <- read.table(paste(path.data, "m_LC_nmfi_and_bsi_diagnosis_v2_maf.tsv", sep=""), 
-                      header=TRUE, stringsAsFactors=FALSE, fill=TRUE, sep="\t", quote="", 
-                      na.strings="")
-object.uplc_neg <- read.table(paste(path.data, "m_UPLC_NEG_nmfi_and_bsi_diagnosis_v2_maf.tsv", sep=""), 
-                            header=TRUE, stringsAsFactors=FALSE, fill=TRUE, sep="\t", quote="", 
-                            na.strings="")
-object.uplc_pos <- read.table(paste(path.data, "m_UPLC_POS_nmfi_and_bsi_diagnosis_v2_maf.tsv", sep=""), 
-                            header=TRUE, stringsAsFactors=FALSE, fill=TRUE, sep="\t", quote="", 
-                            na.strings="")
-
-meta.uplc_neg <- read.table(paste(path.data, "a_UPLC_NEG_nmfi_and_bsi_diagnosis.txt", sep=""), 
-                      header=TRUE, stringsAsFactors=FALSE, fill=TRUE)
-meta.uplc_pos <- read.table(paste(path.data, "a_UPLC_POS_nmfi_and_bsi_diagnosis.txt", sep=""), 
-                            header=TRUE, stringsAsFactors=FALSE, fill=TRUE)
-
-samp.uplc_neg <- meta.uplc_neg$Sample.Name[match(sub("X", "", colnames(object.uplc_neg)[
-  22:ncol(object.uplc_neg)]), meta.uplc_neg$MS.Assay.Name)]
-samp.uplc_pos <- meta.uplc_pos$Sample.Name[match(sub("X", "", colnames(object.uplc_pos)[
-  22:ncol(object.uplc_pos)]), meta.uplc_pos$MS.Assay.Name)]
[…]

### 2019-02-11 code/grEBEN_metabol2.R

@@ -1,204 +1,0 @@
-##############################  preamble  #############################
-# grEBEN on real data                                                 #
-# version: 01                                                         #
-# author: Magnus Munch                                                #
-# created: 04-04-2017                                                 #
-# last edited: 04-04-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#######################################################################
-
-##########################  data description  #########################
-# 248 samples of cases of Alzheimers and controls with measurements   #
-# on 230 metabolites.                                                 #
-#######################################################################
-
-### paths
-path.data <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/data/" ,"~/EBEN/data/"))
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/code/" ,"~/EBEN/code/"))
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/results/" ,"~/EBEN/results/"))
-
-### libraries
-library(glmnet)
-library(penalized)
-library(GRridge)
-library(pROC)
-library(Biobase)
-
-### loading data
-load(paste(path.data, "ESetMbolCSFPR2.Rdata", sep=""))
-
-### loading functions 
-source(paste(path.code, "grVBEM.R", sep=""))
-
-# cross-validate to find penalty parameters
-cv.pen <- function(x, y, intercept, psel) {
-  
-  n <- nrow(x)
-  seq.alpha <- seq(0.01, 0.99, length.out=50)
-  seq.lambda <- seq.df <- seq.cvm <- numeric(length(seq.alpha))
-  
[…]

### 2019-02-11 code/grEBEN_methylation_lung_cancer.R

@@ -1,245 +1,0 @@
-################################### preamble ###################################
-# lung cancer data from Putri                                                  #
-# version: 01                                                                  #
-# author: Magnus M?nch                                                         #
-# created: 27-11-2017                                                          #
-# last edited: 27-11-2017                                                      #
-################################################################################
-
-##################################### notes ####################################
-#                                                                              #
-################################################################################
-
-### paths
-path.data <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/data/lung_cancer_methylation",
-                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/data/lung_cancer_methylation/")
-path.code <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/code/",
-                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/code/")
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-
-### loading libraries
-library(GRridge)
-library(glmnet)
-library(pROC)
-library(limma)
-
-### loading function
-source(paste(path.code, "grVBEM.R", sep=""))
-
-### loading data
-load(paste(path.data, "datSputum.Rdata", sep=""))
-
-### data preparation
-resp <- as.numeric(pd$Sample_Group=="LungCancer")
-rownames(pd) <- pd$Sample_Name
-rownames(AnnoFile) <- AnnoFile$Name
-eset <- ExpressionSet(betaPreProc, as(pd, "AnnotatedDataFrame"), 
[…]

### 2019-02-11 code/grVBEM.R

@@ -1,1248 +1,0 @@
-##############################  preamble  #############################
-# grMCEM implemented in R and C++                                     #
-# version: 01                                                         #
-# author: Magnus M?nch                                                #
-# created: 14-03-2017                                                 #
-# last edited: 17-10-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-# 28-03-2017: Added multiple partitions                               #
-# 15-03-2017: First implementation of group-regularized elastic net   #
-#             with lambda1 penalty multiplier, the square root of     #
-#             lambda2 penalty multiplier                              #
-#######################################################################
-
-# paths
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin", "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/code/" ,
-                                 "~/EBEN/code/"))
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin", "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/" ,
-                                "~/EBEN/results/"))
-
-### libraries
-library(Rcpp)
-library(glmnet)
-
-### functions
-# source function for variational Bayes
-sourceCpp(paste(path.code, "ENVB2.cpp", sep=""))
-
-# various functions of the elastic net prior
-dtau <- function(x, lambda1, lambda2, log=FALSE) {
-  scale <- 8*lambda2/lambda1^2
-  if(log) {
-    dens <- -(log(0.5) - 0.5*log(x*(x > 1)) - 0.5*log(scale) - 0.5*log(pi) - x/scale - 
-                pnorm(-sqrt(2/scale), log.p=TRUE))
-  } else {
[…]

### 2019-02-11 code/grVBEM_channel.R

@@ -1,190 +1,0 @@
-##############################  preamble  #############################
-# grMCEM on real data                                                 #
-# version: 01                                                         #
-# author: Magnus Münch                                                #
-# created: 21-03-2017                                                 #
-# last edited: 21-03-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-# 21-03-2017: Results are used in presentation for channel meeting    #
-#             2017                                                    #
-#######################################################################
-
-##########################  data description  #########################
-# A deep sequencing analysis on small non-coding ribonucleic acid     #
-# (miRNAseq) was performed on 56 samples (24 women with high-grade    #
-# cervical intraepithelial neoplasia (CIN3) and 32 healthy women) for #
-# the purpose of finding relevant screening markers for cervical      #
-# cancer screening. The next generation sequencing analysis resulted  #
-# in 2,576 transcripts. The data was normalized and pre-processed,    #
-# rendering 772 transcripts. More detail description of the data sets #
-# and the preprocessing step are available on the supplementary       #
-# material of this following publication "Better diagnostic           #
-# signatures from RNAseq data through use of auxiliary co-data",      #
-# Bioinformatics (2017).                                              #
-#######################################################################
-
-### paths
-path.rcode <- "~/EBEN/code/"
-path.res <- "~/EBEN/results/"
-path.data <- "~/EBEN/data/"
-
-### libraries
-library(glmnet)
-library(penalized)
[…]

### 2019-02-11 code/grVBEM_test.R

@@ -1,950 +1,0 @@
-##############################  preamble  #############################
-# testing grMCEM                                                      #
-# version: 01                                                         #
-# author: Magnus M?nch                                                #
-# created: 15-03-2017                                                 #
-# last edited: 15-03-2017                                             #
-#######################################################################
-
-###############################  notes  ###############################
-#                                                                     #
-#######################################################################
-
-### paths
-path.code <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/code/" ,"~/EBEN/code/"))
-path.graph <- "/Users/magnusmunch/Documents/PhD/EBEN/graphs/"
-path.data <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/data/" ,"~/EBEN/data/"))
-path.res <- as.character(ifelse(Sys.info()[1]=="Darwin","/Users/magnusmunch/Documents/PhD/EBEN/results/" ,"~/EBEN/results/"))
-
-## libraries
-library(GRridge)
-library(mvtnorm)
-library(grpreg)
-library(SGL)
-library(psych)
-library(glmnet)
-
-# source grENVB functions
-source(paste(path.code, "grVBEM.R", sep=""))
-
-### simulation 4 (one partition)
-set.seed(123)
-n <- 200
-p <- 150
-G <- 3
-groups <- rep(1:G, each=p/G)
-x <- matrix(rnorm(n*p), ncol=p, nrow=n)
-lambda1 <- 1
-lambda2 <- 1
-lambdag <- c(0.2, 3, 1/(0.2*3))
-m <- rep(1, n)
-b0 <- rnorm(p + 1)
-sigma0 <- diag(rchisq(p + 1, 1))
[…]

### 2019-02-11 code/gren.R

@@ -1,365 +1,0 @@
-# group-regularized elastic net function
-gren <- function(x, y, m, unpenalized=NULL, partitions=NULL, alpha=0.5, 
-                 lambda=NULL, intercept=TRUE, monotone=NULL, psel=TRUE, 
-                 posterior=FALSE, nfolds=nrow(x), foldid=NULL, trace=TRUE,
-                 control=list(epsilon=0.001, maxit=500, maxit.opt=1000, 
-                              maxit.vb=100)) {
-  
-  # save argument list
-  argum <- formals(gren)
-  
-  # change some input for convenience
-  if(is.data.frame(x)) {
-    names.xr <- colnames(x)
-    x <- as.matrix(x)
-  } else {
-    names.xr <- NULL
-  }
-  if(is.data.frame(unpenalized)) {
-    names.xu <- colnames(unpenalized)
-    if(intercept) {names.xu <- c("Intercept", names.xu)}
-    unpenalized <- as.matrix(unpenalized)
-  } else {
-    names.xu <- NULL
-  }
-  if(is.vector(partitions) & is.atomic(partitions)) {
-    partitions <- list(partition1=partitions)
-  }
-  
-  # check input
-  if(!is.numeric(x) | !is.numeric(y)) {
-    stop("only numerical input data is supported at the moment")
-  } else if(!(is.null(ncol(y)) | ncol(y)==2)) {
-    stop("y is either a vector or matrix with two columns")
-  } else if(!is.null(unpenalized) | !is.numeric(unpenalized)) {
-    stop("only numerical unpenalized data or no unpenalized data is 
-         supported at the moment")
-  } else if(!is.list(partitions)) {
-    stop("partitions should be either a list of partitions or one partition as 
-         a numeric vector")
-  } else if(!is.numeric(alpha) | length(alpha)!=1 | (alpha < 0) | (alpha > 1)) {
-    stop("alpha should be of length one and a value between 0 and 1")
-  } else if(!is.null(lambda) & !(is.vector(lambda) & is.atomic(lambda))) {
-    stop("lambda should be either NULL or a numeric vector")
-  } else if(!is.logical(intercept)) {
-    stop("logical should be either TRUE or FALSE")
-  } else if(!is.null(monotone) & 
-            !(is.logical(monotone) & length(monotone)==length(partitions)) &
-            !(is.list(monotone) & length(monotone)==2 & 
[…]

### 2019-02-11 code/gren_sim2.R

@@ -1,750 +1,0 @@
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-
-library(mvtnorm)
-library(gren)
-library(pROC)
-library(GRridge)
-library(grpreg)
-library(SGL)
-library(irr)
-
-############################# scenario 2
-# similar as model 1 in yuan and lin (2006)
-n <- 100
-ntest <- 1000
-p <- 900
-G <- 300
-rho <- 0.5
-Gactive <- 50
-
-Sigma <- diag(G)
-for(i in 1:G) {
-  for(j in 1:G) {
-    Sigma[i, j] <- rho^abs(i - j)
-  }
-}
-beta.active.mean <- 0.3
-
-nclass <- 3
-q <- c(-Inf, qnorm((c(1:nclass)/nclass)[-nclass]), Inf)
-beta <- numeric(p)
-beta.active <- seq(beta.active.mean*Gactive/(Gactive/2 + 0.5),
-                   beta.active.mean/(Gactive/2 + 0.5), length.out=Gactive)
-for(g in 1:Gactive) {
-  id <- ((g - 1)*nclass + 1):((g - 1)*nclass + nclass - 1)
-  beta[id] <- beta.active[g]
-}
-
-part1 <- rep(c(1:G), each=p/G)
-csel <- c(seq(1, 10, 2), seq(15, 50, 5), seq(60, 140, 10), seq(160, 200, 20))
-nreps <- 100
-psel2 <- kappa2 <- list(grridge=matrix(NA, nrow=nreps, ncol=length(csel)),
-                        enet1=matrix(NA, nrow=nreps, ncol=100),
-                        enet2=matrix(NA, nrow=nreps, ncol=100),
-                        enet3=matrix(NA, nrow=nreps, ncol=100),
-                        gren1=matrix(NA, nrow=nreps, ncol=100),
-                        gren2=matrix(NA, nrow=nreps, ncol=100),
-                        gren3=matrix(NA, nrow=nreps, ncol=100),
-                        sglasso1=matrix(NA, nrow=nreps, ncol=100),
-                        sglasso2=matrix(NA, nrow=nreps, ncol=100),
[…]

### 2019-02-11 code/gren_sim5.R

@@ -1,600 +1,0 @@
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-
-library(mvtnorm)
-library(gren)
-library(pROC)
-library(GRridge)
-library(grpreg)
-library(SGL)
-library(irr)
-
-##################### scenario 5
-n <- 100
-p <- 1000
-ntest <- 1000
-
-G <- 10
-rho <- 0.5
-Sigma <- diag(p)
-for(i in 1:p) {
-  for(j in 1:p) {
-    Sigma[i, j] <- rho^abs(i - j)
-  }
-}
-
-beta.group <- c(rep(0, 8), 0.2, 0.5)
-q.zero <- 0.85
-
-part1 <- rep(c(1:G), each=p/G)
-csel <- c(seq(1, 8, 1), seq(10, 20, 2), seq(25, 60, 5), seq(70, 140, 10))
-nreps <- 100
-psel5 <- kappa5 <- list(grridge=vector("list", nreps),
-                        enet1=vector("list", nreps),
-                        enet2=vector("list", nreps),
-                        enet3=vector("list", nreps),
-                        gren1=vector("list", nreps),
-                        gren2=vector("list", nreps),
-                        gren3=vector("list", nreps),
-                        sglasso1=vector("list", nreps),
-                        sglasso2=vector("list", nreps),
-                        sglasso3=vector("list", nreps),
-                        cmcp1=vector("list", nreps),
-                        cmcp2=vector("list", nreps),
-                        cmcp3=vector("list", nreps),
-                        gelasso1=vector("list", nreps),
-                        gelasso2=vector("list", nreps),
-                        gelasso3=vector("list", nreps))
-auc5 <- mse5 <- briers5 <- list(ridge=numeric(nreps),
-                                grridge=vector("list", nreps),
-                                enet1=vector("list", nreps),
[…]

### 2019-02-11 code/penalty_parameter_estimation.R

@@ -1,268 +1,0 @@
-################################### preamble ###################################
-# penalty parameter selection simulations                                      #
-# version: 01                                                                  #
-# author: Magnus M?nch                                                         #
-# created: 01-12-2017                                                          #
-# last edited: 01-12-2017                                                      #
-################################################################################
-
-##################################### notes ####################################
-#                                                                              #
-################################################################################
-
-### paths
-path.data <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/data/lung_cancer_methylation",
-                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/data/lung_cancer_methylation/")
-path.code <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/code/",
-                    "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/code/")
-path.res <- ifelse(as.character(Sys.info()[1])!="Darwin", "~/EBEN/results/",
-                   "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/results/")
-path.graph <- "/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/"
-
-### loading libraries
-library(glmnet)
-library(psych)
-library(mvtnorm)
-
-### simulations
-set.seed(3002)
-n <- c(50, 100)
-p <- 1000
-k <- c(20, 50, 100)
-pblock <- 20
-rho <- c(0, 0.7, 0.9)
-settings <- expand.grid(n=n, p=p, k=k, pblock=pblock, rho=rho)
-
-nreps <- 50
[…]

### 2019-02-19 docs/supplement.Rnw
lymph node example worked out

@@
 \begin{document}
+
+% load packages and set options
+<<settings, include=FALSE, echo=FALSE>>=
+library(cambridge)
+opts_knit$set(base.dir="../figs")
+opts_chunk$set(fig.align='center', echo=FALSE, fig.path="../figs/", cache=FALSE,
+               message=FALSE)
+@
+
+% read figure code
+<<figures, include=FALSE>>=
+read_chunk("../code/figures.R")
+@
+
   \maketitle
   \noindent
-1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
+  1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
+  research institute, VU University Medical Center, PO Box 7057, 1007 MB
   Amsterdam, The Netherlands\\
@@
 \end{figure}
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
+	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
+	\label{fig:col_bar}
+\end{figure}
+
+Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}.
+All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS.
+In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC.
+This is especially true for the larger models.
+\texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models.
+A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly.
+This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation.
+In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used.
+Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting.
+Note that models with few features are often desirable for clinical implementation.
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
+	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
+	\label{fig:col_per}
+\end{figure}
+
+\section{Application to RNAseq in lymph node cancer}\label{sec:applymph}
+
+CoRF data:
+
+<<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Cross-validated AUC in lymph node metastasis example, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+@
 

### 2019-02-19 code/figures.R

@@ -3,10 +3,10 @@
 # ---- boxplot_metabolomics_alzheimer_res1_auc ---- 
 library(pROC)
 library(Biobase)
-labels1 <- c(unique(as.character(feat$Platform))[-c(4, 5)], "Oxidative Stress")
 res1 <- read.table("results/metabolomics_alzheimer_res1.csv", 
                    stringsAsFactors=FALSE)
 load("data/ESetMbolCSFPR2.Rdata")
+feat <- fData(ESetMbolCSFPR2)
 pheno <- pData(ESetMbolCSFPR2)
 pheno.apoe <- pheno[(pheno$D_diag_name=="Probable AD" & pheno$APOE=="E4YES") |
                       (pheno$D_diag_name=="Subjectieve klachten" & 
@@ -17,6 +17,7 @@ auc1 <- sapply(res1[grepl("pred", rownames(res1)), ], function(s) {
   pROC::auc(alzheim.apoe, s)})
 psel1 <- colMeans(res1[grepl("psel", rownames(res1)), ])
 
+labels1 <- c(unique(as.character(feat$Platform))[-c(4, 5)], "Oxidative Stress")
 plot(psel1, auc1)
 boxplot(res1[grepl("psel", rownames(res1)), ])
 
@@ -24,18 +25,98 @@ barplot(rbind(fit1.gren1$lambdag$part, fit1.grridge$lambdamults$part),
         beside=TRUE)
 abline(h=1, lty=2)
 
+# ---- lines_rnaseq_lymph_node_metastasis_res1_auc ----
+library(CoRF)
+library(pROC)
+library(grpreg)
+load("../results/rnaseq_lymph_node_metastasis_fit1.Rdata")
+res1 <- read.table("../results/rnaseq_lymph_node_metastasis_res1.csv", 
+                   stringsAsFactors=FALSE)
 
-
-
+methods1 <- c("ridge", "grridge", "gren", "enet", "cmcp", "gel")
+pred1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="pred", ])
+psel1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="psel", ])
+ytest <- as.numeric(levels(RespValidation))[RespValidation]
 auc1 <- apply(pred1, 2, function(m) {pROC::auc(ytest, m)})
+plot.data1 <- lapply(c(t(outer(methods1[-c(1, 2)], 1:3, paste, sep=""))), 
[…]

### 2019-02-19 code/rnaseq_lymph_node_metastasis.R

@@ -10,6 +10,16 @@ if(!("CoRF" %in% installed.packages())) {
                  auth_token=Sys.getenv("GITHUB_PAT"))
 }
 
+if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
+   substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
+  if(!("devtools" %in% installed.packages())) {
+    install.packages("devtools")
+  }
+  library(devtools)
+  install_github("magnusmunch/gren/rpackage", local=FALSE,
+                 auth_token=Sys.getenv("GITHUB_PAT"))
+}
+
 ### libraries
 library(CoRF)
 library(GRridge)
@@ -43,6 +53,7 @@ corpart0 <- grridge2gren(CreatePartition(CDT$Corrs, ngroup=5, decreasing=TRUE))
 #roepmanpart <- CDT$RoepmanGenes+1
 partLNM <- list(corr=corpart0, pv=pvpart0)
 
+################################### model 1 ####################################
 ### fitting the models
 nfolds <- 1
 set.seed(2018)
@@ -73,53 +84,49 @@ fit1.gel2 <- cv.grpreg(x, y, as.numeric(as.factor(paste(corpart0, pvpart0))),
 fit1.gel3 <- cv.grpreg(x, y, as.numeric(as.factor(paste(corpart0, pvpart0))), 
                        penalty="gel", family="binomial", alpha=0.95)
 
+save(fit1.grridge, fit1.gren1, fit1.gren2, fit1.gren3, 
+     fit1.cmcp1, fit1.cmcp2, fit1.cmcp3, fit1.gel1, fit1.gel2,
+     fit1.gel3, file="results/rnaseq_lymph_node_metastasis_fit1.Rdata")
+
 ### obtaining predictions for test data
 ytest <- as.numeric(levels(RespValidation))[RespValidation]
 xtest <- ValidationData[, selectgenes]
 pred1 <- data.frame(ridge=predict.grridge(fit1.grridge, t(xtest))[, 1], 
                     grridge=predict.grridge(fit1.grridge, t(xtest))[, 2],
-                    gren1=as.numeric(predict(fit1.gren1, xtest, 
-                                             type="groupreg")),
-                    gren2=as.numeric(predict(fit1.gren2, xtest, 
-                                             type="groupreg")), 
-                    gren3=as.numeric(predict(fit1.gren3, xtest, 
[…]

### 2019-02-19 code/functions.R
lymph node example worked out

@@ -1,40 +1,53 @@
-### function to find a specific number of features
-froot <- function(lambda, psel, x, y, part, alpha, type) {
-  if(type=="sgl") {
-    fit <- SGL(list(x=x, y=y), part, type="logit", alpha=alpha, standardize=FALSE,
-               lambdas=lambda, nlam=1, min.frac=0.001)
-    out <- psel - sum(fit$beta!=0)
-  } else if(type=="cmcp") {
-    fit <- grpreg(x, y, part, "cMCP", "binomial", lambda=lambda, alpha=alpha)
-    out <- psel - sum(fit$beta[-1, ]!=0)
-  } else if(type=="gel") {
-    fit <- grpreg(x, y, part, "gel", "binomial", lambda=lambda, alpha=alpha)
-    out <- psel - sum(fit$beta[-1, ]!=0)
-  }
+### gives back the number of selected features in grpreg for specific lambda
+froot <- function(lambda, psel, inp) {
+  # fit the model with specific lambda
+  fit <- do.call("grpreg", c(inp, lambda=lambda))
+  
+  # calculate the number of selected features
+  out <- psel - sum(fit$beta[-1, ]!=0)
   return(out)
 }
 
+## grpreg that selects specific number of features
 sel.grpreg <- function(X, y, group=1:ncol(X), 
                        penalty=c("grLasso", "grMCP", "grSCAD", "gel", "cMCP"), 
                        family=c("gaussian", "binomial", "poisson"),
-                       nlambda=100, lambda, 
                        lambda.min={if (nrow(X) > ncol(X)) 1e-4 else .05},
                        log.lambda = TRUE, alpha=1, eps=1e-4, max.iter=10000, 
                        dfmax=p, gmax=length(unique(group)), 
                        gamma=ifelse(penalty == "grSCAD", 4, 3), tau = 1/3, 
-                       group.multiplier, warn=TRUE, returnX = FALSE) {
+                       group.multiplier=rep(1, length(unique(group))), 
+                       warn=TRUE, returnX = FALSE, psel) {
   
+  # store the arguments
+  # args <- as.list(sys.call())
+  inp <- lapply(as.list(match.call(definition=sys.function(),
+                                   call=sys.call()))[-1], eval)
+  psel <- inp$psel
+  inp <- inp[-length(inp)]
+
+  # find range of lambdas using built-in grpreg functions
[…]

### 2019-02-19 code/metabolomics_alzheimer.R

@@ -14,7 +14,6 @@ if(!("gren" %in% installed.packages())) {
 library(gren)
 library(GRridge)
 library(grpreg)
-library(SGL)
 library(Biobase)
 
 ### load data
@@ -27,13 +26,22 @@ pheno.apoe <- pheno[(pheno$D_diag_name=="Probable AD" & pheno$APOE=="E4YES") |
                       (pheno$D_diag_name=="Subjectieve klachten" & 
                          pheno$APOE=="E4NO"), ]
 
-
 metabol <- t(exprs(ESetMbolCSFPR2))
 metabol.apoe <- metabol[(pheno$D_diag_name=="Probable AD" & 
                            pheno$APOE=="E4YES") |
                           (pheno$D_diag_name=="Subjectieve klachten" & 
                              pheno$APOE=="E4NO"), ]
-      
+
+### randomly split in test and train data
+set.seed(2019)
+id.train <- c(sample(which(pheno.apoe$D_diag_name=="Probable AD"), 
+                     size=floor(sum(pheno.apoe$D_diag_name=="Probable AD")/2)),
+              sample(which(pheno.apoe$D_diag_name=="Subjectieve klachten"), 
+                     size=floor(sum(
+                       pheno.apoe$D_diag_name=="Subjectieve klachten")/2)))
+train <- pheno.apoe[id.train, ]     
+test <- pheno.apoe[-id.train, ]     
+
 ### create co-data
 feat <- fData(ESetMbolCSFPR2)
 
@@ -356,10 +364,6 @@ fit3.gren2 <- gren(x, y, partitions=list(part=part), alpha=0.5, trace=FALSE)
 fit3.gren3 <- gren(x, y, partitions=list(part=part), alpha=0.95, trace=FALSE)
 
 fit3.grridge <- grridge(t(x), y, list(part=split(1:p, part)))
-
-# fit3.sgl1 <- cvSGL(list(x=x, y=y), part, type="logit", alpha=0.05)
-# fit3.sgl2 <- cvSGL(list(x=x, y=y), part, type="logit", alpha=0.5)
-# fit3.sgl3 <- cvSGL(list(x=x, y=y), part, type="logit", alpha=0.95)
 
 fit3.cmcp1 <- cv.grpreg(x, y, part, penalty="cMCP", family="binomial", 
                         alpha=0.05)

### 2019-02-19 code/metabolomics_alzheimer.R
fixed issue of control check warning

@@ -39,8 +39,6 @@ id.train <- c(sample(which(pheno.apoe$D_diag_name=="Probable AD"),
               sample(which(pheno.apoe$D_diag_name=="Subjectieve klachten"), 
                      size=floor(sum(
                        pheno.apoe$D_diag_name=="Subjectieve klachten")/2)))
-train <- pheno.apoe[id.train, ]     
-test <- pheno.apoe[-id.train, ]     
 
 ### create co-data
 feat <- fData(ESetMbolCSFPR2)
@@ -53,30 +51,35 @@ part.RSDqc <- CreatePartition(feat$RSDqc, ngroup=3)
 quality <- rep(c(1:3), sapply(part.RSDqc, length))[unlist(part.RSDqc)]
 
 # sds gives groups based on the standard deviations of the features
-part.sds.apoe <- CreatePartition(apply(metabol.apoe, 1, sd), ngroup=3)
-sds.apoe <- rep(c(1:3), sapply(part.sds.apoe, length))[unlist(part.sds.apoe)]
+part.sds.apoe.train <- CreatePartition(apply(metabol.apoe[id.train, ], 1, sd), 
+                                       ngroup=3)
+sds.apoe.train <- rep(c(1:3), sapply(part.sds.apoe.train, 
+                                     length))[unlist(part.sds.apoe.train)]
 
 # degree gives the degree classes of the differential network
 degree <- NetworkDegreeClass[, 2]
 
 ### transform data
 alzheim.apoe <- as.numeric(pheno.apoe$D_diag_name) - 1
-metabol.apoe.scaled <- scale(metabol.apoe)
 
 ################################### model 1 ####################################
 ### fitting the models
 set.seed(2018)
-y <- alzheim.apoe
-x <- metabol.apoe.scaled
+y.train <- alzheim.apoe[id.train]
+x.train <- scale(metabol.apoe[id.train, ])
+y.test <- alzheim.apoe[-id.train]
+x.test <- scale(metabol.apoe[-id.train, ])
 part <- platformcode
-n <- nrow(x)
-p <- ncol(x)
+p <- ncol(x.train)
 
-fit1.gren1 <- gren(x, y, partitions=list(part=part), alpha=0.05, trace=FALSE)
-fit1.gren2 <- gren(x, y, partitions=list(part=part), alpha=0.5, trace=FALSE)
[…]

### 2019-02-19 rpackage/R/gren.R
fixed issue of control check warning

@@ -24,16 +24,16 @@
   }
   
   # if only one control is set, replace the rest
-  if("epsilon"!=names(control)) {
+  if(!("epsilon" %in% names(control))) {
     control$epsilon <- 0.001
   }
-  if("maxit"!=names(control)) {
+  if(!("maxit" %in% names(control))) {
     control$maxit <- 500
   }
-  if("maxit.opt"!=names(control)) {
+  if(!("maxit.opt" %in% names(control))) {
     control$maxit.opt <- 1000
   }
-  if("maxit.vb"!=names(control)) {
+  if(!("maxit.vb" %in% names(control))) {
     control$maxit.vb <- 100
   }
   

### 2019-02-19 docs/manuscript.Rnw
exchanged cross-validation from metabolomics_alzheimer.R example with train/test data setup

@@
 \begin{document}
+
+% load packages and set options
+<<settings, include=FALSE, echo=FALSE>>=
+library(cambridge)
+opts_knit$set(base.dir="../figs")
+opts_chunk$set(fig.align='center', echo=FALSE, fig.path="../figs/", cache=FALSE,
+               message=FALSE)
+@
+
+% read figure code
+<<figures, include=FALSE>>=
+read_chunk("../code/figures.R")
+@
 
@@
 
-	\section{Application}\label{sec:app}
+	\section{Application to metabolomics in alzheimer}\label{sec:app}
+
+<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Cross-validated AUC in alzheimer example using metabolomics class co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+@
+
+<<lines_metabolomics_alzheimer_res2_auc, fig.cap="Cross-validated AUC in alzheimer example using quality score co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+@
+
+<<lines_metabolomics_alzheimer_res3_auc, fig.cap="Cross-validated AUC in alzheimer example using node degree co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+@
+
+<<lines_metabolomics_alzheimer_res4_auc, fig.cap="Cross-validated AUC in alzheimer example using all co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+@
+
+<<barplots_metabolomics_alzheimer_res4_auc, fig.cap="Estimated penalty multipliers in alzheimer example using all co-data for a) metabolomics class, b) quality score, and c) node degree", out.width="100%", fig.asp=1>>=
+@
 

### 2019-02-19 code/rnaseq_lymph_node_metastasis.R

@@ -127,6 +127,5 @@ psel1 <- c(ridge=p, grridge=p,
            gel3=colSums(fit1.gel1$fit$beta[-1, ]!=0))
 
 res1 <- rbind(pred1, psel1)
-rownames(res1) <- c(paste0("pred", c(1:length(ytest))), 
-                    paste0("psel", 1:nfolds))
+rownames(res1) <- c(paste0("pred", c(1:length(ytest))), paste0("psel", 1))
 write.table(res1, file="results/rnaseq_lymph_node_metastasis_res1.csv")

### 2019-02-19 code/figures.R
exchanged cross-validation from metabolomics_alzheimer.R example with train/test data setup

@@ -1,29 +1,440 @@
 ################################## manuscript ##################################
 # ---- figures ----
-# ---- boxplot_metabolomics_alzheimer_res1_auc ---- 
+# ---- lines_metabolomics_alzheimer_res1_auc ----
 library(pROC)
-library(Biobase)
-res1 <- read.table("results/metabolomics_alzheimer_res1.csv", 
+library(grpreg)
+load("../results/metabolomics_alzheimer_fit1.Rdata")
+res1 <- read.table("../results/metabolomics_alzheimer_res1.csv", 
                    stringsAsFactors=FALSE)
-load("data/ESetMbolCSFPR2.Rdata")
+
+methods2 <- c("ridge", "grridge", "gren", "enet", "sgl", "cmcp", "gel")
+pred1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="pred", ])
+psel1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="psel", ])
+auc1 <- as.matrix(res1[substr(rownames(res1), 1, 3)=="auc", ])
+plot.data1 <- lapply(c(t(outer(methods2[-c(1, 2)], 1:3, paste, sep=""))), 
+                     function(m) {aggregate(auc1[, grepl(m, colnames(auc1))], 
+                                            list(psel=psel1[, grepl(m, colnames(
+                                              psel1))]), mean)})
+cv.psel1 <- c(sum(coef(fit1.gren1$freq.model$groupreg, 
+                       s=fit1.gren1$lambda)[-1, ]!=0),
+              sum(coef(fit1.gren2$freq.model$groupreg, 
+                       s=fit1.gren2$lambda)[-1, ]!=0),
+              sum(coef(fit1.gren3$freq.model$groupreg, 
+                       s=fit1.gren3$lambda)[-1, ]!=0),
+              sum(coef(fit1.gren1$freq.model$regular, 
+                       s=fit1.gren1$lambda)[-1, ]!=0),
+              sum(coef(fit1.gren2$freq.model$regular, 
+                       s=fit1.gren2$lambda)[-1, ]!=0),
+              sum(coef(fit1.gren3$freq.model$regular, 
[…]

### 2019-02-19 code/metabolomics_alzheimer.R

@@ -1,7 +1,8 @@
 #!/usr/bin/env Rscript
 
 ### installation of packages
-if(!("gren" %in% installed.packages())) {
+if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
+   substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
   if(!("devtools" %in% installed.packages())) {
     install.packages("devtools")
   }
@@ -15,6 +16,7 @@ library(gren)
 library(GRridge)
 library(grpreg)
 library(Biobase)
+library(SGL)
 
 ### load data
 load("data/ESetMbolCSFPR2.Rdata")
@@ -48,13 +50,14 @@ platformcode <- feat$PlatformCode2
 
 # quality gives groups based on a quality measure of the assayd features
 part.RSDqc <- CreatePartition(feat$RSDqc, ngroup=3)
-quality <- rep(c(1:3), sapply(part.RSDqc, length))[unlist(part.RSDqc)]
+quality <- rep(c(1:3), times=sapply(part.RSDqc, length))[
+  order(unlist(part.RSDqc))]
 
 # sds gives groups based on the standard deviations of the features
 part.sds.apoe.train <- CreatePartition(apply(metabol.apoe[id.train, ], 1, sd), 
                                        ngroup=3)
-sds.apoe.train <- rep(c(1:3), sapply(part.sds.apoe.train, 
-                                     length))[unlist(part.sds.apoe.train)]
+sds.apoe.train <- rep(c(1:3), sapply(part.sds.apoe.train, length))[
+  order(unlist(part.sds.apoe.train))]
 
 # degree gives the degree classes of the differential network
 degree <- NetworkDegreeClass[, 2]
@@ -65,432 +68,366 @@ alzheim.apoe <- as.numeric(pheno.apoe$D_diag_name) - 1
 ################################### model 1 ####################################
 ### fitting the models
 set.seed(2018)
-y.train <- alzheim.apoe[id.train]
-x.train <- scale(metabol.apoe[id.train, ])
-y.test <- alzheim.apoe[-id.train]
-x.test <- scale(metabol.apoe[-id.train, ])
[…]

### 2019-02-20 docs/supplement.Rnw
added barplots of multiplier estimates to metabolomics_alzheimer.R example

@@
+% set options
+<<settings, include=FALSE, echo=FALSE>>=
+opts_knit$set(root.dir="..", base.dir="../figs")
+opts_chunk$set(fig.align='center', fig.path="../figs/", 
+               echo=FALSE, cache=FALSE, message=FALSE)
+@
+
+% read figure code
+<<figures, include=FALSE>>=
+read_chunk("code/figures.R")
+@
+
 \documentclass[a4paper,hidelinks]{article}
-\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}
+\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
+rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
+bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
+dsfont,parskip}
 
@@
 \setlength{\parindent}{0pt}
-\title{Adaptive group-regularized logistic elastic net regression Supplementary material}
+\title{Adaptive group-regularized logistic elastic net regression Supplementary 
+material}
 \date{}
-\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}
+\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
+\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
+Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}
 
 \begin{document}
-
-% load packages and set options
-<<settings, include=FALSE, echo=FALSE>>=
-library(cambridge)
-opts_knit$set(base.dir="../figs")
-opts_chunk$set(fig.align='center', echo=FALSE, fig.path="../figs/", cache=FALSE,
-               message=FALSE)
-@
-
-% read figure code
-<<figures, include=FALSE>>=
-read_chunk("../code/figures.R")
-@
 
@@
 
-\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes
+  \textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; 
+  Variational Bayes
 
-\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}
+  \textbf{Software available from}: 
+  \url{https://CRAN.R-project.org/package=gren}
 
   \section{Introduction}
-This document contains supplementary material to the paper `Adaptive group-regularized logistic elastic net regression'.
-Section \ref{sec:enetprior} contains extra details on the elastic net prior distribution.
-Section \ref{sec:objectivefunction} gives the objective function that is maximised in the variational Bayes procedure introduced in Section 3 of the Main Document.
-Section \ref{sec:polyagamma} gives details on the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main Document.
-The derivations of the variational posterior used in the Main Document are shown in Section \ref{sec:variationaldistributions}.
-Section \ref{sec:computationalcomplexity} shows how we may reduce the computational complexity of our variational Bayes procedure.
-Section \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected joint log likelihood that is maximised to estimate the penalty multipliers as explained in Section 3 of the Main Document.
-The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}.
-We apply our method to a study on microRNAs in colorectal cancer data \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} .
-Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.
+  This document contains supplementary material to the paper `Adaptive 
+  group-regularized logistic elastic net regression'. Section 
+  \ref{sec:enetprior} contains extra details on the elastic net prior 
+  distribution. Section \ref{sec:objectivefunction} gives the objective 
+  function that is maximised in the variational Bayes procedure introduced in
+  Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on
+  the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main
+  Document. The derivations of the variational posterior used in the Main
+  Document are shown in Section \ref{sec:variationaldistributions}. Section
+  \ref{sec:computationalcomplexity} shows how we may reduce the computational
+  complexity of our variational Bayes procedure. Section 
+  \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected
+  joint log likelihood that is maximised to estimate the penalty multipliers as
+  explained in Section 3 of the Main Document. The full procedure of the
+  proposed method is shown in Section \ref{sec:fullestimationprocedure}. We
+  apply our method to a study on microRNAs in colorectal cancer data 
+  \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} . Finally, we
+  introduce the ridge and lasso versions of the proposed method in Section
+  \ref{sec:evbridgelasso}.
 
@@
   \subsection{Density function}
-The density function of $\beta$, drawn from an elastic net prior distribution, is given by:
+  The density function of $\beta$, drawn from an elastic net prior distribution,
+  is given by:
   $$
-f(\beta) = g(\lambda_1,\lambda_2) \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\].
+  f(\beta) = g(\lambda_1,\lambda_2) \exp\[-\frac{1}{2} (\lambda_1 |\beta| + 
+  \lambda_2 \beta^2)\].
   $$
@@
   $$
-g(\lambda_1,\lambda_2) =\left\{ \int_{-\infty}^{\infty} \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta \right\}^{-1}= \frac{\sqrt{\lambda_2}}{2} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1},
+  g(\lambda_1,\lambda_2) =\left\{ \int_{-\infty}^{\infty} \exp\[-\frac{1}{2} 
+  (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta \right\}^{-1}= 
+  \frac{\sqrt{\lambda_2}}{2} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \)
+  \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1},
   $$
-with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution functions of $x$, respectively.
+  with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution 
+  functions of $x$, respectively.
 
@@
   \begin{align*}
-\E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
-\V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta - \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} - \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
+  \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\
+  -\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
+  \V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot
+  \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta -
+  \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} -
+  \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 
+  \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
   \end{align*}
@@
 
+\section*{Session info}
+<<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=
+devtools::session_info()
+@
+
 \end{document}

### 2019-02-20 docs/manuscript.Rnw
added barplots of multiplier estimates to metabolomics_alzheimer.R example

@@
+% set options
+<<settings, include=FALSE, echo=FALSE>>=
+opts_knit$set(root.dir="..", base.dir="../figs")
+opts_chunk$set(fig.align='center', fig.path="../figs/", 
+               echo=FALSE, cache=FALSE, message=FALSE)
+@
+
+% read figure code
+<<figures, include=FALSE>>=
+read_chunk("code/figures.R")
+@
+
 \documentclass[a4paper,hidelinks]{article}
@@
 \begin{document}
-
-% load packages and set options
-<<settings, include=FALSE, echo=FALSE>>=
-library(cambridge)
-opts_knit$set(base.dir="../figs")
-opts_chunk$set(fig.align='center', echo=FALSE, fig.path="../figs/", cache=FALSE,
-               message=FALSE)
-@
-
-% read figure code
-<<figures, include=FALSE>>=
-read_chunk("../code/figures.R")
-@
 
@@
 
-	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as a computationally more convenient scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$.
-Using this result, we write the generalised elastic net model in its Bayesian form as:
+	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as
+	a computationally more convenient scale mixture of normals, with mixing 
+	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
+	\tr$. Using this result, we write the generalised elastic net model in its 
+	Bayesian form as:
 	\begin{subequations}\label{eq:enmodel2}
 		\begin{align}
-		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i \bbeta ) \right),  \\
-		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
-		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right).
+		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
+		\bbeta ) \right),  \\
+		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
+		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
+		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
+		}{\lambda_1^2}, \left(1,\infty \right) \right).
 		\end{align}
 	\end{subequations}
-	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$.
-In this Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.
+	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
+	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
+	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
+	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
+	play the role of the hyperparameters in a Bayesian hierarchical model.
 
@@
 	\subsection{Empirical Bayes}\label{seq:empiricalbayes}
-	If the penalty parameters are known, estimation of the frequentist elastic net model parameters or finding the posterior of the generalized elastic net model is feasible with small adjustments of the available algorithms \cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers \cite[]{li_bayesian_2010}.
-Determining these penalty parameters, however, is not straightforward.
+	If the penalty parameters are known, estimation of the frequentist elastic net
+	model parameters or finding the posterior of the generalized elastic net model
+	is feasible with small adjustments of the available algorithms 
+	\cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers
+	\cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is 
+	not straightforward.
 
-	In the frequentist elastic net without group-wise penalisation, two main strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by cross-validation over a two-dimensional grid of values or (ii) re-parametrise the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 \lambda_2 + \lambda_1}$ and $\lambda = 2 \lambda_2 + \lambda_1$, fix the proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global penalty parameter $\lambda$.
-Strategy (i) is advised by \cite{waldron_optimized_2011}, while (ii) is proposed by \cite{friedman_regularization_2010}.
-In the generalised elastic net setting, strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, respectively.
-$K$-fold cross validation over $D$ values then results in $K \cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate.
-Typically $K$ is set to 5, 10, or to the number of samples $n$, while $D$ is in the order of $100$, so that even for small $G$, the number of models to estimate is prohibitively large.
+	In the frequentist elastic net without group-wise penalisation, two main 
+	strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by 
+	cross-validation over a two-dimensional grid of values or (ii) re-parametrise 
+	the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 
+	\lambda_2 + \lambda_1}$ and $\lambda = 2 \lambda_2 + \lambda_1$, fix the 
+	proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global 
+	penalty parameter $\lambda$. Strategy (i) is advised by 
+	\cite{waldron_optimized_2011}, while (ii) is proposed by 
+	\cite{friedman_regularization_2010}. In the generalised elastic net setting, 
+	strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, 
+	respectively. $K$-fold cross validation over $D$ values then results in $K 
+	\cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate. Typically $K$ is 
+	set to 5, 10, or to the number of samples $n$, while $D$ is in the order of 
+	$100$, so that even for small $G$, the number of models to estimate is 
+	prohibitively large.
 
-	In the Bayesian framework, estimation of penalty parameters may be avoided by the addition of a hyperprior to the model hierarchy.
-The hyperprior takes the uncertainty in the penalty parameters into account by integrating over them.
-This approach introduces two issues.
-Firstly, the choice of hyperprior is not straightforward.
-Many authors suggest a hyperprior from the gamma family of distributions \cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, but the precise parametrisation of this gamma prior is not so obvious.
-A second issue is the loss of correspondence between the  Bayesian and frequentist elastic net.
-If the ultimate goal is feature selection, this correspondence may be exploited through the automatic feature selection property of the frequentist elastic net.
-Endowing the penalty parameters with a hyperprior obstructs their point estimation and, consequently, impedes automatic feature selection.
-Therefore, to circumvent the problem of hyperprior choice and allow for feature selection by the frequentist elastic net, we propose to estimate the penalty parameters by empirical Bayes.
+	In the Bayesian framework, estimation of penalty parameters may be avoided by 
+	the addition of a hyperprior to the model hierarchy. The hyperprior takes the 
+	uncertainty in the penalty parameters into account by integrating over them. 
+	This approach introduces two issues. Firstly, the choice of hyperprior is not 
+	straightforward. Many authors suggest a hyperprior from the gamma family of 
+	distributions 
+	\cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, 
+	but the precise parametrisation of this gamma prior is not so obvious. A 
+	second issue is the loss of correspondence between the  Bayesian and 
+	frequentist elastic net. If the ultimate goal is feature selection, this 
+	correspondence may be exploited through the automatic feature selection 
+	property of the frequentist elastic net. Endowing the penalty parameters with 
+	a hyperprior obstructs their point estimation and, consequently, impedes 
+	automatic feature selection. Therefore, to circumvent the problem of 
+	hyperprior choice and allow for feature selection by the frequentist elastic 
+	net, we propose to estimate the penalty parameters by empirical Bayes.
 
-	Many forms of empirical Bayes exist, the most formal one being maximisation of the marginal likelihood with respect to the hyperparameters.
-The resulting hyperparameter estimates are then plugged into the prior.
-The marginal likelihood is often introduced as a measure of model evidence given the observed data and is computed by integrating the product of likelihood and prior with respect to the model parameters.
-In the case of the elastic net introduced in (\ref{eq:enmodel2}) this results in the following marginal empirical Bayes posterior for $\bbeta$:
+	Many forms of empirical Bayes exist, the most formal one being maximisation of
+	the marginal likelihood with respect to the hyperparameters. The resulting 
+	hyperparameter estimates are then plugged into the prior. The marginal 
+	likelihood is often introduced as a measure of model evidence given the 
+	observed data and is computed by integrating the product of likelihood and 
+	prior with respect to the model parameters. In the case of the elastic net 
+	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
+	empirical Bayes posterior for $\bbeta$:
 	\begin{align}
-	p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = \frac{\int_{\btau} \L (\y ; \bbeta) \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \, d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
-	\hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
-	& = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. \label{eq:mmlestimate}
+	p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) \pi_{\hat{\blambda}
+	} (\bbeta) }{p_{\hat{\blambda}}(\y)} = \frac{\int_{\btau} \L (\y ; \bbeta) 
+	\pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \, d\btau}{p_
+	{\hat{\blambda}}(\y)}, \nonumber \\
+	\hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
+	\underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, 
+	\bbeta, \btau) \, d\bbeta d\btau \nonumber \\
+	& = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta) 
+	\pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. 
+	\label{eq:mmlestimate}
 	\end{align}
-	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the elastic net.
-In the omics setting the integrals are also high-dimensional, in which case numerical and Monte Carlo approximation methods become tedious and computationally expensive.
-Moreover, Laplace approximation is known to suffer from low accuracy in many high-dimensional settings \cite[]{shun_laplace_1995}.
-In \cite{casella_empirical_2001} an EM algorithm is described that estimates the hyperparameters.
-This EM algorithm iteratively maximises the expected joint log likelihood, such that the sequence:
+	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
+	elastic net. In the omics setting the integrals are also high-dimensional, in 
+	which case numerical and Monte Carlo approximation methods become tedious and 
+	computationally expensive. Moreover, Laplace approximation is known to suffer 
+	from low accuracy in many high-dimensional settings 
+	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
+	is described that estimates the hyperparameters. This EM algorithm iteratively
+	maximises the expected joint log likelihood, such that the sequence:
 	\begin{equation}\label{eq:ebEM}
-	\blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
+	\blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
+	\log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
 	\end{equation}
-	converges to a local maximum of the marginal likelihood.
-The difficulty herein is in the calculation of the expected joint log likelihood.
+	converges to a local maximum of the marginal likelihood. The difficulty herein
+	is in the calculation of the expected joint log likelihood. 
 	\cite{casella_empirical_2001} suggests to approximate the expectation by its Monte Carlo expectation.
@@
 	\section{Application to metabolomics in alzheimer}\label{sec:app}
+	\subsection{Co-data separately included}
+<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Cross-validated AUC in alzheimer example using metabolomics class co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+@
 
-<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Cross-validated AUC in alzheimer example using metabolomics class co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+<<barplots_metabolomics_alzheimer_res1_auc, fig.cap="Estimated penalty multipliers in alzheimer example using metabolomics class co-data", out.width="50%", fig.asp=1>>=
 @
@@
 
+<<barplots_metabolomics_alzheimer_res2_auc, fig.cap="Estimated penalty multipliers in alzheimer example using quality score co-data", out.width="50%", fig.asp=1>>=
+@
+
 <<lines_metabolomics_alzheimer_res3_auc, fig.cap="Cross-validated AUC in alzheimer example using node degree co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
@@
 
+<<barplots_metabolomics_alzheimer_res3_auc, fig.cap="Estimated penalty multipliers in alzheimer example using node degree co-data", out.width="50%", fig.asp=1>>=
+@
+
+\subsection{Co-data included simultaneously}
 <<lines_metabolomics_alzheimer_res4_auc, fig.cap="Cross-validated AUC in alzheimer example using all co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
@@
 
+	\section*{Session info}
+<<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=
+devtools::session_info()
+@
+
 \end{document}

### 2019-02-20 code/figures.R

@@ -3,8 +3,8 @@
 # ---- lines_metabolomics_alzheimer_res1_auc ----
 library(pROC)
 library(grpreg)
-load("../results/metabolomics_alzheimer_fit1.Rdata")
-res1 <- read.table("../results/metabolomics_alzheimer_res1.csv", 
+load("results/metabolomics_alzheimer_fit1.Rdata")
+res1 <- read.table("results/metabolomics_alzheimer_res1.csv", 
                    stringsAsFactors=FALSE)
 
 methods2 <- c("ridge", "grridge", "gren", "enet", "sgl", "cmcp", "gel")
@@ -97,11 +97,39 @@ points(t(sapply(1:length(plot.data1), function(m) {
   col=rep(col1[-c(1, 2)], each=3), pch=1)
 par(mfrow=c(1, 1))
 
+# ---- barplots_metabolomics_alzheimer_res1_auc ----
+library(Biobase)
+load("data/ESetMbolCSFPR2.Rdata")
+load("results/metabolomics_alzheimer_fit1.Rdata")
+
+methods1 <- c("grridge", paste0("gren", 1:3))
+col1 <- grey.colors(length(methods1), start=0.3, end=0.9, gamma=2.2, alpha=NULL)
+lty1 <- c(1:4)
+labels1 <- list(platform=replace(unique(as.character(fData(
+  ESetMbolCSFPR2)$Platform))[-5], 4, "Oxidative Stress"))
+
+plot.data1 <- lapply(1:length(fit1.gren1$lambdag), function(s) {
+  rbind(fit1.grridge$lambdamults[[s]], fit1.gren1$lambdag[[s]],
+        fit1.gren2$lambdag[[s]], fit1.gren3$lambdag[[s]])})
+opar <- par(no.readonly=TRUE)
+par(mar=opar$mar*c(1, 1.3, 1, 1))
+barplot(plot.data1[[1]], beside=TRUE, col=col1,
+        legend.text=c(methods1[1],
+                      expression("gren, "~alpha==0.05, "gren, "~alpha==0.5,
+                                 "gren, "~alpha==0.95), "no co-data"),
+        args.legend=list(x="topleft", fill=c(col1, NA),
+                         border=c(rep(1, length(methods1)), NA),
+                         lty=c(rep(NA, length(methods1)), 2),
+                         seg.len=1, merge=TRUE),
[…]

### 2019-02-20 code/metabolomics_alzheimer.R

@@ -17,6 +17,7 @@ library(GRridge)
 library(grpreg)
 library(Biobase)
 library(SGL)
+library(grpregOverlap)
 
 ### load data
 load("data/ESetMbolCSFPR2.Rdata")
@@ -67,7 +68,7 @@ alzheim.apoe <- as.numeric(pheno.apoe$D_diag_name) - 1
 
 ################################### model 1 ####################################
 ### fitting the models
-set.seed(2018)
+set.seed(2019)
 ytrain <- alzheim.apoe[id.train]
 xtrain <- scale(metabol.apoe[id.train, ])
 ytest <- alzheim.apoe[-id.train]
@@ -351,6 +352,8 @@ part.grridge <- list(platformcode=split(1:p, platformcode),
                      quality=split(1:p, quality),
                      degree=split(1:p, degree))
 part.gl <- as.numeric(factor(paste(platformcode, quality, degree)))
+part.ogl <- unlist(lapply(part.grridge, function(part) {
+  lapply(part, function(s) {colnames(xtrain)[s]})}), recursive=FALSE)
 
 fit4.gren1 <- gren(xtrain, ytrain, partitions=part.gren, alpha=0.05, 
                    trace=TRUE)
@@ -382,9 +385,24 @@ fit4.gel2 <- cv.grpreg(xtrain, ytrain, part.gl, penalty="gel",
 fit4.gel3 <- cv.grpreg(xtrain, ytrain, part.gl, penalty="gel", 
                        family="binomial", alpha=0.95)
 
+fit4.ocmcp1 <- cv.grpregOverlap(xtrain, ytrain, part.ogl, penalty="cMCP", 
+                                family="binomial", alpha=0.05)
+fit4.ocmcp2 <- cv.grpregOverlap(xtrain, ytrain, part.ogl, penalty="cMCP", 
+                                family="binomial", alpha=0.5)
+fit4.ocmcp3 <- cv.grpregOverlap(xtrain, ytrain, part.ogl, penalty="cMCP", 
+                                family="binomial", alpha=0.95)
+
+fit4.ogel1 <- cv.grpregOverlap(xtrain, ytrain, part.ogl, penalty="gel", 
+                                family="binomial", alpha=0.05)
+fit4.ogel2 <- cv.grpregOverlap(xtrain, ytrain, part.ogl, penalty="gel", 
[…]

### 2019-02-20 code/micrornaseq_colorectal_cancer.R
made start with micrornaseq_colorectal_cancer.R analysis

@@ -40,7 +40,7 @@ diff.threegroup <- miRNA %in% miRNA.BFDR + miRNA %in% miRNA.TumMirs + 1
 benefit <- as.numeric(resp) - 1
 
 ### mirna data
-micrornas <- scale(t(mirnormcen_resp))
+micrornas <- t(mirnormcen_resp)
 colnames(micrornas) <- miRNA
 
 ### randomly split in test and train data (30% and 70%, respectively)
@@ -52,20 +52,23 @@ id.train <- c(sample(which(benefit==0), size=floor(sum(benefit==0)*0.7)),
 ### fitting the models
 set.seed(2019)
 ytrain <- benefit[id.train]
-xtrain <- micrornas[id.train, ]
+# we remove the constant micrornas
+xtrain <- scale(micrornas[id.train, ])[, apply(micrornas[id.train, ], 2, sd)!=0]
 utrain <- unpenal[id.train, ]
 ytest <- benefit[-id.train]
-xtest <- scale(micrornas[-id.train, ])
+xtest <- scale(micrornas[-id.train, ])[, apply(micrornas[id.train, ], 2, sd)!=0]
+# set the constant micrornas to 0
+xtest[is.nan(xtest)] <- 0
 utest <- utrain[-id.train, ]
-part <- diff.threegroup
+part <- diff.threegroup[apply(micrornas[id.train, ], 2, sd)!=0]
 p <- ncol(xtrain)
 
 fit1.gren1 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.05, 
-                   standardize=FALSE, trace=FALSE)
+                   standardize=TRUE, trace=FALSE)
 fit1.gren2 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.5, 
-                   standardize=FALSE, trace=FALSE)
+                   standardize=TRUE, trace=FALSE)
 fit1.gren3 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.95, 
-                   standardize=FALSE, trace=FALSE)
+                   standardize=TRUE, trace=FALSE)
 
 fit1.grridge <- grridge(t(xtrain), ytrain, list(part=split(1:p, part)))
 
@@ -92,101 +95,116 @@ fit1.gel3 <- cv.grpreg(xtrain, ytrain, part, penalty="gel", family="binomial",
 
[…]

### 2019-02-20 rpackage/R/gren.R

@@ -1,6 +1,6 @@
 # group-regularized elastic net function
 gren <- function(x, y, m=rep(1, nrow(x)), unpenalized=NULL, partitions=NULL, 
-                 alpha=0.5, lambda=NULL, standardize=FALSE, intercept=TRUE, 
+                 alpha=0.5, lambda=NULL, standardize=TRUE, intercept=TRUE, 
                  monotone=NULL, psel=TRUE, compare=TRUE, posterior=FALSE, 
                  nfolds=nrow(x), foldid=NULL, trace=TRUE,
                  init=list(lambdag=NULL, mu=NULL, sigma=NULL, 

### 2019-02-20 docs/manuscript.Rnw
made start with micrornaseq_colorectal_cancer.R analysis

@@
 	\section*{Software}
-	The method is available as an \texttt{R} package from \url{https://CRAN.R-project.org/package=gren}.
+	A (stable) \texttt{R} package is available from 
+	\url{https://CRAN.R-project.org/package=gren}.
 
 	\section*{Supplementary Material}
-	Supplementary Material is available online from \url{http://pub.math.leidenuniv.nl/~munchmm/}.
+	Supplementary Material is available online from 
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All materials needed to   
+	recreate this document are available from 
+	\url{https://github.com/magnusmunch/gren}.
 

### 2019-02-27 docs/manuscript.Rnw
changed the applications in manuscript.Rnw and supplement.Rnw

@@
 \title{Adaptive group-regularized logistic elastic net regression}
-\date{}
+\date{\today}
 \author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
@@
 
-	\section{Introduction}
+	\section{Introduction}\label{sec:introduction}
 	Prediction from high-dimensional data is a challenge common to many fields of
@@
 	An example of such a study with extra information available is investigated in
-	Section \ref{sec:app}. INTRODUCE DATA. This problem is not straightforward to
-	solve with the standard methods available in literature. At first sight, such
+	Sections [! section refs here]. The first one is concerned with treatment 
+	response prediction for colorectal cancer patients, based on sequenced
+	microRNAs. In the second example we attempt to predict lymph node metastases,
+	using RNAseq data. In both settings we have additional information on the 
+	features in the form of a partitioning available. Furthermore, in both
+	examples data is sparse, so we would like to 
+	include the additional information to enhance predictive performance of our
+	models. Inclusion of the partitions is not straightforward using standard 
+	methods available in literature. At first sight, such
 	an \textit{a priori} grouping of the features suggests the group lasso
@@
 	approximation to this model and use this novel approximation in the empirical
-	Bayes estimation of multiple, group-specific penalty parameters. In Sections
-	\ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a
-	simulation study and a real data example, respectively. In Section 
+	Bayes estimation of multiple, group-specific penalty parameters. In Section
+	\ref{sec:simulations} and Sections [! section ref here] and 
+	[! section ref here], we demonstrate the approach in a
+	simulation study and a two real data examples, respectively. In Section 
 	\ref{sec:discussion} we conclude by both discussing some differences and
@@
 	\begin{align}
-	p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) \pi_{\hat{\blambda}
-	} (\bbeta) }{p_{\hat{\blambda}}(\y)} = \frac{\int_{\btau} \L (\y ; \bbeta) 
-	\pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \, d\btau}{p_
-	{\hat{\blambda}}(\y)}, \nonumber \\
+	  p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) 
+	  \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = 
+	  \frac{\int_{\btau} \L (\y ; \bbeta) 
+	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
+	  d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
 	  \hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
@@
 	is in the calculation of the expected joint log likelihood. 
-	\cite{casella_empirical_2001} suggests to approximate the expectation by its Monte Carlo expectation.
-Although elegant and simple, this method requires a converged MCMC sample from the posterior for every iteration: a computationally intensive procedure.
-\cite{roy_selection_2017} introduce generalised importance sampling for the Bayesian elastic net, such that we just need a limited, pre-specified number of MCMC chains.
-However, this still requires several converged chains, so is not feasible in many high-dimensional omics settings.
-We propose to tackle this problem by approximating the expectation in (\ref{eq:ebEM}) using variational Bayes.
+	\cite{casella_empirical_2001} suggests to approximate the expectation by its 
+	Monte Carlo expectation. Although elegant and simple, this method requires a 
+	converged MCMC sample from the posterior for every iteration: a 
+	computationally intensive procedure. \cite{roy_selection_2017} introduce
+	generalised importance sampling for the Bayesian elastic net, such that we
+	just need a limited, pre-specified number of MCMC chains. However, this still
+	requires several converged chains, so is not feasible in many high-dimensional
+	omics settings. We propose to tackle this problem by approximating the
+	expectation in (\ref{eq:ebEM}) using variational Bayes.
 
@@
 	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
-It has successfully been applied in a wide range of applications, including genetic association studies \cite[]{carbonetto_scalable_2012} and gene network reconstruction \cite[]{leday_gene_2017}.
-In variational Bayes, the posterior is approximated by a tractable form and estimated by optimizing a lower bound on the marginal likelihood of this model (see Section 3 of the SM for the lower bound of the proposed model).
-For an extensive introduction and concise review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, respectively.
+	It has successfully been applied in a wide range of applications, including 
+	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
+	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
+	is approximated by a tractable form and estimated by optimizing a lower bound 
+	on the marginal likelihood of this model (see Section 3 of the SM for the 
+	lower bound of the proposed model). For an extensive introduction and concise 
+	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
+	respectively.
 
-	To simplify the computations of our variational approximation, we follow \cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for $i=1, \dots, n$.
-Conditional on $\bbeta$, the $\omega_i$ are independent of the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more details).
-We augment Model (\ref{eq:enmodel2}) with:
+	To simplify the computations of our variational approximation, we follow
+	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
+	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
+	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more 
+	details). We augment Model (\ref{eq:enmodel2}) with:
 	\begin{equation}\label{eq:logmodel}
@@
 	\end{equation}
-	Our variational Bayes approximation to the posterior distribution of (\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of parameters. We choose the blocks such that:
+	Our variational Bayes approximation to the posterior distribution of 
+	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
+	parameters. We choose the blocks such that:
 	\begin{equation}\label{eq:varbayesapproximation}
-	p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
+	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
+	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
 	\end{equation}
-	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, calculus of variations gives the optimal distributions $q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} [\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the Kullback-Leibler divergence of the posterior to the approximate distribution \cite[]{neal_view_1998}.
-The approximation in (\ref{eq:varbayesapproximation}) renders both the posterior parameter calculations and the expected joint log likelihood as introduced in (\ref{eq:ebEM}) tractable.
+	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
+	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
+	calculus of variations gives the optimal distributions 
+	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
+	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
+	Kullback-Leibler divergence of the posterior to the approximate distribution 
+	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
+	renders both the posterior parameter calculations and the expected joint log 
+	likelihood as introduced in (\ref{eq:ebEM}) tractable.
 
-	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal distributions in our variational Bayes implementation for the model parameters as:
+	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
+	distributions in our variational Bayes implementation for the model parameters
+	as:
 	\begin{equation}\label{eq:varbayesdistr}
-	q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)\text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
+	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
+	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
+	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
+	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
 	\end{equation}
-	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian distribution (See SM Section 5 for the derivations).
-The so-called variational parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we update them by:
+	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
+	distribution (See SM Section 5 for the derivations). The so-called variational
+	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
+	update them by:
 	\begin{subequations}\label{eq:VBupdateequations}
 		\begin{align}
-		\bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
+		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
+		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
+		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
+		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
+		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
 		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
-		c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
-		\chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
+		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
+		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
+		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
+		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
 		\end{align}
 	\end{subequations}
-	until convergence.
-Here, $\bLambda'$ is a diagonal matrix with entries $\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & \cdots & m_n \end{bmatrix} \tr$.
-Furthermore, $\mathbf{A}_{jj}$ and $\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th element of a column vector, respectively.
-Naive calculation of the variational parameters is computationally expensive.
-In Section 6 of the SM we show that informed calculation of the parameters results in a significant reduction of computational complexity.
+	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
+	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
+	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
+	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
+	element of a column vector, respectively. Naive calculation of the variational
+	parameters is computationally expensive. In Section 6 of the SM we show that 
+	informed calculation of the parameters results in a significant reduction of 
+	computational complexity.
 
 	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
-	Variational Bayes was shown to underestimate the posterior variance of the parameters, both numerically and theoretically, in several settings \cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,wang_inadequacy_2005}.
-This coincides with our experience that the global penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated.
-To prevent overestimation we use the parametrisation of the elastic net in \cite{friedman_regularization_2010} as introduced in Section \ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by cross-validation of the regular elastic net model, such that the overall penalisation is determined by cross-validation of only $\lambda$.
-By combining cross-validation of the global penalty parameter $\lambda$ with empirical Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more robust to underestimation of the variational posterior variances.
-The remaining issue is the choice of $\alpha$.
-\cite{hastie_glmnet_2016} recommend to either fix $\alpha$ \textit{a priori}, or compare the results for several choices of $\alpha$.
-We recommend the latter.
-Otherwise we note that, in our experience, the choice $\alpha=0.5$ often gives good results.
+	Variational Bayes was shown to underestimate the posterior variance of the 
+	parameters, both numerically and theoretically, in several settings 
+	\cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,
+	wang_inadequacy_2005}. This coincides with our experience that the global 
+	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To 
+	prevent overestimation we use the parametrisation of the elastic net in 
+	\cite{friedman_regularization_2010} as introduced in Section 
+	\ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
+	cross-validation of the regular elastic net model, such that the overall 
+	penalisation is determined by cross-validation of only $\lambda$. By combining
+	cross-validation of the global penalty parameter $\lambda$ with empirical 
+	Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} 
+	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more 
+	robust to underestimation of the variational posterior variances. The 
+	remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend
+	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
+	choices of $\alpha$. We recommend the latter. Otherwise we note that, in our 
+	experience, the choice $\alpha=0.5$ often gives good results.
 
-	For estimating the penalty multipliers, the intractable posterior expectation in (\ref{eq:ebEM}) is approximated by the variational posterior:
+	For estimating the penalty multipliers, the intractable posterior expectation
+	in (\ref{eq:ebEM}) is approximated by the variational posterior:
 	$$
-	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \] = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
+	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \]
+	= \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
+	\lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
 	$$
-	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full derivation).
-The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j \in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$.
-An estimate of the new penalty multipliers is now given by:
+	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full 
+	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
+	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
+	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
+	of the new penalty multipliers is now given by:
 	\begin{subequations}\label{eq:mmlupdateequation}
 		\begin{align}
-		\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
+		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
+		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
 		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
@@
 	\end{subequations}
-	Although the solution to (\ref{eq:mmlupdateequation}) is not available in closed form, this convex problem is easily solved by a numerical optimisation routine.
-The full procedure is summarized in Section 8 of the SM.
+	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
+	closed form, this convex problem is easily solved by a numerical optimisation 
+	routine. The full procedure is summarized in Section 8 of the SM.
 
 	\subsection{Feature selection}
-	Feature selection is often desirable in high-dimensional prediction problems and omics data problems are no exception.
-A selection of biomarkers may lead to a large cost reduction by supporting targeted assays.
-Bayesian feature selection is often done by inspection of posterior credible intervals.
-Similarly as in frequentist hypothesis testing, we may select a feature if zero is not contained in the credible interval.
-However, the Bayesian lasso's credible intervals (a special case of the elastic net) are known to suffer from low frequentist coverage in sparse settings \cite[]{castillo_bayesian_2015}.
-We often assume sparsity of the features in omics research, so selection by credible intervals is generally inconsistent.
+	Feature selection is often desirable in high-dimensional prediction problems 
+	and omics data problems are no exception. A selection of biomarkers may lead 
+	to a large cost reduction by supporting targeted assays. Bayesian feature
+	selection is often done by inspection of posterior credible intervals. 
+	Similarly as in frequentist hypothesis testing, we may select a feature if 
+	zero is not contained in the credible interval. However, the Bayesian lasso's
+	credible intervals (a special case of the elastic net) are known to suffer 
+	from low frequentist coverage in sparse settings 
+	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
+	omics research, so selection by credible intervals is generally inconsistent. 
 	We therefore propose to select features in the frequentist paradigm.
 
-	As shortly touched upon in Section \ref{seq:empiricalbayes}, frequentist feature selection is trivial after estimation of the penalty multipliers.
-We simply plug the estimated penalty parameters into some frequentist elastic net algorithm that allows for differential penalization.
-In our own package \texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} \cite[]{friedman_regularization_2010}.
-In the frequentist elastic net, feature selection is then done automatically.
-Furthermore, to select a specific number of features, we simply adjust the global $\lambda$ until we select the desired number of features.
+	As shortly touched upon in Section \ref{seq:empiricalbayes}, frequentist 
+	feature selection is trivial after estimation of the penalty multipliers. We 
+	simply plug the estimated penalty parameters into some frequentist elastic net
+	algorithm that allows for differential penalization. In our own package 
+	\texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} 
+	\cite[]{friedman_regularization_2010}. In the frequentist elastic net, feature
+	selection is then done automatically. Furthermore, to select a specific number
+	of features, we simply adjust the global $\lambda$ until we select the desired
+	number of features.
 
 	\section{Simulations}\label{sec:simulations}
-	\subsection{Setup}\label{sec:setup}
-	To assess variable selection and predictive performance we conduct a simulation study in which we compare \texttt{gren} to the regular elastic net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, group mimimax concave penalty (\texttt{grMCP}) \cite[]{breheny_penalized_2009}, and the group exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}.
-\texttt{GRridge} is similar to \texttt{gren} in the sense that it estimates group-specific penalty multipliers.
-The two main differences with \texttt{gren} are (i) the absence of an $L_1$-norm penalty and (ii) the estimation procedure.
-The other methods are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} and not adaptive on the group level.
-In contrast to the original group lasso, all of them are able to select single features, instead of just on the group level.
+	To assess variable selection and predictive performance we conducted a 
+	simulation study in which we compare \texttt{gren} to the regular elastic 
+	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
+	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
+	group mimimax concave penalty (\texttt{grMCP}) 
+	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
+	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
+	\texttt{gren} in the sense that it estimates group-specific penalty 
+	multipliers. The two main differences with \texttt{gren} are (i) the absence 
+	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
+	are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} 
+	and not adaptive on the group level. In contrast to the original group lasso,
+	all of them are able to select single features, instead of just on the group 
+	level.
 
-	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, we fix the proportion of $L_1$-norm penalty to three different values: $\alpha \in \{0.05, 0.5, 0.95 \}$.
-The first setting closely resembles the ridge setting, where $\alpha=0$, while the third setting is similar to the lasso with $\alpha=1$.
-For the methods that do automatic variable selection (all methods except the regular ridge and \texttt{GRridge}), we estimate the models for a range of model sizes.
-
-	Predictive performance is measured by area under the receiver operator curve (AUC) and Brier skill score (BSS).
-AUC gives the area under the curve of sensitivity versus $1-$ specificity and equals the probability that the classifier ranks a randomly chosen case higher than a randomly chosen non-case.
-The BSS is a normalised version of the Brier score: $\text{BSS} = 1 - \frac{\sum_{i=1}^n (y_i - \hat{y}_i)^2}{\sum_{i=1}^n (y_i - \bar{y})^2}$,
-	where $\hat{y}_i=\expit(\x_i \tr \hat{\bbeta})$ is the prediction for observation $i$ and $\bar{y}=n^{-1} \sum_{i=1}^n y_i$.
-The BSS measures the fraction of Brier score improvement compared to a non-informative model.
-Quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960} and parameter estimation by mean squared error (MSE).
-Both are presented in SM Section 10.
+	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, 
+	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
+	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
+	such that the first setting closely resembles the ridge setting, 
+	where $\alpha=0$, while the third setting is similar to the lasso with 
+	$\alpha=1$. In \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, $\alpha$
+	determines the proportion of group-level penalization ($\alpha=0$) compared to
+	$L_1$-norm penalty on the feature level ($\alpha=1$). For the methods that do 
+	automatic variable selection, we estimate 
+	the models for a range of model sizes.
 
@@
 	\begin{enumerate}
-		\item differential signal between the groups and uniformly distributed model parameters,
+		\item differential signal between the groups and uniformly distributed model
+		parameters,
 		\item a large number of small groups of features,
-		\item no differential signal between the groups, but strong correlations within groups of features,
-		\item differential signal between the groups and heavy-tailed distributed model parameters, and
+		\item no differential signal between the groups, but strong correlations 
+		within groups of features,
+		\item differential signal between the groups and heavy-tailed distributed 
+		model parameters, and
 		\item a very sparse setting with no signal in some of the groups.
 	\end{enumerate}
-	Full descriptions of the scenarios are given in Section 10 of the SM.
-In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform.
-In the second scenario, we expect \texttt{gren} and \texttt{GRridge} to overfit due to the large number of hyperparameters to estimate.
-The rest of the methods are conjectured to perform better, due to a lack of overfitting.
-In scenario three, we expect all methods to perform badly due to the high correlation between the features.
-In scenario three
+	Full descriptions of the scenarios are given in Section [! ref to section] of 
+	the SM. In the first scenario, the distribution of the model parameters is far
+	off from the assumed ridge/elastic net/lasso distributions, so we assume that 
+	all methods will underperform. In the second scenario, we expect \texttt{gren}
+	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
+	estimate. The rest of the methods are conjectured to perform better, due to a
+	lack of overfitting. In scenario three, we expect all methods to perform badly
+	due to the high correlation between the features. The \texttt{gren} fourth 
+	scenario is conjectured to give a better \texttt{gren} performance, due to the
+	correctly assumed heavier-tailed distribution and differential signal between
+	the groups. Scenario five is expected to lead to a bad performance in all 
+	models, because of the sparseness of the truth and lack of differential 
+	signal.
 
-	In all scenarios we created a training set of $n=100$ samples to estimate the models and a test set of $n_{\text{test}}=1000$ samples on which we compute the performance measures.
-The features $\x_i$, $i=1, \dots, n$, are sampled from a $p$-dimensional Gaussian distribution centred at zero, where we set $p=1000$.
-We let the covariance matrix of the features vary over the scenarios.
-Likewise for the model parameters $\beta_j$, $j=1, \dots, p$.
-We simulate the outcome data $y_i$, $i=1, \dots, n$, from a binary logistic model: $y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$.
-To mitigate the influence of random variation in the simulations, we repeated every scenario 100 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.
+	In all scenarios we created a training set of $n=100$ samples to estimate the 
+	models and a test set of $n_{\text{test}}=1000$ samples on which we compute 
+	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
+	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
+	$p=1000$. We let the covariance matrix of the features vary over the 
+	scenarios. Likewise for the model parameters $\beta_j$, $j=1, \dots, p$. We 
+	simulate the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
+	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
+	of random variation in the simulations, we repeated every scenario 100 times 
+	and report estimated penalty parameters and predictive performance measures 
+	area under the receiver operator curve (AUC) and Brier skill score (BSS).
 
-	\subsection{Results}\label{sec:results}
+	[! results here]
 
-	\section{Application to metabolomics in alzheimer}\label{sec:app}
-	\subsection{Co-data separately included}
-<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Cross-validated AUC in alzheimer example using metabolomics class co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+	Additional results are presented in Section [! section ref here] of the SM.
+
+	\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
+  The presented data is from a deep sequencing analysis on microRNAs 
+  \cite[]{neerincx_combination_2018}. The study was done in 88 treated 
+  colorectal cancer patients, with the aim of predicting treatment response,
+  coded as either non-progressive/remission (70 patients) or progressive 
+  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
+  remained. In addition to the 2114 microRNAs, four clinical covariates were
+  included in the analysis: prior use of adjuvant therapy (binary),
+  the type of systemic treatment regimen (ternary), age, and primary tumor
+  differentiation (binary). These clinical covariates are included as 
+  unpenalized features in the model.
+
+  In a preliminary experiment on different subjects, the microRNA expression 
+  levels of metastatic colorectal tumour tissue were compared to normal 
+  non-colorectal tissue and primary colorectal tumour tissue was compared to
+  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 221
+  microRNAs that were differentially expressed in both comparisons ($\text{FDR}
+  \leq 0.05$), versus 1893 not differentially expressed microRNAs. We expect
+  that incorporation of this partitioning enhances therapy response 
+  classification, because tumor-specific miRNAs might be more relevant than 
+  non-specific ones. In addition, we divided the differentially expressed
+  microRNAs even further into 127 highly differentially expressed microRNAs
+  ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs 
+  ($0.001 < \text{FDR} \leq 0.05$).  We will refer to this second partitioning
+  as the three-group setting, as opposed to the first, the two-group setting.
+
+  We split the data into 61 training, and 27 test instances and
+  compared the performance of \texttt{gren} on the test set to ridge, 
+  \texttt{GRridge}, elastic net, and the composite minimax
+  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. Further
+  comparisons to the sparse
+  group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and the group
+  exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015} are presented in 
+  the SM [! section ref here]. We compared the methods on AUC. More performance
+  measures are presented in the SM [! section ref here].
+  For the elastic net methods (including \texttt{gren}) and \texttt{cMCP}, we 
+  set $\alpha \in \{ 0.05, 0.5, 0.95\}$. We experienced numerical issues with 
+  \texttt{GRridge} when cross-validating 
+  the global penalty parameter on the training data. We therefore estimated this
+  global parameter on the full data for \texttt{GRridge}.
+
+  The results are presented in Figure 
+  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. 
+  The penalty multipliers estimated by 
+  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
+  expressed group receives the lowest penalty, followed by the medium expressed
+  group, while the non-expressed group receives the strongest penalty. 
+  \texttt{gren} seems to outperform the other methods for a range of model 
+  sizes. This is especially pronounced for the larger models. The performance of
+  \texttt{cmcp} is somewhat unstable. We noticed this is in general
+  for the group-lasso and its extensions (see SM [! section ref here]). Overall,
+  it seems as though the inclusion of the extra information on the features 
+  benefited predictive performance: both \texttt{GRridge} and \texttt{gren} 
+  outperform their respective non-group-regularized versions.
+<<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
 @
 
-<<barplots_metabolomics_alzheimer_res1_auc, fig.cap="Estimated penalty multipliers in alzheimer example using metabolomics class co-data", out.width="50%", fig.asp=1>>=
+  \section{Application to RNAseq in lymph node cancer}\label{sec:applymph}
+  In this application, the aim is to predict lymph node metastasis in patients 
+  with HPV negative oral cancer using sequenced RNAs from TCGA 
+  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
+  normalised TCGA RNASeqv2 profiles of head-and-neck squamous cell 
+  carcinomas on 133 patients with HPV-negative tumours in the oral cavity. Of 
+  these patients, 76 had lymph node metastasis, and 57 did not. For a thorough 
+  introduction of these data, see \cite{te_beest_improved_2017}.
+
+  As co-data we considered two sources: (a) the cis-correlation between TCGA DNA
+  copy number measurements and the RNASeqv2 data on the same patients, 
+  quantified through Kendall's $\tau$ and binned into five groups, and (b) 
+  p-values from an external microarray data set \cite[]{mes_prognostic_2017}, 
+  binned into five groups. We expect features 
+  with a positive Kendall's $\tau$ to be more important in metastasis prediction   
+  \cite[]{masayesva_gene_2004}. Likewise, we expect features with a low 
+  p-value to be more important.
+
+  We split our data into 120 training instances and 97 test instances, trained
+  the models on the training set and estimated AUC on the test set. Other 
+  performance measures are presented in the SM Section [! section ref here]. We
+  compared the same models as in Section \ref{sec:appcolon}. In this 
+  application, the co-data consists of two partitions with overlapping groups.
+  \texttt{cMCP} is not designed to deal with overlapping groups, so we 
+  cross-tabulate the two co-data partitions to create one paritioning with 25 
+  groups for \texttt{cMCP}. In addition, we compare to an extension of 
+  \texttt{cMCP} that is able to handle multiple groupings 
+  \cite[]{zeng_overlapping_2016}. We present the results in Figure 
+  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. 
+  Additional comparisons to \texttt{SGL}, \texttt{gel} and 
+  the overlapping group version of \texttt{gel} \cite[]{zeng_overlapping_2016}
+  \cite[]{breheny_group_2015} are found in the SM Section [! section ref here].
+<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.", out.width="100%", fig.asp=1>>=
 @
 
-<<lines_metabolomics_alzheimer_res2_auc, fig.cap="Cross-validated AUC in alzheimer example using quality score co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
-@
-
-<<barplots_metabolomics_alzheimer_res2_auc, fig.cap="Estimated penalty multipliers in alzheimer example using quality score co-data", out.width="50%", fig.asp=1>>=
-@
-
-<<lines_metabolomics_alzheimer_res3_auc, fig.cap="Cross-validated AUC in alzheimer example using node degree co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
-@
-
-<<barplots_metabolomics_alzheimer_res3_auc, fig.cap="Estimated penalty multipliers in alzheimer example using node degree co-data", out.width="50%", fig.asp=1>>=
-@
-
-\subsection{Co-data included simultaneously}
-<<lines_metabolomics_alzheimer_res4_auc, fig.cap="Cross-validated AUC in alzheimer example using all co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
-@
-
-<<barplots_metabolomics_alzheimer_res4_auc, fig.cap="Estimated penalty multipliers in alzheimer example using all co-data for a) metabolomics class, b) quality score, and c) node degree", out.width="100%", fig.asp=1>>=
-@
+  We present additional applications and comparisons in the SM Sections 
+  [! sections ref here].
 
 	\section{Discussion}\label{sec:discussion}
-	In a taxonomy of Bayesian methods, the proposed method may be considered a local shrinkage model, as opposed to the global-local shrinkage priors that \cite{polson_local_2012,bernardo_shrink_2011} discuss.
-They characterise certain desirable properties of these global-local shrinkage priors in high dimensions, which, for example, the horseshoe possesses \cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}.
-In our case, global shrinkage would imply adding another hyperprior for the global $\lambda_1$ and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters.
-We argue however, that if the groups are informative, the empirical Bayes estimation of the (semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than full Bayes shrinkage of the global penalty parameters, because the latter does not use any known structure to model the variability in the hyperparameters.
-Nonetheless, an interesting direction of future research is the extension of the group-regularized elastic net to a group-regularized horseshoe model, since the horseshoe has been shown to handle sparsity well and render better coverage of credibility intervals than lasso-type priors \cite[]{van_der_pas_horseshoe_2014}.
+	In a taxonomy of Bayesian methods, the proposed method may be considered a 
+	local shrinkage model, as opposed to the global-local shrinkage priors that 
+	\cite{polson_local_2012,bernardo_shrink_2011} discuss. They characterise 
+	certain desirable properties of these global-local shrinkage priors in high 
+	dimensions, which, for example, the horseshoe possesses 
+	\cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}. In our case, global 
+	shrinkage would imply adding another hyperprior for the global $\lambda_1$ 
+	and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters. We argue however,
+	that if the groups are informative, the empirical Bayes estimation of the 
+	(semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than 
+	full Bayes shrinkage of the global penalty parameters, because the latter does
+	not use any known structure to model the variability in the hyperparameters. 
+	Nonetheless, an interesting direction of future research is the extension of 
+	the group-regularized elastic net to a group-regularized horseshoe model, 
+	since the horseshoe has been shown to handle sparsity well and render better
+	coverage of credibility intervals than lasso-type priors 
+	\cite[]{van_der_pas_horseshoe_2014}.
 
-	Although our method can be considered weakly adaptive, it is different from the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net \cite[]{zou_adaptive_2009} in the sense that it adapts to external information rather than to the data.
-It also differs in the scale of adaptation: in the adaptive lasso and elastic net the adaptive weights are feature specific, while in our case they are estimated on the group level, rendering the adaptation more robust.
-As can be seen from the simulations and real data example in Sections \ref{sec:simulations} and \ref{sec:app}, adaptation to external data may be beneficial for prediction and feature selection.
-We believe that this is due to the `borrowing of information' effect: estimates that are believed to be similar are shrunken in the same way, yielding overall, better estimates.
+	Although our method can be considered weakly adaptive, it is different from 
+	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
+	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
+	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
+	rather than to the data. It also differs in the scale of adaptation: in the 
+	adaptive lasso and elastic net the adaptive weights are feature specific,
+	while in our case they are estimated on the group level, rendering the 
+	adaptation more robust. As can be seen from the simulations and real data 
+	example in Sections \ref{sec:simulations} and [! section ref here], 
+	adaptation to 
+	external data may be beneficial for prediction and feature selection. We 
+	believe that this is due to the `borrowing of information' effect: estimates 
+	that are believed to be similar are shrunken in the same way, yielding 
+	overall, better estimates.
 
-	Another obvious comparison is to the group lasso \cite[]{meier_group_2008,yuan_model_2006}.
-Although it is similar in the sense that it shrinks on the group level, the group lasso is built upon an entirely different philosophy: it sets whole groups of features to zero.
-The intended application of such a group-wise penalty is to small interpretable groups of features, like, for example, dummies of a categorical variable.
-Another difference between \texttt{gren} and the group lasso is the number of penalty parameters.
-\texttt{gren} estimates one parameter per group, while the group lasso estimates one overall penalty parameter; it is thereby less flexible in differential shrinkage of the parameters.
-In addition, in the simulations and data applications of \cite{van_de_wiel_better_2016}, group lasso prediction performed inferior to \texttt{GRridge} prediction.
+	Another obvious comparison is to the group lasso \cite[]{meier_group_2008,
+	yuan_model_2006}. Although it is similar in the sense that it shrinks on the 
+	group level, the group lasso is built upon an entirely different philosophy: 
+	it sets whole groups of features to zero. The intended application of such a 
+	group-wise penalty is to small interpretable groups of features, like, for 
+	example, dummies of a categorical variable. Another difference between 
+	\texttt{gren} and the group lasso is the number of penalty parameters. 
+	\texttt{gren} estimates one parameter per group, while the group lasso
+	estimates one overall penalty parameter; it is thereby less flexible in 
+	differential shrinkage of the parameters. In addition, in the simulations and
+	data applications of \cite{van_de_wiel_better_2016}, group lasso prediction 
+	performed inferior to \texttt{GRridge} prediction.
 
-	A possible weak point of the proposed method is the double EM loop.
-The double loop increases the chance of ending up in a local optimum.
-In the application discussed above we investigated the occurrence of multiple optima, but never encountered them.
-This does not guarantee that local optima do not occur, but it provides some evidence that local optima are not ubiquitous.
-Local optima can often be avoided by an informed choice of starting values.
-In our experience, reasonable starting values are obtained by running a group-regularized ridge regression and using the estimated penalty multipliers as starting values for \texttt{gren}.
-In addition, we note that even a local optimum might give `good enough' predictions in many practical settings.
+	A possible weak point of the proposed method is the double EM loop. The double
+	loop increases the chance of ending up in a local optimum. In the application
+	discussed above we investigated the occurrence of multiple optima, but never
+	encountered them. This does not guarantee that local optima do not occur, but 
+	it provides some evidence that local optima are not ubiquitous. Local optima 
+	can often be avoided by an informed choice of starting values. In our 
+	experience, reasonable starting values are obtained by running a 
+	group-regularized ridge regression and using the estimated penalty multipliers
+	as starting values for \texttt{gren}. In addition, we note that even a local
+	optimum might give `good enough' predictions in many practical settings.
 
@@
 	\section*{Funding}
-	This research has received funding from the European Research Council under ERC Grant Agreement 320637.
+	This research has received funding from the European Research Council under 
+	ERC Grant Agreement 320637.
 

### 2019-02-27 code/figures.R

@@ -3,15 +3,16 @@
 # ---- lines_metabolomics_alzheimer_res1_auc ----
 library(pROC)
 library(grpreg)
+library(sp)
 load("results/metabolomics_alzheimer_fit1.Rdata")
 res1 <- read.table("results/metabolomics_alzheimer_res1.csv", 
                    stringsAsFactors=FALSE)
 
-methods2 <- c("ridge", "grridge", "gren", "enet", "sgl", "cmcp", "gel")
+methods1 <- c("ridge", "grridge", "gren", "enet", "sgl", "cmcp", "gel")
 pred1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="pred", ])
 psel1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="psel", ])
 auc1 <- as.matrix(res1[substr(rownames(res1), 1, 3)=="auc", ])
-plot.data1 <- lapply(c(t(outer(methods2[-c(1, 2)], 1:3, paste, sep=""))), 
+plot.data1 <- lapply(c(t(outer(methods1[-c(1, 2)], 1:3, paste, sep=""))), 
                      function(m) {aggregate(auc1[, grepl(m, colnames(auc1))], 
                                             list(psel=psel1[, grepl(m, colnames(
                                               psel1))]), mean)})
@@ -39,7 +40,7 @@ cv.psel1 <- c(sum(coef(fit1.gren1$freq.model$groupreg,
               psel1[, grepl("gel1", colnames(psel1))][fit1.gel1$min],
               psel1[, grepl("gel2", colnames(psel1))][fit1.gel2$min],
               psel1[, grepl("gel3", colnames(psel1))][fit1.gel3$min])
-col1 <- c(1:length(methods2))
+col1 <- bpy.colors(length(methods1), cutoff.tail=0.1)
 lty1 <- c(1:4)
 
 par(mfrow=c(1, 2))
@@ -65,13 +66,13 @@ points(t(sapply(1:length(plot.data1), function(m) {
   subset(plot.data1[[m]], psel==cv.psel1[m])})), 
   col=rep(col1[-c(1, 2)], each=3), pch=1)
 legend("bottomright", 
-       legend=c(methods2, expression(alpha==0.05, alpha==0.5, alpha==0.95),
+       legend=c(methods1, expression(alpha==0.05, alpha==0.5, alpha==0.95),
                 "CV model"), 
        fill=c(col1, rep(NA, 4)),
[…]

### 2019-02-27 code/rnaseq_lymph_node_metastasis.R

@@ -1,15 +1,6 @@
 #!/usr/bin/env Rscript
 
 ### installation of packages
-if(!("CoRF" %in% installed.packages())) {
-  if(!("devtools" %in% installed.packages())) {
-    install.packages("devtools")
-  }
-  library(devtools)
-  install_github("DennisBeest/CoRF", local=FALSE,
-                 auth_token=Sys.getenv("GITHUB_PAT"))
-}
-
 if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
    substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
   if(!("devtools" %in% installed.packages())) {
@@ -24,8 +15,10 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 library(CoRF)
 library(GRridge)
 library(gren)
-library(grpreg)
+library(grpregOverlap)
+library(SGL)
 library(pROC)
+
 
 ### load data
 data(LNM_Example)
@@ -55,63 +48,96 @@ partLNM <- list(corr=corpart0, pv=pvpart0)
 
 ################################### model 1 ####################################
 ### fitting the models
-nfolds <- 1
-set.seed(2018)
-y <- RespTrain
-x <- Tr
+set.seed(2019)
+ytrain <- RespTrain
+xtrain <- Tr
+ytest <- as.numeric(levels(RespValidation))[RespValidation]
+xtest <- ValidationData[, selectgenes]
 part <- partLNM
-n <- nrow(x)
-p <- ncol(x)
+part.ogl <- unlist(lapply(list(corr=split(1:p, part$corr), 
+                               pv=split(1:p, part$pv)), function(part) {
+                                 lapply(part, function(s) {
+                                   colnames(xtrain)[s]})}), recursive=FALSE)
+n <- nrow(xtrain)
+p <- ncol(xtrain)
 
-fit1.gren1 <- gren(x, y, partitions=part, alpha=0.05, trace=FALSE)
-fit1.gren2 <- gren(x, y, partitions=part, alpha=0.5, trace=FALSE)
-fit1.gren3 <- gren(x, y, partitions=part, alpha=0.95, trace=FALSE)
+fit1.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE)
[…]

### 2019-02-27 docs/response.Rnw
changed the applications in manuscript.Rnw and supplement.Rnw

\documentclass{letter}

\newenvironment{itquote}
{\begin{quote}\itshape}
{\end{quote}}

\signature{Magnus M\"unch}
\begin{document}
  \begin{letter}{}
    \opening{Dear editor,}

    Thank you for the opportunity to resubmit our manuscript. We realise it has
    taken us an extroardinarily long time to respond, but given the extensive 
    nature of the revision, we hope that you will still consider this 
    resubmission.

    Below we formulate our response to your and the reviewers' comments
    point by point. We hope that these responses and the accompanying changes in
    the manuscript and supplementary material address the raised issues, such 
    that the manuscript is deemed appropriate for publication in Biostatistics.

    \textbf{Reply to editor}
    \begin{itquote}
      I would be willing to consider a revision of your work if you 
      made a substantial effort to rework the simulation examples and practical 
      case studies to demonstrate clearly unique cases where your method works 
      and others don't (as well as the reverse case!).
    \end{itquote} 
    The simulations were extended to five different scenarios. In three
    scenarios, we expect all methods to underperform. In one scenario, we expect
    \texttt{gren} to underperform compared to its competitors and in one 
    scenario we expect \texttt{gren} to outperform its competitors. We have 
    included 4/5 applications in the manuscript and supplementary material. In
    one application \texttt{gren} underperforms compared to the others. We argue
    as for why this happens and deduce certain scenarios in which \texttt{gren}
    should not be used.

    \begin{itquote}
      This comparison should include, not just the EN and Grouped 
      Lasso, but other methods that approach this same problem as well. There 
      would need to be clear empirical demonstration on real data of a unique 
      advantage of this approach for it to be a good fit at Biostatistics.
    \end{itquote}
    We included comparisons to the sparse group lasso (\texttt{SGL}) by 
    Simon et al. (2013), the group mimimax concave penalty
    (\texttt{grMCP}) by Breheny and Huang (2009), and the group exponential
    lasso (\texttt{gel}) by Breheny (2015). \texttt{gren} outperforms these in 
    a number of comparisons. Our approach is based on a different philosophy
    than the group lasso and its extensions. Group weights are adaptively 
    estimated from the data and not based on one common penalty, as is the case
    in the group lasso and its extensions.

    \textbf{Reply to reviewer 1}
    \begin{itquote}
      However, in my opinion the scientific contribution and novelty 
      of the methodology is just incremental. Elastic net itself is designed to 
      do the grouping. So, I am not sure why putting group penalty elastic net
      will be better than other group penalty methods like group lasso
    \end{itquote}
    The grouping property of elastic net refers to a theoretical advantage over
    the lasso in dealing with correlated features. It does not refer to the
    explicit inclusion of external information in the form of groups of 
    features and is therefore unrelated to the topic of the manuscript.

    \begin{itquote}
      Also Bayesian logistic regression can be formulated based on 
      Polya-Gamma latent variables (Polson et al.) and that need to be 
      incorporated in your modeling strategy.
    \end{itquote}
    The Polya-Gamma parametrization is actually an important part of our method.
    It is introduced in equation (6) in the Section 3.2 and extensively covered 
    in Section 4 of the supplementary material.

    \begin{itquote}
      Why no comparison between the proposed method and at least the 
      group lasso which assigns the group penalty.
    \end{itquote}
    We included comparisons to the group lasso and several of its extensions in
    the simulations and data applications.

    \begin{itquote}
      The model contains several tuning/prior parameters. Sensitivity 
      analysis need to be provided for that.
    \end{itquote}
    We show in one of our applications that randomly assigning the features
    to the groups does not lead to bias in the tuning parameter estimates, 
    suggesting that the method is robust (at least in this application).

    \begin{itquote}
      How much your model is scalable? EM is notoriously slow.
    \end{itquote}
    We included a comparison to other methods in terms of computation time on 
    one of the applications in the discussion. The comparison shows that our 
    method is competitive and outperforms some commonly used methods in terms of
    computation time.

    \begin{itquote}
      The authors have also missed several relevant references and 
      should put some more focus on the literature review.
    \end{itquote}
    We have included one of the suggested references in the introduction. 
    However, we feel that the other suggested references are not relevant and
    have thus not included them.

    \textbf{Reply to reviewer 2}
    \begin{itquote}
      This paper considered the use of group elastic net method. 
      Compared to Meier et al (2008, JRSS B) who used group lasso penalty, you 
      proposed to use the group elastic net. This extension is fine, but the 
      novelty is rather little in this sense.
    \end{itquote}
    The group lasso is built on an eniterely different philosophy.
    Estimation is very different and not adaptive. Group lasso was designed for
    small interpretable groups of features, such as dummy variables. Our method
    is designed for larger groups of features, where the grouping is based on 
    external information that may informative for the prediction problem at 
    hand. The group lasso has one penalty parameter for all groups and estimates
    it by cross-validation. Our method uses a different penalty parameter per 
    group of features and estimates them by empirical Bayes.

    \begin{itquote}
      In (2.2), the penalization parameters $\lambda_1$ , $\lambda_2$ 
      and $\lambda_g$ appears in the forms of $\lambda_1 \sqrt{\lambda_g'}$ and 
      $\lambda_2 \lambda_g'$. A question is, is there is an identifiability 
      issue here? If not, why not?
    \end{itquote}
    There is not, since we fix $\lambda_1$ and $\lambda_2$ beforehand through 
    cross-validation. This leaves $G$ penalties to estimate for $G$ groups with
    empirical Bayes, where each group consists of many features.

    \begin{itquote}
      The computational load and time: As commented in the Discussion,
      the proposed method uses the double EM loop, which not only increases the
      chance of ending up in a local optimum but also can be very time consuming
      in computation. It would be helpful by commenting on the computation time,
      for example, in the real data analysis.
    \end{itquote}
    We included comparisons in terms of computation time in the discussion that 
    show that our method is competitive in terms of computation time. In 
    addition, we investigated different starting values for our algorithm in the
    applications presented in the manuscript and supplementary material. In none
    of the applications did this lead to different optimisation results.

    \closing{Yours Faithfully,}
    \encl{manuscript.pdf, supplement.pdf}

  \end{letter}
\end{document}

### 2019-02-27 docs/supplement.Rnw
changed the applications in manuscript.Rnw and supplement.Rnw

@@
   \begin{align*}
-  \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\
+  \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[
   -\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
@@
 
-\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
-\subsection{Partitioning based on differential expression}\label{sec:mirna}
-The data set is from a deep sequencing analysis on microRNA \cite[]{neerincx_combination_2018}.
-The study was done in 88 treated colorectal cancer patients, with the aim of classifying treatment response, coded as either non-progressive/remission (70 patients) or progressive (18 patients).
-After pre-processing and normalisation, 2114 microRNAs remained.
-In addition to the 2114 microRNAs we incorporated 4 unpenalized clinical covariates into the analysis: prior use of adjuvant therapy (binary), the type of systemic treatment regimen (ternary), age, and primary tumor differentiation (binary).
+  \section{Application to MicroRNAs in colorectal cancer}\label{sec:colorectal}
+  In this section we present additional results to Section [! section ref here]
+  in the Main Document.
 
-In a preliminary experiment on different subjects, the microRNA expression levels of metastatic colorectal tumour tissue were compared to normal non-colorectal tissue and primary colorectal tumour tissue was compared to primary colorectal tumour tissue \cite[]{neerincx_mir_2015}.
-This yielded 221 microRNAs that were differentially expressed in both comparisons ($\text{FDR} \leq 0.05$), versus 1893 not differentially expressed microRNAs.
-We expect that incorporation of this partitioning enhances therapy response classification, because tumor-specific miRNAs might be more relevant than non-specific ones.
-In addition, we divided the differentially expressed microRNAs even further into 127 highly differentially expressed microRNAs ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs ($0.001 < \text{FDR} \leq 0.05$).
-We will refer to this second partitioning as the three-group setting, as opposed to the first, the two-group setting.
+  \subsection{Random groups}
+  Considering that the features may be partitioned into many groups, with one 
+  parameter per group, we have to be aware of overfitting risks. We investigated
+  this using the data introduced in Section [! section ref here] of the Main 
+  Document, randomly dividing 
+  the features into three groups. We fixed the group sizes to the group sizes 
+  used in the three-group setting in Section [! section ref here] of the Main
+  Docuemtn. Under this random 
+  partitioning of the features, we expect all penalty multipliers to be 
+  estimated as one if no overfitting occurs.
 
-We compared \texttt{gren} to ridge and elastic net regression, and \texttt{GRridge}.
-For the elastic net methods (including \texttt{gren}) we set $\alpha \in \{ 0.05, 0.5, 0.95\}$.
-The estimated penalty multipliers are according to expectation: in all group-regularized methods, the 221 differentially expressed microRNAs receive the smallest penalty (Figure \ref{fig:col_bar}a).
-In the three-group setting the pattern is again as expected (Figure \ref{fig:col_bar}b): the highly expressed group, receives the lowest penalty, followed by the medium expressed group, while the non-expressed group receives the strongest penalty.
-\texttt{GRridge} is not able to distinguish between the medium and non-differentially expressed groups of microRNAs.
+  We compared the estimated multipliers to the estimates by \texttt{GRridge}. 
+  Since these results depend on one specific randomisation of the groups, we 
+  repeated the procedure 100 times and present the results in [! figure ref]. 
+  From this figure we see that the estimates for \texttt{gren} are close to one.
+  The estimates by \texttt{GRridge} show much more variation. Additionally, 
+  \texttt{GRridge} gives slightly biased estimates: the median penalty parameter
+  estimates are [! medians here] for the three groups. In contrast, the median 
+  estimates of \texttt{gren} are [! medians here].
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
-	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
-	\label{fig:col_bar}
-\end{figure}
-Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}.
-All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS.
-In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC.
-This is especially true for the larger models.
-\texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models.
-A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly.
-This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation.
-In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used.
-Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting.
-Note that models with few features are often desirable for clinical implementation.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
-	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
-	\label{fig:col_per}
-\end{figure}
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
-	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
-	\label{fig:col_bar}
-\end{figure}
+  [! figure here]
 
-Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}.
-All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS.
-In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC.
-This is especially true for the larger models.
-\texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models.
-A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly.
-This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation.
-In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used.
-Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting.
-Note that models with few features are often desirable for clinical implementation.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
-	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
-	\label{fig:col_per}
-\end{figure}
+  We repeated the simulation with ten evenly sized groups. Figure 
+  [! figure ref here] presents the estimated penalty multipliers for 100 random 
+  splits of the microRNA features into 10 evenly sized groups for 
+  \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$. 
+  From the Figure we see that \texttt{GRridge} shows more variation around one 
+  then the other models. In addition, the median penalty multipliers for the 
+  three \texttt{gren} models are all between [! medians here], while the 
+  \texttt{GRridge} median estimates range from [! medians here]. This indicates 
+  that \texttt{GRridge} suffers from a slight bias in the estimation, while the 
+  \texttt{gren} estimates appear to be unbiased.
 
-\section{Application to RNAseq in lymph node cancer}\label{sec:applymph}
+  [! figure here]
 
-CoRF data:
+  \subsection{Stability of selection}
+  One common criticism on the elastic net, and especially the lasso, is the 
+  instability of feature selection. That is, different data sets yield different
+  sets of selected features. To investigate the stability of feature selection 
+  by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA 
+  data \cite[]{neerincx_combination_2018}, where we stratified by the target of 
+  prediction, treatment response. 
+  For each of these bootstrap samples we estimated the regular elastic
+  net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected
+  25 features (including the 5 unpenalized covariates). We calculated the size 
+  of all ${50}\choose{2}$ intersections of selected features and present the 
+  results in Figure [! figure ref here].
 
-<<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Cross-validated AUC in lymph node metastasis example, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
+  [! figure here]
+
+  \section{Application to metabolomics in alzheimer}\label{sec:alzheimer}
+	We applied our method to an Alzheimer's dataset from 
+	\cite{de_leeuw_blood-based_2017}. The
+	data consist of 87 diagnosed Alzheimer's patients with the $\epsilon$4
+	allele of the APOE gene and 87 subjects with subjective un-diagnosed cognitive
+	decline, without the $\epsilon$4 allele. The aim of the study was to diagnose 
+	subjects with Alzheimer's disease based on their metabolic profile. After 
+	pre-processing, 230 metabolite expression levels were obtained. To enhance
+	classification performance, we included two sources of co-data: (a) the 
+	relative standard devitation of the measurements (RSD), a commonly 
+	used quality score (smaller is better) for metabolites, binned into five 
+	groups, and (b) the node degree of the metabolites, based on a differential 
+	network analysis of the metabolites, binned into three categories: 0 degree, 
+	positive, lower than average degree, and postive, higher than average degree. 
+	We expect a lower RSD and a higher node degree to be indicative of an 
+	important metabolite and therefore receive less penalty.
+
+	To estimate performance measures AUC and BSS, we split the data into 120 
+	training instances and 54 test instances. 
+	The training data is used to fit the models, while the test data
+	is used to estimate the performance measures on. We included the different 
+	co-data types seperately and combined, which resulted in the estimated 
+	multipliers and performance measures presented in Figure 
+	\ref{fig:barplot_lines_metabolomics_alzheimer_res5_auc}.
+<<barplot_lines_metabolomics_alzheimer_res5_auc, fig.cap="Results for the alzheimer example with (a) the estimated penalty multipliers for the quality score, (b) the estimated penalty multipliers for the network degree, (c) the estimated AUC, and (d) the estimated Brier skill score.", out.width="100%", fig.asp=1>>=
 @
 
-\subsection{Random groups}
-Considering that the features may be partitioned into many groups, with one parameter per group, we have to be aware of overfitting risks.
-We investigated this using the data introduced in Section \ref{sec:mirna}, randomly dividing the features into three groups.
-We fixed the group sizes to the group sizes used in the three-group setting in Section \ref{sec:mirna}.
-Under this random partitioning of the features, we expect all penalty multipliers to be estimated as one if overfitting does not occur.
+  In this example, \texttt{GRridge} and \texttt{gren} underperform due to a 
+  small number of features to learn the penalty parameters from. The other 
+  methods estimate one global penalty parameter, so do not suffer from the 
+  lower dimensionality of the problem.
 
-We compared the estimated multipliers to the estimates by \texttt{GRridge}.
-Since these results depend on one specific randomisation of the groups, we repeated the procedure 100 times and present the results in Figure \ref{fig:random1}.
-From this figure we see that the estimates for \texttt{gren} are close to one.
-The estimates by \texttt{GRridge} show much more variation.
-Additionally, \texttt{GRridge} gives slightly biased estimates: the median penalty parameter estimates are 1.16, 1, and 0.85 for the three groups.
-In contrast, the median estimates of \texttt{gren} are 1, 1, and 0.99.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.5\linewidth]{grEBEN_mirseq_Maarten_res6_boxplot.png}
-	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 3 random groups in colorectal miRNA data.}
-	\label{fig:random1}
-\end{figure}
+  \section{Application to microRNAs in cervical cancer}
+  A deep sequencing analysis on small non-coding ribonucleic acid (microRNAseq) 
+  was performed on 24 women with high-grade cervical intraepithelial 
+  neoplasia (CIN3) and 32 healthy women for the purpose of finding relevant 
+  screening markers for cervical cancer screening. The next generation 
+  sequencing analysis resulted in 2,576 transcripts. The data was normalized 
+  and pre-processed, rendering 752 transcripts. More details of the
+  data sets and the pre-processing steps are found in the supplementary   
+  material of \cite{novianti_better_2017}.
 
-We repeated the simulation with ten evenly sized groups.
-Figure \ref{fig:random2} presents the estimated penalty multipliers for 100 random splits of the microRNA features into 10 evenly sized groups for \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
-From the Figure we see that \texttt{GRridge} shows more variation around one then the other models.
-In addition, the median penalty multipliers for the three \texttt{gren} models are all between 0.99 and 1.02, while the \texttt{GRridge} median estimates range from 1.04 to 1.09.
-This indicates that \texttt{GRridge} suffers from a slight bias in the estimation, while the \texttt{gren} estimates appear to be unbiased.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.9\linewidth]{grEBEN_mirseq_Maarten_res8_boxplot.png}
-	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 10 random groups in colorectal miRNA data.}
-	\label{fig:random2}
-\end{figure}
+  The 752 microRNAs were divided in three classes based on their conservation 
+  status. Conservation status refers to whether the microRNAs are only found in
+  humans or are conserved across different species. They come in three classes:
+  (a) non-conserved (535), (b) conserved in mammals (70), and (c) 
+  conserved across most vertebrates (147). We expect (c) to contain the most 
+  important microRNAs, followed by (b) and (a), because these microRNAs have not
+  been lost during evolution and might therefore regulate important processes.
 
-\subsection{Stability of selection}
-One common criticism on the elastic net, and especially the lasso, is the instability of feature selection.
-That is, different data sets yield different sets of selected features.
-To investigate the stability of feature selection by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA data \cite[]{neerincx_combination_2018}, where we stratified by treatment response.
-For each of these bootstrap samples we estimated the regular elastic net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected 25 features (including the 5 unpenalized covariates).
-We calculated the size of all ${50}\choose{2}$ intersections of selected features and present the results in Figure \ref{fig:bootstrap}.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.9\linewidth]{gren_mirseq_Maarten_overlap.png}
-	\caption{Overlap between stratified bootstrap samples for the elastic net models with (a) $\alpha=0.05$, (b) $\alpha=0.5$, and (c) $\alpha=0.95$.}
-	\label{fig:bootstrap}
-\end{figure}
-From this figure it can be seen that stability of feature selection in \texttt{gren} is higher than in the regular elastic net for all three $\alpha$'s.
+<<barplots_micrornaseq_cervical_cancer_res1_auc, fig.cap="Estimated penalty multipliers for conservation status co-data", out.width="50%", fig.asp=2/3>>=
+@
+
+<<lines_micrornaseq_cervical_cancer_res1_auc, fig.cap="Cross-validated AUC in cervicl cancer example", out.width="50%", fig.asp=1>>=
+@
 

### 2019-02-27 code/metabolomics_alzheimer.R

@@ -50,7 +50,7 @@ feat <- fData(ESetMbolCSFPR2)
 # platformcode gives the type of metabolites
 platformcode <- feat$PlatformCode2
 
-# quality gives groups based on a quality measure of the assayd features
+# quality gives groups based on a quality measure of the assayed features
 part.RSDqc <- CreatePartition(feat$RSDqc, ngroup=3)
 quality <- rep(c(1:3), times=sapply(part.RSDqc, length))[
   order(unlist(part.RSDqc))]
@@ -160,7 +160,7 @@ write.table(res1, file="results/metabolomics_alzheimer_res1.csv")
 
 ################################### model 2 ####################################
 ### fitting the models
-set.seed(2018)
+set.seed(2019)
 ytrain <- alzheim.apoe[id.train]
 xtrain <- scale(metabol.apoe[id.train, ])
 ytest <- alzheim.apoe[-id.train]
@@ -251,7 +251,7 @@ write.table(res2, file="results/metabolomics_alzheimer_res2.csv")
 
 ################################### model 3 ####################################
 ### fitting the models
-set.seed(2018)
+set.seed(2019)
 ytrain <- alzheim.apoe[id.train]
 xtrain <- scale(metabol.apoe[id.train, ])
 ytest <- alzheim.apoe[-id.train]
@@ -342,7 +342,7 @@ write.table(res3, file="results/metabolomics_alzheimer_res3.csv")
 
 ################################### model 4 ####################################
 ### fitting the models
-set.seed(2018)
+set.seed(2019)
 ytrain <- alzheim.apoe[id.train]
 xtrain <- scale(metabol.apoe[id.train, ])
 ytest <- alzheim.apoe[-id.train]
@@ -462,3 +462,128 @@ res4 <- rbind(pred4, psel4, auc4)
 rownames(res4) <- c(paste0("pred", c(1:length(ytest))), paste0("psel", 1),
                     paste0("auc", 1))
 write.table(res4, file="results/metabolomics_alzheimer_res4.csv")
+
+
+################################### model 5 ####################################
+### fitting the models
+set.seed(2019)
+ytrain <- alzheim.apoe[id.train]
+xtrain <- scale(metabol.apoe[id.train, ])
+ytest <- alzheim.apoe[-id.train]
[…]

### 2019-02-27 code/micrornaseq_cervical_cancer.R

@@ -1,0 +1,132 @@
+#!/usr/bin/env Rscript
+
+### installation of packages
+if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
+   substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
+  if(!("devtools" %in% installed.packages())) {
+    install.packages("devtools")
+  }
+  library(devtools)
+  install_github("magnusmunch/gren/rpackage", local=FALSE,
+                 auth_token=Sys.getenv("GITHUB_PAT"))
+}
+
+### libraries
+library(gren)
+library(GRridge)
+library(grpreg)
+library(SGL)
+
+### load data
+load("old_data/mirsData.RData")
+
+### filter out the features with less than 2 unique counts
+unique.counts <- apply(t(as.matrix(mirsData$countData)), 2, function(s) {
+  length(unique(s))})
+select <- which(unique.counts > 2)
+filtered.micrornas <- t(as.matrix(mirsData$countData))[, select]
+
+### transform the data
+micrornas <- sqrt(filtered.micrornas + 3/8) - sqrt(3/8)
+cervical <- as.numeric(as.character(mirsData$response)=="CIN3")
+
+### conservation status of the filtered microRNAs
+conservation <- (as.numeric(levels(mirsData$conservation))[
+  mirsData$conservation] + 1)[select]
+# 1 is not-conserved, 2 is conserved across mammals, 3 broadly conserved across
+# most vertebrates
+
+################################### model 1 ####################################
+### fitting the models
+set.seed(2019)
+y <- cervical
+x <- scale(micrornas)
+part <- conservation
+p <- ncol(x)
+n <- nrow(x)
+
+psel <- c(seq(1, 5, 1), seq(7, 15, 2), seq(18, 30, 3), seq(34, 50, 4))
+
+fit1.gren1 <- gren(x, y, partitions=list(part=part), alpha=0.05, 
+                   standardize=TRUE, trace=FALSE, psel=psel)
+fit1.gren2 <- gren(x, y, partitions=list(part=part), alpha=0.5, 
+                   standardize=TRUE, trace=FALSE, psel=psel)
[…]

### 2019-02-27 code/micrornaseq_colorectal_cancer.R

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 
-### installation of packages
+### installation of gren if updated on github
 if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
    substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
   if(!("devtools" %in% installed.packages())) {
@@ -59,18 +59,31 @@ ytest <- benefit[-id.train]
 xtest <- scale(micrornas[-id.train, ])[, apply(micrornas[id.train, ], 2, sd)!=0]
 # set the constant micrornas to 0
 xtest[is.nan(xtest)] <- 0
-utest <- utrain[-id.train, ]
-part <- diff.threegroup[apply(micrornas[id.train, ], 2, sd)!=0]
+utest <- unpenal[-id.train, ]
+part <- diff.twogroup[apply(micrornas[id.train, ], 2, sd)!=0]
 p <- ncol(xtrain)
+u <- ncol(utrain)
 
-fit1.gren1 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.05, 
-                   standardize=TRUE, trace=FALSE)
-fit1.gren2 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.5, 
-                   standardize=TRUE, trace=FALSE)
-fit1.gren3 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.95, 
-                   standardize=TRUE, trace=FALSE)
+fit1.gren1 <- gren(xtrain, ytrain, unpenalized=utrain, 
+                   partitions=list(part=part), alpha=0.05, standardize=TRUE, 
+                   trace=FALSE)
+fit1.gren2 <- gren(xtrain, ytrain, unpenalized=utrain, 
+                   partitions=list(part=part), alpha=0.5, standardize=TRUE, 
+                   trace=FALSE)
+fit1.gren3 <- gren(xtrain, ytrain, unpenalized=utrain, 
+                   partitions=list(part=part), alpha=0.95, standardize=TRUE, 
+                   trace=FALSE)
 
-fit1.grridge <- grridge(t(xtrain), ytrain, list(part=split(1:p, part)))
+init1.grridge <- grridge(t(scale(micrornas)), benefit, 
+                         list(part=split(1:ncol(micrornas), diff.twogroup)),
+                         unpenal= ~ 1 + adjth2 + thscheme2 + thscheme3 + age + 
+                           pcrcdiff3, 
[…]

### 2019-03-01 code/figures.R
ran micrornaseq_colorectal_cancer.R 10 random groups and stability selection

@@ -1,5 +1,197 @@
 ################################## manuscript ##################################
-# ---- figures ----
+# ---- barplot_lines_micrornaseq_colorectal_cancer_res2_auc ----
+library(sp)
+load("results/micrornaseq_colorectal_cancer_fit2.Rdata")
+res <- read.table("results/micrornaseq_colorectal_cancer_res2.csv", 
+                  stringsAsFactors=FALSE)
+pred <- as.matrix(res[substr(rownames(res), 1, 4)=="pred", ])
+psel <- as.matrix(res[substr(rownames(res), 1, 4)=="psel", ])
+auc <- as.matrix(res[substr(rownames(res), 1, 3)=="auc", ])
+
+methods1 <- c("grridge", expression("gren, "~alpha==0.05, "gren, "~alpha==0.5,
+                                    "gren, "~alpha==0.95), "no co-data")
+methods2 <- c("ridge", "grridge", "gren1", "gren2", "enet1", "cmcp1")
+col1 <- grey.colors(length(methods1) - 1, start=0.3, end=0.9, gamma=2.2, 
+                    alpha=NULL)
+col2 <- bpy.colors(length(methods2), cutoff.tail=0.1)
+lty1 <- 2
+lty2 <- c(1, 2)
+labels1 <- list(diff.threegroup=expression("FDR">0.05,
+                                           atop(NA, atop(textstyle(
+                                             0.05>=phantom(0)),
+                                             textstyle("FDR" > 0.001))),
+                                           "FDR"<=0.001))
+labels2 <- expression("ridge", "GRridge", "gren,"~alpha==0.05, 
+                      "gren,"~alpha==0.5, "enet"~alpha==0.05, 
+                      "cMCP"~alpha==0.05)
+
+plot.data1 <- lapply(1:length(fit2.gren1$lambdag), function(s) {
+  rbind(fit2.grridge$lambdamults[[s]], fit2.gren1$lambdag[[s]],
+        fit2.gren2$lambdag[[s]], fit2.gren3$lambdag[[s]])})
+plot.data2 <- lapply(methods2, function(m) {
+  aggregate(auc[, grepl(m, colnames(auc))], 
+            list(psel=psel[, grepl(m, colnames(psel))]), mean)})
+names(plot.data2) <- methods2
+
+opar <- par(no.readonly=TRUE)
[…]

### 2019-03-01 code/micrornaseq_colorectal_cancer.R

@@ -408,3 +408,174 @@ for(k in 1:nsplits) {
 res3 <- multipliers
 rownames(res3) <- paste0("split", c(1:nsplits))
 write.table(res3, file="results/micrornaseq_colorectal_cancer_res3.csv")
+
+################################### model 3 ####################################
+###  create 100 random splits of features into 3 groups, same size as FDR groups
+set.seed(2019)
+nsplits <- 100
+y <- benefit
+x <- scale(micrornas)
+unpenal <- unpenal
+p <- ncol(x)
+u <- ncol(unpenal)
+part <- diff.threegroup
+
+multipliers <- data.frame(grridge=matrix(NA, nrow=nsplits, 
+                                         ncol=length(unique(part))),
+                          gren1=matrix(NA, nrow=nsplits, 
+                                       ncol=length(unique(part))),
+                          gren2=matrix(NA, nrow=nsplits, 
+                                       ncol=length(unique(part))),
+                          gren3=matrix(NA, nrow=nsplits, 
+                                       ncol=length(unique(part))))
+
+for(k in 1:nsplits) {
+  cat(paste("split ", k, "\n"))
+  set.seed(2019 + k)
+  
+  part.train <- sample(part)
+  
+  fit3.gren1 <- gren(x, y, unpenalized=unpenal, 
+                     partitions=list(part=part.train), alpha=0.05, 
+                     standardize=TRUE, trace=FALSE)
+  fit3.gren2 <- gren(x, y, unpenalized=unpenal, 
+                     partitions=list(part=part.train), alpha=0.5, 
+                     standardize=TRUE, trace=FALSE)
+  fit3.gren3 <- gren(x, y, unpenalized=unpenal, 
+                     partitions=list(part=part.train), alpha=0.95, 
+                     standardize=TRUE, trace=FALSE)
+  
+  fit3.grridge <- grridge(t(x), y, list(part=split(1:p, part.train)),
+                          unpenal= ~ 1 + adjth2 + thscheme2 + thscheme3 + age + 
+                            pcrcdiff3, dataunpen=as.data.frame(unpenal))
+  
+  multipliers[k, c(1:length(unique(part)))] <- fit3.grridge$lambdamults$part
[…]

### 2019-03-01 rpackage/R/gren.R

@@ -221,6 +221,7 @@ gren <- function(x, y, m=rep(1, nrow(x)), unpenalized=NULL, partitions=NULL,
   # starting values for the model parameters
   if(is.null(init$mu) | is.null(init$sigma) | is.null(init$chi) | 
      is.null(init$ci)) {
+    
     fit.start <- glmnet(x=x, y=ymat, family="binomial", alpha=0, lambda=lambda, 
                         standardize=FALSE, intercept=intercept, 
                         penalty.factor=c(rep(0, u), lambdamultvecold))

### 2019-03-01 docs/manuscript.Rnw
ran micrornaseq_colorectal_cancer.R 10 random groups and stability selection

@@
 
-	An example of such a study with extra information available is investigated in
+	Examples of such a study with extra information available are investigated in
 	Sections [! section refs here]. The first one is concerned with treatment 
@@
   (18 patients). After pre-processing and normalisation, 2114 microRNAs 
-  remained. In addition to the 2114 microRNAs, four clinical covariates were
-  included in the analysis: prior use of adjuvant therapy (binary),
-  the type of systemic treatment regimen (ternary), age, and primary tumor
-  differentiation (binary). These clinical covariates are included as 
-  unpenalized features in the model.
+  remained. In addition to the 2114 microRNAs, four clinical unpenalized 
+  covariates were included in the analysis: prior use of adjuvant therapy 
+  (binary), the type of systemic treatment regimen (ternary), age, and primary 
+  tumor differentiation (binary).
 
@@
   non-colorectal tissue and primary colorectal tumour tissue was compared to
-  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 221
-  microRNAs that were differentially expressed in both comparisons ($\text{FDR}
-  \leq 0.05$), versus 1893 not differentially expressed microRNAs. We expect
-  that incorporation of this partitioning enhances therapy response 
+  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 127
+  microRNAs that were differentially expressed in both comparisons with
+  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 not 
+  differentially expressed microRNAs. We expect that incorporation of this 
+  partitioning enhances therapy response 
   classification, because tumor-specific miRNAs might be more relevant than 
-  non-specific ones. In addition, we divided the differentially expressed
-  microRNAs even further into 127 highly differentially expressed microRNAs
-  ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs 
-  ($0.001 < \text{FDR} \leq 0.05$).  We will refer to this second partitioning
-  as the three-group setting, as opposed to the first, the two-group setting.
+  non-specific ones.
 
-  We split the data into 61 training, and 27 test instances and
-  compared the performance of \texttt{gren} on the test set to ridge, 
-  \texttt{GRridge}, elastic net, and the composite minimax
-  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. Further
-  comparisons to the sparse
+  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
+  elastic net, and to the best performing
+  group lasso extension, the composite minimax
+  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. 
+  For the methods with an extra tuning parameter 
+  $\alpha$, we picked the one (or two) that gave the best performance and refer
+  the reader to SM Section [! section ref here] for more comparisons to 
+  different $\alpha$s and to the sparse
   group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and the group
-  exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015} are presented in 
-  the SM [! section ref here]. We compared the methods on AUC. More performance
-  measures are presented in the SM [! section ref here].
-  For the elastic net methods (including \texttt{gren}) and \texttt{cMCP}, we 
-  set $\alpha \in \{ 0.05, 0.5, 0.95\}$. We experienced numerical issues with 
+  exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}.
+
+  To estimate performance, we split the data into 61 training, and 27 test 
+  instances stratified by treatment response. We estimated the models on the 
+  training data and calculated AUC on the test data. We present AUC for a range 
+  of model sizes, together with the estimated 
+  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
+  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. More 
+  performance measures comparisons are presented in Section [! section ref here]
+  of the SM. In addition, we investigated the sensitivity of
+  the multiplier estimation using this data set in SM Section 
+  [! section ref here]. A small sidenote: We experienced numerical issues with 
   \texttt{GRridge} when cross-validating 
@@
 
-  The results are presented in Figure 
-  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. 
   The penalty multipliers estimated by 
@@
   sizes. This is especially pronounced for the larger models. The performance of
-  \texttt{cmcp} is somewhat unstable. We noticed this is in general
+  \texttt{cMCP} is somewhat unstable. We noticed this is in general
   for the group-lasso and its extensions (see SM [! section ref here]). Overall,
@@
   benefited predictive performance: both \texttt{GRridge} and \texttt{gren} 
-  outperform their respective non-group-regularized versions.
+  outperform their respective non-group-regularized versions, albeit it only
+  slightly for \texttt{GRridge}.
 <<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
@@
   In this application, the aim is to predict lymph node metastasis in patients 
-  with HPV negative oral cancer using sequenced RNAs from TCGA 
+  with HPV-negative oral cancer using sequenced RNAs from TCGA 
   \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
@@
   carcinomas on 133 patients with HPV-negative tumours in the oral cavity. Of 
-  these patients, 76 had lymph node metastasis, and 57 did not. For a thorough 
-  introduction of these data, see \cite{te_beest_improved_2017}.
+  these patients, 76 suffered from lymph node metastasis, while 57 did not. 
+  For a thorough introduction of these data, see \cite{te_beest_improved_2017}.
 
-  As co-data we considered two sources: (a) the cis-correlation between TCGA DNA
-  copy number measurements and the RNASeqv2 data on the same patients, 
-  quantified through Kendall's $\tau$ and binned into five groups, and (b) 
-  p-values from an external microarray data set \cite[]{mes_prognostic_2017}, 
-  binned into five groups. We expect features 
-  with a positive Kendall's $\tau$ to be more important in metastasis prediction   
-  \cite[]{masayesva_gene_2004}. Likewise, we expect features with a low 
-  p-value to be more important.
+  As co-data we considered two sources: (a) the cis-correlation between the
+  the RNASeqv2 data and DNA copy numbers on the same patients, 
+  quantified by Kendall's $\tau$ and binned into five groups, and (b) 
+  $p$-values from an external microarray data set described in 
+  \cite{mes_prognostic_2017}, again binned into five groups. We expect features 
+  with a large positive Kendall's $\tau$ to be more important in 
+  metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
+  features with low $p$-values to be more important.
 
-  We split our data into 120 training instances and 97 test instances, trained
-  the models on the training set and estimated AUC on the test set. Other 
-  performance measures are presented in the SM Section [! section ref here]. We
-  compared the same models as in Section \ref{sec:appcolon}. In this 
-  application, the co-data consists of two partitions with overlapping groups.
-  \texttt{cMCP} is not designed to deal with overlapping groups, so we 
-  cross-tabulate the two co-data partitions to create one paritioning with 25 
-  groups for \texttt{cMCP}. In addition, we compare to an extension of 
-  \texttt{cMCP} that is able to handle multiple groupings 
-  \cite[]{zeng_overlapping_2016}. We present the results in Figure 
-  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. 
-  Additional comparisons to \texttt{SGL}, \texttt{gel} and 
-  the overlapping group version of \texttt{gel} \cite[]{zeng_overlapping_2016}
-  \cite[]{breheny_group_2015} are found in the SM Section [! section ref here].
-<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.", out.width="100%", fig.asp=1>>=
+  We compared \texttt{gren} to the same methods as in Section 
+  \ref{sec:appcolon}. However, the co-data consists of two partitions with 
+  overlapping groups, so we use an extension of \texttt{cMCP} that is able to 
+  handle multiple groupings \cite[]{zeng_overlapping_2016}. We split the data 
+  into 120 training instances and 97 test instances, trained
+  the models on the training set and estimated AUC on the test set. We present 
+  the results in Figure 
+  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Other 
+  comparisons and performance measures are presented in the SM Section 
+  [! section ref here].
+
+  The estimated penalty multipliers follow the expected pattern. Larger 
+  cis-correlations receive smaller penalties. Likewise, smaller $p$-values are
+  also penalized less. This pattern is slightly less pronounced in \texttt{gren}
+  with $\alpha=0.95$ than in the other methods. The predictive performance of
+  \texttt{gren} is slighty better than the other methods for a range of model
+  sizes. We also note that the overlapping group \texttt{cMCP} is much more
+  stable than the regular \texttt{cMCP} (see in SM [! section ref here]), 
+  where we use as grouping the cross-tabulation of the two paritions. In this 
+  example, ridge and \texttt{GRridge} perform almost identical.
+<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.
+OcMCP refers to the overlapping group cMCP.", out.width="100%", fig.asp=1>>=
 @
-
-  We present additional applications and comparisons in the SM Sections 
-  [! sections ref here].
 
@@
 	while in our case they are estimated on the group level, rendering the 
-	adaptation more robust. As can be seen from the simulations and real data 
-	example in Sections \ref{sec:simulations} and [! section ref here], 
-	adaptation to 
+	adaptation more robust. As can be seen from both the simulations in Section 
+	\ref{sec:simulations} and the application examples with different types of 
+	outcome data (prognostic and diagnostic) and marker types (RNAseq, microRNAs, 
+	and metabolomics) in Sections [! section refs here], and the SM
+	Sections [! section refs here], adaptation to 
 	external data may be beneficial for prediction and feature selection. We 
@@
 	A (stable) \texttt{R} package is available from 
-	\url{https://CRAN.R-project.org/package=gren}.
+	\url{https://CRAN.R-project.org/package=gren}. All materials needed to   
+	recreate this document are available from 
+	\url{https://github.com/magnusmunch/gren}.
 
@@
 	Supplementary Material is available online from 
-	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All materials needed to   
-	recreate this document are available from 
-	\url{https://github.com/magnusmunch/gren}.
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}.
 
 	\section*{Acknowledgements}
-	\textit{Conflict of Interest}: None declared.
+	\textit{Conflict of Interest}: None declared. We thank the editors and 
+	referees for their useful suggestions and comments.
 

### 2019-03-01 docs/supplement.Rnw
ran micrornaseq_colorectal_cancer.R 10 random groups and stability selection

@@
   parameter per group, we have to be aware of overfitting risks. We investigated
-  this using the data introduced in Section [! section ref here] of the Main 
-  Document, randomly dividing 
-  the features into three groups. We fixed the group sizes to the group sizes 
-  used in the three-group setting in Section [! section ref here] of the Main
-  Docuemtn. Under this random 
+  this using the data introduced in Section [! section ref here] of the MD, 
+  randomly dividing the features into three groups. We fixed the group sizes to 
+  the group sizes used in the MD. Under this random 
   partitioning of the features, we expect all penalty multipliers to be 
@@
 
+<<>>=
+res <- read.table("results/micrornaseq_colorectal_cancer_res3.csv", 
+                  stringsAsFactors=FALSE)
+medians <- round(apply(res, 2, median), 2)
+@
   We compared the estimated multipliers to the estimates by \texttt{GRridge}. 
   Since these results depend on one specific randomisation of the groups, we 
-  repeated the procedure 100 times and present the results in [! figure ref]. 
+  repeated the procedure 100 times and present the results in Figure 
+  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers}. 
   From this figure we see that the estimates for \texttt{gren} are close to one.
-  The estimates by \texttt{GRridge} show much more variation. Additionally, 
-  \texttt{GRridge} gives slightly biased estimates: the median penalty parameter
-  estimates are [! medians here] for the three groups. In contrast, the median 
-  estimates of \texttt{gren} are [! medians here].
-
-  [! figure here]
+  The estimates by \texttt{GRridge} show much more variation. We found no strong
+  indication of bias, with median estimated penalty parameters 
+  \Sexpr{medians[c(1:3)]} for
+  \texttt{GRridge} and \Sexpr{medians[c(4:6)]}; \Sexpr{medians[c(7:9)]},
+  and \Sexpr{medians[c(10:12)]} for \texttt{gren} with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
+<<boxplots_micrornaseq_colorectal_cancer_res3_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 3 random groups of features.", out.width="50%", fig.asp=1>>=
+@
 
@@
   \texttt{gren} estimates appear to be unbiased.
-
-  [! figure here]
+<<boxplots_micrornaseq_colorectal_cancer_res4_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
+@
 
@@
   [! figure here]
+
+  \section{Application to RNAseq in lymph node cancer}
+  \texttt{cMCP} is not designed to deal with overlapping groups, so we 
+  cross-tabulate the two co-data partitions to create one paritioning with 25 
+  groups for \texttt{cMCP}.  
+  Additional comparisons to \texttt{SGL}, \texttt{gel} and 
+  the overlapping group version of \texttt{gel} \cite[]{zeng_overlapping_2016}
+  \cite[]{breheny_group_2015} are found in the SM Section [! section ref here].
 

### 2019-03-06 docs/supplement.Rnw
stability selection in parallel

@@
 opts_chunk$set(fig.align='center', fig.path="../figs/", 
-               echo=FALSE, cache=FALSE, message=FALSE)
+               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
 @
@@
 
+  \subsection{Additional comparisons}
+
   \subsection{Random groups}
@@
   \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers}. 
-  From this figure we see that the estimates for \texttt{gren} are close to one.
-  The estimates by \texttt{GRridge} show much more variation. We found no strong
+  From this figure we see that the estimates for \texttt{gren} are relatively 
+  close to one.
+  The estimates by \texttt{GRridge} show more variation. We found no strong
   indication of bias, with median estimated penalty parameters 
@@
 
+<<>>=
+res <- read.table("results/micrornaseq_colorectal_cancer_res4.csv", 
+                  stringsAsFactors=FALSE)
+medians <- round(sapply(c("grridge", paste0("gren", 1:3)), function(s) {
+  median(as.matrix(res[, grepl(s, colnames(res))]))}), 2)
+@
   We repeated the simulation with ten evenly sized groups. Figure 
-  [! figure ref here] presents the estimated penalty multipliers for 100 random 
+  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res4_multipliers} presents 
+  the estimated penalty multipliers for 100 random 
   splits of the microRNA features into 10 evenly sized groups for 
@@
   From the Figure we see that \texttt{GRridge} shows more variation around one 
-  then the other models. In addition, the median penalty multipliers for the 
-  three \texttt{gren} models are all between [! medians here], while the 
-  \texttt{GRridge} median estimates range from [! medians here]. This indicates 
-  that \texttt{GRridge} suffers from a slight bias in the estimation, while the 
-  \texttt{gren} estimates appear to be unbiased.
+  then the other models. There is no indication of bias: the median penalty 
+  multiplier for 
+  \texttt{GRridge} is \Sexpr{medians[1]}, while the median multipliers in
+  \texttt{gren} are \Sexpr{medians[c(2:4)]} for 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
 <<boxplots_micrornaseq_colorectal_cancer_res4_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
@@
 
+  \section{Choice of $\alpha$}	
+	Choice of $\alpha$ does matter in Figure 
+	\ref{fig:lines_colorectal_lymph_node_gren_auc}.
+<<lines_colorectal_lymph_node_gren_auc, fig.cap="Estimated AUC for gren in the (a) colorectal cancer example and the (b) lymph node metastasis example.", out.width="100%", fig.asp=1/2>>=
+@
+
 \section{Empirical-variational Bayes for the ridge and lasso}\label{sec:evbridgelasso}

### 2019-03-06 code/micrornaseq_colorectal_cancer.R

@@ -18,6 +18,9 @@ library(grpreg)
 library(Biobase)
 library(SGL)
 library(grpregOverlap)
+library(foreach)
+library(doParallel)
+library(microbenchmark)
 
 ### load data
 load("data/forMagnusN88.Rdata")
@@ -64,25 +67,25 @@ part <- diff.twogroup[apply(micrornas[id.train, ], 2, sd)!=0]
 p <- ncol(xtrain)
 u <- ncol(utrain)
 
-fit1.gren1 <- gren(xtrain, ytrain, unpenalized=utrain, 
-                   partitions=list(part=part), alpha=0.05, standardize=TRUE, 
+fit1.gren1 <- gren(xtrain, ytrain, unpenalized=utrain,
+                   partitions=list(part=part), alpha=0.05, standardize=TRUE,
                    trace=FALSE)
-fit1.gren2 <- gren(xtrain, ytrain, unpenalized=utrain, 
-                   partitions=list(part=part), alpha=0.5, standardize=TRUE, 
+fit1.gren2 <- gren(xtrain, ytrain, unpenalized=utrain,
+                   partitions=list(part=part), alpha=0.5, standardize=TRUE,
                    trace=FALSE)
-fit1.gren3 <- gren(xtrain, ytrain, unpenalized=utrain, 
-                   partitions=list(part=part), alpha=0.95, standardize=TRUE, 
+fit1.gren3 <- gren(xtrain, ytrain, unpenalized=utrain,
+                   partitions=list(part=part), alpha=0.95, standardize=TRUE,
                    trace=FALSE)
 
-init1.grridge <- grridge(t(scale(micrornas)), benefit, 
+init1.grridge <- grridge(t(scale(micrornas)), benefit,
                          list(part=split(1:ncol(micrornas), diff.twogroup)),
-                         unpenal= ~ 1 + adjth2 + thscheme2 + thscheme3 + age + 
-                           pcrcdiff3, 
+                         unpenal= ~ 1 + adjth2 + thscheme2 + thscheme3 + age +
+                           pcrcdiff3,
                          optl=NULL, dataunpen=as.data.frame(unpenal))
 
 fit1.grridge <- grridge(t(xtrain), ytrain, list(part=split(1:p, part)),
-                        unpenal= ~ 1 + adjth2 + thscheme2 + thscheme3 + age + 
[…]

### 2019-03-06 docs/manuscript.Rnw
stability selection in parallel

@@
 
-	Examples of such a study with extra information available are investigated in
-	Sections [! section refs here]. The first one is concerned with treatment 
+	Two such studies are introduced in Sections \ref{sec:colorectal} and
+	\ref{sec:lymphnode}. The first study is concerned with treatment 
 	response prediction for colorectal cancer patients, based on sequenced
-	microRNAs. In the second example we attempt to predict lymph node metastases,
-	using RNAseq data. In both settings we have additional information on the 
-	features in the form of a partitioning available. Furthermore, in both
-	examples data is sparse, so we would like to 
-	include the additional information to enhance predictive performance of our
-	models. Inclusion of the partitions is not straightforward using standard 
-	methods available in literature. At first sight, such
-	an \textit{a priori} grouping of the features suggests the group lasso
-	\cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which
-	penalises and selects features at the group level. There are two limitations 
-	to the group lasso: (a) it selects groups as a whole instead of single 
-	features and (b) it is not adaptive: the amount of penalization a group
-	receives depends on a `global' penalty parameter common to all groups.
-	Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013}
-	partly deal with (a), but do not address (b).
+	microRNAs. In the second study we attempt to predict lymph node metastases,
+	using RNAseq data. In both settings additional information on the 
+	features is available in the form of a partitioning. Furthermore, in both
+	examples data is sparse, so any extra included information may help to boost
+	predictive performance. Inclusion of such feature partitions in 
+	the analysis is not straightforward with standard methods.
+
+	At first sight, such an \textit{a priori} grouping of the features suggests 
+	the group lasso
+	\cite[]{meier_group_2008, yuan_model_2006} or one of its extensions, because
+	the group lasso penalises and selects features at the group level. 
+	There are two limitations to this approach: the group lasso (a) selects groups
+	as a whole instead of single 
+	features and (b) does not penalise adaptively: all groups are penalised 
+	equally. Extensions such as the sparse group lasso 
+	\cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
 
@@
 	outcome data in omics research, we focus on the logistic elastic net.
+
 	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
@@
 	possibility of adaptive inclusion of external information on the features to
-	aid in prediction and feature selection. We would like to emphasize that our
-	method is different from recent developments in grouped penalized regression
-	such as \cite{dondelinger_joint_nodate}, where the grouping is over the
-	observations as opposed to the features
+	aid in prediction and feature selection. Bayesian versions of support vector 
+	machines have been used in classification problems as well 
+	\cite[]{chakraborty_bayesian_2011}, but just as VBVS, these methods lack 
+	adaptive inclusion of external information on the features. We emphasize that 
+	our method is different from recent developments in grouped penalized 
+	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
+	features, the observations are partitioned.
 
-	In line with the above, our proposed method is (a) subjective: the few
-	remaining tuning parameters may be cross-validated, (b) adaptive: prior
-	information is automatically and subjectively weighed with respect to its
-	informativeness, (c) fast compared to full Bayesian analysis or 
-	cross-validation, and (d) requires minimal input by the user: only the data 
-	and grouping of the features is required.
+	In line with the above, our proposed method is (a) subjective: the 
+	hyper-parameters are estimated from the data and not chosen on objective 
+	grounds, (b) adaptive: prior information is automatically and subjectively 
+	weighed with respect to its informativeness, (c) fast compared to full 
+	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
+	user: only the data and grouping of the features is required.
 
-	We introduce the Bayesian logistic (generalised) elastic net model in Section
+	The rest of the paper is structured as follows. We introduce the Bayesian 
+	logistic generalised elastic net model in Section
 	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
@@
 	Bayes estimation of multiple, group-specific penalty parameters. In Section
-	\ref{sec:simulations} and Sections [! section ref here] and 
-	[! section ref here], we demonstrate the approach in a
-	simulation study and a two real data examples, respectively. In Section 
-	\ref{sec:discussion} we conclude by both discussing some differences and
+	\ref{sec:simulations} we compare the method in a simulation study. In Sections
+	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
+	the two previously introduced data sets. In Section \ref{sec:discussion} 
+	we conclude by both discussing some differences and
 	parallels between the proposed method, termed \texttt{gren}, and other methods
@@
 	\section{Model}\label{sec:model}
-	\subsection{Logistic elastic net regression}\label{sec:logitmodel}
 	In logistic regression the outcome variables are assumed to be binary or sums
@@
 
-	In elastic net regression, the penalised likelihood is maximised to yield
-	parameter estimates:
-	$$
-	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
-	\frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
-	$$
-	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} 
-	y_1 & \cdots & y_n \end{bmatrix} \tr$. The likelihood term is thus 
-	complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the
-	parameters, multiplied by respective penalty parameters $\lambda_1$ and
-	$\lambda_2$. We have scaled the penalty parameters by $\frac{1}{2}$, to make
-	them correspond with the Bayesian version in the following. The maximiser of
-	the penalised likelihood corresponds to the posterior mode of a Bayesian
-	elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
-	\begin{equation}\label{eq:elasticnetprior}
-	  \beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1
-	  |\beta_j| + \lambda_2 \beta_j^2)\].
-	\end{equation}
-	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is 
-	dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the 
-	Supplementary Material (SM), along with further details on the elastic net
-	prior). The elastic net combines $L_1$ and $L_2$-norm penalisation such that
-	the model parameters are shrunken towards zero. The $L_1$-norm may set some of
-	the estimates exactly to zero, thus automatically selecting features. The 
-	$L_2$-norm ensures that correlated features behave similarly, as one would
-	generally require. We extend the elastic net model to the generalised elastic
-	net to allow for different penalty parameters per group of features.
-
-	\subsection{The generalised elastic net}\label{sec:genen}
 	Assume we have a partitioning of the features in $G$ groups, such that each
 	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
-	$g$ for $g=1, \dots, G$. To apply differential penalisation we include a 
-	group-specific weight $w_g \in \mathbb{R}_{>0}$. The model parameters are now
-	estimated by:
+	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
+	group-specific weight. In generalised elastic net regression, the penalised 
+	likelihood is maximised to yield parameter estimates:
 	\begin{align}
-	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
 	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
@@
 	  \label{eq:weightedlik}\\
-	  & = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} 
+	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
 	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
@@
 	\end{align}
-	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific
-	weights may be interpreted as penalty multipliers. Throughout the following we
-	assume that the geometric mean of the multipliers, weighted by their 
+	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
+	observed data $\y = \begin{bmatrix} 
+	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
+	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
+	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
+	group-specific weights may be interpreted as penalty multipliers. Note that
+	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.
+
+  Throughout the following we assume that the geometric mean of the multipliers, 
+	weighted by their 
 	respective group sizes, is one, such that the average shrinkage of the model
-	parameters is determined by the `global' $\lambda_1$ and $\lambda_2$. That is,
+	parameters is determined by $\lambda_1$ and $\lambda_2$. That is,
 	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
@@
 	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
-	(\ref{eq:weightedlik}).
+	(\ref{eq:weightedlik}). The elastic net combines $L_1$ and $L_2$-norm 
+	penalisation such that the model parameters are shrunken towards zero. 
+	The $L_1$-norm may set some of the estimates exactly to zero, thus 
+	automatically selecting features. The $L_2$-norm ensures that correlated 
+	features behave similarly, as one would generally require.
 
-	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as
+	The maximiser of the penalised likelihood in the elastic net, corresponds 
+	to the posterior mode of a 
+	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
+	\cite{li_bayesian_2010} show that the elastic net prior 
+	(see SM Section [! section reference here] for details on this prior) may be 
+	written as
 	a computationally more convenient scale mixture of normals, with mixing 
@@
 	is approximated by a tractable form and estimated by optimizing a lower bound 
-	on the marginal likelihood of this model (see Section 3 of the SM for the 
+	on the marginal likelihood of this model (see Section [! section ref here] 
+	of the SM for the 
 	lower bound of the proposed model). For an extensive introduction and concise 
@@
 	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
-	choices of $\alpha$. We recommend the latter. Otherwise we note that, in our 
-	experience, the choice $\alpha=0.5$ often gives good results.
+	choices of $\alpha$. We recommend the latter.
 
-	For estimating the penalty multipliers, the intractable posterior expectation
-	in (\ref{eq:ebEM}) is approximated by the variational posterior:
+	To estimate the penalty multipliers, the intractable posterior expectation
+	in (\ref{eq:ebEM}) is approximated using the variational posterior:
 	$$
@@
 	Feature selection is often desirable in high-dimensional prediction problems 
-	and omics data problems are no exception. A selection of biomarkers may lead 
+	and omics data problems are no exception. For example, a selection of 
+	biomarkers may lead 
 	to a large cost reduction by supporting targeted assays. Bayesian feature
@@
 
-	\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
-  The presented data is from a deep sequencing analysis on microRNAs 
-  \cite[]{neerincx_combination_2018}. The study was done in 88 treated 
-  colorectal cancer patients, with the aim of predicting treatment response,
-  coded as either non-progressive/remission (70 patients) or progressive 
+	\section{Application to microRNAs in colorectal cancer}
+	\label{sec:colorectal}
+  As shortly touched upon in Section \ref{sec:introduction}, the data is from a 
+  deep sequencing analysis on microRNAs 
+  \cite[]{neerincx_combination_2018}. This study in 88 treated 
+  colorectal cancer patients was conducted with the aim of predicting treatment 
+  response, coded as either non-progressive/remission (70 patients) or 
+  progressive 
   (18 patients). After pre-processing and normalisation, 2114 microRNAs 
@@
   $\alpha$, we picked the one (or two) that gave the best performance and refer
-  the reader to SM Section [! section ref here] for more comparisons to 
-  different $\alpha$s and to the sparse
-  group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and the group
-  exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}.
+  the reader to SM Section [! section ref here] for more comparisons with 
+  different $\alpha$s and comparisons to the sparse
+  group lasso by \cite{simon_sparse-group_2013} and group
+  exponential lasso \cite[]{breheny_group_2015}.
 
-  To estimate performance, we split the data into 61 training, and 27 test 
-  instances stratified by treatment response. We estimated the models on the 
+  To estimate performance, we split the data into 61 training and 27 test 
+  instances, stratified by treatment response. We estimated the models on the 
   training data and calculated AUC on the test data. We present AUC for a range 
@@
   \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. More 
-  performance measures comparisons are presented in Section [! section ref here]
+  performance measures are presented in Section [! section ref here]
   of the SM. In addition, we investigated the sensitivity of
-  the multiplier estimation using this data set in SM Section 
-  [! section ref here]. A small sidenote: We experienced numerical issues with 
+  the multiplier estimation using in SM Section 
+  [! section ref here]. We experienced numerical issues with 
   \texttt{GRridge} when cross-validating 
@@
   expressed group receives the lowest penalty, followed by the medium expressed
-  group, while the non-expressed group receives the strongest penalty. 
+  group, while the non-expressed group receives the strongest penalty 
+  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}a). 
   \texttt{gren} seems to outperform the other methods for a range of model 
-  sizes. This is especially pronounced for the larger models. The performance of
-  \texttt{cMCP} is somewhat unstable. We noticed this is in general
-  for the group-lasso and its extensions (see SM [! section ref here]). Overall,
-  it seems as though the inclusion of the extra information on the features 
-  benefited predictive performance: both \texttt{GRridge} and \texttt{gren} 
-  outperform their respective non-group-regularized versions, albeit it only
-  slightly for \texttt{GRridge}.
+  sizes in terms of AUC 
+  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}b). 
+  This is especially pronounced for the larger models. The performance of
+  \texttt{cMCP} is somewhat unstable. We noticed that this is in general the 
+  case for the group-lasso and its extensions (see SM [! section ref here]). 
+  Overall, it seems as though the inclusion of the extra information on the 
+  features benefited predictive performance: both \texttt{GRridge} and 
+  \texttt{gren} outperform their respective non-group-regularized versions, 
+  albeit it only slightly for \texttt{GRridge}.
 <<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
@@
 
-  \section{Application to RNAseq in lymph node cancer}\label{sec:applymph}
-  In this application, the aim is to predict lymph node metastasis in patients 
+  \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
+  The aim of this study is to predict lymph node metastasis in patients 
   with HPV-negative oral cancer using sequenced RNAs from TCGA 
   \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
-  normalised TCGA RNASeqv2 profiles of head-and-neck squamous cell 
-  carcinomas on 133 patients with HPV-negative tumours in the oral cavity. Of 
-  these patients, 76 suffered from lymph node metastasis, while 57 did not. 
-  For a thorough introduction of these data, see \cite{te_beest_improved_2017}.
+  3096 transformed and normalised TCGA RNASeqv2 profiles of head-and-neck 
+  squamous cell carcinomas on 133 patients with HPV-negative tumours in the oral
+  cavity. Of these patients, 76 suffered from lymph node metastasis, while 57 
+  did not. For a thorough introduction of these data, see 
+  \cite{te_beest_improved_2017}.
 
-  As co-data we considered two sources: (a) the cis-correlation between the
-  the RNASeqv2 data and DNA copy numbers on the same patients, 
-  quantified by Kendall's $\tau$ and binned into five groups, and (b) 
-  $p$-values from an external microarray data set described in 
+  As external feature information, we considered two sources: (a) the 
+  cis-correlation between the RNASeqv2 data and TCGA DNA copy numbers on 
+  the same patients, 
+  quantified by Kendall's $\tau$ and binned into five groups. We also considered
+  (b) $p$-values from an external microarray data set described in 
   \cite{mes_prognostic_2017}, again binned into five groups. We expect features 
@@
   We compared \texttt{gren} to the same methods as in Section 
-  \ref{sec:appcolon}. However, the co-data consists of two partitions with 
-  overlapping groups, so we use an extension of \texttt{cMCP} that is able to 
-  handle multiple groupings \cite[]{zeng_overlapping_2016}. We split the data 
-  into 120 training instances and 97 test instances, trained
-  the models on the training set and estimated AUC on the test set. We present 
-  the results in Figure 
+  \ref{sec:colorectal}. However, because the co-data consists of two partitions 
+  with overlapping groups, we use an extension of \texttt{cMCP} that allows for
+  overlapping groups \cite[]{zeng_overlapping_2016}.
+  We estimated AUC for the methods on an independent validation set of 97 
+  HPV-negative oral cancer patients \cite[]{mes_prognostic_2017}. The features 
+  in this validation set were
+  microarrays, normalised to account for a different scale of the transformed 
+  RNAseq data. We present the results on this validation set in Figure 
   \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Other 
@@
 
-  The estimated penalty multipliers follow the expected pattern. Larger 
-  cis-correlations receive smaller penalties. Likewise, smaller $p$-values are
-  also penalized less. This pattern is slightly less pronounced in \texttt{gren}
-  with $\alpha=0.95$ than in the other methods. The predictive performance of
+  The estimated penalty multipliers follow the expected pattern 
+  (Figure \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}a): 
+  larger cis-correlations receive smaller penalties. 
+  Likewise, smaller $p$-values are penalized less. This pattern is slightly 
+  less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the other 
+  methods. The predictive performance of
   \texttt{gren} is slighty better than the other methods for a range of model
@@
   stable than the regular \texttt{cMCP} (see in SM [! section ref here]), 
-  where we use as grouping the cross-tabulation of the two paritions. In this 
-  example, ridge and \texttt{GRridge} perform almost identical.
-<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.
-OcMCP refers to the overlapping group cMCP.", out.width="100%", fig.asp=1>>=
+  where we create one paritioning by cross-tabulation of the two paritions. 
+  In this example, ridge and \texttt{GRridge} perform almost identical.
+<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.", out.width="100%", fig.asp=1>>=
 @
@@
 	adaptation more robust. As can be seen from both the simulations in Section 
-	\ref{sec:simulations} and the application examples with different types of 
-	outcome data (prognostic and diagnostic) and marker types (RNAseq, microRNAs, 
-	and metabolomics) in Sections [! section refs here], and the SM
-	Sections [! section refs here], adaptation to 
-	external data may be beneficial for prediction and feature selection. We 
-	believe that this is due to the `borrowing of information' effect: estimates 
-	that are believed to be similar are shrunken in the same way, yielding 
-	overall, better estimates.
+	\ref{sec:simulations} and the applications to different outcomes 
+	(prognostic or diagnostic) and marker types 
+	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
+	\ref{sec:lymphnode}, and SM Sections [! section refs here], adaptation to 
+	external data may be beneficial for prediction and feature selection. 
+	We believe that this is due to the `borrowing of information' effect: 
+	estimates of feature effects that behave similarly are shrunken similarly, 
+	yielding overall, better estimates.
 
-	Another obvious comparison is to the group lasso \cite[]{meier_group_2008,
-	yuan_model_2006}. Although it is similar in the sense that it shrinks on the 
+	A common criticism of the lasso (and elastic net) is unstable feature 
+	selection: different data sets often lead to wildly differing sets of selected
+	features. We investigated stability selection on the data from Section
+	\ref{sec:colorectal} by taking bootstraps samples from the data, estimating
+	\texttt{gren} and looking at the overlap between selected features over 
+	different bootstrap samples. We compared it to the regular elastic net and 
+	found that inclusion of extra feature information increases selection 
+	stability (see SM Section [! section ref here]). In addition, we investigated 
+	the stability of penalty multiplier estimation by examining the distribution
+	of estimated penalty parameters from 100 random partitionings of the features
+  (see SM Section [! section ref here]). We found that the penalty parameters
+  tend to cluster around one, as is expected with random groups, and are 
+  unbiased.
+
+	As shortly touched upon in Section \ref{sec:introduction}, another obvious 
+	comparison is to the group lasso \cite[]{meier_group_2008,
+	yuan_model_2006} and its extensions. Although it is similar in the sense that 
+	it shrinks on the 
 	group level, the group lasso is built upon an entirely different philosophy: 
-	it sets whole groups of features to zero. The intended application of such a 
-	group-wise penalty is to small interpretable groups of features, like, for 
-	example, dummies of a categorical variable. Another difference between 
+	its intended application is to small interpretable groups of features, like, 
+	for example, dummies of a categorical variable. Another difference between 
 	\texttt{gren} and the group lasso is the number of penalty parameters. 
@@
 	group-regularized ridge regression and using the estimated penalty multipliers
-	as starting values for \texttt{gren}. In addition, we note that even a local
-	optimum might give `good enough' predictions in many practical settings.
+	as starting values for \texttt{gren}. Alternatively, a skeptical user may set 
+	the starting values for all penalty multipliers to one, thereby starting from
+	the regular elastic net and adaptively adjusting the penalty parameters to a
+	group-regularized model.
+
+<<>>=
+bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench2.csv", 
+                     stringsAsFactors=FALSE)
+bench2 <- read.table("results/rnaseq_lymph_node_metastasis_bench1.csv", 
+                     stringsAsFactors=FALSE)
+times1 <- round(bench1$time*10^(-9), 2)
+times2 <- round(bench2$time*10^(-9), 2)
+session <- gsub("_", ".", devtools::session_info()$platform$system)
+@
+  We have investigated the computation time of the algorithm and compared to the
+  other methods introduced in Sections \ref{sec:colorectal}--\ref{sec:lymphnode}
+  on the respective datasets. All timings were obtained on a Macbook Pro 2016 
+  running \Sexpr{session}. 
+  The timings of \texttt{gren} are competitive. On the colorectal cancer example
+  we achieved the following times:
+  \Sexpr{times1[c(1:3)]} seconds on the colorectal cancer example, 
+  with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively. \texttt{GRridge} is 
+  slightly faster with \Sexpr{times1[4]} seconds, while \texttt{SGL} was slightly
+  slower: \Sexpr{times1[c(5:7)]} seconds, with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
+  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
+  On the lymph node cancer example \texttt{gren} achieved 
+  \Sexpr{times2[c(1:3)]}, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
+  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
+  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+  \texttt{cMCP} (\Sexpr{times2[c(8:10)]} seconds) and \texttt{gel} 
+  (\Sexpr{times2[c(11:13)]} seconds) are faster than \texttt{gren}. The 
+  overlapping group methods achieved \Sexpr{times2[c(14:16)]} seconds
+  (\texttt{ocMCP}) and \Sexpr{times2[c(17:19)]} seconds (\texttt{ogel}), for 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
+  We note that a considerable speedup may be achieved by 
+  doing one EB update for every few VB updates, instead of running the VB EM 
+  until convergence. This works in practice, because an EM tends to take a few 
+  initial large steps towards the optimum, followed by a series of smaller 
+  steps. We argue that The first few iterations of the VB EM render the 
+  approximate expectation in [! equation ref here] accurate enough for a 
+  reliable EB step. In addition, this is the approach often taken by more
+  computationally oriented disciplines. For example, in many machine learning 
+  domains, the problem is viewed as purely algorithmic [! references here]. The
+  the EB and VB 
+  steps optimisation steps are interpreted as steps along the gradient in the 
+  larger combined space of model- and hyper-parameters.
 
@@
 	A (stable) \texttt{R} package is available from 
-	\url{https://CRAN.R-project.org/package=gren}. All materials needed to   
-	recreate this document are available from 
-	\url{https://github.com/magnusmunch/gren}.
+	\url{https://CRAN.R-project.org/package=gren}.
 
@@
 	Supplementary Material is available online from 
-	\url{http://pub.math.leidenuniv.nl/~munchmm/}.
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All materials needed to   
+	recreate this document are available from 
+	\url{https://github.com/magnusmunch/gren}.
 

### 2019-03-06 code/figures.R

@@ -34,7 +34,8 @@ plot.data2 <- lapply(methods2, function(m) {
 names(plot.data2) <- methods2
 
 opar <- par(no.readonly=TRUE)
-par(mfrow=c(1, 2), mar=opar$mar*c(1, 1.3, 1, 0.3), cex=0.70)
+par(mar=opar$mar*c(1, 1.3, 1, 1))
+layout(matrix(rep(c(1, 1, 2, 2), 2), 2, 4, byrow=TRUE))
 barplot(plot.data1[[1]], beside=TRUE, col=col1,
         legend.text=methods1,
         args.legend=list(x="bottomleft", fill=c(col1, NA),
@@ -46,7 +47,7 @@ barplot(plot.data1[[1]], beside=TRUE, col=col1,
 abline(h=1, lty=2)
 
 plot(plot.data2[[3]], type="l", xlab="Number of selected features", ylab="AUC", 
-     main="(b)", ylim=range(auc), xlim=c(0, 500), col=col2[3], 
+     main="(b)", ylim=range(auc), xlim=c(0, 600), col=col2[3], 
      lty=lty2[1])
 lines(plot.data2[[4]], col=col2[4], lty=lty2[1])
 lines(plot.data2[[5]], col=col2[5], lty=lty2[1])
@@ -91,7 +92,7 @@ labels1 <- list(corr=expression(tau > 0.54,
                                             textstyle(p <= 2.1e-02))), 
                               p > 2.1e-02))
 labels2 <- expression("ridge", "GRridge", "gren,"~alpha==0.05, 
-                      "enet"~alpha==0.05, "OcMCP,"~alpha==0.05)
+                      "enet"~alpha==0.05, "cMCP,"~alpha==0.05)
 
 plot.data1 <- lapply(1:length(fit1.gren1$lambdag), function(s) {
   rbind(fit1.grridge$lambdamults[[s]], fit1.gren1$lambdag[[s]],
@@ -128,6 +129,53 @@ abline(h=plot.data2[[1]][, 2], col=col2[1], lty=lty2[2])
 abline(h=plot.data2[[2]][, 2], col=col2[2], lty=lty2[2])
 legend("bottomright", legend=labels2, col=col2, 
        lty=c(rep(lty2[2], 2), rep(lty2[1], 4)), bg = "white")
+par(opar)
+
+# ---- lines_colorectal_lymph_node_gren_auc ----
+library(sp)
[…]

### 2019-03-06 code/rnaseq_lymph_node_metastasis.R

@@ -61,54 +61,61 @@ part.ogl <- unlist(lapply(list(corr=split(1:p, part$corr),
 n <- nrow(xtrain)
 p <- ncol(xtrain)
 
-fit1.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE)
-fit1.gren2 <- gren(xtrain, ytrain, partitions=part, alpha=0.5, trace=FALSE)
-fit1.gren3 <- gren(xtrain, ytrain, partitions=part, alpha=0.95, trace=FALSE)
+bench1 <- microbenchmark(
+fit1.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE),
+fit1.gren2 <- gren(xtrain, ytrain, partitions=part, alpha=0.5, trace=FALSE),
+fit1.gren3 <- gren(xtrain, ytrain, partitions=part, alpha=0.95, trace=FALSE),
 
 fit1.grridge <- grridge(t(xtrain), ytrain, list(corr=split(1:p, part$corr), 
-                                                pv=split(1:p, part$pv)))
+                                                pv=split(1:p, part$pv))), {
 
 fit1.sgl1 <- cvSGL(list(x=xtrain, y=ytrain), as.numeric(as.factor(paste(
   corpart0, pvpart0))), type="logit", alpha=0.05)
-fit1.sgl1$fit$type <- "logit"
+fit1.sgl1$fit$type <- "logit"}, {
 fit1.sgl2 <- cvSGL(list(x=xtrain, y=ytrain), as.numeric(as.factor(paste(
   corpart0, pvpart0))), type="logit", alpha=0.5)
-fit1.sgl2$fit$type <- "logit"
+fit1.sgl2$fit$type <- "logit"}, {
 fit1.sgl3 <- cvSGL(list(x=xtrain, y=ytrain), as.numeric(as.factor(paste(
   corpart0, pvpart0))), type="logit", alpha=0.95)
-fit1.sgl3$fit$type <- "logit"
+fit1.sgl3$fit$type <- "logit"},
 
 fit1.cmcp1 <- cv.grpreg(xtrain, ytrain, as.numeric(as.factor(paste(
-  corpart0, pvpart0))), penalty="cMCP", family="binomial", alpha=0.05)
+  corpart0, pvpart0))), penalty="cMCP", family="binomial", alpha=0.05),
[…]

### 2019-03-06 code/simulations.R

@@ -1,0 +1,22 @@
+#!/usr/bin/env Rscript
+
+### installation of gren if updated on github
+if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
+   substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
+  if(!("devtools" %in% installed.packages())) {
+    install.packages("devtools")
+  }
+  library(devtools)
+  install_github("magnusmunch/gren/rpackage", local=FALSE,
+                 auth_token=Sys.getenv("GITHUB_PAT"))
+}
+
+### libraries
+library(gren)
+library(GRridge)
+library(grpreg)
+library(SGL)
+library(pROC)
+library(foreach)
+
+################################# simulation 1 #################################

### 2019-03-06 docs/manuscript.Rnw
added briers to rnaseq_lymph_node_metastasis.R and micrornaseq_colorectal_cancer.R

@@
   penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
-  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. More 
+  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
   performance measures are presented in Section [! section ref here]
@@
   group, while the non-expressed group receives the strongest penalty 
-  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}a). 
+  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
   \texttt{gren} seems to outperform the other methods for a range of model 
   sizes in terms of AUC 
-  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}b). 
+  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
   This is especially pronounced for the larger models. The performance of
@@
   albeit it only slightly for \texttt{GRridge}.
-<<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
+<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
 @
@@
 <<>>=
-bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench2.csv", 
+bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench1.csv", 
                      stringsAsFactors=FALSE)

### 2019-03-06 docs/supplement.Rnw
added briers to rnaseq_lymph_node_metastasis.R and micrornaseq_colorectal_cancer.R

@@
 rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
-bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
+bbm,algorithm,algpseudocode,threeparttable,booktabs,mathtools,
 dsfont,parskip}
@@
 \newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
-\newcommand{\appropto}{\mathrel{\vcenter{
-  \offinterlineskip\halign{\hfil$##$\cr
-    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
-\newcommand{\card}[1]{\text{card} \left( #1 \right)}
-\makeatletter
-\newcommand*{\defeq}{\mathrel{\rlap{%
-			\raisebox{0.3ex}{$\m@th\cdot$}}%
-		\raisebox{-0.3ex}{$\m@th\cdot$}}%
-	=}
-\makeatother
 
@@
 \setlength{\textwidth}{15cm}
-\setlength{\parindent}{0pt}
 \title{Adaptive group-regularized logistic elastic net regression Supplementary 
 material}
-\date{}
+\date{\today}
 \author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
 \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
-Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}
+Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}
 
@@
   \maketitle
+
   \noindent
 	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
-  research institute, VU University Medical Center, PO Box 7057, 1007 MB
-  Amsterdam, The Netherlands\\
-  2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
-  3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\
+	research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
+	Amsterdam, The Netherlands \\*
+	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
+	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom
 
@@
   apply our method to a study on microRNAs in colorectal cancer data 
-  \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} . Finally, we
-  introduce the ridge and lasso versions of the proposed method in Section
+  \cite[]{neerincx_combination_2018} in Section [! section ref here]. Finally, 
+  we introduce the ridge and lasso versions of the proposed method in Section
   \ref{sec:evbridgelasso}.
@@
   \begin{align*}
-  \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[
-  -\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
+    \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot 
+    \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
     \V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot
@@
   Generating samples from the elastic net prior is not straightforward. 
-\cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$:
+  \cite{li_bayesian_2010} show that the elastic net prior may be written as a 
+  scale mixture of normals, with mixing parameter 
+  $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$: 
   \begin{align*}
-\beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} \frac{\tau - 1}{\tau} \right), \\
-\tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right),
+    \beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} 
+    \frac{\tau - 1}{\tau} \right), \\
+    \tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2},
+    \left(1,\infty \right) \right),
   \end{align*}
-where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$.
-So with an efficient truncated gamma sampler available we may sample from the prior using any standard software package that includes a Gaussian sampling routine.
-More specifically, we require a a sample from the $\mathcal{TG} \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ distribution.
-The CDF of this distribution is given by:
+  where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
+  distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with
+  an efficient truncated gamma sampler available we may sample from the prior 
+  using any standard software package that includes a Gaussian sampling routine.
+  More specifically, we require a a sample from the $\mathcal{TG} 
+  \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ 
+  distribution. The CDF of this distribution is given by:
   $$
@@
     0, & \text{if }\tau < 1 \\
-1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}{\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
+    1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}
+    {\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
   \end{dcases}
@@
 
-Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for $\tau \geq 1$, the quantile function may be written as a function of the probability $p \in (0,1)$:
+  Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for 
+  $\tau \geq 1$, the quantile function may be written as a function of the 
+  probability $p \in (0,1)$:
   $$
-Q_{T}(p) =	F^{-1}_{T}(p) = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - p) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2,
+  Q_{T}(p) =	F^{-1}_{T}(p) = \frac{4 \lambda_2}{\lambda_1^2} 
+  \Phi^{-1} \[ (1 - p) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2,
   $$
-where $\Phi^{-1}(p)$ is the quantile function of the standard normal distribution, i.e, the probit function.
-Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now:
+  where $\Phi^{-1}(p)$ is the quantile function of the standard normal 
+  distribution, i.e, the probit function. Employing inverse transform sampling 
+  for the truncated gamma distribution, an elastic net prior sampling scheme is 
+  now:
   \begin{algorithmic}
 	  \State{Generate $U \sim \mathcal{U} (0,1)$}
-	\State{Set $\tau = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - u) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2$}
+	  \State{Set $\tau = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - u) 
+	    \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2$}
 	  \State{Set $\nu^2= \frac{\tau - 1}{\tau \lambda_2}$}
@@
   \section{Objective function}\label{sec:objectivefunction}
-The lower bound on the marginal likelihood, or evidence lower bound (ELBO), at iteration $t$ is given by:
+  The lower bound on the marginal likelihood, or evidence lower bound (ELBO), 
+  at iteration $t$ is given by:
   \begin{subequations}
     \begin{align}
-	\text{ELBO} (Q^{(t)}) & = \E_{Q^{(t)}} [\log \L (\y, \bomega, \bbeta, \btau)] - \E_{Q^{(t)}} [\log Q (\bomega, \bbeta, \btau)] \\
-	& = \E_{Q^{(t)}} [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \label{eq:elboomega} \\
-	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\bbeta|\btau)] - \E_{Q^{(t)}} [\log q(\bbeta)] \label{eq:elbobeta}\\
-	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\btau)] - \E_{Q^{(t)}} [\log q(\btau)] \label{eq:elbotau}.
+      \text{ELBO} (Q^{(t)}) & = \E_{Q^{(t)}} [\log \L (\y, \bomega, \bbeta, 
+      \btau)] - \E_{Q^{(t)}} [\log Q (\bomega, \bbeta, \btau)] \\
+      & = \E_{Q^{(t)}} [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} 
+      [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} 
+      [\log q(\bomega)] \label{eq:elboomega} \\
+      & \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\bbeta|\btau)] - \E_{Q^{(t)}} 
+      [\log q(\bbeta)] \label{eq:elbobeta}\\
+      & \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\btau)] - \E_{Q^{(t)}} 
+      [\log q(\btau)] \label{eq:elbotau}.
     \end{align}
@@
   $$
-h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}\frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}\exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
+  h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}
+  \frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}
+  \exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
   $$
-Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$.
-Furthermore, in the following, all expectations and variances are with respect to $Q^{(t)}$.
-Next we write out the three terms in (\ref{eq:elboomega}):
+  Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, 
+  in the following, all expectations and variances are with respect to 
+  $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):
   \begin{subequations} \label{eq:elboomega2}
     \begin{align}
-	\E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\}, \\
-	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
-	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] \E [\omega_i] \\
-	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
-	\E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n c_i^2 \E [\omega_i] \\
+      \E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + 
+      \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)]
+      \right\}, \\
+      \E [\log \pi(\bomega | \bbeta)] & = \m \tr \E 
+      \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - 
+      \frac{1}{2} \m \tr \X \E [\bbeta] \\
+      & \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n 
+      [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] 
+      \E [\omega_i] \\
+      & \,\,\,\,\,+ \sum_{i=1}^n \E 
+      \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
+      \E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - 
+      \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} 
+      \sum_{i=1}^n c_i^2 \E [\omega_i] \\
       & \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
@@
   where all mathematical operations on vectors and matrices are element wise. 
-After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
+  After combining the terms in (\ref{eq:elboomega2}) and substituting the 
+  expectations and variances given in the Main Document, several terms cancel to
+  give:
   \begin{subequations} \label{eq:elboomega3}
 	  \begin{align}
-	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
-	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
+	    \E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} 
+	    [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ 
+	    & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) 
+	    \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - 
+	    \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
+	    & \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} 
+	    \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - 
+	    ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
     \end{align}
@@
 
-Inspection of the updating equations in the Main Document, learns us that the last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that we are left with just (\ref{eq:elboomega4}).
-After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
+  Inspection of the updating equations in the Main Document, learns us that the 
+  last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
+  we are left with just (\ref{eq:elboomega4}). After a change of variables 
+  $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
   \begin{subequations} \label{eq:elbobeta2}
 	  \begin{align}
-	\E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - \frac{p}{2} \log (2 \pi) \\
-	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) (\bmu_j^2 + \bSigma_{jj}) \\
-	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
-	\E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - \frac{p}{2} \log |\bSigma| .
+	    \E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G 
+	    |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - 
+	    \frac{p}{2} \log (2 \pi) \\
+	    & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in 
+	    \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) 
+	    (\bmu_j^2 + \bSigma_{jj}) \\
+	    & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} 
+	    \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
+	    \E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - 
+	    \frac{p}{2} \log |\bSigma| .
 	  \end{align}
   \end{subequations}
-We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):
+  We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for 
+  (\ref{eq:elbobeta}):
   \begin{subequations} \label{eq:elbobeta3}
 	   \begin{align}
-	\E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} [\log q (\bbeta )] \\ 
-	& = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + \frac{p}{2} \\
-	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
-	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
+	     \E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} 
+	     [\log q (\bbeta )] \\ 
+	     & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + 
+	     \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + 
+	     \frac{p}{2} \\
+	     & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g 
+	     \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 
+	     \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
+	     & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + 
+	     \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
 	  \end{align}
@@
 We describe our implementation here for two partitions of the features.
-To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively.
+To this end, let $(\G_1(1), \dots,\allowbreak \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively.
 Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively.
@@
   \section{Application to MicroRNAs in colorectal cancer}\label{sec:colorectal}
-  In this section we present additional results to Section [! section ref here]
-  in the Main Document.
-
   \subsection{Additional comparisons}
+  We compared \texttt{gren} to two methods in addition to the ones in the MD: 
+  the sparse group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and 
+  the group exponential lasso (\texttt{gel}) by \cite{breheny_group_2015}. For
+  all methods with a tuning parameter $\alpha$, we set 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ and present the AUC on the test set in 
+  Figure \ref{fig:lines_micrornaseq_colorectal_cancer_res1_auc}.
+<<lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated AUC for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
+@
 
@@
 <<>>=
-res <- read.table("results/micrornaseq_colorectal_cancer_res3.csv", 
+res <- read.table("results/micrornaseq_colorectal_cancer_res2.csv", 
                   stringsAsFactors=FALSE)
@@
   repeated the procedure 100 times and present the results in Figure 
-  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers}. 
+  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res2_multipliers}. 
   From this figure we see that the estimates for \texttt{gren} are relatively 
@@
   $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
-<<boxplots_micrornaseq_colorectal_cancer_res3_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 3 random groups of features.", out.width="50%", fig.asp=1>>=
+<<boxplots_micrornaseq_colorectal_cancer_res2_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 3 random groups of features.", out.width="50%", fig.asp=1>>=
 @
@@
 <<>>=
-res <- read.table("results/micrornaseq_colorectal_cancer_res4.csv", 
+res <- read.table("results/micrornaseq_colorectal_cancer_res3.csv", 
                   stringsAsFactors=FALSE)
@@
   We repeated the simulation with ten evenly sized groups. Figure 
-  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res4_multipliers} presents 
+  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers} presents 
   the estimated penalty multipliers for 100 random 
@@
   $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
-<<boxplots_micrornaseq_colorectal_cancer_res4_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
+<<boxplots_micrornaseq_colorectal_cancer_res3_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
 @
@@
   25 features (including the 5 unpenalized covariates). We calculated the size 
-  of all ${50}\choose{2}$ intersections of selected features and present the 
+  of all $\genfrac(){0pt}{2}{50}{2}$ intersections of selected features and present the 
   results in Figure [! figure ref here].
@@
   \section{Application to RNAseq in lymph node cancer}
-  \texttt{cMCP} is not designed to deal with overlapping groups, so we 
-  cross-tabulate the two co-data partitions to create one paritioning with 25 
-  groups for \texttt{cMCP}.  
-  Additional comparisons to \texttt{SGL}, \texttt{gel} and 
-  the overlapping group version of \texttt{gel} \cite[]{zeng_overlapping_2016}
-  \cite[]{breheny_group_2015} are found in the SM Section [! section ref here].
+  In addition to the results presented in Section [! section ref here] of the 
+  MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 
+  \cite{simon_sparse-group_2013} and the group exponential lasso (\texttt{gel}) 
+  by \cite{breheny_group_2015}. The original group lasso methods are not 
+  designed to deal with overlapping groups, so in addition to using the versions
+  that allow for overlapping groups, we 
+  cross-tabulated the two co-data partitions to create one paritioning with 25 
+  groups for \texttt{cMCP} and \texttt{gel}.
+  We use $\alpha \in \{0.05, 0.5, 0.95 \}$ for the
+  methods that require this extra tuning parameter. The AUCs calculated on the
+  validation set are presented in Figure 
+  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_auc}
+<<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated AUC for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
+@
 
@@
 
-  \section{Choice of $\alpha$}	
+  \section{Choice of \texorpdfstring{$\alpha$}{alpha}}	
 	Choice of $\alpha$ does matter in Figure 

### 2019-03-06 code/figures.R

@@ -1,8 +1,8 @@
 ################################## manuscript ##################################
-# ---- barplot_lines_micrornaseq_colorectal_cancer_res2_auc ----
+# ---- barplot_lines_micrornaseq_colorectal_cancer_res1_auc ----
 library(sp)
-load("results/micrornaseq_colorectal_cancer_fit2.Rdata")
-res <- read.table("results/micrornaseq_colorectal_cancer_res2.csv", 
+load("results/micrornaseq_colorectal_cancer_fit1.Rdata")
+res <- read.table("results/micrornaseq_colorectal_cancer_res1.csv", 
                   stringsAsFactors=FALSE)
 pred <- as.matrix(res[substr(rownames(res), 1, 4)=="pred", ])
 psel <- as.matrix(res[substr(rownames(res), 1, 4)=="psel", ])
@@ -16,18 +16,17 @@ col1 <- grey.colors(length(methods1) - 1, start=0.3, end=0.9, gamma=2.2,
 col2 <- bpy.colors(length(methods2), cutoff.tail=0.1)
 lty1 <- 2
 lty2 <- c(1, 2)
-labels1 <- list(diff.threegroup=expression("FDR">0.05,
-                                           atop(NA, atop(textstyle(
-                                             0.05>=phantom(0)),
-                                             textstyle("FDR" > 0.001))),
-                                           "FDR"<=0.001))
+labels1 <- list(diff.expr=expression("FDR">0.05,
+                                     atop(NA, atop(textstyle(0.05>=phantom(0)),
+                                                   textstyle("FDR" > 0.001))),
+                                     "FDR"<=0.001))
 labels2 <- expression("ridge", "GRridge", "gren,"~alpha==0.05, 
                       "gren,"~alpha==0.5, "enet"~alpha==0.05, 
                       "cMCP"~alpha==0.05)
 
-plot.data1 <- lapply(1:length(fit2.gren1$lambdag), function(s) {
-  rbind(fit2.grridge$lambdamults[[s]], fit2.gren1$lambdag[[s]],
-        fit2.gren2$lambdag[[s]], fit2.gren3$lambdag[[s]])})
[…]

### 2019-03-06 code/micrornaseq_colorectal_cancer.R

@@ -12,12 +12,11 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### libraries
+library(Biobase)
 library(gren)
 library(GRridge)
 library(grpreg)
-library(Biobase)
 library(SGL)
-library(grpregOverlap)
 library(foreach)
 library(doParallel)
 library(microbenchmark)
@@ -28,16 +27,12 @@ load("data/forMagnusN88.Rdata")
 ### create model matrix for unpenalized covariates
 unpenal <- model.matrix(~ adjth + thscheme + age + pcrcdiff, data=datfr)[, -1]
 
-### create partitioning based on FDR <= 0.05
-diff.twogroup <- rep(c(2, 1), times=unlist(lapply(partkeep$TS, length)))[
-  order(unlist(partkeep$TS))]
-
 ### create partitioning based on FDR <= 0.05 and FDR <= 0.001
 miRNA.BFDR <- as.character(TumMirs$miRNA)[TumMirs$BFDR_PNminP < 0.001]
 miRNA.TumMirs <- as.character(TumMirs$miRNA)
 miRNA <- as.character(sapply(rownames(mirnormcen_resp), function(s) {
   strsplit(s, split=" ")[[1]][1]}))
-diff.threegroup <- miRNA %in% miRNA.BFDR + miRNA %in% miRNA.TumMirs + 1
+diff.expr <- miRNA %in% miRNA.BFDR + miRNA %in% miRNA.TumMirs + 1
 
 ### target vector
 benefit <- as.numeric(resp) - 1
@@ -52,7 +47,7 @@ id.train <- c(sample(which(benefit==0), size=floor(sum(benefit==0)*0.7)),
               sample(which(benefit==1), size=floor(sum(benefit==1)*0.7)))
 
 ################################### model 1 ####################################
-### fitting the models
+### preparing data
 set.seed(2019)
 ytrain <- benefit[id.train]
 # we remove the constant micrornas
@@ -63,270 +58,144 @@ xtest <- scale(micrornas[-id.train, ])[, apply(micrornas[id.train, ], 2, sd)!=0]
 # set the constant micrornas to 0
 xtest[is.nan(xtest)] <- 0
 utest <- unpenal[-id.train, ]
-part <- diff.twogroup[apply(micrornas[id.train, ], 2, sd)!=0]
[…]

### 2019-03-06 code/rnaseq_lymph_node_metastasis.R

@@ -13,12 +13,11 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 
 ### libraries
 library(CoRF)
-library(GRridge)
 library(gren)
+library(GRridge)
 library(grpregOverlap)
 library(SGL)
-library(pROC)
-
+library(microbenchmark)
 
 ### load data
 data(LNM_Example)
@@ -32,7 +31,6 @@ Tr <- TrainData[, selectgenes]
 n <- nrow(Tr)
 
 grridge2gren <- function(partit){
-  #partit <- TS
   vec <- array(dim=length(unlist(partit)))
   for(i in 1:length(partit)){
     vec[partit[[i]]] <- i
@@ -42,8 +40,6 @@ grridge2gren <- function(partit){
 pvpart0 <- grridge2gren(CreatePartition(CDT$pvalsVUmc, ngroup=5,
                                         decreasing=FALSE))
 corpart0 <- grridge2gren(CreatePartition(CDT$Corrs, ngroup=5, decreasing=TRUE))
-#sds
-#roepmanpart <- CDT$RoepmanGenes+1
 partLNM <- list(corr=corpart0, pv=pvpart0)
 
 ################################### model 1 ####################################
@@ -61,114 +57,116 @@ part.ogl <- unlist(lapply(list(corr=split(1:p, part$corr),
 n <- nrow(xtrain)
 p <- ncol(xtrain)
 
-bench1 <- microbenchmark(
-fit1.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE),
-fit1.gren2 <- gren(xtrain, ytrain, partitions=part, alpha=0.5, trace=FALSE),
-fit1.gren3 <- gren(xtrain, ytrain, partitions=part, alpha=0.95, trace=FALSE),
+bench <- microbenchmark(
+fit.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE),
+fit.gren2 <- gren(xtrain, ytrain, partitions=part, alpha=0.5, trace=FALSE),
+fit.gren3 <- gren(xtrain, ytrain, partitions=part, alpha=0.95, trace=FALSE),
 
-fit1.grridge <- grridge(t(xtrain), ytrain, list(corr=split(1:p, part$corr), 
+fit.grridge <- grridge(t(xtrain), ytrain, list(corr=split(1:p, part$corr), 
[…]

### 2019-03-06 code/simulations.R

@@ -43,13 +43,13 @@ beta <- numeric(p)
 
 part <- rep(c(1:G), each=p/G)
 nreps <- 1
+ncores <- min(detectCores() - 1, nreps)
 
 ### analysis splits in parallel
-ncores <- min(detectCores() - 1, nreps)
 cluster <- makeCluster(ncores, type="FORK")
 registerDoParallel(cluster)
 
-out <- foreach(k=c(1:nreps)) %do% {
+res1 <- foreach(k=c(1:nreps)) %dopar% {
   set.seed(2019 + k)
   
   for(g in 1:G) {

### 2019-03-07 code/micrornaseq_cervical_cancer.R
included briers in micrornaseq_cervical_cancer.R

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 
-### installation of packages
+### installation of gren if updated on github
 if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
    substr(packageDescription("gren")$GithubSHA1, 1, 7)) {
   if(!("devtools" %in% installed.packages())) {
@@ -18,7 +18,7 @@ library(grpreg)
 library(SGL)
 
 ### load data
-load("old_data/mirsData.RData")
+load("data/mirsData.RData")
 
 ### filter out the features with less than 2 unique counts
 unique.counts <- apply(t(as.matrix(mirsData$countData)), 2, function(s) {
@@ -45,18 +45,18 @@ part <- conservation
 p <- ncol(x)
 n <- nrow(x)
 
-psel <- c(seq(1, 5, 1), seq(7, 15, 2), seq(18, 30, 3), seq(34, 50, 4))
+csel <- c(seq(1, 5, 1), seq(7, 15, 2), seq(18, 30, 3), seq(34, 50, 4))
 
-fit1.gren1 <- gren(x, y, partitions=list(part=part), alpha=0.05, 
-                   standardize=TRUE, trace=FALSE, psel=psel)
-fit1.gren2 <- gren(x, y, partitions=list(part=part), alpha=0.5, 
-                   standardize=TRUE, trace=FALSE, psel=psel)
-fit1.gren3 <- gren(x, y, partitions=list(part=part), alpha=0.95, 
-                   standardize=TRUE, trace=FALSE, psel=psel)
+fit.gren1 <- gren(x, y, partitions=list(part=part), alpha=0.05, 
+                  standardize=TRUE, trace=FALSE, psel=csel)
+fit.gren2 <- gren(x, y, partitions=list(part=part), alpha=0.5, 
+                  standardize=TRUE, trace=FALSE, psel=csel)
+fit.gren3 <- gren(x, y, partitions=list(part=part), alpha=0.95, 
+                  standardize=TRUE, trace=FALSE, psel=csel)
 
-fit1.grridge <- grridge(t(x), y, list(part=split(1:p, part)))
+fit.grridge <- grridge(t(x), y, list(part=split(1:p, part)))
 
-save(fit1.grridge, fit1.gren1, fit1.gren2, fit1.gren3, 
+save(fit.grridge, fit.gren1, fit.gren2, fit.gren3, 
[…]

### 2019-03-07 docs/supplement.Rnw
included briers in micrornaseq_cervical_cancer.R

@@
   all methods with a tuning parameter $\alpha$, we set 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ and present the AUC on the test set in 
-  Figure \ref{fig:lines_micrornaseq_colorectal_cancer_res1_auc}.
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ and present the AUC and Brier skill score 
+  on the test set in Figures 
+  \ref{fig:lines_micrornaseq_colorectal_cancer_res1_auc}--
+  \ref{fig:lines_micrornaseq_colorectal_cancer_res1_briers}.
 <<lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated AUC for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_micrornaseq_colorectal_cancer_res1_briers, fig.cap="Estimated Brier skill score for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
 @

### 2019-03-07 code/figures.R

@@ -131,6 +131,110 @@ legend("bottomright", legend=labels2, col=col2,
 par(opar)
 
 ################################## supplement ##################################
+# ---- lines_micrornaseq_colorectal_cancer_res1_auc ----
+library(sp)
+res <- read.table("results/micrornaseq_colorectal_cancer_res1.csv", 
+                  stringsAsFactors=FALSE)
+
+pred <- as.matrix(res[substr(rownames(res), 1, 4)=="pred", ])
+psel <- as.matrix(res[substr(rownames(res), 1, 4)=="psel", ])
+auc <- as.matrix(res[substr(rownames(res), 1, 3)=="auc", ])
+
+methods <- c("ridge", "grridge", paste0("gren", c(1:3)), paste0("enet", c(1:3)),
+             paste0("sgl", c(1:3)), paste0("cmcp", c(1:3)), 
+             paste0("gel", c(1:3)))
+col <- bpy.colors(length(methods), cutoff.tail=0.1)
+lty <- 1:length(methods)
+labels <- expression("ridge", "GRridge", "gren,"~alpha==0.05, 
+                     "gren,"~alpha==0.5, "gren,"~alpha==0.95, 
+                     "enet,"~alpha==0.05, "enet,"~alpha==0.5, 
+                     "enet,"~alpha==0.95, "SGL,"~alpha==0.05, "SGL,"~alpha==0.5, 
+                     "SGL,"~alpha==0.95, "cMCP,"~alpha==0.05, 
+                     "cMCP,"~alpha==0.5, "cMCP,"~alpha==0.95, 
+                     "gel,"~alpha==0.05, "gel,"~alpha==0.5, "gel,"~alpha==0.95)
+
+plot.data <- lapply(methods, function(m) {
+  aggregate(auc[, grepl(m, colnames(auc))], 
+            list(psel=psel[, grepl(m, colnames(psel))]), mean)})
+names(plot.data) <- methods
+
+opar <- par(no.readonly=TRUE)
+par(mar=opar$mar*c(1, 1.3, 1, 1))
+plot(plot.data[[3]], type="l", xlab="Number of selected features", ylab="AUC", 
+     main="", ylim=range(auc), xlim=c(0, 500), col=col[3], 
+     lty=lty[3])
+lines(plot.data[[4]], col=col[4], lty=lty[4])
[…]

### 2019-03-07 code/figures.R
all barplots, auc, and briers figures in supplement

@@ -297,8 +297,7 @@ boxplot(multipliers ~ method, data=plot.data, names=labels,
 abline(h=1, lty=2, lwd=1.5)
 par(opar)
 
-# ---- hist_micrornaseq_colorectal_cancer_res5_overlap ----
-
+# ---- hist_micrornaseq_colorectal_cancer_res4_overlap ----
 
 # ---- lines_rnaseq_lymph_node_metastasis_res1_auc ----
 library(sp)
@@ -389,14 +388,15 @@ labels <- expression("ridge", "GRridge", "gren,"~alpha==0.05,
                      "Ogel,"~alpha==0.5, "Ogel,"~alpha==0.95)
 
 plot.data <- lapply(methods, function(m) {
-  aggregate(auc[, grepl(m, colnames(auc))], 
+  aggregate(briers[, grepl(m, colnames(briers))], 
             list(psel=psel[, grepl(m, colnames(psel))]), mean)})
 names(plot.data) <- methods
 
 opar <- par(no.readonly=TRUE)
 par(mar=opar$mar*c(1, 1.3, 1, 1))
-plot(plot.data[[3]], type="l", xlab="Number of selected features", ylab="AUC", 
-     main="", ylim=range(auc), xlim=c(0, 500), col=col[3], 
+plot(plot.data[[3]], type="l", xlab="Number of selected features", 
+     ylab="Brier skill score", 
+     main="", ylim=range(briers), xlim=c(0, 500), col=col[3], 
      lty=lty[3])
 lines(plot.data[[4]], col=col[4], lty=lty[4])
 lines(plot.data[[5]], col=col[5], lty=lty[5])
@@ -424,649 +424,15 @@ legend("bottomright", legend=labels, col=col,
        lty=lty, bg = "white")
 par(opar)
 
-
-
-
-# ---- lines_colorectal_lymph_node_gren_auc ----
-library(sp)
-load("results/rnaseq_lymph_node_metastasis_fit1.Rdata")
-res1 <- read.table("results/rnaseq_lymph_node_metastasis_res1.csv", 
-                   stringsAsFactors=FALSE)
-res2 <- read.table("results/micrornaseq_colorectal_cancer_res2.csv", 
-                   stringsAsFactors=FALSE)
-pred1 <- as.matrix(res1[substr(rownames(res1), 1, 4)=="pred", ])
[…]

### 2019-03-07 docs/supplement.Rnw
all barplots, auc, and briers figures in supplement

@@
 	  \begin{align}
-	\E [\log \pi (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{5 p }{2} \log 2 - \frac{p}{2} \log \pi - \frac{p}{2} - \frac{p \lambda_1^2}{8 \lambda_2} \\
-	& \,\,\,\,\, - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) - \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} - \frac{1}{2} \sum_{j=1}^p \E [\log ( \psi_j + 1)], \\
-	\E [\log q (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{3p}{2} \log 2 -\frac{p}{2} \log \pi -\frac{p}{2} - \frac{\lambda_1}{2 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} \\
-	& \,\,\,\,\, + \frac{1}{4} \sum_{j=1}^p \log \chi_j - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j].
+	    \E [\log \pi (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - 
+	    \frac{5 p }{2} \log 2 - \frac{p}{2} \log \pi - \frac{p}{2} - 
+	    \frac{p \lambda_1^2}{8 \lambda_2} \\
+	    & \,\,\,\,\, - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) -
+	    \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} - 
+	    \frac{1}{2} \sum_{j=1}^p \E [\log ( \psi_j + 1)], \\
+	    \E [\log q (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 -
+	    \frac{3p}{2} \log 2 -\frac{p}{2} \log \pi -\frac{p}{2} - 
+	    \frac{\lambda_1}{2 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} \\
+	    & \,\,\,\,\, + \frac{1}{4} \sum_{j=1}^p \log \chi_j - 
+	    \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j].
 	  \end{align}
@@
 	    \E_{Q^{(t)}} & [\log \pi (\btau)] - \E_{Q^{(t)}} [\log q (\btau)] 
-	\\ & = p \log 2 - \frac{p \lambda_1^2}{8 \lambda_2} - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)} \\
-	& \,\,\,\,\, + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log ( \psi_j + 1)].
+	    \\ & = p \log 2 - \frac{p \lambda_1^2}{8 \lambda_2} - 
+	    p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) + 
+	    \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} -
+	    \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)} \\
+	    & \,\,\,\,\, + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] -
+	    \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log ( \psi_j + 1)].
 	  \end{align}
@@
 
-Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
+  Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and
+  (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
   \begin{align*}
-\text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2} \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
-& + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
-& + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
+    \text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + 
+    \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2}
+    \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
+    & + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} 
+    \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} 
+    \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) 
+    \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
+    & + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} -
+    \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
   \end{align*}
-Here, proportionality is with respect to the variational and penalty parameters.
+  Here, proportionality is with respect to the variational and penalty 
+  parameters.
 
   \section{The P\'{o}lya-Gamma parametrisation}\label{sec:polyagamma}
-In Section 3 of the Main Document latent P\'{o}lya-Gamma distributed variables are introduced.
-Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma variables are introduced into the logistic regression model under a Gaussian prior.
-Here, we explain how introduction of these latent variables simplifies the variational Bayes calculations.
+  In Section 3 of the Main Document latent P\'{o}lya-Gamma distributed variables
+  are introduced. Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma
+  variables are introduced into the logistic regression model under a Gaussian
+  prior. Here, we explain how introduction of these latent variables simplifies
+  the variational Bayes calculations.
 
-In \cite{polson_bayesian_2013}, the central identity for this parametrisation is given in their Theorem 1:
+  In \cite{polson_bayesian_2013}, the central identity for this parametrisation
+  is given in their Theorem 1:
   \begin{equation}\label{eq:intid}
-\frac{\exp(\psi)^{a}}{[1 + \exp(\psi)]^{b}} = 2^{-b} \exp(\kappa \psi) \int_0^{\infty} \exp(-\omega \psi^2 /2) \pi(\omega) d \omega,
+    \frac{\exp(\psi)^{a}}{[1 + \exp(\psi)]^{b}} = 2^{-b} \exp(\kappa \psi)
+    \int_0^{\infty} \exp(-\omega \psi^2 /2) \pi(\omega) d \omega,
   \end{equation}
-where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed.
-This holds for all $b > 0$ and $a \in \mathbb{R}$.
-After setting $a=y_i$, $b=m_i$, and $\psi=\x_i \tr \bbeta$, we may consider the right-hand side of (\ref{eq:intid}) as an un-normalised joint density in $(\bbeta, \omega)$.
-The likelihood contribution of observation $i$ for $\bbeta$ conditional on $\omega_i$ may now be written as:
+  where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed. This
+  holds for all $b > 0$ and $a \in \mathbb{R}$. After setting $a=y_i$, $b=m_i$,
+  and $\psi=\x_i \tr \bbeta$, we may consider the right-hand side of 
+  (\ref{eq:intid}) as an un-normalised joint density in $(\bbeta, \omega)$. The
+  likelihood contribution of observation $i$ for $\bbeta$ conditional on
+  $\omega_i$ may now be written as:
   $$
-\L_i(\bbeta | \omega_i) \propto \exp[\kappa_i \x_i \tr \bbeta - \omega_i (\x_i \tr \bbeta)^2/2].
+  \L_i(\bbeta | \omega_i) \propto \exp[\kappa_i \x_i \tr \bbeta - 
+  \omega_i (\x_i   \tr \bbeta)^2/2].
   $$
@@
   $$
-p(\bbeta | \bomega, \y) \propto \pi(\bbeta) \prod_{i=1}^n \L_i(\bbeta | \omega_i) = \pi(\bbeta) \exp \[-\frac{1}{2}(\bOmega^{-1} \bkappa - \X \bbeta) \tr \bOmega (\bOmega^{-1} \bkappa - \X \bbeta)\],
+  p(\bbeta | \bomega, \y) \propto \pi(\bbeta) \prod_{i=1}^n 
+  \L_i(\bbeta | \omega_i) = \pi(\bbeta) 
+  \exp \[-\frac{1}{2}(\bOmega^{-1} \bkappa - \X \bbeta) \tr 
+  \bOmega (\bOmega^{-1} \bkappa - \X \bbeta)\],
   $$
-where $\bOmega = \diag (\omega_i)$, for $i=1, \dots, n$. Consequently, with $\pi(\bbeta)$ a Gaussian prior, $p(\bbeta | \bomega, \y)$ is also in the Gaussian family.
+  where $\bOmega = \diag (\omega_i)$, for $i=1, \dots, n$. Consequently, with
+  $\pi(\bbeta)$ a Gaussian prior, $p(\bbeta | \bomega, \y)$ is also in the
+  Gaussian family.
 
-Furthermore, by treating the integrand in (\ref{eq:intid}) as an un-normalized density over $(\psi, \phi)$, we find the distribution of $\omega$ conditional on $\bbeta$,
+  Furthermore, by treating the integrand in (\ref{eq:intid}) as an un-normalized
+  density over $(\psi, \phi)$, we find the distribution of $\omega$ conditional
+  on $\bbeta$,
   $$
-p(\omega | \psi) = \frac{\exp(-\omega \psi^2/2) \pi(\omega)}{\int_0^{\infty} \exp(-\omega \psi^2/2) \pi(\omega) d\omega},
+  p(\omega | \psi) = \frac{\exp(-\omega \psi^2/2) \pi(\omega)}{\int_0^{\infty}
+  \exp(-\omega \psi^2/2) \pi(\omega) d\omega},
   $$
-distributed as $\omega | \psi \sim \mathcal{PG}(b,\psi)$ \cite[]{polson_bayesian_2013}, with expectation $\E[ \omega ] = \frac{b}{2\psi} \tanh (\psi/2)$.
-We substitute again $b=m_i$ and $\psi=\x_i \tr \bbeta$, and set the prior $\pi(\bbeta)$ to a Gaussian with mean $\m$ and covariance $\Vm$ to arrive at the following conditional posteriors:
+  distributed as $\omega | \psi \sim \mathcal{PG}(b,\psi)$ 
+  \cite[]{polson_bayesian_2013}, with expectation $\E[ \omega ] = 
+  \frac{b}{2\psi} \tanh (\psi/2)$. We substitute again $b=m_i$ and $\psi=\x_i 
+  \tr \bbeta$, and set the prior $\pi(\bbeta)$ to a Gaussian with mean $\m$ and
+  covariance $\Vm$ to arrive at the following conditional posteriors:
   \begin{subequations}\label{eq:gibbssampler}
@@
   with $\bkappa = \begin{bmatrix} \kappa_1 \dots \kappa_n \end{bmatrix} \tr$.
-Equations (\ref{eq:gibbssampler}) constitute the Gibbs sampler used in \cite{polson_bayesian_2013}.
-However, from the corresponding variational Bayes derivations in Section \ref{sec:variationaldistributions}, it is obvious that the parametrisation above is also useful in the variational Bayes approximation to the posterior distribution.
+  Equations (\ref{eq:gibbssampler}) constitute the Gibbs sampler used in
+  \cite{polson_bayesian_2013}. However, from the corresponding variational Bayes
+  derivations in Section \ref{sec:variationaldistributions}, it is obvious that
+  the parametrisation above is also useful in the variational Bayes 
+  approximation to the posterior distribution.
 
@@
   \begin{align}
-q^*_{\bbeta} (\bbeta) & {\propto} \exp \{ \E_{\bomega, \btau} [\log p (\bbeta, \bomega, \btau, \y)]\} {\propto} \exp \{\log \L (\y ; \bbeta) + \E_{\bomega} [\log \pi (\bomega | \bbeta)] + \E_{\btau} [ \log \pi (\bbeta | \btau)] \}, \label{eq:qbeta} \\
-q^*_{\bomega} (\bomega) & {\propto} \exp \{ \E_{\bbeta, \btau} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi (\bomega | \bbeta)] \},\label{eq:qomega} \\
-q^*_{\tau} (\btau) & {\propto} \exp \{ \E_{\bbeta, \bomega} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi ( \bbeta | \btau)] + \log \pi (\btau) \}. \label{eq:qtau}
+    q^*_{\bbeta} (\bbeta) & {\propto} \exp \{ \E_{\bomega, \btau} 
+    [\log p (\bbeta, \bomega, \btau, \y)]\} {\propto} \exp 
+    \{\log \L (\y ; \bbeta) + \E_{\bomega} [\log \pi (\bomega | \bbeta)] + 
+    \E_{\btau} [ \log \pi (\bbeta | \btau)] \}, \label{eq:qbeta} \\
+    q^*_{\bomega} (\bomega) & {\propto} \exp \{ \E_{\bbeta, \btau} 
+    [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} 
+    [\log \pi (\bomega | \bbeta)] \},\label{eq:qomega} \\
+    q^*_{\tau} (\btau) & {\propto} \exp \{ \E_{\bbeta, \bomega} 
+    [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} 
+    [\log \pi ( \bbeta | \btau)] + \log \pi (\btau) \}. \label{eq:qtau}
   \end{align}
 
-For convenience we work on the log-scale and occasionally do a change of variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three terms on the right-hand side of (\ref{eq:qbeta}):
+  For convenience we work on the log-scale and occasionally do a change of
+  variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three
+  terms on the right-hand side of (\ref{eq:qbeta}):
   \begin{align*}
-\log \L (\y ; \bbeta) & = \sum_{i=1}^n \log f (y_i ; \bbeta) \propto \sum_{i=1}^n \log \left\{ \frac{\exp(\x_i \tr \bbeta)^{y_i}}{[1 + \exp(\x_i \tr \bbeta)]^{m_i}} \right\} \\
+    \log \L (\y ; \bbeta) & = \sum_{i=1}^n \log f (y_i ; \bbeta) \propto 
+    \sum_{i=1}^n \log \left\{ \frac{\exp(\x_i \tr \bbeta)^{y_i}}{[1 + 
+    \exp(\x_i \tr \bbeta)]^{m_i}} \right\} \\
     & = \y \tr \X \bbeta - \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)], \\
-\E_{\bomega} [\log \pi (\bomega | \bbeta)] & \propto \sum_{i=1}^n \E_{\omega_i} [\log \pi (\omega_i | \bbeta)] \\ & \propto \sum_{i=1}^n \E_{\omega_i} \( \log \left\{ \cosh\( \frac{\x_i \tr \bbeta}{2}\)^{m_i} \exp \[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i \]\right\} \) \\
-& \propto \sum_{i=1}^n m_i \log \cosh \( \frac{\x_i \tr \bbeta }{2} \) - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta \\ & \propto \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)] - \frac{1}{2} \m \tr \X \bbeta - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta, \\
-\E_{\btau} [ \log \pi (\bbeta | \btau)] & = \sum_{j=1}^p \E_{\tau_j} [ \log \pi (\beta_j | \tau_j)] \propto \frac{\lambda_2}{2} \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} \E \( \frac{\tau_j}{\tau_j - 1} \) \\
-& = \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} [1 + \E ( \psi_j^{-1} )] = - \frac{1}{2} \lambda_2 \bbeta \tr \bLambda' (\I + \Z) \bbeta,
+    \E_{\bomega} [\log \pi (\bomega | \bbeta)] & \propto \sum_{i=1}^n 
+    \E_{\omega_i} [\log \pi (\omega_i | \bbeta)] \\ & \propto \sum_{i=1}^n
+    \E_{\omega_i} \( \log \left\{ \cosh\( \frac{\x_i \tr \bbeta}{2}\)^{m_i} 
+    \exp \[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i \]\right\} \) \\
+    & \propto \sum_{i=1}^n m_i \log \cosh \( \frac{\x_i \tr \bbeta }{2} \) -
+    \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta \\ & \propto 
+    \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)] - \frac{1}{2} \m \tr 
+    \X \bbeta - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta, \\
+    \E_{\btau} [ \log \pi (\bbeta | \btau)] & = \sum_{j=1}^p \E_{\tau_j} 
+    [ \log \pi (\beta_j | \tau_j)] \propto \frac{\lambda_2}{2} 
+    \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} \E \( \frac{\tau_j}{\tau_j - 1} \) \\
+    & = \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} [1 + \E ( \psi_j^{-1} )] = -
+    \frac{1}{2} \lambda_2 \bbeta \tr \bLambda' (\I + \Z) \bbeta,
   \end{align*}
-where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega = \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' = \diag (\lambda'_{g(j)})$. 
-Since the difficult $\sum_{i=1}^n m_i \log[1 + \exp(\x_i \tr \bbeta)]$ term appears in both the $\log \L (\y ; \bbeta)$ and $\pi (\omega_i | \bbeta)$ part, it cancels, thereby justifying the introduction of the $\omega_i$.
+  where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega =
+  \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' =
+  \diag (\lambda'_{g(j)})$. 
+  Since the difficult $\sum_{i=1}^n m_i \log[1 + \exp(\x_i \tr \bbeta)]$ term
+  appears in both the $\log \L (\y ; \bbeta)$ and $\pi (\omega_i | \bbeta)$ 
+  part, it cancels, thereby justifying the introduction of the $\omega_i$.
   Finally, we have: 
@@
   25 features (including the 5 unpenalized covariates). We calculated the size 
-  of all $\genfrac(){0pt}{2}{50}{2}$ intersections of selected features and present the 
-  results in Figure [! figure ref here].
+  of all $\genfrac(){0pt}{2}{50}{2}$ intersections of selected features and
+  present the results in Figure [! figure ref here].
 
@@
   We use $\alpha \in \{0.05, 0.5, 0.95 \}$ for the
-  methods that require this extra tuning parameter. The AUCs calculated on the
-  validation set are presented in Figure 
-  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_auc}
+  methods that require this extra tuning parameter. The AUCs and Brier skill 
+  scores calculated on the
+  validation set are presented in Figures 
+  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_auc}--
+  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_briers}
 <<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated AUC for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_rnaseq_lymph_node_metastasis_res1_briers, fig.cap="Estimated Briers skill score for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
 @
@@
 
-	To estimate performance measures AUC and BSS, we split the data into 120 
-	training instances and 54 test instances. 
+	To estimate performance measures AUC and Briers skill score, we split the data
+	into 120 training instances and 54 test instances. 
 	The training data is used to fit the models, while the test data
-	is used to estimate the performance measures on. We included the different 
-	co-data types seperately and combined, which resulted in the estimated 
-	multipliers and performance measures presented in Figure 
-	\ref{fig:barplot_lines_metabolomics_alzheimer_res5_auc}.
-<<barplot_lines_metabolomics_alzheimer_res5_auc, fig.cap="Results for the alzheimer example with (a) the estimated penalty multipliers for the quality score, (b) the estimated penalty multipliers for the network degree, (c) the estimated AUC, and (d) the estimated Brier skill score.", out.width="100%", fig.asp=1>>=
+	is used to estimate the performance measures on. This resulted in the 
+	estimated multipliers and performance measures presented in Figures 
+	\ref{fig:barplot_metabolomics_alzheimer_res1}--
+	\ref{fig:lines_metabolomics_alzheimer_res1_briers}.
+<<barplot_metabolomics_alzheimer_res1, fig.cap="Estimated penalty multiplies for the (a) quality score and the (b) network degree in the alzheimer example.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Estimated AUC for the alzheimer example.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_metabolomics_alzheimer_res1_briers, fig.cap="Estimated Brier skill score for the alzheimer example.", out.width="100%", fig.asp=1>>=
 @
@@
 
-<<barplots_micrornaseq_cervical_cancer_res1_auc, fig.cap="Estimated penalty multipliers for conservation status co-data", out.width="50%", fig.asp=2/3>>=
+<<barplot_micrornaseq_cervical_cancer_res1, fig.cap="Estimated penalty multiplies for conservation status in cervical cancer example.", out.width="50%", fig.asp=1>>=
 @
 
-<<lines_micrornaseq_cervical_cancer_res1_auc, fig.cap="Cross-validated AUC in cervicl cancer example", out.width="50%", fig.asp=1>>=
+<<lines_micrornaseq_cervical_cancer_res1_auc, fig.cap="Cross-validated AUC in cervical cancer example", out.width="100%", fig.asp=1>>=
 @
 
-  \section{Choice of \texorpdfstring{$\alpha$}{alpha}}	
-	Choice of $\alpha$ does matter in Figure 
-	\ref{fig:lines_colorectal_lymph_node_gren_auc}.
-<<lines_colorectal_lymph_node_gren_auc, fig.cap="Estimated AUC for gren in the (a) colorectal cancer example and the (b) lymph node metastasis example.", out.width="100%", fig.asp=1/2>>=
+<<lines_micrornaseq_cervical_cancer_res1_briers, fig.cap="Cross-validated Brier skill score in cervical cancer example", out.width="100%", fig.asp=1>>=
 @

### 2019-03-07 code/micrornaseq_colorectal_cancer.R

@@ -303,7 +303,7 @@ part <- diff.expr
 ################################################################################
 
 ### analyse samples in parallel
-ncores <- min(detectCores() - 1, nsamples)
+ncores <- min(detectCores(), nsamples)
 cluster <- makeCluster(ncores, type="FORK")
 registerDoParallel(cluster)
 

### 2019-03-08 docs/supplement.Rnw
parallelised simulations.R simulation 1

@@
   of all $\genfrac(){0pt}{2}{50}{2}$ intersections of selected features and
-  present the results in Figure [! figure ref here].
-
-  [! figure here]
+  present the results in Figure 
+  \ref{fig:hist_micrornaseq_colorectal_cancer_res4_overlap}.
+<<hist_micrornaseq_colorectal_cancer_res4_overlap, fig.cap="Stability of feature selection in colorectal cancer example for (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=2/3>>=
+@
 

### 2019-03-08 code/figures.R

@@ -298,6 +298,32 @@ abline(h=1, lty=2, lwd=1.5)
 par(opar)
 
 # ---- hist_micrornaseq_colorectal_cancer_res4_overlap ----
+library(sp)
+res <- read.table("results/micrornaseq_colorectal_cancer_res4.csv", 
+                  stringsAsFactors=FALSE)
+
+methods <- colnames(res)
+labels <- c("gren", "enet")
+col <- bpy.colors(length(labels), cutoff.tail=0.1)
+
+plot.data <- sapply(c(1:3), function(m) {
+  sapply(1:max(res), function(s) {
+    c(sum(res[[paste0("gren", m)]]==s), sum(res[[paste0("enet", m)]]==s))/
+      nrow(res)})}, simplify=FALSE)
+
+opar <- par(no.readonly=TRUE)
+par(mar=opar$mar*c(1, 1.3, 1, 1))
+layout(matrix(c(rep(c(1, 1, 2, 2), 2), rep(c(0, 3, 3, 0), 2)), 
+              nrow=4, ncol=4, byrow=TRUE))
+barplot(plot.data[[1]], beside=TRUE, xlab="Number of overlapping features",
+        ylab="Density", main="(a)", col=col, names.arg=1:max(res), 
+        legend.text=labels, args.legend=list(x="topright", fill=col, 
+                                             border=col))
+barplot(plot.data[[2]], beside=TRUE, xlab="Number of overlapping features",
+        ylab="Density", main="(b)", col=col, names.arg=1:max(res))
+barplot(plot.data[[3]], beside=TRUE, xlab="Number of overlapping features",
+        ylab="Density", main="(c)", col=col, names.arg=1:max(res))
+par(opar)
 
 # ---- lines_rnaseq_lymph_node_metastasis_res1_auc ----
 library(sp)

### 2019-03-08 code/simulations.R
parallelised simulations.R simulation 1

@@ -12,7 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-parallel <- FALSE
+parallel <- TRUE
 
 ### libraries
 library(gren)
@@ -44,7 +44,7 @@ b <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
 beta <- numeric(p)
 
 part <- rep(c(1:G), each=p/G)
-nreps <- 2
+nreps <- 100
 
 ################################################################################
 ## Error in { :                                                                #
@@ -52,7 +52,8 @@ nreps <- 2
 ################################################################################
 
 ### analysis splits in parallel
-ncores <- min(detectCores() - 1, nreps)
+ncores <- 50
+# ncores <- min(detectCores() - 1, nreps)
 cluster <- makeForkCluster(ncores)
 if(parallel) {
   registerDoParallel(cluster)

### 2019-03-12 code/simulations.R
ran simulations.R (with errors)

@@ -12,7 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-parallel <- TRUE
+parallel <- FALSE
 
 ### libraries
 library(gren)
@@ -61,7 +61,7 @@ if(parallel) {
   registerDoSEQ()
 }
 
-res <- foreach(k=c(1:nreps)) %dopar% {
+res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   set.seed(2019 + k)
   
   for(g in 1:G) {
@@ -175,9 +175,14 @@ res <- foreach(k=c(1:nreps)) %dopar% {
   
 }
 if(parallel) {stopCluster(cluster)}
+save(res, file="results/simulations_res1.Rdata")
+
 test <- res
 test <- sapply(c("psel", "auc", "briers", "mse", "kappa", "mults"), function(s) {
   sapply(test, function(m) {m[[s]]}, simplify=FALSE)}, simplify=FALSE)
+
+str(test$psel)
+
 
 
 res <- sapply(c("psel", "auc", "briers", "mse", "kappa", "mults"), function(s) {
@@ -189,4 +194,3 @@ sapply(res$psel, function(s) {s[grep("gren1", names(s))]}, simplify=FALSE)
 
 
 
-save(results1, file=paste(path.res, "gren_sim1_res1.Rdata", sep=""))

### 2019-03-12 docs/manuscript.Rnw
gren.R small bug fix

@@
 	information on the molecular features involved (e.g., the Cancer Gene Census
-	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. the location of 
-	a gene on the chromosome). Although this information can rarely be directly
+	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. a gene's 
+	chromosome location). Although this information can rarely be directly
 	included in the statistical analysis, it may still be useful and informative
@@
 	machines have been used in classification problems as well 
-	\cite[]{chakraborty_bayesian_2011}, but just as VBVS, these methods lack 
-	adaptive inclusion of external information on the features. We emphasize that 
-	our method is different from recent developments in grouped penalized 
-	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
-	features, the observations are partitioned.
+	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
+	adaptive inclusion of external information on the features.
 
@@
 	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
-	user: only the data and grouping of the features is required.
+	user: only the data and grouping of the features is required. We emphasize that 
+	our method is different from recent developments in grouped penalized 
+	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
+	features, the observations are partitioned.
 
@@
 	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
-	the two previously introduced data sets. In Section \ref{sec:discussion} 
+	two data sets. In Section \ref{sec:discussion} 
 	we conclude by both discussing some differences and
@@
 	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
-	group-specific weight. In generalised elastic net regression, the penalised 
+	group-specific weight. In a generalised elastic net regression, the penalised 
 	likelihood is maximised to yield parameter estimates:
@@
 	\cite{li_bayesian_2010} show that the elastic net prior 
-	(see SM Section [! section reference here] for details on this prior) may be 
-	written as
+	(see Supplementary Material (SM) Section [! section reference here] for 
+	details on this prior) may be written as
 	a computationally more convenient scale mixture of normals, with mixing 
@@
 	Feature selection is often desirable in high-dimensional prediction problems 
-	and omics data problems are no exception. For example, a selection of 
-	biomarkers may lead 
-	to a large cost reduction by supporting targeted assays. Bayesian feature
+	and omics data problems are no exception. For example, biomarker selection 
+	may lead 
+	to a large reduction in costs by supporting targeted assays. Bayesian feature
 	selection is often done by inspection of posterior credible intervals. 
 	Similarly as in frequentist hypothesis testing, we may select a feature if 
-	zero is not contained in the credible interval. However, the Bayesian lasso's
-	credible intervals (a special case of the elastic net) are known to suffer 
+	zero is not in the credible interval. However, the Bayesian lasso's 
+	(a special case of the elastic net) credible intervals  are known to suffer 
 	from low frequentist coverage in sparse settings 
@@
 	To assess variable selection and predictive performance we conducted a 
-	simulation study in which we compare \texttt{gren} to the regular elastic 
+	simulation study in which we compared \texttt{gren} to the regular elastic 
 	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
 	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
-	group mimimax concave penalty (\texttt{grMCP}) 
+	composite mimimax concave penalty (\texttt{cMCP}) 
 	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
@@
 	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
-	are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} 
-	and not adaptive on the group level. In contrast to the original group lasso,
-	all of them are able to select single features, instead of just on the group 
-	level.
+	are extensions of the group lasso and not adaptive on the group level. 
+	However, in contrast to the original group lasso,
+	they select single features, instead of just whole groups.
 
-	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, 
+	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{cMCP}, 
 	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
@@
 	where $\alpha=0$, while the third setting is similar to the lasso with 
-	$\alpha=1$. In \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, $\alpha$
-	determines the proportion of group-level penalization ($\alpha=0$) compared to
-	$L_1$-norm penalty on the feature level ($\alpha=1$). For the methods that do 
-	automatic variable selection, we estimate 
+	$\alpha=1$. In \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}, $\alpha$
+	determines the ratio of group-level penalization ($\alpha=0$) to 
+	feature level ($\alpha=1$) penalization. For the methods that do 
+	automatic feature selection, we estimate 
 	the models for a range of model sizes.
@@
 	\end{enumerate}
-	Full descriptions of the scenarios are given in Section [! ref to section] of 
-	the SM. In the first scenario, the distribution of the model parameters is far
-	off from the assumed ridge/elastic net/lasso distributions, so we assume that 
-	all methods will underperform. In the second scenario, we expect \texttt{gren}
+	Full descriptions of the scenarios are given in Section [! section ref here] 
+	of 
+	the SM. In the first scenario, the distribution of model parameters is far
+	from the assumed ridge/elastic net/lasso distributions, so we expect
+	all methods to underperform. In the second scenario, we expect \texttt{gren}
 	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
-	estimate. The rest of the methods are conjectured to perform better, due to a
-	lack of overfitting. In scenario three, we expect all methods to perform badly
-	due to the high correlation between the features. The \texttt{gren} fourth 
-	scenario is conjectured to give a better \texttt{gren} performance, due to the
-	correctly assumed heavier-tailed distribution and differential signal between
-	the groups. Scenario five is expected to lead to a bad performance in all 
-	models, because of the sparseness of the truth and lack of differential 
+	estimate, while we expect the other methods to perform better, due to a
+	lack of overfitting. In scenario three, we expect all methods to underperform
+	due to the high correlations between the features. The fourth 
+	scenario is conjectured to yield a better \texttt{gren} performance, since
+	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
+	incorporate a differential signal between
+	the groups. Scenario five is expected to yield a bad performance for all 
+	methods, because of the sparse truth and lack of differential 
 	signal.
@@
 	In all scenarios we created a training set of $n=100$ samples to estimate the 
-	models and a test set of $n_{\text{test}}=1000$ samples on which we compute 
+	models and a test set of $n_{\text{test}}=1000$ samples to compute 
 	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
 	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
-	$p=1000$. We let the covariance matrix of the features vary over the 
-	scenarios. Likewise for the model parameters $\beta_j$, $j=1, \dots, p$. We 
-	simulate the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
+	$p=1000$. We varied the covariance matrix of the features and simulation of 
+	the model parameters $\beta_j$, $j=1, \dots, p$ between the
+	scenarios (see SM Section [! section ref here] for the exact 
+	parametrisations). We 
+	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
 	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
-	of random variation in the simulations, we repeated every scenario 100 times 
+	of random variation, we repeated each scenario 100 times 
 	and report estimated penalty parameters and predictive performance measures 
-	area under the receiver operator curve (AUC) and Brier skill score (BSS).
+	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
+	Table [! table ref here].
+
+	- description of results
 
@@
 
-	Additional results are presented in Section [! section ref here] of the SM.
+	Additional results and performance measures are presented in Section 
+	[! section ref here] of the SM.
 
@@
 	\label{sec:colorectal}
-  As shortly touched upon in Section \ref{sec:introduction}, the data is from a 
-  deep sequencing analysis on microRNAs 
-  \cite[]{neerincx_combination_2018}. This study in 88 treated 
-  colorectal cancer patients was conducted with the aim of predicting treatment 
-  response, coded as either non-progressive/remission (70 patients) or 
-  progressive 
+  We investigated the performance of \texttt{gren} on data from a 
+  deep sequencing study in microRNAs 
+  \cite[]{neerincx_combination_2018}. The aim of the study was to predict 
+  treatment response in 88 treated colorectal cancer patients, coded as either
+  non-progressive/remission (70 patients) or progressive 
   (18 patients). After pre-processing and normalisation, 2114 microRNAs 
-  remained. In addition to the 2114 microRNAs, four clinical unpenalized 
+  remained. In addition to the microRNAs, four unpenalized clinical
   covariates were included in the analysis: prior use of adjuvant therapy 
@@
   microRNAs that were differentially expressed in both comparisons with
-  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 not 
-  differentially expressed microRNAs. We expect that incorporation of this 
-  partitioning enhances therapy response 
-  classification, because tumor-specific miRNAs might be more relevant than 
-  non-specific ones.
+  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 
+  microRNAs that were not differentially expressed. We expect that incorporation
+  of this partitioning enhances therapy response 
+  classification, because tumor-specific microRNAs are likely to be more 
+  relevant than non-specific microRNAs.
 
   We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
-  elastic net, and to the best performing
-  group lasso extension, the composite minimax
-  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. 
-  For the methods with an extra tuning parameter 
-  $\alpha$, we picked the one (or two) that gave the best performance and refer
-  the reader to SM Section [! section ref here] for more comparisons with 
-  different $\alpha$s and comparisons to the sparse
-  group lasso by \cite{simon_sparse-group_2013} and group
-  exponential lasso \cite[]{breheny_group_2015}.
+  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
+  three methods, we only present \texttt{cMCP} here; the results for 
+  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
+  [! section ref here]. For the methods with a
+  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
+  performance and refer
+  the reader to SM Section [! section ref here] for comparisons with 
+  different $\alpha$'s.
 
@@
   instances, stratified by treatment response. We estimated the models on the 
-  training data and calculated AUC on the test data. We present AUC for a range 
-  of model sizes, together with the estimated 
+  training data and calculated AUC on the test data. We present the AUC for a
+  range of model sizes, together with the estimated 
   penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
@@
   of the SM. In addition, we investigated the sensitivity of
-  the multiplier estimation using in SM Section 
+  the multiplier estimation on this dataset in SM Section 
   [! section ref here]. We experienced numerical issues with 
   \texttt{GRridge} when cross-validating 
-  the global penalty parameter on the training data. We therefore estimated this
+  the global $\lambda$ on the training data. We therefore estimated this
   global parameter on the full data for \texttt{GRridge}.
@@
   This is especially pronounced for the larger models. The performance of
-  \texttt{cMCP} is somewhat unstable. We noticed that this is in general the 
-  case for the group-lasso and its extensions (see SM [! section ref here]). 
-  Overall, it seems as though the inclusion of the extra information on the 
+  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
+  all group lasso extensions that we investigated.
+  (SM [! section ref here]). Overall, it seems as though the inclusion of the
+  extra information on the 
   features benefited predictive performance: both \texttt{GRridge} and 
@@
   albeit it only slightly for \texttt{GRridge}.
-<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
+<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
 @
@@
   \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
-  The aim of this study is to predict lymph node metastasis in patients 
-  with HPV-negative oral cancer using sequenced RNAs from TCGA 
+  The aim of our second example study was to predict lymph node metastasis in
+  oral cancer patients using sequenced RNAs from TCGA 
   \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
@@
 
-  As external feature information, we considered two sources: (a) the 
-  cis-correlation between the RNASeqv2 data and TCGA DNA copy numbers on 
-  the same patients, 
-  quantified by Kendall's $\tau$ and binned into five groups. We also considered
-  (b) $p$-values from an external microarray data set described in 
-  \cite{mes_prognostic_2017}, again binned into five groups. We expect features 
+  We considered two sources of external feature information: (a) the 
+  cis-correlations between the RNASeqv2 data and TCGA DNA copy numbers on 
+  the same patients, quantified by Kendall's $\tau$ and binned into five groups.
+  In addition, we used (b) $p$-values from a microarray data set
+  described in \cite{mes_prognostic_2017}, again binned into five groups. 
+  We expected features 
   with a large positive Kendall's $\tau$ to be more important in 
   metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
-  features with low $p$-values to be more important.
+  features with low external $p$-values to be more important.
 
   We compared \texttt{gren} to the same methods as in Section 
-  \ref{sec:colorectal}. However, because the co-data consists of two partitions 
-  with overlapping groups, we use an extension of \texttt{cMCP} that allows for
+  \ref{sec:colorectal}. However, since the feature information
+  consists of two partitions with overlapping groups, we used an extension of
+  \texttt{cMCP} that allows for
   overlapping groups \cite[]{zeng_overlapping_2016}.
   We estimated AUC for the methods on an independent validation set of 97 
-  HPV-negative oral cancer patients \cite[]{mes_prognostic_2017}. The features 
+  patients \cite[]{mes_prognostic_2017}. The features 
   in this validation set were
-  microarrays, normalised to account for a different scale of the transformed 
-  RNAseq data. We present the results on this validation set in Figure 
-  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Other 
+  microarray expression levels, normalised to account
+  for a difference in scale compared to the RNAseq training data. We present the
+  results on this validation set, together with the estimated penalty 
+  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
+  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
   comparisons and performance measures are presented in the SM Section 
@@
   larger cis-correlations receive smaller penalties. 
-  Likewise, smaller $p$-values are penalized less. This pattern is slightly 
-  less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the other 
-  methods. The predictive performance of
+  Likewise, smaller external $p$-values are penalized less. This pattern is
+  slightly less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the
+  other methods. The predictive performance of
   \texttt{gren} is slighty better than the other methods for a range of model
-  sizes. We also note that the overlapping group \texttt{cMCP} is much more
-  stable than the regular \texttt{cMCP} (see in SM [! section ref here]), 
-  where we create one paritioning by cross-tabulation of the two paritions. 
-  In this example, ridge and \texttt{GRridge} perform almost identical.
-<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.", out.width="100%", fig.asp=1>>=
+  sizes. In this example, ridge and \texttt{GRridge} perform almost 
+  identical, while \texttt{gren} outperforms the regular elastic net.
+<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the lymph node metastasis example.", out.width="100%", fig.asp=1>>=
 @
@@
 	\section{Discussion}\label{sec:discussion}
-	In a taxonomy of Bayesian methods, the proposed method may be considered a 
+	In a taxonomy of Bayesian methods, \texttt{gren} may be considered a 
 	local shrinkage model, as opposed to the global-local shrinkage priors that 
@@
 
-	Although our method can be considered weakly adaptive, it is different from 
+	Although our method can be considered weakly adaptive and is essentially 
+	a reweighted elastic net, it is different from 
 	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
@@
 	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
-	rather than to the data. It also differs in the scale of adaptation: in the 
-	adaptive lasso and elastic net the adaptive weights are feature specific,
-	while in our case they are estimated on the group level, rendering the 
-	adaptation more robust. As can be seen from both the simulations in Section 
-	\ref{sec:simulations} and the applications to different outcomes 
+	rather than to the primary data. It also differs in the scale of adaptation: 
+	in the adaptive lasso and elastic net the adaptive weights are feature 
+	specific, while in our case they are estimated on the group level, rendering
+	the adaptation more robust. As can be seen from both the simulations in 
+	Section \ref{sec:simulations} and the applications to different outcomes 
 	(prognostic or diagnostic) and marker types 
@@
 
-	A common criticism of the lasso (and elastic net) is unstable feature 
-	selection: different data sets often lead to wildly differing sets of selected
-	features. We investigated stability selection on the data from Section
-	\ref{sec:colorectal} by taking bootstraps samples from the data, estimating
-	\texttt{gren} and looking at the overlap between selected features over 
-	different bootstrap samples. We compared it to the regular elastic net and 
-	found that inclusion of extra feature information increases selection 
-	stability (see SM Section [! section ref here]). In addition, we investigated 
-	the stability of penalty multiplier estimation by examining the distribution
+	A common criticism of the lasso (and elastic net) is its instability of 
+	feature 
+	selection: different data instances lead to different sets of selected
+	features. We investigated the stability of selection on the data from Section
+	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
+	the original data and calculated the sizes of all intersections of selected
+	features. We compared \texttt{gren} to the 
+	regular elastic net and found that the inclusion of extra information
+	increased the selection stability (see SM Section [! section ref here]). 
+	In addition, we investigated 
+	the variance and bias in penalty multiplier estimation by examining the
+	distribution
 	of estimated penalty parameters from 100 random partitionings of the features
   (see SM Section [! section ref here]). We found that the penalty parameters
-  tend to cluster around one, as is expected with random groups, and are 
-  unbiased.
+  tend to cluster around one, as expected for an unbiased estimation method.
 
-	As shortly touched upon in Section \ref{sec:introduction}, another obvious 
+	As touched upon in Section \ref{sec:introduction}, another obvious 
 	comparison is to the group lasso \cite[]{meier_group_2008,
@@
 	Supplementary Material is available online from 
-	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All materials needed to   
-	recreate this document are available from 
-	\url{https://github.com/magnusmunch/gren}.
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All results may be recreated
+	from \url{https://github.com/magnusmunch/gren}.
 

### 2019-03-12 rpackage/R/gren.R

@@ -254,7 +254,8 @@ gren <- function(x, y, m=rep(1, nrow(x)), unpenalized=NULL, partitions=NULL,
   vb.conv <- logical(0)
   conv <- FALSE
   iter1 <- 0
-  if(trace) {cat("\n", "Estimating penalty multipliers by empirical Bayes", "\n", sep="")}
+  if(trace) {cat("\n", "Estimating penalty multipliers by empirical Bayes", 
+                 "\n", sep="")}
   srt <- proc.time()[3]
   while(!conv & (iter1 < control$maxit)) {
     
@@ -414,7 +415,8 @@ gren <- function(x, y, m=rep(1, nrow(x)), unpenalized=NULL, partitions=NULL,
     
       # estimate final models using lambda sequence
       fit.final <- glmnet(x, y=ymat, family="binomial", alpha=alpha, 
-                          lambda=sel.out, standardize=FALSE, intercept=intercept, 
+                          lambda=sel.out, standardize=FALSE, 
+                          intercept=intercept, 
                           penalty.factor=c(rep(0, u), lambdamultvec))
     }
   }

### 2019-03-12 code/functions.R
gren.R small bug fix

@@ -33,19 +33,30 @@ sel.grpreg <- function(X, y, group=1:ncol(X),
   init.lambda <- grpreg:::setupLambda(XG$X, yy, XG$g, family, penalty, alpha, 
                                       lambda.min, log.lambda, 2, XG$m)
   
-  # find the lambdas that give the specified psels
-  sel.lambda <- numeric(length(psel))
-  for(i in 1:length(psel)) {
-    fit.root <- suppressWarnings(uniroot(froot, interval=init.lambda, 
-                                         psel=psel[i], inp=inp))
-    sel.lambda[i] <- fit.root$root
+  # check whether smallest lambda gives largest psel, otherwise, this psel
+  # not possible
+  init.fit <- do.call("grpreg", c(inp, lambda=init.lambda, nlambda=2))
+  pselfit <- psel[psel <= max(colSums(as.matrix(init.fit$beta[-1, ])!=0))]
+  
+  # if no model has fewer selected features than any psel, we return initial
+  # model
+  if(length(pselfit) > 0) {
+    # find the lambdas that give the specified psels
+    sel.lambda <- numeric(length(pselfit))
+    for(i in 1:length(pselfit)) {
+      fit.root <- suppressWarnings(uniroot(froot, interval=init.lambda, 
+                                           psel=pselfit[i], inp=inp))
+      sel.lambda[i] <- fit.root$root
+    }
+  
+    # final fit with the found lambdas
+    inp[["lambda"]] <- sel.lambda
+    inp[["nlambda"]] <- max(length(sel.lambda), 2)
+  
+    fit <- do.call("grpreg", inp)
+  } else {
+    fit <- init.fit
   }
-  
-  # final fit with the found lambdas
-  inp[["lambda"]] <- sel.lambda
-  inp[["nlambda"]] <- length(sel.lambda)
-  
-  fit <- do.call("grpreg", inp)
   return(fit)
   
 }

### 2019-03-12 code/micrornaseq_colorectal_cancer.R

@@ -349,7 +349,7 @@ selnames <- foreach(k=c(1:nsamples), .errorhandling="pass") %dopar% {
 if(parallel) {stopCluster(cluster)}
 
 ################################## DEBGUGGING ##################################
-test1 <- selnames[-46]
+test1 <- selnames[-46]
 test2 <- sapply(c(paste0("gren", 1:3), paste0("enet", 1:3)), function(m) {
   sapply(test1, function(s) {s[[grep(m, names(s))]]}, simplify=FALSE)},
   simplify=FALSE)

### 2019-03-12 code/simulations.R

@@ -44,6 +44,7 @@ b <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
 beta <- numeric(p)
 
 part <- rep(c(1:G), each=p/G)
+csel <- 2^c(1:8)
 nreps <- 100
 
 ################################################################################
@@ -80,29 +81,26 @@ res <- foreach(k=c(1:nreps)) %dopar% {
   fit.gren1 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.05, 
                     standardize=TRUE, trace=FALSE, psel=csel)
   fit.gren2 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.5, 
-                    standardize=TRUE, trace=FALSE)
+                    standardize=TRUE, trace=FALSE, psel=csel)
   fit.gren3 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.95, 
-                    standardize=TRUE, trace=FALSE)
+                    standardize=TRUE, trace=FALSE, psel=csel)
   
   fit.grridge <- grridge(t(xtrain), ytrain, list(part=split(1:p, part)))
   
-  fit.sgl1 <- SGL(list(x=xtrain, y=ytrain), part, type="logit", alpha=0.05)
-  fit.sgl2 <- SGL(list(x=xtrain, y=ytrain), part, type="logit", alpha=0.5)
-  fit.sgl3 <- SGL(list(x=xtrain, y=ytrain), part, type="logit", alpha=0.95)
+  fit.cmcp1 <- sel.grpreg(xtrain, ytrain, part, penalty="cMCP", 
+                          family="binomial", alpha=0.05, psel=csel)
+  fit.cmcp2 <- sel.grpreg(xtrain, ytrain, part, penalty="cMCP", 
+                          family="binomial", 
+                          alpha=0.5, psel=csel)
+  fit.cmcp3 <- sel.grpreg(xtrain, ytrain, part, penalty="cMCP", 
+                          family="binomial", alpha=0.95, psel=csel)
   
-  fit.cmcp1 <- grpreg(xtrain, ytrain, part, penalty="cMCP", family="binomial", 
-                      alpha=0.05)
-  fit.cmcp2 <- grpreg(xtrain, ytrain, part, penalty="cMCP", family="binomial", 
-                      alpha=0.5)
-  fit.cmcp3 <- grpreg(xtrain, ytrain, part, penalty="cMCP", family="binomial", 
[…]

### 2019-03-19 docs/manuscript.Rnw
fixed all cross-document references

@@
 amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
-titlesec,multicol,hyperref,verbatim,threeparttable,booktabs,dsfont,parskip}
-
-% Add history information for the article if required
-%\history{Received August 1, 2010;
-%	revised October 1, 2010;
-%	accepted for publication November 1, 2010}
+titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
+xr-hyper,hyperref}
+\externaldocument[sm-]{supplement}
 
@@
 		regression to estimate both the model and penalty parameters in an 
-		approximate empirical-variational Bayes framework. Simulations and an
-		application to a colon cancer microRNA study show that, if the partitioning
-		of the features is informative, classification performance and feature
+		approximate empirical-variational Bayes framework. Simulations and
+		applications in two different molecular marker and cancer type studies show
+		that, if the partitioning of the features is informative, classification 
+		performance and feature
 		selection are indeed enhanced.}
@@
 	\cite{li_bayesian_2010} show that the elastic net prior 
-	(see Supplementary Material (SM) Section [! section reference here] for 
+	(see Supplementary Material (SM) Section \ref{sm-sec:enetprior} for 
 	details on this prior) may be written as
@@
 	is approximated by a tractable form and estimated by optimizing a lower bound 
-	on the marginal likelihood of this model (see Section [! section ref here] 
-	of the SM for the 
+	on the marginal likelihood of this model (see Section 
+	\ref{sm-sec:objectivefunction} of the SM for the 
 	lower bound of the proposed model). For an extensive introduction and concise 
@@
 	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
-	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more 
+	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
+	\ref{sm-sec:polyagamma} of the SM for more 
 	details). We augment Model (\ref{eq:enmodel2}) with:
@@
 	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
-	distribution (See SM Section 5 for the derivations). The so-called variational
+	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
+	for the derivations). The so-called variational
 	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
@@
 	element of a column vector, respectively. Naive calculation of the variational
-	parameters is computationally expensive. In Section 6 of the SM we show that 
+	parameters is computationally expensive. In Section 
+	\ref{sm-sec:computationalcomplexity} of the SM we show that 
 	informed calculation of the parameters results in a significant reduction of 
@@
 	in (\ref{eq:ebEM}) is approximated using the variational posterior:
-	$$
-	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \]
+	\begin{equation}\label{eq:approximatedloglikelihood}
+	  \E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | 
+	  \blambda'^{(k)} \]
 	  = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
 	  \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
-	$$
-	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full 
+	\end{equation}
+	where $C$ is constant in $\blambda'$ (see SM Section 
+	\ref{sm-sec:expectedloglikelihood} for the full 
 	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
@@
 	closed form, this convex problem is easily solved by a numerical optimisation 
-	routine. The full procedure is summarized in Section 8 of the SM.
+	routine. The full procedure is summarized in Section 
+	\ref{sm-sec:fullestimationprocedure} of the SM.
 
@@
 	\end{enumerate}
-	Full descriptions of the scenarios are given in Section [! section ref here] 
-	of 
+	Full descriptions of the scenarios are given in Section 
+	\ref{sm-sec:simulations} of 
 	the SM. In the first scenario, the distribution of model parameters is far
 	from the assumed ridge/elastic net/lasso distributions, so we expect
-	all methods to underperform. In the second scenario, we expect \texttt{gren}
+	all methods to underperform. In the second scenario, which is similar to
+	the simulations presented in the original group lasso paper 
+	\cite[]{yuan_model_2006}, we expect \texttt{gren}
 	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
@@
 
-	[! results here]
+<<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
+@
 
 	Additional results and performance measures are presented in Section 
-	[! section ref here] of the SM.
+	\ref{sm-sec:simulations} of the SM.
 
@@
   performance and refer
-  the reader to SM Section [! section ref here] for comparisons with 
+  the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
   different $\alpha$'s.
@@
   \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
-  performance measures are presented in Section [! section ref here]
+  performance measures are presented in Section \ref{sm-sec:colorectal}
   of the SM. In addition, we investigated the sensitivity of
   the multiplier estimation on this dataset in SM Section 
-  [! section ref here]. We experienced numerical issues with 
+  \ref{sm-sec:colorectal}. We experienced numerical issues with 
   \texttt{GRridge} when cross-validating 
@@
   all group lasso extensions that we investigated.
-  (SM [! section ref here]). Overall, it seems as though the inclusion of the
-  extra information on the 
+  (SM Section \ref{sm-sec:colorectal}). Overall, it seems as though the 
+  inclusion of the extra information on the 
   features benefited predictive performance: both \texttt{GRridge} and 
@@
   comparisons and performance measures are presented in the SM Section 
-  [! section ref here].
+  \ref{sm-sec:lymphnode}.
 
@@
 	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
-	\ref{sec:lymphnode}, and SM Sections [! section refs here], adaptation to 
+	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
+	\ref{sm-sec:cervical}, adaptation to 
 	external data may be beneficial for prediction and feature selection. 
@@
 	regular elastic net and found that the inclusion of extra information
-	increased the selection stability (see SM Section [! section ref here]). 
+	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
 	In addition, we investigated 
@@
 	of estimated penalty parameters from 100 random partitionings of the features
-  (see SM Section [! section ref here]). We found that the penalty parameters
+  (see SM Section \ref{sm-sec:colorectal}). We found that the penalty parameters
   tend to cluster around one, as expected for an unbiased estimation method.
@@
 	comparison is to the group lasso \cite[]{meier_group_2008,
-	yuan_model_2006} and its extensions. Although it is similar in the sense that 
-	it shrinks on the 
-	group level, the group lasso is built upon an entirely different philosophy: 
+	yuan_model_2006} and its extensions. Although the group lasso is similar in 
+	the sense that it shrinks on the 
+	group level, it is built upon an entirely different philosophy: 
 	its intended application is to small interpretable groups of features, like, 
@@
 	estimates one overall penalty parameter; it is thereby less flexible in 
-	differential shrinkage of the parameters. In addition, in the simulations and
-	data applications of \cite{van_de_wiel_better_2016}, group lasso prediction 
-	performed inferior to \texttt{GRridge} prediction.
+	differential shrinkage of the parameters. In addition, we show in our 
+	simulations and data applications that \texttt{gren} is competitive and often
+	even superior to (extensions of) the group lasso.
 
@@
 @
-  We have investigated the computation time of the algorithm and compared to the
-  other methods introduced in Sections \ref{sec:colorectal}--\ref{sec:lymphnode}
-  on the respective datasets. All timings were obtained on a Macbook Pro 2016 
-  running \Sexpr{session}. 
+  The double EM loop runs the danger of excessive computation time. We have   
+  compared computation times to the
+  other methods introduced in Sections \ref{sec:colorectal}--
+  \ref{sec:lymphnode}. All timings were obtained on a Macbook 
+  Pro 2016 running \Sexpr{session}. 
   The timings of \texttt{gren} are competitive. On the colorectal cancer example
   we achieved the following times:
-  \Sexpr{times1[c(1:3)]} seconds on the colorectal cancer example, 
-  with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively. \texttt{GRridge} is 
+  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
+  respectively. \texttt{GRridge} is 
   slightly faster with \Sexpr{times1[4]} seconds, while \texttt{SGL} was slightly
@@
   (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
-  On the lymph node cancer example \texttt{gren} achieved 
-  \Sexpr{times2[c(1:3)]}, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
+  The results for the lymph node cancer example are similar. \texttt{gren} 
+  achieved 
+  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
+  respectively.
   \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
@@
   $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
-  \texttt{cMCP} (\Sexpr{times2[c(8:10)]} seconds) and \texttt{gel} 
-  (\Sexpr{times2[c(11:13)]} seconds) are faster than \texttt{gren}. The 
-  overlapping group methods achieved \Sexpr{times2[c(14:16)]} seconds
-  (\texttt{ocMCP}) and \Sexpr{times2[c(17:19)]} seconds (\texttt{ogel}), for 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
-  We note that a considerable speedup may be achieved by 
-  doing one EB update for every few VB updates, instead of running the VB EM 
+  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
+  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}. 
+  We note that a considerable speedup may be achieved by iterating between 
+  single EB and single VB updates, instead of running the VB EM 
   until convergence. This works in practice, because an EM tends to take a few 
   initial large steps towards the optimum, followed by a series of smaller 
-  steps. We argue that The first few iterations of the VB EM render the 
-  approximate expectation in [! equation ref here] accurate enough for a 
-  reliable EB step. In addition, this is the approach often taken by more
-  computationally oriented disciplines. For example, in many machine learning 
-  domains, the problem is viewed as purely algorithmic [! references here]. The
-  the EB and VB 
-  steps optimisation steps are interpreted as steps along the gradient in the 
-  larger combined space of model- and hyper-parameters.
+  steps. We argue that the first few iterations of the VB EM render the 
+  approximate expectation in \ref{eq:approximatedloglikelihood}
+  accurate enough for a 
+  reliable EB step. This approach is often used in more 
+  computational disciplines, where the problem is viewed as purely algorithmic 
+  and the EB and VB iterations are interpreted as optimisation steps along the 
+  gradient in the larger combined model- and hyper-parameters space.
 

### 2019-03-19 code/simulations.R

@@ -50,14 +50,8 @@ part <- rep(c(1:G), each=p/G)
 csel <- 2^c(1:8)
 nreps <- 100
 
-################################################################################
-## Error in { :                                                                #
-##     task 1 failed - ""logitNest" not resolved from current namespace (SGL)" #
-################################################################################
-
 ### analysis splits in parallel
-ncores <- 50
-# ncores <- min(detectCores() - 1, nreps)
+ncores <- min(detectCores() - 1, nreps)
 cluster <- makeForkCluster(ncores)
 if(parallel) {
   registerDoParallel(cluster)
@@ -169,53 +163,157 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
 if(parallel) {stopCluster(cluster)}
 save(res, file="results/simulations_res1.Rdata")
 
-################################## DEBUGGING ###################################
-load("results/simulations_res1.Rdata")
-test1 <- res
-test2 <- sapply(c("psel", "auc", "briers", "mse", "kappa", "mults"), 
-                function(s) {
-                  sapply(test1, function(m) {
-                    m[[s]]}, simplify=FALSE)}, simplify=FALSE)
-
-test3 <- sapply(c(paste0("gren", 1:3), paste0("enet", 1:3),
-                  paste0("cmcp", 1:3), paste0("gel", 1:3)), function(m) {
-                    cbind(rowMeans(sapply(1:length(test2$psel), function(s) {
-                      vec1 <- rep(NA, 100);
-                      vec2 <- test2$psel[[s]][grep(m, names(test2$psel[[s]]))]
-                      replace(vec1, 1:length(vec2), vec2)}), na.rm=TRUE),
-                      rowMeans(sapply(1:length(test2$auc), function(s) {
-                        vec1 <- rep(NA, 100);
-                        vec2 <- test2$auc[[s]][grep(m, names(test2$auc[[s]]))]
-                        replace(vec1, 1:length(vec2), vec2)}), na.rm=TRUE))},
-                simplify=FALSE)
-
-grep("cmcp1", test1[[1]]$auc)
-
[…]

### 2019-03-19 rpackage/R/predict.gren.R

@@ -1,6 +1,6 @@
 # function to make predictions with the gren object
 predict.gren <- function(object, newx, unpenalized=NULL, s=NULL, 
-                         type=c("groupreg", "regular")) {
+                         type=c("groupreg", "regular"), ...) {
   if(class(object)!="gren") {
     stop("object should be a gren fit")
   } else if(!is.null(s) & !is.numeric(s)) {

### 2019-03-19 docs/supplement.Rnw
fixed all cross-document references

@@
 \usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
-rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
+rotating,caption,subcaption,natbib,appendix,titlesec,multicol,verbatim,
 bbm,algorithm,algpseudocode,threeparttable,booktabs,mathtools,
-dsfont,parskip}
+dsfont,parskip,xr-hyper,hyperref}
+\externaldocument[md-]{manuscript}
 
@@
 
-  \section{Introduction}
+  \section{Introduction}\label{sec:introduction}
   This document contains supplementary material to the paper `Adaptive 
@@
   $$
-\log q^*_{\bbeta} (\bbeta) = \bkappa \tr \X \bbeta - \frac{1}{2} \bbeta \tr (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z) \bbeta,
-$$
-which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma \X \tr \bkappa$.
+  \log q^*_{\bbeta} (\bbeta) = \bkappa \tr \X \bbeta - \frac{1}{2} \bbeta \tr 
+  (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z) \bbeta,
+  $$
+  which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a 
+  Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + 
+  \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma 
+  \X \tr \bkappa$.
 
@@
   \begin{align*}
-\log q^*_{\bomega} (\bomega) & = \E_{\bbeta} [\log \pi (\bomega | \bbeta)] = \sum_{i=1}^n \E_{\bbeta} [\log \pi (\omega_i | \bbeta)] \\ & = \sum_{i=1}^n \E_{\bbeta} \( m_i \log \left\{ \cosh \( \frac{\x_i \tr \bbeta}{2}\) \exp\[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i\] \cdot f(\omega_i | m_i, 0) \right\} \) \\
-& = \sum_{i=1}^n \E_{\bbeta} \[ m_i \log \cosh \( \frac{\x_i \tr \bbeta}{2}\) \] - \frac{1}{2} \sum_{i=1}^n \omega_i \E[ (\x_i \tr \bbeta)^2] + \sum_{i=1}^n \log f(\omega_i | m_i, 0) \\ 
-& \propto \sum_{i=1}^n m_i \log \cosh \left\{ \frac{\sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}}{2}\right\} - \frac{1}{2} \sum_{i=1}^n \omega_i \left\{ [\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \right\} \\
+    \log q^*_{\bomega} (\bomega) & = \E_{\bbeta} [\log \pi (\bomega | \bbeta)] =
+    \sum_{i=1}^n \E_{\bbeta} [\log \pi (\omega_i | \bbeta)] \\ & = \sum_{i=1}^n
+    \E_{\bbeta} \( m_i \log \left\{ \cosh \( \frac{\x_i \tr \bbeta}{2}\) 
+    \exp\[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i\] \cdot f(\omega_i | m_i, 0) 
+    \right\} \) \\ 
+    & = \sum_{i=1}^n \E_{\bbeta} \[ m_i \log \cosh 
+    \( \frac{\x_i \tr \bbeta}{2}\) \] - \frac{1}{2} \sum_{i=1}^n \omega_i 
+    \E[ (\x_i \tr \bbeta)^2] + \sum_{i=1}^n \log f(\omega_i | m_i, 0) \\ 
+    & \propto \sum_{i=1}^n m_i \log \cosh \left\{ \frac{\sqrt{[\x_i \tr 
+    \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}}{2}\right\} - 
+    \frac{1}{2} \sum_{i=1}^n \omega_i \left\{ [\x_i \tr \E(\bbeta)]^2 + 
+    \x_i \tr \V[\bbeta] \right\} \\ 
     & \,\,\, + \sum_{i=1}^n \log f(\omega_i | m_i, 0),
   \end{align*}
-where proportionality is with respect to $\bomega$.
-If exponentiated, this is a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.
-
-The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated as:
-\begin{align*}
-\E_{\bbeta} [\log \pi ( \bbeta | \btau)] & = \sum_{j=1}^p \E_{\bbeta_j} [\log \pi(\beta_j | \tau_j )] \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}, \\
-\log \pi (\btau) & = \sum_{j=1}^p \log \pi (\tau_j) \propto - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j. 
+  where proportionality is with respect to $\bomega$. If exponentiated, this is 
+  a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables
+  \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr 
+  \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.
+
+  The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated 
+  as:
+  \begin{align*}
+    \E_{\bbeta} [\log \pi ( \bbeta | \btau)] & = \sum_{j=1}^p \E_{\bbeta_j} 
+    [\log \pi(\beta_j | \tau_j )] \propto \frac{1}{2} \sum_{j=1}^p \log 
+    \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2)
+    \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}, \\
+    \log \pi (\btau) & = \sum_{j=1}^p \log \pi (\tau_j) \propto - 
+    \frac{1}{2} \sum_{j=1}^p \log \tau_j - 
+    \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j. 
   \end{align*}
@@
   \begin{align*}
-\log q^*_{\tau} (\btau) & \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}  - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j \\
-& = \frac{1}{2} \sum_{j=1}^p \log [(\tau_j - 1)^{-1}] - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2}\tau_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}\),
-\end{align*}
-where proportionality is with respect to $\btau$. Again applying the change of variables $\psi_j = \tau_j - 1$, we have:
-$$
-\log q^*_{\psi} (\psi) \propto \frac{1}{2} \sum_{j=1}^p \log \psi_j^{-1} - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2} \psi_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \psi_j^{-1}\),
-$$
-which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$ random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + \V(\beta_j)]$.
+    \log q^*_{\tau} (\btau) & \propto \frac{1}{2} \sum_{j=1}^p 
+    \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p 
+    \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}  - 
+    \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p 
+    \frac{\lambda_1^2}{8 \lambda_2} \tau_j \\
+    & = \frac{1}{2} \sum_{j=1}^p \log [(\tau_j - 1)^{-1}] - 
+    \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2}\tau_j + 
+    [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} 
+    \frac{\tau_j}{\tau_j - 1}\),
+  \end{align*}
+  where proportionality is with respect to $\btau$. Again applying the change of
+  variables $\psi_j = \tau_j - 1$, we have:
+  $$
+  \log q^*_{\psi} (\psi) \propto \frac{1}{2} \sum_{j=1}^p \log \psi_j^{-1} -
+  \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2} \psi_j + 
+  [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \psi_j^{-1}\),
+  $$
+  which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$
+  random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + 
+  \V(\beta_j)]$.
 
   \section{Computational complexity}\label{sec:computationalcomplexity}
-Calculation of $\bSigma$ involves the inversion of a $(p \times p)$-dimensional matrix, which occurs at every iteration of the algorithm.
-If naively done, every inversion is of computational complexity $\mathcal{O}(p^3)$.
-With $p$ large, this is a serious computational bottleneck of the algorithm, especially with possibly many iterations.
-However, inspection of the updating equations in (3.9) in the main document, reveals that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the $c_i$ and $\chi_j$.
-In Section 2.2 of the main document we show that for the EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, and the diagonal of $\bSigma$.
-Here we provide details on how to compute these with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the computational cost of the algorithm.
-If necessary, we may calculate the full $\bSigma$ once after the algorithm has converged.
-
-In the following we slightly change notation and let $\diag(\A)$, with $\A$ a $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$ to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$.
-We start by calculating the complexity of $\diag(\bSigma)$.
-To this end, let $\H = \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z$.
-Now, by the Woodbury identity, we may decompose $\diag(\bSigma)$ as:
+  Calculation of $\bSigma$ involves the inversion of a 
+  $(p \times p)$-dimensional matrix, which occurs at every iteration of the
+  algorithm. If naively done, every inversion is of computational complexity
+  $\mathcal{O}(p^3)$. With $p$ large, this is a serious computational bottleneck
+  of the algorithm, especially with possibly many iterations. However, 
+  inspection of the updating equations in (3.9) in the main document, reveals
+  that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i 
+  \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the
+  $c_i$ and $\chi_j$. In Section 2.2 of the main document we show that for the 
+  EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, 
+  and the diagonal of $\bSigma$. Here we provide details on how to compute these
+  with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the
+  computational cost of the algorithm. If necessary, we may calculate the full
+  $\bSigma$ once after the algorithm has converged.
+
+  In the following we slightly change notation and let $\diag(\A)$, with $\A$ a
+  $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$
+  to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} 
+  \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$. We start by calculating the
+  complexity of $\diag(\bSigma)$. To this end, let $\H = \lambda_2 \bLambda' +
+  \lambda_2 \bLambda' \Z$. Now, by the Woodbury identity, we may decompose 
+  $\diag(\bSigma)$ as:
   \begin{subequations}\label{eq:diagsigma}
 	  \begin{align}
-	\diag(\bSigma) ~& = \diag(\H^{-1}) - \diag[\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1}] \\
-	& = \diag(\H^{-1}) - [(\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1}) \circ \H^{-1} \X \tr] \cdot \ones_{n \times 1},
+	    \diag(\bSigma) ~& = \diag(\H^{-1}) - \diag[\H^{-1} \X \tr \bOmega 
+	    (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1}] \\
+	    & = \diag(\H^{-1}) - [(\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr 
+	    \bOmega)^{-1}) \circ \H^{-1} \X \tr] \cdot \ones_{n \times 1},
 	  \end{align}
   \end{subequations}
-where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is a $p \times n$-dimensional matrix of ones.
-By recognising that $\H$ is a diagonal matrix, we can post-multiply it by a $p \times n$-dimensional matrix with computational complexity $\mathcal{O}(np)$, which leads to a total complexity of $\mathcal{O}(n^2p)$ for $\diag(\bSigma)$.
-Next we consider the calculation of $\bmu$.
-Again using the Woodbury identity, we may write $\bmu$ as:
+  where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is
+  a $p \times n$-dimensional matrix of ones. By recognising that $\H$ is a
+  diagonal matrix, we can post-multiply it by a $p \times n$-dimensional matrix
+  with computational complexity $\mathcal{O}(np)$, which leads to a total
+  complexity of $\mathcal{O}(n^2p)$ for $\diag(\bSigma)$. Next we consider the
+  calculation of $\bmu$. Again using the Woodbury identity, we may write $\bmu$
+  as:
   \begin{align}\label{eq:mu}
-\bmu = \bSigma \X \tr \bkappa = [\H^{-1} \X \tr - \H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1} \X \tr] \bkappa,
-\end{align}
-which is, by a similar argument as before, again an operation of complexity $\mathcal{O}(n^2p)$.
-
-Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$ as:
-\begin{align*}
-\c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
-& = \sqrt{\diag \[\X \bSigma \X \tr\] +  \diag \[\X \bmu \bmu \tr \X \tr\]} \\
-& = \sqrt{ [(\X \bSigma) \circ \X ] \cdot \ones_{p \times 1} + (\X \bmu) \circ (\X \bmu)}.
-\end{align*}
-The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$, Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}(n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.
-
-The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p \end{bmatrix} \tr$ are rewritten as:
+    \bmu = \bSigma \X \tr \bkappa = [\H^{-1} \X \tr - \H^{-1} \X \tr \bOmega 
+    (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1} \X \tr] \bkappa,
+  \end{align}
+  which is, by a similar argument as before, again an operation of complexity
+  $\mathcal{O}(n^2p)$.
+
+  Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n
+  \end{bmatrix} \tr$ as:
+  \begin{align*}
+    \c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & 
+    \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
+    & = \sqrt{\diag \[\X \bSigma \X \tr\] +  
+    \diag \[\X \bmu \bmu \tr \X \tr\]} \\
+    & = \sqrt{ [(\X \bSigma) \circ \X ] \cdot \ones_{p \times 1} + 
+    (\X \bmu) \circ (\X \bmu)}.
+  \end{align*}
+  The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$,
+  Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}
+  (n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.
+
+  The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p 
+  \end{bmatrix} \tr$ are rewritten as:
   \begin{subequations}\label{eq:chicomp}
 	  \begin{align}
-	\bm{\chi} & = \begin{bmatrix} \lambda'_{g(1)} \lambda_2 (\bSigma_{11} + \bmu_1^2) & \cdots & \lambda'_{g(p)} \lambda_2 (\bSigma_{pp} + \bmu_p^2) \end{bmatrix} \tr\\
+	    \bm{\chi} & = \begin{bmatrix} \lambda'_{g(1)} \lambda_2 (\bSigma_{11} +
+	    \bmu_1^2) & \cdots & \lambda'_{g(p)} \lambda_2 (\bSigma_{pp} + \bmu_p^2)
+	    \end{bmatrix} \tr\\
 	    & = \lambda_2 \bLambda' \[ \diag (\bSigma ) + \diag (\bmu \bmu \tr )\] \\
@@
   \end{subequations}
- By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we see that the computation of $\bm{\chi}$ consists of one matrix diagonal of $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two Hadamard products of $\mathcal{O}(p)$.
-Computation of $\bm{\chi}$ is therefore of complexity of $\mathcal{O}(n^2p)$.
-
-\section{Variational Bayes expected joint log likelihood}\label{sec:expectedloglikelihood}
-Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then the variational Bayes expected joint log likelihood is derived as follows:
-\begin{align*}
-\E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ] &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}] \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}(\beta_j | \tau_j) | \blambda'^{(k)}] \\
-& \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\
-& \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
+  By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we
+  see that the computation of $\bm{\chi}$ consists of one matrix diagonal of
+  $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two
+  Hadamard products of $\mathcal{O}(p)$. Computation of $\bm{\chi}$ is therefore
+  of complexity of $\mathcal{O}(n^2p)$.
+
+  \section{Variational Bayes expected joint log likelihood}
+  \label{sec:expectedloglikelihood}
+  Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then
+  the variational Bayes expected joint log likelihood is derived as follows:
+  \begin{align*}
+    \E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ]
+    &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}]
+    \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}
+    (\beta_j | \tau_j) | \blambda'^{(k)}] \\
+    & \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  
+    \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G 
+    \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1}
+    \beta_j^2 | \blambda'^{(k)} \) \\ 
+    & \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - 
+    \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
   \end{align*}
@@
   \begin{align*}
-d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\
-& = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} ( \beta_j | \blambda'^{(k)} ) \] \\
-& = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
+    d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1}
+    \beta_j^2 | \blambda'^{(k)} \) \\
+    & = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} 
+    ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} 
+    ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} 
+    ( \beta_j | \blambda'^{(k)} ) \] \\
+    & = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + 
+    (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} 
+    \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
   \end{align*}
@@
   \section{Full estimation procedure}\label{sec:fullestimationprocedure}
-The expectation and maximisation steps introduced in the Main Document result in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
+  The expectation and maximisation steps introduced in the Main Document result
+  in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
   \begin{algorithm}
-	\caption{Group-regularized empirical Bayes elastic net}\label{alg:summaryalg}
+    \caption{Group-regularized empirical Bayes elastic net}
+    \label{alg:summaryalg}
 	  \begin{algorithmic}[h]
@@
 		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
-		\State{Estimate $\lambda$ by cross-validation of the regular elastic net model}
-		\While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| > \epsilon_1$}
-		\While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
-		\State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ for $j=1, \dots, p$ using (3.9) in the Main Document}
+		  \State{Estimate $\lambda$ by cross-validation of the regular elastic net
+		  model}
+		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
+		  \epsilon_1$}
+		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
+		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
+		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
+		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
+		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
+		  for $j=1, \dots, p$ using (3.9) in the Main Document}
 		  \EndWhile
@@
   \subsection{Unpenalized covariates}
-Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix $\X$ with a column of ones.
-Penalization of such an intercept parameter is not desirable \cite[]{Hastie_Elements_2009}.
-Additionally, it is often desirable to include unpenalized covariates in the model.
-Common examples of such  covariates in clinical research are patient characteristics, such as age, BMI, and sex.
-In the following we assume the intercept to be included in the unpenalized covariates.
-
-To include unpenalized covariates, we divide the model matrix into two parts $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$ unpenalized variables and $\X_r$ are the $r$ penalized variables.
-Let $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with $u$ zero columns and $u$ zero rows.
-Then we have for the current estimate of the covariance matrix $\bSigma$:
+  Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix
+  $\X$ with a column of ones. Penalization of such an intercept parameter is not
+  desirable \cite[]{Hastie_Elements_2009}. Additionally, it is often desirable 
+  to include unpenalized covariates in the model. Common examples of such 
+  covariates in clinical research are patient characteristics, such as age, BMI,
+  and sex. In the following we assume the intercept to be included in the
+  unpenalized covariates.
+
+  To include unpenalized covariates, we divide the model matrix into two parts
+  $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$
+  unpenalized variables and $\X_r$ are the $r$ penalized variables. Let
+  $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with
+  $u$ zero columns and $u$ zero rows. Then we have for the current estimate of
+  the covariance matrix $\bSigma$:
   \begin{align}\label{eq:unpencov}
-\bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
+    \bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + 
+    \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
     & = 
@@
       \X_u \tr \bOmega \X_u & \X_u \tr \bOmega \X_r \\
-\X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
+      \X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' +
+      \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
     \end{bmatrix}^{-1}.
   \end{align}
-With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension $p \times p$.
-Inversion of this matrix is done efficiently by applying the Woodbury identity.
+  With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion
+  renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r +
+  \lambda_2 \bLambda' + 
+  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension 
+  $p \times p$. Inversion of this matrix is done efficiently by applying the
+  Woodbury identity.
 
   \subsection{Monotonicity of the penalty parameters}
-Enforcing monotonicity of the penalty multipliers is desirable in some settings.
-An example of such a setting is if we have $p$-values from a previous, related study available.
-A $p$-values-based partitioning of the variables may, \textit{a priori}, be expected to render penalty multipliers that increase monotonically with $p$-value.
-That is, larger $p$-values may be expected to yield at least as large penalty multipliers as variables with smaller $p$-values.
-We propose to include this \textit{a priori} assumption by requiring the penalty multipliers to increase monotonically with $p$-value, thereby also stabilizing their estimates.
-
-A natural way of enforcing monotonicity is to extend the constraint in (10) in the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$.
-The problem is still convex and may be numerically solved.
-From experience, however, we note that in combination with this constrained optimisation, the EM algorithm described in Section~3 in the Main Document often converges to a local optimum close to the initialisation.
-We therefore enforce monotonicity through a \textit{post hoc} isotonic regression on the penalty multipliers after every optimisation step.
+  Enforcing monotonicity of the penalty multipliers is desirable in some
+  settings. An example of such a setting is if we have $p$-values from a
+  previous, related study available. A $p$-values-based partitioning of the
+  variables may, \textit{a priori}, be expected to render penalty multipliers
+  that increase monotonically with $p$-value. That is, larger $p$-values may be
+  expected to yield at least as large penalty multipliers as variables with
+  smaller $p$-values. We propose to include this \textit{a priori} assumption by
+  requiring the penalty multipliers to increase monotonically with $p$-value,
+  thereby also stabilizing their estimates.
+
+  A natural way of enforcing monotonicity is to extend the constraint in (10) in
+  the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$. The problem is
+  still convex and may be numerically solved. From experience, however, we note
+  that in combination with this constrained optimisation, the EM algorithm
+  described in Section~3 in the Main Document often converges to a local optimum
+  close to the initialisation. We therefore enforce monotonicity through a
+  \textit{post hoc} isotonic regression on the penalty multipliers after every
+  optimisation step.
 
   \subsection{Multiple partitions}
-In many cases the features may be partitioned in more than one way.
-For example, we may have both information on annotation and $p$-values from a previous study available.
-A naive way of incorporating multiple partitions is to cross-tabulate the partitions and create a separate group for every combination.
-This poses two problems: (i) The number of penalty parameters increases exponentially with the number of partitions and (ii) some of these combinations may contain only few features, so that the estimation procedure becomes unstable.
-We propose to stabilise the procedure and keep the number of parameters to estimate manageable by modelling the penalty parameters multiplicatively.
-
-We describe our implementation here for two partitions of the features.
-To this end, let $(\G_1(1), \dots,\allowbreak \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively.
-Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively.
-In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively.
-The generalised frequentist elastic net estimator and corresponding conditional prior are now:
+  In many cases the features may be partitioned in more than one way. For
+  example, we may have both information on annotation and $p$-values from a
+  previous study available. A naive way of incorporating multiple partitions is
+  to cross-tabulate the partitions and create a separate group for every
+  combination. This poses two problems: (i) The number of penalty parameters
+  increases exponentially with the number of partitions and (ii) some of these
+  combinations may contain only few features, so that the estimation procedure
+  becomes unstable. We propose to stabilise the procedure and keep the number of
+  parameters to estimate manageable by modelling the penalty parameters
+  multiplicatively.
+
+  We describe our implementation here for two partitions of the features. To
+  this end, let $(\G_1(1), \dots,\allowbreak \G_1(G_1))$ and $(\G_2(1), \dots,
+  \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups
+  respectively. Furthermore, in the following we assume that the empty sum and
+  empty product evaluate to 0 and 1, respectively. In this two-partition setting
+  we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$
+  and $\lambda''_{g_2}$, respectively. The generalised frequentist elastic net
+  estimator and corresponding conditional prior are now:
   \begin{subequations}\label{eq:multpart}
 	  \begin{align}
-	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
-	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
+	    \hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  
+	    \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} 
+	    \sqrt{\lambda'_{g_1} \lambda''_{g_2}} 
+	    \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j|
+	    - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1}
+	    \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ 
+	    \cap \G_2(g_2)}}} \beta_j^2, \\
+	    \bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} 
+	    \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} 
+	    \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2}
+	    \frac{\tau_j - 1}{\tau_j} \right).
 	  \end{align}
   \end{subequations}
-After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
+  After switching to the parametrisation in \cite{friedman_regularization_2010},
+  the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 &
+  \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix}
+  \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
   \begin{subequations}\label{eq:estmultpart}
 	  \begin{alignat}{2}
-	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
+	    \blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}
+	    {\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| 
+	    \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| 
+	    \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} 
+	    \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} 
+	    d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       
+	    & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} 
+	    (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
 	  \end{alignat}
   \end{subequations}
-with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$.
-The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine.
-This method naturally generalises to more than two partitions of the data.
-
-In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers in a multiplicative way, mainly for convenience.
-First, there are computational reasons: The separation of the square root in the $L_1$-norm penalty term into two terms facilitates the numerical estimation.
-Additionally, multiplicative modelling of the partitions allows for more flexibility in the estimates.
-To see this consider the obvious alternative: additive partition-specific multipliers.
-Since the penalty multipliers are strictly positive, the constraint on the geometric mean makes one large additive penalty estimate difficult to compensate with a smaller estimate, thereby impairing flexibility of the model.
-Large multiplicative penalty multipliers are easier to compensate for by smaller penalty multipliers, rendering this the less restrictive option for modelling the partitions.
-
-  \section{Simulations}
-The quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960}.
-This kappa is measures similarity between two sets of categorical values.
-One set contains the feature selection indicators: $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
+  with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\
+  \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha
+  \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. The optimisation in
+  (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some
+  numerical optimisation routine. This method naturally generalises to more than
+  two partitions of the data.
+
+  In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers
+  in a multiplicative way, mainly for convenience. First, there are 
+  computational reasons: The separation of the square root in the $L_1$-norm
+  penalty term into two terms facilitates the numerical estimation. 
+  Additionally, multiplicative modelling of the partitions allows for more
+  flexibility in the estimates. To see this consider the obvious alternative:
+  additive partition-specific multipliers. Since the penalty multipliers are
+  strictly positive, the constraint on the geometric mean makes one large
+  additive penalty estimate difficult to compensate with a smaller estimate,
+  thereby impairing flexibility of the model. Large multiplicative penalty
+  multipliers are easier to compensate for by smaller penalty multipliers,
+  rendering this the less restrictive option for modelling the partitions.
+
+  \section{Simulations}\label{sec:simulations}
+  The quality of feature selection is assessed by Cohen's kappa 
+  \cite[]{cohen_coefficient_1960}. This kappa is measures similarity between two
+  sets of categorical values. One set contains the feature selection indicators:
+  $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains
+  the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
   We calculate Cohen's $\kappa$ as:
 	\begin{align*}
-	\kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}\sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
-	\end{align*}
-	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and expected correctly identified features, respectively.
-A positive or negative kappa indicate better and worse feature selection than expected by chance, respectively.
-In addition, we evaluate estimation accuracy by mean squared error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
+	  \kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}
+	  \sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} 
+	  \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + 
+	  \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
+	\end{align*}
+	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and
+	expected correctly identified features, respectively. A positive or negative
+	kappa indicate better and worse feature selection than expected by chance,
+	respectively. In addition, we evaluate estimation accuracy by mean squared
+	error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
 
   \subsection{Scenario 1}
-	In this sparse scenario we assume differential signal between 5 groups of features. The average signal in a group increases by a factor 1.6 with group number:
-	\begin{align*}
-	\bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & \beta_{50}^1 \end{matrix}}_{50}} & \cdots & \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr, \\
+	In this sparse scenario we assume differential signal between 5 groups of
+	features. The average signal in a group increases by a factor 1.6 with group
+	number:
+	\begin{align*}
+	  \bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} 
+	  \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} 
+	  & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & 
+	  \beta_{50}^1 \end{matrix}}_{50}} & \cdots & 
+	  \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0}
+	  \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} 
+	  \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr,
+	  \\
 	  & \, \\
-	\beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots, 5  \text{, } j=1, \dots, 50,
-	\end{align*}
-	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$.
-The model parameters clearly do not follow the relatively heavy-tailed elastic net distribution.
-Because of this mismatch between the assumed and true distributions, we expect poor performance by \texttt{gren}.
-Furthermore, we introduce correlation between the features by having covariance elements $\bSigma_{ij} = 0.5^{|i-j|}$.
+	  \beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots,
+	  5  \text{, } j=1, \dots, 50,
+	\end{align*}
+	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. The
+	model parameters clearly do not follow the relatively heavy-tailed elastic net
+	distribution. Because of this mismatch between the assumed and true 
+	distributions, we expect poor performance by \texttt{gren}. Furthermore, we
+	introduce correlation between the features by having covariance elements
+	$\bSigma_{ij} = 0.5^{|i-j|}$.
 
 	\subsection{Scenario 2}
-	The second scenario contains many groups of features and is geared towards the non-adaptive group-regularized methods.
-These methods are able to incorporate the grouping of the features, but were originally designed for small interpretable groups, such as dummies of a categorical variable.
-\texttt{gren} on the other hand, relies on large groups of features to accurately estimate group-specific penalty weights.
-The estimation of many penalty weights in small groups is inefficient and computationally challenging.
-We therefore expect \texttt{gren} to under-perform compared to the other methods.
-
-	For this scenario we follow the second simulation example in \cite{meier_group_2008}.
-They set $n=100$, $G=250$, and $p=1000$, such that we have 4 features per group.
-The first 10 groups are active and those 40 model parameters are set to 0.2.
-The predictors are simulated from a $p$-dimensional Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.
+	The second scenario contains many groups of features and is geared towards the
+	non-adaptive group-regularized methods. These methods are able to incorporate
+	the grouping of the features, but were originally designed for small 
+	interpretable groups, such as dummies of a categorical variable. \texttt{gren}
+	on the other hand, relies on large groups of features to accurately estimate
+	group-specific penalty weights. The estimation of many penalty weights in
+	small groups is inefficient and computationally challenging. We therefore
+	expect \texttt{gren} to under-perform compared to the other methods.
+
+	For this scenario we follow the second simulation example in 
+	\cite{meier_group_2008}. They set $n=100$, $G=250$, and $p=1000$, such that we
+	have 4 features per group. The first 10 groups are active and those 40 model
+	parameters are set to 0.2. The predictors are simulated from a $p$-dimensional
+	Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.
 
 	\subsection{Scenario 3}
-	A third scenario is modelled by considering a situation in which there is no differential signal between the groups, but the correlation within groups is very high.
-We conjecture that this leads to overestimation of effect within the groups by \texttt{gren} and consequently, performance will suffer.
-
-	To that end we sample the predictors in $G=5$ separate blocks of 200 variables, corresponding to the partitioning.
-The variances of the features are set to one, while all covariances are 0.7.
-The 1000 model parameters are simulated in two steps.
-First, we draw 1000 parameters from the elastic net distribution, parametrised as in \cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$.
-The second step consists of setting the (in absolute value) smallest 100 parameters in a group to zero.
-This results in a total of 500 zero and 500 non-zero parameters, evenly distributed over the five groups.
+	A third scenario is modelled by considering a situation in which there is no
+	differential signal between the groups, but the correlation within groups is
+	very high. We conjecture that this leads to overestimation of effect within
+	the groups by \texttt{gren} and consequently, performance will suffer.
+
+	To that end we sample the predictors in $G=5$ separate blocks of 200 
+	variables, corresponding to the partitioning. The variances of the features
+	are set to one, while all covariances are 0.7. The 1000 model parameters are
+	simulated in two steps. First, we draw 1000 parameters from the elastic net
+	distribution, parametrised as in \cite{friedman_regularization_2010}, where
+	$\alpha=0.5$ and $\lambda=100$. The second step consists of setting the 
+	(in absolute value) smallest 100 parameters in a group to zero. This results
+	in a total of 500 zero and 500 non-zero parameters, evenly distributed over
+	the five groups.
 
 	\subsection{Scenario 4}
-	This scenario is similar to Scenario 3, except each groups penalty parameter is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the covariance matrix of the predictors is block diagonal, with block sizes $25 \times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$.
-We expect \texttt{gren} to pick up this differential signal and outperform the other methods.
+	This scenario is similar to Scenario 3, except each groups penalty parameter
+	is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the
+	covariance matrix of the predictors is block diagonal, with block sizes $25
+	\times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$.
+	We expect \texttt{gren} to pick up this differential signal and outperform the
+	other methods.
 
 	\subsection{Scenario 5}
-	In the last scenario we create a sparse true model that contains some differential signal. We impl
+	In the last scenario we create a sparse true model that contains some 
+	differential signal. We impl
 
@@
   \section{Application to RNAseq in lymph node cancer}
+  \label{sec:lymphnode}
   In addition to the results presented in Section [! section ref here] of the 
@@
 
-  \section{Application to microRNAs in cervical cancer}
+  \section{Application to microRNAs in cervical cancer}\label{sec:cervical}
   A deep sequencing analysis on small non-coding ribonucleic acid (microRNAseq) 
@@
 
-\section{Empirical-variational Bayes for the ridge and lasso}\label{sec:evbridgelasso}
+  \section{Empirical-variational Bayes for the ridge and lasso}
+  \label{sec:evbridgelasso}
   \subsection{Ridge regression}
-Ridge regression is a special case of the elastic net, that does not include an $L_1$-norm on the parameters. Its estimator is given by:
-$$
-\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_2}{2} \norm{\bbeta}_2.
-$$
-The Bayesian version of ridge regression is regression under a Gaussian prior. The generalised version of this prior is:
-$$
-\bbeta \sim \prod_{g=1}^G \prod_{j \in \G(g)} \N(0,(\lambda_2 \lambda'_g)^{-1}).
-$$
-Although the posterior is analytically calculated in the linear case, in the logistic case we must approximate it iteratively. The variational distributions in this case are:
+  Ridge regression is a special case of the elastic net, that does not include
+  an $L_1$-norm on the parameters. Its estimator is given by:
+  $$
+  \hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+  \frac{\lambda_2}{2} \norm{\bbeta}_2.
+  $$
+  The Bayesian version of ridge regression is regression under a Gaussian prior.
+  The generalised version of this prior is:
+  $$
+  \bbeta \sim \prod_{g=1}^G \prod_{j \in \G(g)} 
+  \N(0,(\lambda_2 \lambda'_g)^{-1}).
+  $$
+  Although the posterior is analytically calculated in the linear case, in the
+  logistic case we must approximate it iteratively. The variational 
+  distributions in this case are:
   \begin{align*}
@@
   \end{align*}
-The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net and are updated, together with $\bSigma$, until convergence:
-$$
-\bSigma^{t+1} = (\X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda')^{-1}\text{, with } \bOmega^{(t)} = \diag \[ \frac{m_i}{2 c^{(t)}_i} \tanh \( \frac{c^{(t)}_i}{2} \) \].
-$$
-
-Updating the penalty parameters is done by solving the following (convex) constraint optimisation problem:
-\begin{align*}
-\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+  The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net
+  and are updated, together with $\bSigma$, until convergence:
+  $$
+  \bSigma^{t+1} = (\X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda')^{-1}
+  \text{, with } \bOmega^{(t)} = \diag \[ \frac{m_i}{2 c^{(t)}_i} 
+  \tanh \( \frac{c^{(t)}_i}{2} \) \].
+  $$
+
+  Updating the penalty parameters is done by solving the following (convex)
+  constraint optimisation problem:
+  \begin{align*}
+    \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G
+    |\G(g)| \log (\lambda'_g) - \frac{\lambda_2}{2} \sum_{g=1}^G 
+    \lambda'_g d^{(k)}_g \\
     & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
@@
   \subsection{Lasso regression}
-In lasso regression, another special case of the elastic net, we have the following estimator:
-$$
-\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1.
-$$
-The Bayesian counterpart is regression under a Laplace prior on the model parameters. The generalised Bayesian prior may decomposed as:
+  In lasso regression, another special case of the elastic net, we have the
+  following estimator:
+  $$
+  \hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+  \frac{\lambda_1}{2} \norm{\bbeta}_1.
+  $$
+  The Bayesian counterpart is regression under a Laplace prior on the model
+  parameters. The generalised Bayesian prior may decomposed as:
   \begin{align*}
     \bbeta | \btau & \sim \prod_{j=1}^p \mathcal{N} \left(0,\tau_j \right), \\
-\btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \text{Exp} \left( \frac{\sqrt{\lambda'_g}\lambda_1^2}{8} \right).
-\end{align*}
-
-The variational distributions are similar as in the elastic net, except that we do not re-parametrise the $\tau_j$:
+    \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \text{Exp} 
+    \left( \frac{\sqrt{\lambda'_g}\lambda_1^2}{8} \right).
+  \end{align*}
+
+  The variational distributions are similar as in the elastic net, except that
+  we do not re-parametrise the $\tau_j$:
   \begin{align*}
@@
     q^*_{\bomega} (\bomega) & \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i), \\
-q^*_{\btau}(\btau) & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{GIG} \(\frac{1}{2}, \phi_g, \chi_j\).
-\end{align*}
-The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic net, while the $\phi_g$ are given by:
+    q^*_{\btau}(\btau) & \sim \prod_{g=1}^G \prod_{j \in \G(g)} 
+    \mathcal{GIG} \(\frac{1}{2}, \phi_g, \chi_j\).
+  \end{align*}
+  The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic
+  net, while the $\phi_g$ are given by:
   $$
@@
   $$
-The following parameters are updated, together with the unchanged $c_i$ and $\bmu$, until convergence:
-\begin{align*}
-\bSigma^{(t + 1)} &= ( \X \tr \bOmega^{(t)} \X + \Z^{(t)})^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\(\sqrt{\frac{\phi_g(j)}{\chi_j^{(t)}}}\), \\
-\chi_j^{(t + 1)} &= \bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \text{, for } j=1, \dots, p.
-\end{align*}
-
-Updating the penalty parameters is done by solving the following (convex) constraint problem:
-\begin{align*}
-\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_1^2}{8} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+  The following parameters are updated, together with the unchanged $c_i$ and
+  $\bmu$, until convergence:
+  \begin{align*}
+    \bSigma^{(t + 1)} &= ( \X \tr \bOmega^{(t)} \X + \Z^{(t)})^{-1}, \\ 
+    & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) 
+    \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } 
+    \Z^{(t)}=\diag\(\sqrt{\frac{\phi_g(j)}{\chi_j^{(t)}}}\), \\
+    \chi_j^{(t + 1)} &= \bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 
+    \text{, for } j=1, \dots, p.
+  \end{align*}
+
+  Updating the penalty parameters is done by solving the following (convex)
+  constraint problem:
+  \begin{align*}
+    \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \sum_{g=1}^G |\G(g)|
+    \log (\lambda'_g) - \frac{\lambda_1^2}{8} \sum_{g=1}^G 
+    \lambda'_g d^{(k)}_g \\
     & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
@@
   $$
-d^{(k)}_g = \sum_{j \in \G(g)} \( \frac{1}{\phi_g^{(k)}} + \frac{\chi^{(k)}_j}{\phi_g^{(k)}} \).
-$$ 
-Note that the $\phi_g$ are updated after every penalty multiplier update iteration as well, in contrast to the elastic net.
+  d^{(k)}_g = \sum_{j \in \G(g)} \( \frac{1}{\phi_g^{(k)}} + 
+  \frac{\chi^{(k)}_j}{\phi_g^{(k)}} \).
+  $$ 
+  Note that the $\phi_g$ are updated after every penalty multiplier update  
+  iteration as well, in contrast to the elastic net.
 

### 2019-03-19 code/figures.R

@@ -1,4 +1,80 @@
 ################################## manuscript ##################################
+# ---- lines_simulations_res1_auc_briers ----
+library(sp)
+load(file="results/simulations_res1.Rdata")
+plot.data <- c(sapply(c("grridge", "ridge"), function(m) {
+  unname(cbind(NA, mean(sapply(res, function(s) {
+    s$auc[grep(m, names(s$auc))[1]]}), na.rm=TRUE),
+    mean(sapply(res, function(s) {
+      s$auc[grep(m, names(s$briers))[1]]}), na.rm=TRUE)))}, simplify=FALSE),
+  sapply(c(paste0("gren", 1:3), paste0("enet", 1:3),
+           paste0("cmcp", 1:3), paste0("gel", 1:3)), function(m) {
+             cbind(rowMeans(sapply(res, function(s) {
+               vec <- s$psel[grep(m, names(s$psel))]
+               replace(rep(NA, 8), 1:length(vec), vec)}), 
+               na.rm=TRUE),
+               rowMeans(sapply(res, function(s) {
+                 vec <- s$auc[grep(m, names(s$auc))]
+                 replace(rep(NA, 8), 1:length(vec), vec)}), 
+                 na.rm=TRUE),
+               rowMeans(sapply(res, function(s) {
+                 vec <- s$briers[grep(m, names(s$briers))]
+                 replace(rep(NA, 8), 1:length(vec), vec)}), 
+                 na.rm=TRUE))}, simplify=FALSE))
+
+methods <- c("GRridge", "ridge", "gren", "enet", "cmcp", "gel")
+labels <- c(methods, expression(alpha==0.05), expression(alpha==0.5), 
+            expression(alpha==0.95))
+
+col <- bpy.colors(length(methods), cutoff.tail=0.1)
+lty <- c(1:3)
+
+xlim <- range(sapply(plot.data[-c(1, 2)], function(s) {s[, 1]}), na.rm=TRUE)
+ylim1 <- range(sapply(plot.data, function(s) {s[, 2]}), na.rm=TRUE)
+ylim2 <- range(sapply(plot.data[-c(9, 10, 11)], function(s) {s[, 3]}), 
[…]

### 2019-03-19 rpackage/R/coef.gren.R

@@ -1,5 +1,5 @@
 # retrieve coefficients from gren estimated model
-coef.gren <- function(object, s=NULL, type=c("groupreg", "regular")) {
+coef.gren <- function(object, s=NULL, type=c("groupreg", "regular"), ...) {
   if(class(object)!="gren") {
     stop("object should be a gren fit") 
   } else if (!is.null(s) & !is.numeric(s)) {

### 2019-03-22 docs/supplement.Rnw
textual changes

@@
   function that is maximised in the variational Bayes procedure introduced in
-  Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on
-  the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main
-  Document. The derivations of the variational posterior used in the Main
-  Document are shown in Section \ref{sec:variationaldistributions}. Section
+  Section \ref{md-sec:estimation} of the Main Document (MD). Section
+  \ref{sec:polyagamma} gives details on
+  the P{\'o}lya-Gamma parametrisation introduced in Section 
+  \ref{md-sec:estimation} of the MD. The
+  derivations of the variational posterior used in the MD are shown in Section
+  \ref{sec:variationaldistributions}. Section
   \ref{sec:computationalcomplexity} shows how we may reduce the computational
@@
   joint log likelihood that is maximised to estimate the penalty multipliers as
-  explained in Section 3 of the Main Document. The full procedure of the
-  proposed method is shown in Section \ref{sec:fullestimationprocedure}. We
-  apply our method to a study on microRNAs in colorectal cancer data 
-  \cite[]{neerincx_combination_2018} in Section [! section ref here]. Finally, 
+  explained in Section \ref{md-sec:estimation} of the MD. The full procedure of
+  the
+  proposed method is shown in Section \ref{sec:fullestimationprocedure}. 
+  Additional results to the applications in the MD Sections 
+  \ref{md-sec:colorectal}--\ref{md-sec:lymphnode}. We
+  apply our method to a study on metabolomics in Alzheimer's in Section
+  \ref{sec:alzheimer} and to study in microRNAs in cervical cancer data 
+  in Section \ref{sec:cervical}. Finally, 
   we introduce the ridge and lasso versions of the proposed method in Section
@@
   After combining the terms in (\ref{eq:elboomega2}) and substituting the 
-  expectations and variances given in the Main Document, several terms cancel to
+  expectations and variances, several terms cancel to
   give:
@@
 
-  Inspection of the updating equations in the Main Document, learns us that the 
+  Inspection of the updating equations in the MD, learns us that the 
   last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
@@
   \section{The P\'{o}lya-Gamma parametrisation}\label{sec:polyagamma}
-  In Section 3 of the Main Document latent P\'{o}lya-Gamma distributed variables
+  In Section \ref{md-sec:estimation} of the MD latent P\'{o}lya-Gamma
+  distributed variables
   are introduced. Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma
@@
   of the algorithm, especially with possibly many iterations. However, 
-  inspection of the updating equations in (3.9) in the main document, reveals
+  inspection of the updating equations in (\ref{md-eq:VBupdateequations})
+  in the MD, reveals
   that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i 
   \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the
-  $c_i$ and $\chi_j$. In Section 2.2 of the main document we show that for the 
+  $c_i$ and $\chi_j$. In Section \ref{md-sec:estimation} of the MD we show that 
+  for the 
   EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, 
@@
   \section{Full estimation procedure}\label{sec:fullestimationprocedure}
-  The expectation and maximisation steps introduced in the Main Document result
+  The expectation and maximisation steps introduced in the Main Document Section
+  \ref{md-sec:estimation} result
   in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
@@
 		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
-		  for $j=1, \dots, p$ using (3.9) in the Main Document}
+		  for $j=1, \dots, p$ using (\ref{md-eq:VBupdateequations}) in the MD}
 		  \EndWhile
-		  \State{Update $\blambda'$ by (3.10) in the Main Document}
+		  \State{Update $\blambda'$ by (\ref{md-eq:mmlupdateequation}) in the MD}
 		  \EndWhile
@@
   that in combination with this constrained optimisation, the EM algorithm
-  described in Section~3 in the Main Document often converges to a local optimum
+  described in Section \ref{md-sec:estimation} in the MD often converges to a
+  local optimum
   close to the initialisation. We therefore enforce monotonicity through a
@@
   \subsection{Additional comparisons}
-  We compared \texttt{gren} to two methods in addition to the ones in the MD: 
+  We compared \texttt{gren} to two methods in addition to the ones in the MD
+  Section \ref{md-sec:colorectal}: 
   the sparse group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and 
@@
 
-  \subsection{Random groups}
+  \subsection{Random groups}\label{sec:randomgroups}
   Considering that the features may be partitioned into many groups, with one 
   parameter per group, we have to be aware of overfitting risks. We investigated
-  this using the data introduced in Section [! section ref here] of the MD, 
+  this using the data introduced in Section \ref{md-sec:colorectal} of the MD, 
   randomly dividing the features into three groups. We fixed the group sizes to 
@@
   \label{sec:lymphnode}
-  In addition to the results presented in Section [! section ref here] of the 
+  In addition to the results presented in Section \ref{md-sec:lymphnode} of the 
   MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 

### 2019-03-22 docs/manuscript.Rnw
textual changes

@@
 	logistic generalised elastic net model in Section
-	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
+	\ref{sec:model}. In Section \ref{sec:estimation} we derive a variational
 	approximation to this model and use this novel approximation in the empirical
@@
 
-	\section{Estimation}\label{seq:estimation}
-	\subsection{Empirical Bayes}\label{seq:empiricalbayes}
+	\section{Estimation}\label{sec:estimation}
+	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
 	If the penalty parameters are known, estimation of the frequentist elastic net
@@
 
-	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
+	\subsection{Empirical-variational Bayes}\label{sec:empvarbayes}
 	Variational Bayes was shown to underestimate the posterior variance of the 
@@
 	\cite{friedman_regularization_2010} as introduced in Section 
-	\ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
+	\ref{sec:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
 	cross-validation of the regular elastic net model, such that the overall 
@@
 
-	As shortly touched upon in Section \ref{seq:empiricalbayes}, frequentist 
+	As shortly touched upon in Section \ref{sec:empiricalbayes}, frequentist 
 	feature selection is trivial after estimation of the penalty multipliers. We 
@@
 	\section{Simulations}\label{sec:simulations}
-	To assess variable selection and predictive performance we conducted a 
+	To assess classification and variable selection performance we conducted a 
 	simulation study in which we compared \texttt{gren} to the regular elastic 
@@
 	feature level ($\alpha=1$) penalization. For the methods that do 
-	automatic feature selection, we estimate 
-	the models for a range of model sizes.
+	automatic feature selection, we estimate models for a range of model sizes.
 
@@
 	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
-	estimate, while we expect the other methods to perform better, due to a
-	lack of overfitting. In scenario three, we expect all methods to underperform
+	estimate. The group lasso extensions are expected to perform better: they 
+	estimate just one penalty parameter, so are less sensitive to overfitting.
+	In scenario three, we expect all methods to underperform
 	due to the high correlations between the features. The fourth 
-	scenario is conjectured to yield a better \texttt{gren} performance, since
+	scenario is expected to yield a better \texttt{gren} performance, since
 	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
-	incorporate a differential signal between
+	detect differential signal between
 	the groups. Scenario five is expected to yield a bad performance for all 
@@
 	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
-	$p=1000$. We varied the covariance matrix of the features and simulation of 
+	$p=900$ in scenario 2 and $p=1000$ in the other scenarios. We varied the 
+	covariance matrix of the features and simulation of 
 	the model parameters $\beta_j$, $j=1, \dots, p$ between the
-	scenarios (see SM Section [! section ref here] for the exact 
+	scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
 	parametrisations). We 
@@
 	of random variation, we repeated each scenario 100 times 
-	and report estimated penalty parameters and predictive performance measures 
+	and report predictive performance measures 
 	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
-	Table [! table ref here].
+	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
+	Additional results and performance measures are presented in Section 
+	\ref{sm-sec:simulations} of the SM.
 
@@
 @
-
-	Additional results and performance measures are presented in Section 
-	\ref{sm-sec:simulations} of the SM.
 
@@
 
-	Although our method can be considered weakly adaptive and is essentially 
-	a reweighted elastic net, it is different from 
+	Although our method is essentially a reweighted elastic net and can be
+	considered weakly adaptive, it is different from 
 	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
@@
 
-	A common criticism of the lasso (and elastic net) is its instability of 
-	feature 
-	selection: different data instances lead to different sets of selected
-	features. We investigated the stability of selection on the data from Section
-	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
-	the original data and calculated the sizes of all intersections of selected
-	features. We compared \texttt{gren} to the 
-	regular elastic net and found that the inclusion of extra information
-	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
-	In addition, we investigated 
-	the variance and bias in penalty multiplier estimation by examining the
-	distribution
-	of estimated penalty parameters from 100 random partitionings of the features
-  (see SM Section \ref{sm-sec:colorectal}). We found that the penalty parameters
-  tend to cluster around one, as expected for an unbiased estimation method.
-
-	As touched upon in Section \ref{sec:introduction}, another obvious 
+	As touched upon in Section \ref{sec:introduction}, an obvious 
 	comparison is to the group lasso \cite[]{meier_group_2008,
@@
 	simulations and data applications that \texttt{gren} is competitive and often
-	even superior to (extensions of) the group lasso.
+	superior to (extensions of) the group lasso.
+
+	A common criticism of the lasso (and elastic net) is its instability of 
+	feature 
+	selection: different data instances lead to different sets of selected
+	features. We investigated the stability of selection on the data from Section
+	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
+	the original data and calculated the sizes of all intersections of selected
+	features. We compared \texttt{gren} to the 
+	regular elastic net and found that the inclusion of extra information
+	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
+	In addition, we investigated 
+	the bias and variance of the penalty multipliers by examination of the
+	distribution of estimated penalty parameters from 100 random partitionings 
+	of the features
+  (see SM Section \ref{sm-sec:colorectal}). We found that the penalty parameters
+  tend to cluster around one, as expected for an unbiased estimation method, and
+  that the variance of the penalty multipliers is less than for the similar
+  \texttt{GRridge} method.
 
 	A possible weak point of the proposed method is the double EM loop. The double
-	loop increases the chance of ending up in a local optimum. In the application
-	discussed above we investigated the occurrence of multiple optima, but never
-	encountered them. This does not guarantee that local optima do not occur, but 
-	it provides some evidence that local optima are not ubiquitous. Local optima 
-	can often be avoided by an informed choice of starting values. In our 
+	loop increases the chance of ending up in a local optimum. In the applications
+	discussed in Sections \ref{sec:colorectal}--
+	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
+	\ref{sm-sec:cervical}, we investigated the occurrence of multiple optima, 
+	but never encountered them. This does not guarantee that local optima do not
+	occur, but 
+	it provides some evidence that local optima are not ubiquitous. In any case,
+	bad local optima can often be avoided by an informed choice of starting 
+	values. In our 
 	experience, reasonable starting values are obtained by running a 
@@
 	the regular elastic net and adaptively adjusting the penalty parameters to a
-	group-regularized model.
+	group-regularized model. The latter is the default approach in our \texttt{R}
+	package
 
@@
   accurate enough for a 
-  reliable EB step. This approach is often used in more 
-  computational disciplines, where the problem is viewed as purely algorithmic 
-  and the EB and VB iterations are interpreted as optimisation steps along the 
-  gradient in the larger combined model- and hyper-parameters space.
+  reliable EB step.
 
@@
 	\section*{Session info}
+
 <<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=

### 2019-03-22 docs/supplement_biostatistics.Rnw
all simulations except 2 finished, second version biostatistics manuscript initiated

--- 2019-02-05 docs/supplement.Rnw
@@
-\documentclass[a4paper,hidelinks]{article}
-\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}
+% set options
+<<settings, include=FALSE, echo=FALSE>>=
+opts_knit$set(root.dir="..", base.dir="../figs")
+opts_chunk$set(fig.align='center', fig.path="../figs/", 
+               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
+@
 
-% vectors and matrices
+% read figure code
+<<figures, include=FALSE>>=
+read_chunk("code/figures.R")
+@
+
+\documentclass[oupdraft]{bio}
+\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
+\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable}
+
+% Add history information for the article if required
+%\history{Received August 1, 2010;
+%	revised October 1, 2010;
+%	accepted for publication November 1, 2010}
+
+% objects
+\newcommand{\x}{\mathbf{x}}
 \newcommand{\bbeta}{\bm{\beta}}
+\newcommand{\X}{\mathbf{X}}
+\newcommand{\y}{\mathbf{y}}
+\newcommand{\G}{\mathcal{G}}
 \newcommand{\btau}{\bm{\tau}}
-\newcommand{\balpha}{\bm{\alpha}}
+\newcommand{\blambda}{\bm{\lambda}}
 \newcommand{\bomega}{\bm{\omega}}
-\newcommand{\bOmega}{\bm{\Omega}}
-\newcommand{\bSigma}{\bm{\Sigma}}
-\newcommand{\bkappa}{\bm{\kappa}}
 \newcommand{\btheta}{\bm{\theta}}
 \newcommand{\bmu}{\bm{\mu}}
+\newcommand{\bSigma}{\bm{\Sigma}}
 \newcommand{\bpsi}{\bm{\psi}}
-\newcommand{\blambda}{\bm{\lambda}}
+\newcommand{\bOmega}{\bm{\Omega}}
+\newcommand{\Z}{\mathbf{Z}}
 \newcommand{\bLambda}{\bm{\Lambda}}
-\newcommand{\bphi}{\bm{\phi}}
-\newcommand{\Y}{\mathbf{Y}}
-\newcommand{\y}{\mathbf{y}}
-\newcommand{\X}{\mathbf{X}}
-\newcommand{\x}{\mathbf{x}}
-\newcommand{\I}{\mathbf{I}}
-\newcommand{\T}{\mathbf{T}}
-\newcommand{\Z}{\mathbf{Z}}
-\newcommand{\W}{\mathbf{W}}
-\renewcommand{\O}{\mathbf{O}}
-\newcommand{\B}{\mathbf{B}}
-\renewcommand{\H}{\mathbf{H}}
-\newcommand{\U}{\mathbf{U}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\Vm}{\mathbf{V}}
-\newcommand{\rv}{\mathbf{r}}
-\newcommand{\A}{\mathbf{A}}
-\newcommand{\Q}{\mathbf{Q}}
-\newcommand{\Sm}{\mathbf{S}}
-\newcommand{\0}{\bm{0}}
-\newcommand{\tx}{\tilde{\mathbf{x}}}
-\newcommand{\hy}{\hat{y}}
-\newcommand{\tm}{\tilde{m}}
-\newcommand{\tkappa}{\tilde{\kappa}}
 \newcommand{\m}{\mathbf{m}}
-\newcommand{\C}{\mathbf{C}}
-\renewcommand{\v}{\mathbf{v}}
-\newcommand{\g}{\mathbf{g}}
-\newcommand{\s}{\mathbf{s}}
-\renewcommand{\c}{\mathbf{c}}
-\newcommand{\ones}{\mathbf{1}}
 
 % functions and operators
+\newcommand{\tr}{^{\text{T}}}
+\newcommand{\expit}{\text{expit}}
+\newcommand{\argmax}{\text{argmax} \,}
 \renewcommand{\L}{\mathcal{L}}
-\newcommand{\G}{\mathcal{G}}
-\renewcommand{\P}{\mathcal{P}}
-\newcommand{\argmax}{\text{argmax} \,}
-\newcommand{\expit}{\text{expit}}
-\newcommand{\erfc}{\text{erfc}}
 \newcommand{\E}{\mathbb{E}}
-\newcommand{\V}{\mathbb{V}}
-\newcommand{\tr}{^{\text{T}}}
 \newcommand{\diag}{\text{diag}}
-\newcommand{\KL}{D_{\text{KL}}}
 \newcommand{\trace}{\text{tr}}
-\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
-\newcommand{\appropto}{\mathrel{\vcenter{
-  \offinterlineskip\halign{\hfil$##$\cr
-    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
-\newcommand{\card}[1]{\text{card} \left( #1 \right)}
-\makeatletter
-\newcommand*{\defeq}{\mathrel{\rlap{%
-			\raisebox{0.3ex}{$\m@th\cdot$}}%
-		\raisebox{-0.3ex}{$\m@th\cdot$}}%
-	=}
-\makeatother
-
-% distributions
-\newcommand{\N}{\text{N}}
-\newcommand{\TG}{\text{TG}}
-\newcommand{\Bi}{\text{Binom}}
-\newcommand{\PG}{\text{PG}}
-\newcommand{\Mu}{\text{Multi}}
-\newcommand{\GIG}{\text{GIG}}
-\newcommand{\IGauss}{\text{IGauss}}
-\newcommand{\Un}{\text{U}}
 
@@
 \renewcommand{\]}{\right]}
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-
-% settings
-\graphicspath{{../figs/}}
-\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
-\setlength{\textwidth}{15cm}
-\setlength{\parindent}{0pt}
-\title{Adaptive group-regularized logistic elastic net regression Supplementary material}
-\date{}
-\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
 
 \begin{document}
+
+	% Title of paper
+	\title{Adaptive group-regularized logistic elastic net regression 
+	Supplementary material}
+
+	% List of authors, with corresponding author marked by asterisk
+	\author{MAGNUS M. M\"UNCH$^{\ast 1,2}$, CAREL F.W. PEETERS$^{1}$, AAD W. VAN 
+	DER VAART$^{2}$, \\* MARK A. VAN DE WIEL$^{1,3}$\\[4pt]
+	% Author addresses
+	\textit{1. Department of Epidemiology \& Biostatistics, Amsterdam Public 
+	Health research institute, Amsterdam University medical centers, PO Box 7057,
+	1007 MB Amsterdam, The Netherlands \\*
+	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
+	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom}
+	\\[2pt]
+	% E-mail address for correspondence
+	{m.munch@vumc.nl}}
+
+	% Running headers of paper:
+	\markboth%
+	% First field is the short list of authors
+	{M. M. M\"unch and others}
+	% Second field is the short title of the paper
+	{Adaptive group-regularized logistic elastic net regression}
+
 	\maketitle
-\noindent
-1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
-Amsterdam, The Netherlands\\
-2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
-3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\
 
-\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes
+	% Add a footnote for the corresponding author if one has been
+	% identified in the author list
+	\footnotetext{To whom correspondence should be addressed.}
 
-\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}
-
-\section{Introduction}
-This document contains supplementary material to the paper `Adaptive group-regularized logistic elastic net regression'.
-Section \ref{sec:enetprior} contains extra details on the elastic net prior distribution.
-Section \ref{sec:objectivefunction} gives the objective function that is maximised in the variational Bayes procedure introduced in Section 3 of the Main Document.
-Section \ref{sec:polyagamma} gives details on the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main Document.
-The derivations of the variational posterior used in the Main Document are shown in Section \ref{sec:variationaldistributions}.
-Section \ref{sec:computationalcomplexity} shows how we may reduce the computational complexity of our variational Bayes procedure.
-Section \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected joint log likelihood that is maximised to estimate the penalty multipliers as explained in Section 3 of the Main Document.
-The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}.
-We apply our method to a study on microRNAs in colorectal cancer data \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} .
-Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.
+	\section{Introduction}\label{sec:introduction}
+  This document contains supplementary material to the paper `Adaptive 
+  group-regularized logistic elastic net regression'. Section 
+  \ref{sec:enetprior} contains extra details on the elastic net prior 
+  distribution. Section \ref{sec:objectivefunction} gives the objective 
+  function that is maximised in the variational Bayes procedure introduced in
+  Section \ref{md-sec:estimation} of the Main Document (MD). Section
+  \ref{sec:polyagamma} gives details on
+  the P{\'o}lya-Gamma parametrisation introduced in Section 
+  \ref{md-sec:estimation} of the MD. The
+  derivations of the variational posterior used in the MD are shown in Section
+  \ref{sec:variationaldistributions}. Section
+  \ref{sec:computationalcomplexity} shows how we may reduce the computational
+  complexity of our variational Bayes procedure. Section 
+  \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected
+  joint log likelihood that is maximised to estimate the penalty multipliers as
+  explained in Section \ref{md-sec:estimation} of the MD. The full procedure of
+  the
+  proposed method is shown in Section \ref{sec:fullestimationprocedure}. 
+  Additional results to the applications in the MD Sections 
+  \ref{md-sec:colorectal}--\ref{md-sec:lymphnode}. We
+  apply our method to a study on metabolomics in Alzheimer's in Section
+  \ref{sec:alzheimer} and to study in microRNAs in cervical cancer data 
+  in Section \ref{sec:cervical}. Finally, 
+  we introduce the ridge and lasso versions of the proposed method in Section
+  \ref{sec:evbridgelasso}.
 
@@
   \begin{align*}
-\E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
-\V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta - \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} - \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
+    \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot 
+    \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
+    \V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot
+    \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta -
+    \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} -
+    \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 
+    \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
   \end{align*}
@@
   where all mathematical operations on vectors and matrices are element wise. 
-After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
+  After combining the terms in (\ref{eq:elboomega2}) and substituting the 
+  expectations and variances, several terms cancel to
+  give:
   \begin{subequations} \label{eq:elboomega3}
 	  \begin{align}
-	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
-	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
+	    \E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} 
+	    [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ 
+	    & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) 
+	    \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - 
+	    \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
+	    & \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} 
+	    \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - 
+	    ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
     \end{align}
@@
 
-Inspection of the updating equations in the Main Document, learns us that the last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that we are left with just (\ref{eq:elboomega4}).
-After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
+  Inspection of the updating equations in the MD, learns us that the 
+  last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
+  we are left with just (\ref{eq:elboomega4}). After a change of variables 
+  $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
   \begin{subequations} \label{eq:elbobeta2}
 	  \begin{align}
-	\E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - \frac{p}{2} \log (2 \pi) \\
-	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) (\bmu_j^2 + \bSigma_{jj}) \\
-	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
-	\E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - \frac{p}{2} \log |\bSigma| .
+	    \E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G 
+	    |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - 
+	    \frac{p}{2} \log (2 \pi) \\
+	    & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in 
+	    \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) 
+	    (\bmu_j^2 + \bSigma_{jj}) \\
+	    & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} 
+	    \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
+	    \E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - 
+	    \frac{p}{2} \log |\bSigma| .
 	  \end{align}
   \end{subequations}
-We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):
+  We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for 
+  (\ref{eq:elbobeta}):
   \begin{subequations} \label{eq:elbobeta3}
 	   \begin{align}
-	\E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} [\log q (\bbeta )] \\ 
-	& = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + \frac{p}{2} \\
-	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
-	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
+	     \E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} 
+	     [\log q (\bbeta )] \\ 
+	     & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + 
+	     \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + 
+	     \frac{p}{2} \\
+	     & \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g 
+	     \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 
+	     \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
+	     & \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + 
+	     \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
 	  \end{align}
@@
 
-Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
+  Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and
+  (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:
   \begin{align*}
-\text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2} \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
-& + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
-& + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
+    \text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + 
+    \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2}
+    \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
+    & + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} 
+    \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} 
+    \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) 
+    \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
+    & + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} -
+    \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
   \end{align*}
-Here, proportionality is with respect to the variational and penalty parameters.
+  Here, proportionality is with respect to the variational and penalty 
+  parameters.
 
   \section{The P\'{o}lya-Gamma parametrisation}\label{sec:polyagamma}
-In Section 3 of the Main Document latent P\'{o}lya-Gamma distributed variables are introduced.
-Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma variables are introduced into the logistic regression model under a Gaussian prior.
-Here, we explain how introduction of these latent variables simplifies the variational Bayes calculations.
+  In Section \ref{md-sec:estimation} of the MD latent P\'{o}lya-Gamma
+  distributed variables
+  are introduced. Likewise, in \cite{polson_bayesian_2013}, P\'{o}lya-Gamma
+  variables are introduced into the logistic regression model under a Gaussian
+  prior. Here, we explain how introduction of these latent variables simplifies
+  the variational Bayes calculations.
 
-In \cite{polson_bayesian_2013}, the central identity for this parametrisation is given in their Theorem 1:
+  In \cite{polson_bayesian_2013}, the central identity for this parametrisation
+  is given in their Theorem 1:
   \begin{equation}\label{eq:intid}
-\frac{\exp(\psi)^{a}}{[1 + \exp(\psi)]^{b}} = 2^{-b} \exp(\kappa \psi) \int_0^{\infty} \exp(-\omega \psi^2 /2) \pi(\omega) d \omega,
+    \frac{\exp(\psi)^{a}}{[1 + \exp(\psi)]^{b}} = 2^{-b} \exp(\kappa \psi)
+    \int_0^{\infty} \exp(-\omega \psi^2 /2) \pi(\omega) d \omega,
   \end{equation}
-where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed.
-This holds for all $b > 0$ and $a \in \mathbb{R}$.
-After setting $a=y_i$, $b=m_i$, and $\psi=\x_i \tr \bbeta$, we may consider the right-hand side of (\ref{eq:intid}) as an un-normalised joint density in $(\bbeta, \omega)$.
-The likelihood contribution of observation $i$ for $\bbeta$ conditional on $\omega_i$ may now be written as:
+  where $\kappa=a - b/2$ and $\omega$ is $\mathcal{PG}\(b,0\)$ distributed. This
+  holds for all $b > 0$ and $a \in \mathbb{R}$. After setting $a=y_i$, $b=m_i$,
+  and $\psi=\x_i \tr \bbeta$, we may consider the right-hand side of 
+  (\ref{eq:intid}) as an un-normalised joint density in $(\bbeta, \omega)$. The
+  likelihood contribution of observation $i$ for $\bbeta$ conditional on
+  $\omega_i$ may now be written as:
   $$
-\L_i(\bbeta | \omega_i) \propto \exp[\kappa_i \x_i \tr \bbeta - \omega_i (\x_i \tr \bbeta)^2/2].
+  \L_i(\bbeta | \omega_i) \propto \exp[\kappa_i \x_i \tr \bbeta - 
+  \omega_i (\x_i   \tr \bbeta)^2/2].
   $$
@@
   \begin{align}
-q^*_{\bbeta} (\bbeta) & {\propto} \exp \{ \E_{\bomega, \btau} [\log p (\bbeta, \bomega, \btau, \y)]\} {\propto} \exp \{\log \L (\y ; \bbeta) + \E_{\bomega} [\log \pi (\bomega | \bbeta)] + \E_{\btau} [ \log \pi (\bbeta | \btau)] \}, \label{eq:qbeta} \\
-q^*_{\bomega} (\bomega) & {\propto} \exp \{ \E_{\bbeta, \btau} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi (\bomega | \bbeta)] \},\label{eq:qomega} \\
-q^*_{\tau} (\btau) & {\propto} \exp \{ \E_{\bbeta, \bomega} [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} [\log \pi ( \bbeta | \btau)] + \log \pi (\btau) \}. \label{eq:qtau}
+    q^*_{\bbeta} (\bbeta) & {\propto} \exp \{ \E_{\bomega, \btau} 
+    [\log p (\bbeta, \bomega, \btau, \y)]\} {\propto} \exp 
+    \{\log \L (\y ; \bbeta) + \E_{\bomega} [\log \pi (\bomega | \bbeta)] + 
+    \E_{\btau} [ \log \pi (\bbeta | \btau)] \}, \label{eq:qbeta} \\
+    q^*_{\bomega} (\bomega) & {\propto} \exp \{ \E_{\bbeta, \btau} 
+    [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} 
+    [\log \pi (\bomega | \bbeta)] \},\label{eq:qomega} \\
+    q^*_{\tau} (\btau) & {\propto} \exp \{ \E_{\bbeta, \bomega} 
+    [\log p (\bbeta, \bomega, \btau, \y)] \} {\propto} \exp \{ \E_{\bbeta} 
+    [\log \pi ( \bbeta | \btau)] + \log \pi (\btau) \}. \label{eq:qtau}
   \end{align}
 
-For convenience we work on the log-scale and occasionally do a change of variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three terms on the right-hand side of (\ref{eq:qbeta}):
+  For convenience we work on the log-scale and occasionally do a change of
+  variables $\psi_j = \tau_j - 1$. Starting with $\bbeta$ we have for the three
+  terms on the right-hand side of (\ref{eq:qbeta}):
   \begin{align*}
-\log \L (\y ; \bbeta) & = \sum_{i=1}^n \log f (y_i ; \bbeta) \propto \sum_{i=1}^n \log \left\{ \frac{\exp(\x_i \tr \bbeta)^{y_i}}{[1 + \exp(\x_i \tr \bbeta)]^{m_i}} \right\} \\
+    \log \L (\y ; \bbeta) & = \sum_{i=1}^n \log f (y_i ; \bbeta) \propto 
+    \sum_{i=1}^n \log \left\{ \frac{\exp(\x_i \tr \bbeta)^{y_i}}{[1 + 
+    \exp(\x_i \tr \bbeta)]^{m_i}} \right\} \\
     & = \y \tr \X \bbeta - \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)], \\
-\E_{\bomega} [\log \pi (\bomega | \bbeta)] & \propto \sum_{i=1}^n \E_{\omega_i} [\log \pi (\omega_i | \bbeta)] \\ & \propto \sum_{i=1}^n \E_{\omega_i} \( \log \left\{ \cosh\( \frac{\x_i \tr \bbeta}{2}\)^{m_i} \exp \[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i \]\right\} \) \\
-& \propto \sum_{i=1}^n m_i \log \cosh \( \frac{\x_i \tr \bbeta }{2} \) - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta \\ & \propto \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)] - \frac{1}{2} \m \tr \X \bbeta - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta, \\
-\E_{\btau} [ \log \pi (\bbeta | \btau)] & = \sum_{j=1}^p \E_{\tau_j} [ \log \pi (\beta_j | \tau_j)] \propto \frac{\lambda_2}{2} \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} \E \( \frac{\tau_j}{\tau_j - 1} \) \\
-& = \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} [1 + \E ( \psi_j^{-1} )] = - \frac{1}{2} \lambda_2 \bbeta \tr \bLambda' (\I + \Z) \bbeta,
+    \E_{\bomega} [\log \pi (\bomega | \bbeta)] & \propto \sum_{i=1}^n 
+    \E_{\omega_i} [\log \pi (\omega_i | \bbeta)] \\ & \propto \sum_{i=1}^n
+    \E_{\omega_i} \( \log \left\{ \cosh\( \frac{\x_i \tr \bbeta}{2}\)^{m_i} 
+    \exp \[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i \]\right\} \) \\
+    & \propto \sum_{i=1}^n m_i \log \cosh \( \frac{\x_i \tr \bbeta }{2} \) -
+    \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta \\ & \propto 
+    \sum_{i=1}^n m_i \log [1 + \exp(\x_i \tr \bbeta)] - \frac{1}{2} \m \tr 
+    \X \bbeta - \frac{1}{2} \bbeta \tr \X \tr \bOmega \X \bbeta, \\
+    \E_{\btau} [ \log \pi (\bbeta | \btau)] & = \sum_{j=1}^p \E_{\tau_j} 
+    [ \log \pi (\beta_j | \tau_j)] \propto \frac{\lambda_2}{2} 
+    \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} \E \( \frac{\tau_j}{\tau_j - 1} \) \\
+    & = \sum_{j=1}^p \beta_j^2 \lambda'_{g(j)} [1 + \E ( \psi_j^{-1} )] = -
+    \frac{1}{2} \lambda_2 \bbeta \tr \bLambda' (\I + \Z) \bbeta,
   \end{align*}
-where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega = \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' = \diag (\lambda'_{g(j)})$. 
-Since the difficult $\sum_{i=1}^n m_i \log[1 + \exp(\x_i \tr \bbeta)]$ term appears in both the $\log \L (\y ; \bbeta)$ and $\pi (\omega_i | \bbeta)$ part, it cancels, thereby justifying the introduction of the $\omega_i$.
+  where proportionality is with respect to $\bbeta$. Furthermore, $\bOmega =
+  \diag [\E(\omega_i)]]$, $\Z = \diag [ \E (\psi_j^{-1})]$, and $\bLambda' =
+  \diag (\lambda'_{g(j)})$. 
+  Since the difficult $\sum_{i=1}^n m_i \log[1 + \exp(\x_i \tr \bbeta)]$ term
+  appears in both the $\log \L (\y ; \bbeta)$ and $\pi (\omega_i | \bbeta)$ 
+  part, it cancels, thereby justifying the introduction of the $\omega_i$.
   Finally, we have: 
   $$
-\log q^*_{\bbeta} (\bbeta) = \bkappa \tr \X \bbeta - \frac{1}{2} \bbeta \tr (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z) \bbeta,
+  \log q^*_{\bbeta} (\bbeta) = \bkappa \tr \X \bbeta - \frac{1}{2} \bbeta \tr 
+  (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z) \bbeta,
   $$
-which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma \X \tr \bkappa$.
+  which is quadratic in $\bbeta$ and thus recognised as the log-kernel of a 
+  Gaussian distribution with covariance $\bSigma = (\X \tr \bOmega \X + 
+  \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z)^{-1}$ and mean $\bmu=\bSigma 
+  \X \tr \bkappa$.
 
@@
   \begin{align*}
-\log q^*_{\bomega} (\bomega) & = \E_{\bbeta} [\log \pi (\bomega | \bbeta)] = \sum_{i=1}^n \E_{\bbeta} [\log \pi (\omega_i | \bbeta)] \\ & = \sum_{i=1}^n \E_{\bbeta} \( m_i \log \left\{ \cosh \( \frac{\x_i \tr \bbeta}{2}\) \exp\[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i\] \cdot f(\omega_i | m_i, 0) \right\} \) \\
-& = \sum_{i=1}^n \E_{\bbeta} \[ m_i \log \cosh \( \frac{\x_i \tr \bbeta}{2}\) \] - \frac{1}{2} \sum_{i=1}^n \omega_i \E[ (\x_i \tr \bbeta)^2] + \sum_{i=1}^n \log f(\omega_i | m_i, 0) \\ 
-& \propto \sum_{i=1}^n m_i \log \cosh \left\{ \frac{\sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}}{2}\right\} - \frac{1}{2} \sum_{i=1}^n \omega_i \left\{ [\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \right\} \\
+    \log q^*_{\bomega} (\bomega) & = \E_{\bbeta} [\log \pi (\bomega | \bbeta)] =
+    \sum_{i=1}^n \E_{\bbeta} [\log \pi (\omega_i | \bbeta)] \\ & = \sum_{i=1}^n
+    \E_{\bbeta} \( m_i \log \left\{ \cosh \( \frac{\x_i \tr \bbeta}{2}\) 
+    \exp\[-\frac{(\x_i \tr \bbeta)^2}{2} \omega_i\] \cdot f(\omega_i | m_i, 0) 
+    \right\} \) \\ 
+    & = \sum_{i=1}^n \E_{\bbeta} \[ m_i \log \cosh 
+    \( \frac{\x_i \tr \bbeta}{2}\) \] - \frac{1}{2} \sum_{i=1}^n \omega_i 
+    \E[ (\x_i \tr \bbeta)^2] + \sum_{i=1}^n \log f(\omega_i | m_i, 0) \\ 
+    & \propto \sum_{i=1}^n m_i \log \cosh \left\{ \frac{\sqrt{[\x_i \tr 
+    \E(\bbeta)]^2 + \x_i \tr \V[\bbeta]}}{2}\right\} - 
+    \frac{1}{2} \sum_{i=1}^n \omega_i \left\{ [\x_i \tr \E(\bbeta)]^2 + 
+    \x_i \tr \V[\bbeta] \right\} \\ 
     & \,\,\, + \sum_{i=1}^n \log f(\omega_i | m_i, 0),
   \end{align*}
-where proportionality is with respect to $\bomega$.
-If exponentiated, this is a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.
+  where proportionality is with respect to $\bomega$. If exponentiated, this is 
+  a product of tilted and renormalised $\mathcal{PG}(b_i, c_i)$ variables
+  \cite[]{polson_bayesian_2013}, with $b_i=m_i$ and $c_i = \sqrt{[\x_i \tr 
+  \E(\bbeta)]^2 + \x_i \tr \V[\bbeta] \x_i }$.
 
-The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated as:
+  The variational posterior terms of $\btau$ in (\ref{eq:qtau}) are calculated 
+  as:
   \begin{align*}
-\E_{\bbeta} [\log \pi ( \bbeta | \btau)] & = \sum_{j=1}^p \E_{\bbeta_j} [\log \pi(\beta_j | \tau_j )] \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}, \\
-\log \pi (\btau) & = \sum_{j=1}^p \log \pi (\tau_j) \propto - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j. 
+    \E_{\bbeta} [\log \pi ( \bbeta | \btau)] & = \sum_{j=1}^p \E_{\bbeta_j} 
+    [\log \pi(\beta_j | \tau_j )] \propto \frac{1}{2} \sum_{j=1}^p \log 
+    \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2)
+    \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}, \\
+    \log \pi (\btau) & = \sum_{j=1}^p \log \pi (\tau_j) \propto - 
+    \frac{1}{2} \sum_{j=1}^p \log \tau_j - 
+    \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j. 
   \end{align*}
@@
   \begin{align*}
-\log q^*_{\tau} (\btau) & \propto \frac{1}{2} \sum_{j=1}^p \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}  - \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p \frac{\lambda_1^2}{8 \lambda_2} \tau_j \\
-& = \frac{1}{2} \sum_{j=1}^p \log [(\tau_j - 1)^{-1}] - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2}\tau_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}\),
+    \log q^*_{\tau} (\btau) & \propto \frac{1}{2} \sum_{j=1}^p 
+    \log \(\frac{\tau_j}{\tau_j - 1} \) - \frac{1}{2} \sum_{j=1}^p 
+    \E (\beta_j^2) \lambda_2 \lambda'_{g(j)} \frac{\tau_j}{\tau_j - 1}  - 
+    \frac{1}{2} \sum_{j=1}^p \log \tau_j - \sum_{j=1}^p 
+    \frac{\lambda_1^2}{8 \lambda_2} \tau_j \\
+    & = \frac{1}{2} \sum_{j=1}^p \log [(\tau_j - 1)^{-1}] - 
+    \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2}\tau_j + 
+    [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} 
+    \frac{\tau_j}{\tau_j - 1}\),
   \end{align*}
-where proportionality is with respect to $\btau$. Again applying the change of variables $\psi_j = \tau_j - 1$, we have:
+  where proportionality is with respect to $\btau$. Again applying the change of
+  variables $\psi_j = \tau_j - 1$, we have:
   $$
-\log q^*_{\psi} (\psi) \propto \frac{1}{2} \sum_{j=1}^p \log \psi_j^{-1} - \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2} \psi_j + [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \psi_j^{-1}\),
+  \log q^*_{\psi} (\psi) \propto \frac{1}{2} \sum_{j=1}^p \log \psi_j^{-1} -
+  \frac{1}{2} \sum_{j=1}^p \(\frac{\lambda_1^2}{4 \lambda_2} \psi_j + 
+  [\E (\beta_j)^2 + \V(\beta_j)] \lambda_2 \lambda'_{g(j)} \psi_j^{-1}\),
   $$
-which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$ random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + \V(\beta_j)]$.
+  which is the log-kernel of a $\GIG(1/2, \lambda_1^2/(4 \lambda_2), \chi_j)$
+  random variable with $\chi_j=\lambda_2 \lambda'_{g(j)} [\E (\beta_j)^2 + 
+  \V(\beta_j)]$.
 
   \section{Computational complexity}\label{sec:computationalcomplexity}
-Calculation of $\bSigma$ involves the inversion of a $(p \times p)$-dimensional matrix, which occurs at every iteration of the algorithm.
-If naively done, every inversion is of computational complexity $\mathcal{O}(p^3)$.
-With $p$ large, this is a serious computational bottleneck of the algorithm, especially with possibly many iterations.
-However, inspection of the updating equations in (3.9) in the main document, reveals that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the $c_i$ and $\chi_j$.
-In Section 2.2 of the main document we show that for the EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, and the diagonal of $\bSigma$.
-Here we provide details on how to compute these with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the computational cost of the algorithm.
-If necessary, we may calculate the full $\bSigma$ once after the algorithm has converged.
+  Calculation of $\bSigma$ involves the inversion of a 
+  $(p \times p)$-dimensional matrix, which occurs at every iteration of the
+  algorithm. If naively done, every inversion is of computational complexity
+  $\mathcal{O}(p^3)$. With $p$ large, this is a serious computational bottleneck
+  of the algorithm, especially with possibly many iterations. However, 
+  inspection of the updating equations in (\ref{md-eq:VBupdateequations})
+  in the MD, reveals
+  that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i 
+  \tr \bmu$, and $\bmu$, plus the diagonal of $\bSigma$ to directly update the
+  $c_i$ and $\chi_j$. In Section \ref{md-sec:estimation} of the MD we show that 
+  for the 
+  EM updates of $\blambda'$, we only require the $c_i$, the $\chi_j$, $\bmu$, 
+  and the diagonal of $\bSigma$. Here we provide details on how to compute these
+  with complexity $\mathcal{O}(n^2 p)$, thereby considerably reducing the
+  computational cost of the algorithm. If necessary, we may calculate the full
+  $\bSigma$ once after the algorithm has converged.
 
-In the following we slightly change notation and let $\diag(\A)$, with $\A$ a $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$ to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$.
-We start by calculating the complexity of $\diag(\bSigma)$.
-To this end, let $\H = \lambda_2 \bLambda' + \lambda_2 \bLambda' \Z$.
-Now, by the Woodbury identity, we may decompose $\diag(\bSigma)$ as:
+  In the following we slightly change notation and let $\diag(\A)$, with $\A$ a
+  $p \times p$ square matrix, be the operator that extracts the diagonal of $\A$
+  to a $p$-dimensional column vector, e.g., $\diag(\A) = \begin{bmatrix} 
+  \A_{11} & \cdots & \A_{pp} \end{bmatrix} \tr$. We start by calculating the
+  complexity of $\diag(\bSigma)$. To this end, let $\H = \lambda_2 \bLambda' +
+  \lambda_2 \bLambda' \Z$. Now, by the Woodbury identity, we may decompose 
+  $\diag(\bSigma)$ as:
   \begin{subequations}\label{eq:diagsigma}
 	  \begin{align}
-	\diag(\bSigma) ~& = \diag(\H^{-1}) - \diag[\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1}] \\
-	& = \diag(\H^{-1}) - [(\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1}) \circ \H^{-1} \X \tr] \cdot \ones_{n \times 1},
+	    \diag(\bSigma) ~& = \diag(\H^{-1}) - \diag[\H^{-1} \X \tr \bOmega 
+	    (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1}] \\
+	    & = \diag(\H^{-1}) - [(\H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr 
+	    \bOmega)^{-1}) \circ \H^{-1} \X \tr] \cdot \ones_{n \times 1},
 	  \end{align}
   \end{subequations}
-where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is a $p \times n$-dimensional matrix of ones.
-By recognising that $\H$ is a diagonal matrix, we can post-multiply it by a $p \times n$-dimensional matrix with computational complexity $\mathcal{O}(np)$, which leads to a total complexity of $\mathcal{O}(n^2p)$ for $\diag(\bSigma)$.
-Next we consider the calculation of $\bmu$.
-Again using the Woodbury identity, we may write $\bmu$ as:
+  where $\circ$ denotes the Hadamard matrix product and $\ones_{p \times n}$ is
+  a $p \times n$-dimensional matrix of ones. By recognising that $\H$ is a
+  diagonal matrix, we can post-multiply it by a $p \times n$-dimensional matrix
+  with computational complexity $\mathcal{O}(np)$, which leads to a total
+  complexity of $\mathcal{O}(n^2p)$ for $\diag(\bSigma)$. Next we consider the
+  calculation of $\bmu$. Again using the Woodbury identity, we may write $\bmu$
+  as:
   \begin{align}\label{eq:mu}
-\bmu = \bSigma \X \tr \bkappa = [\H^{-1} \X \tr - \H^{-1} \X \tr \bOmega (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1} \X \tr] \bkappa,
+    \bmu = \bSigma \X \tr \bkappa = [\H^{-1} \X \tr - \H^{-1} \X \tr \bOmega 
+    (\I + \X \H^{-1} \X \tr \bOmega)^{-1} \X  \H^{-1} \X \tr] \bkappa,
   \end{align}
-which is, by a similar argument as before, again an operation of complexity $\mathcal{O}(n^2p)$.
+  which is, by a similar argument as before, again an operation of complexity
+  $\mathcal{O}(n^2p)$.
 
-Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$ as:
+  Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n
+  \end{bmatrix} \tr$ as:
   \begin{align*}
-\c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
-& = \sqrt{\diag \[\X \bSigma \X \tr\] +  \diag \[\X \bmu \bmu \tr \X \tr\]} \\
-& = \sqrt{ [(\X \bSigma) \circ \X ] \cdot \ones_{p \times 1} + (\X \bmu) \circ (\X \bmu)}.
+    \c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & 
+    \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
+    & = \sqrt{\diag \[\X \bSigma \X \tr\] +  
+    \diag \[\X \bmu \bmu \tr \X \tr\]} \\
+    & = \sqrt{ [(\X \bSigma) \circ \X ] \cdot \ones_{p \times 1} + 
+    (\X \bmu) \circ (\X \bmu)}.
   \end{align*}
-The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$, Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}(n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.
+  The calculation requires matrix products of complexity $\mathcal{O}(n^2 p)$,
+  Hadamard products of $\mathcal{O}(np)$ and matrix inversions of $\mathcal{O}
+  (n^3)$, such that the total complexity is of $\mathcal{O}(n^2p)$.
 
-The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p \end{bmatrix} \tr$ are rewritten as:
+  The parameters $\bm{\chi} = \begin{bmatrix} \chi_1 & \cdots & \chi_p 
+  \end{bmatrix} \tr$ are rewritten as:
   \begin{subequations}\label{eq:chicomp}
 	  \begin{align}
-	\bm{\chi} & = \begin{bmatrix} \lambda'_{g(1)} \lambda_2 (\bSigma_{11} + \bmu_1^2) & \cdots & \lambda'_{g(p)} \lambda_2 (\bSigma_{pp} + \bmu_p^2) \end{bmatrix} \tr\\
+	    \bm{\chi} & = \begin{bmatrix} \lambda'_{g(1)} \lambda_2 (\bSigma_{11} +
+	    \bmu_1^2) & \cdots & \lambda'_{g(p)} \lambda_2 (\bSigma_{pp} + \bmu_p^2)
+	    \end{bmatrix} \tr\\
 	    & = \lambda_2 \bLambda' \[ \diag (\bSigma ) + \diag (\bmu \bmu \tr )\] \\
@@
   \end{subequations}
- By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we see that the computation of $\bm{\chi}$ consists of one matrix diagonal of $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two Hadamard products of $\mathcal{O}(p)$.
-Computation of $\bm{\chi}$ is therefore of complexity of $\mathcal{O}(n^2p)$.
+  By inserting (\ref{eq:diagsigma}) and (\ref{eq:mu}) into (\ref{eq:chicomp}) we
+  see that the computation of $\bm{\chi}$ consists of one matrix diagonal of
+  $\mathcal{O}(n^2p)$, matrix products of complexity $\mathcal{O}(n^2p)$ and two
+  Hadamard products of $\mathcal{O}(p)$. Computation of $\bm{\chi}$ is therefore
+  of complexity of $\mathcal{O}(n^2p)$.
 
-\section{Variational Bayes expected joint log likelihood}\label{sec:expectedloglikelihood}
-Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then the variational Bayes expected joint log likelihood is derived as follows:
+  \section{Variational Bayes expected joint log likelihood}
+  \label{sec:expectedloglikelihood}
+  Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then
+  the variational Bayes expected joint log likelihood is derived as follows:
   \begin{align*}
-\E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ] &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}] \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}(\beta_j | \tau_j) | \blambda'^{(k)}] \\
-& \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1} \beta_j^2 | \blambda'^{(k)} \) \\
-& \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
+    \E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ]
+    &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}]
+    \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}
+    (\beta_j | \tau_j) | \blambda'^{(k)}] \\
+    & \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  
+    \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G 
+    \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1}
+    \beta_j^2 | \blambda'^{(k)} \) \\ 
+    & \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - 
+    \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
   \end{align*}
@@
   \section{Full estimation procedure}\label{sec:fullestimationprocedure}
-The expectation and maximisation steps introduced in the Main Document result in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
+  The expectation and maximisation steps introduced in the Main Document Section
+  \ref{md-sec:estimation} result
+  in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
   \begin{algorithm}
-	\caption{Group-regularized empirical Bayes elastic net}\label{alg:summaryalg}
+    \caption{Group-regularized empirical Bayes elastic net}
+    \label{alg:summaryalg}
 	  \begin{algorithmic}[h]
@@
 		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
-		\State{Estimate $\lambda$ by cross-validation of the regular elastic net model}
-		\While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| > \epsilon_1$}
-		\While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
-		\State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ for $j=1, \dots, p$ using (3.9) in the Main Document}
+		  \State{Estimate $\lambda$ by cross-validation of the regular elastic net
+		  model}
+		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
+		  \epsilon_1$}
+		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
+		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
+		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
+		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
+		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
+		  for $j=1, \dots, p$ using (\ref{md-eq:VBupdateequations}) in the MD}
 		  \EndWhile
-		\State{Update $\blambda'$ by (3.10) in the Main Document}
+		  \State{Update $\blambda'$ by (\ref{md-eq:mmlupdateequation}) in the MD}
 		  \EndWhile
@@
   \subsection{Unpenalized covariates}
-Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix $\X$ with a column of ones.
-Penalization of such an intercept parameter is not desirable \cite[]{Hastie_Elements_2009}.
-Additionally, it is often desirable to include unpenalized covariates in the model.
-Common examples of such  covariates in clinical research are patient characteristics, such as age, BMI, and sex.
-In the following we assume the intercept to be included in the unpenalized covariates.
+  Inclusion of an intercept $\beta_0$ is achieved by appending the data matrix
+  $\X$ with a column of ones. Penalization of such an intercept parameter is not
+  desirable \cite[]{Hastie_Elements_2009}. Additionally, it is often desirable 
+  to include unpenalized covariates in the model. Common examples of such 
+  covariates in clinical research are patient characteristics, such as age, BMI,
+  and sex. In the following we assume the intercept to be included in the
+  unpenalized covariates.
 
-To include unpenalized covariates, we divide the model matrix into two parts $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$ unpenalized variables and $\X_r$ are the $r$ penalized variables.
-Let $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with $u$ zero columns and $u$ zero rows.
-Then we have for the current estimate of the covariance matrix $\bSigma$:
+  To include unpenalized covariates, we divide the model matrix into two parts
+  $\X = \begin{bmatrix}\X_u & \X_r\end{bmatrix}$, where $\X_u$ are the $u$
+  unpenalized variables and $\X_r$ are the $r$ penalized variables. Let
+  $\bLambda'_*$ and $\Z_*$ be the matrices $\bLambda'$ and $\Z$ prepended with
+  $u$ zero columns and $u$ zero rows. Then we have for the current estimate of
+  the covariance matrix $\bSigma$:
   \begin{align}\label{eq:unpencov}
-\bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
+    \bSigma &=\(\X \tr \bOmega \X + \lambda_2 \bLambda'_* + 
+    \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda'_* \Z_*\)^{-1} \nonumber \\
     & = 
@@
       \X_u \tr \bOmega \X_u & \X_u \tr \bOmega \X_r \\
-\X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
+      \X_r \tr \bOmega \X_u & \X_r \tr \bOmega \X_r + \lambda_2 \bLambda' +
+      \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z
     \end{bmatrix}^{-1}.
   \end{align}
-With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension $p \times p$.
-Inversion of this matrix is done efficiently by applying the Woodbury identity.
+  With the choice of blocks as in (\ref{eq:unpencov}), blockwise inversion
+  renders the largest required matrix inverse $(\X_r \tr \bOmega \X_r +
+  \lambda_2 \bLambda' + 
+  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z)^{-1}$, of dimension 
+  $p \times p$. Inversion of this matrix is done efficiently by applying the
+  Woodbury identity.
 
   \subsection{Monotonicity of the penalty parameters}
-Enforcing monotonicity of the penalty multipliers is desirable in some settings.
-An example of such a setting is if we have $p$-values from a previous, related study available.
-A $p$-values-based partitioning of the variables may, \textit{a priori}, be expected to render penalty multipliers that increase monotonically with $p$-value.
-That is, larger $p$-values may be expected to yield at least as large penalty multipliers as variables with smaller $p$-values.
-We propose to include this \textit{a priori} assumption by requiring the penalty multipliers to increase monotonically with $p$-value, thereby also stabilizing their estimates.
+  Enforcing monotonicity of the penalty multipliers is desirable in some
+  settings. An example of such a setting is if we have $p$-values from a
+  previous, related study available. A $p$-values-based partitioning of the
+  variables may, \textit{a priori}, be expected to render penalty multipliers
+  that increase monotonically with $p$-value. That is, larger $p$-values may be
+  expected to yield at least as large penalty multipliers as variables with
+  smaller $p$-values. We propose to include this \textit{a priori} assumption by
+  requiring the penalty multipliers to increase monotonically with $p$-value,
+  thereby also stabilizing their estimates.
 
-A natural way of enforcing monotonicity is to extend the constraint in (10) in the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$.
-The problem is still convex and may be numerically solved.
-From experience, however, we note that in combination with this constrained optimisation, the EM algorithm described in Section~3 in the Main Document often converges to a local optimum close to the initialisation.
-We therefore enforce monotonicity through a \textit{post hoc} isotonic regression on the penalty multipliers after every optimisation step.
+  A natural way of enforcing monotonicity is to extend the constraint in (10) in
+  the Main Document with $\lambda'_1 \leq \dots \leq \lambda'_G$. The problem is
+  still convex and may be numerically solved. From experience, however, we note
+  that in combination with this constrained optimisation, the EM algorithm
+  described in Section \ref{md-sec:estimation} in the MD often converges to a
+  local optimum
+  close to the initialisation. We therefore enforce monotonicity through a
+  \textit{post hoc} isotonic regression on the penalty multipliers after every
+  optimisation step.
 
   \subsection{Multiple partitions}
-In many cases the features may be partitioned in more than one way.
-For example, we may have both information on annotation and $p$-values from a previous study available.
-A naive way of incorporating multiple partitions is to cross-tabulate the partitions and create a separate group for every combination.
-This poses two problems: (i) The number of penalty parameters increases exponentially with the number of partitions and (ii) some of these combinations may contain only few features, so that the estimation procedure becomes unstable.
-We propose to stabilise the procedure and keep the number of parameters to estimate manageable by modelling the penalty parameters multiplicatively.
+  In many cases the features may be partitioned in more than one way. For
+  example, we may have both information on annotation and $p$-values from a
+  previous study available. A naive way of incorporating multiple partitions is
+  to cross-tabulate the partitions and create a separate group for every
+  combination. This poses two problems: (i) The number of penalty parameters
+  increases exponentially with the number of partitions and (ii) some of these
+  combinations may contain only few features, so that the estimation procedure
+  becomes unstable. We propose to stabilise the procedure and keep the number of
+  parameters to estimate manageable by modelling the penalty parameters
+  multiplicatively.
 
-We describe our implementation here for two partitions of the features.
-To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively.
-Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively.
-In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively.
-The generalised frequentist elastic net estimator and corresponding conditional prior are now:
+  We describe our implementation here for two partitions of the features. To
+  this end, let $(\G_1(1), \dots,\allowbreak \G_1(G_1))$ and $(\G_2(1), \dots,
+  \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups
+  respectively. Furthermore, in the following we assume that the empty sum and
+  empty product evaluate to 0 and 1, respectively. In this two-partition setting
+  we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$
+  and $\lambda''_{g_2}$, respectively. The generalised frequentist elastic net
+  estimator and corresponding conditional prior are now:
   \begin{subequations}\label{eq:multpart}
 	  \begin{align}
-	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
-	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
+	    \hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  
+	    \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} 
+	    \sqrt{\lambda'_{g_1} \lambda''_{g_2}} 
+	    \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j|
+	    - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1}
+	    \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ 
+	    \cap \G_2(g_2)}}} \beta_j^2, \\
+	    \bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} 
+	    \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} 
+	    \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2}
+	    \frac{\tau_j - 1}{\tau_j} \right).
 	  \end{align}
   \end{subequations}
-After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
+  After switching to the parametrisation in \cite{friedman_regularization_2010},
+  the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 &
+  \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix}
+  \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
   \begin{subequations}\label{eq:estmultpart}
 	  \begin{alignat}{2}
-	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
+	    \blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}
+	    {\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| 
+	    \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| 
+	    \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} 
+	    \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} 
+	    d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       
+	    & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} 
+	    (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
 	  \end{alignat}
   \end{subequations}
-with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$.
-The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine.
-This method naturally generalises to more than two partitions of the data.
+  with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\
+  \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha
+  \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. The optimisation in
+  (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some
+  numerical optimisation routine. This method naturally generalises to more than
+  two partitions of the data.
 
-In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers in a multiplicative way, mainly for convenience.
-First, there are computational reasons: The separation of the square root in the $L_1$-norm penalty term into two terms facilitates the numerical estimation.
-Additionally, multiplicative modelling of the partitions allows for more flexibility in the estimates.
-To see this consider the obvious alternative: additive partition-specific multipliers.
-Since the penalty multipliers are strictly positive, the constraint on the geometric mean makes one large additive penalty estimate difficult to compensate with a smaller estimate, thereby impairing flexibility of the model.
-Large multiplicative penalty multipliers are easier to compensate for by smaller penalty multipliers, rendering this the less restrictive option for modelling the partitions.
+  In (\ref{eq:multpart}) we model the two partition-specific penalty multipliers
+  in a multiplicative way, mainly for convenience. First, there are 
+  computational reasons: The separation of the square root in the $L_1$-norm
+  penalty term into two terms facilitates the numerical estimation. 
+  Additionally, multiplicative modelling of the partitions allows for more
+  flexibility in the estimates. To see this consider the obvious alternative:
+  additive partition-specific multipliers. Since the penalty multipliers are
+  strictly positive, the constraint on the geometric mean makes one large
+  additive penalty estimate difficult to compensate with a smaller estimate,
+  thereby impairing flexibility of the model. Large multiplicative penalty
+  multipliers are easier to compensate for by smaller penalty multipliers,
+  rendering this the less restrictive option for modelling the partitions.
 
-\section{Simulations}
-The quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960}.
-This kappa is measures similarity between two sets of categorical values.
-One set contains the feature selection indicators: $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
+  \section{Simulations}\label{sec:simulations}
+  The quality of feature selection is assessed by Cohen's kappa 
+  \cite[]{cohen_coefficient_1960}. This kappa is measures similarity between two
+  sets of categorical values. One set contains the feature selection indicators:
+  $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains
+  the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
   We calculate Cohen's $\kappa$ as:
 	\begin{align*}
-	\kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}\sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
+	  \kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}
+	  \sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} 
+	  \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + 
+	  \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
 	\end{align*}
-	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and expected correctly identified features, respectively.
-A positive or negative kappa indicate better and worse feature selection than expected by chance, respectively.
-In addition, we evaluate estimation accuracy by mean squared error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
+	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and
+	expected correctly identified features, respectively. A positive or negative
+	kappa indicate better and worse feature selection than expected by chance,
+	respectively. In addition, we evaluate estimation accuracy by mean squared
+	error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
 
   \subsection{Scenario 1}
-	In this sparse scenario we assume differential signal between 5 groups of features. The average signal in a group increases by a factor 1.6 with group number:
+	In this sparse scenario we assume differential signal between 5 groups of
+	features. The average signal in a group increases by a factor 1.6 with group
+	number:
 	\begin{align*}
-	\bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & \beta_{50}^1 \end{matrix}}_{50}} & \cdots & \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr, \\
+	  \bbeta & =  \begin{bmatrix} \smash[b]{\underbrace{ \begin{matrix} 
+	  \mathbf{0} & \cdots & \mathbf{0} \end{matrix}}_{150}} 
+	  & \smash[b]{\underbrace{ \begin{matrix} \beta_{1}^1 & \cdots & 
+	  \beta_{50}^1 \end{matrix}}_{50}} & \cdots & 
+	  \smash[b]{\underbrace{ \begin{matrix} \mathbf{0} & \cdots & \mathbf{0}
+	  \end{matrix}}_{150}} & \smash[b]{\underbrace{ \begin{matrix} 
+	  \beta_{1}^5 & \cdots & \beta^5_{50} \end{matrix}}_{50}} \end{bmatrix} \tr,
+	  \\
 	  & \, \\
-	\beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots, 5  \text{, } j=1, \dots, 50,
+	  \beta_{j}^g & \sim \mathcal{U} (0, b_1 \cdot1.6^{g-1}) \text{, } g=1, \dots,
+	  5  \text{, } j=1, \dots, 50,
 	\end{align*}
-	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$.
-The model parameters clearly do not follow the relatively heavy-tailed elastic net distribution.
-Because of this mismatch between the assumed and true distributions, we expect poor performance by \texttt{gren}.
-Furthermore, we introduce correlation between the features by having covariance elements $\bSigma_{ij} = 0.5^{|i-j|}$.
+	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. The
+	model parameters clearly do not follow the relatively heavy-tailed elastic net
+	distribution. Because of this mismatch between the assumed and true 
+	distributions, we expect poor performance by \texttt{gren}. Furthermore, we
+	introduce correlation between the features by having covariance elements
+	$\bSigma_{ij} = 0.5^{|i-j|}$.
 
 	\subsection{Scenario 2}
-	The second scenario contains many groups of features and is geared towards the non-adaptive group-regularized methods.
-These methods are able to incorporate the grouping of the features, but were originally designed for small interpretable groups, such as dummies of a categorical variable.
-\texttt{gren} on the other hand, relies on large groups of features to accurately estimate group-specific penalty weights.
-The estimation of many penalty weights in small groups is inefficient and computationally challenging.
-We therefore expect \texttt{gren} to under-perform compared to the other methods.
+	The second scenario contains many groups of features and is geared towards the
+	non-adaptive group-regularized methods. These methods are able to incorporate
+	the grouping of the features, but were originally designed for small 
+	interpretable groups, such as dummies of a categorical variable. \texttt{gren}
+	on the other hand, relies on large groups of features to accurately estimate
+	group-specific penalty weights. The estimation of many penalty weights in
+	small groups is inefficient and computationally challenging. We therefore
+	expect \texttt{gren} to under-perform compared to the other methods.
 
-	For this scenario we follow the second simulation example in \cite{meier_group_2008}.
-They set $n=100$, $G=250$, and $p=1000$, such that we have 4 features per group.
-The first 10 groups are active and those 40 model parameters are set to 0.2.
-The predictors are simulated from a $p$-dimensional Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.
+	For this scenario we follow the second simulation example in 
+	\cite{meier_group_2008}. They set $n=100$, $G=250$, and $p=1000$, such that we
+	have 4 features per group. The first 10 groups are active and those 40 model
+	parameters are set to 0.2. The predictors are simulated from a $p$-dimensional
+	Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.
 
 	\subsection{Scenario 3}
-	A third scenario is modelled by considering a situation in which there is no differential signal between the groups, but the correlation within groups is very high.
-We conjecture that this leads to overestimation of effect within the groups by \texttt{gren} and consequently, performance will suffer.
+	A third scenario is modelled by considering a situation in which there is no
+	differential signal between the groups, but the correlation within groups is
+	very high. We conjecture that this leads to overestimation of effect within
+	the groups by \texttt{gren} and consequently, performance will suffer.
 
-	To that end we sample the predictors in $G=5$ separate blocks of 200 variables, corresponding to the partitioning.
-The variances of the features are set to one, while all covariances are 0.7.
-The 1000 model parameters are simulated in two steps.
-First, we draw 1000 parameters from the elastic net distribution, parametrised as in \cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$.
-The second step consists of setting the (in absolute value) smallest 100 parameters in a group to zero.
-This results in a total of 500 zero and 500 non-zero parameters, evenly distributed over the five groups.
+	To that end we sample the predictors in $G=5$ separate blocks of 200 
+	variables, corresponding to the partitioning. The variances of the features
+	are set to one, while all covariances are 0.7. The 1000 model parameters are
+	simulated in two steps. First, we draw 1000 parameters from the elastic net
+	distribution, parametrised as in \cite{friedman_regularization_2010}, where
+	$\alpha=0.5$ and $\lambda=100$. The second step consists of setting the 
+	(in absolute value) smallest 100 parameters in a group to zero. This results
+	in a total of 500 zero and 500 non-zero parameters, evenly distributed over
+	the five groups.
 
 	\subsection{Scenario 4}
-	This scenario is similar to Scenario 3, except each groups penalty parameter is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the covariance matrix of the predictors is block diagonal, with block sizes $25 \times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$.
-We expect \texttt{gren} to pick up this differential signal and outperform the other methods.
+	This scenario is similar to Scenario 3, except each groups penalty parameter
+	is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the
+	covariance matrix of the predictors is block diagonal, with block sizes $25
+	\times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$.
+	We expect \texttt{gren} to pick up this differential signal and outperform the
+	other methods.
 
 	\subsection{Scenario 5}
-	In the last scenario we create a sparse true model that contains some differential signal. We impl
+	In the last scenario we create a sparse true model that contains some 
+	differential signal. We impl
 
-\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
-\subsection{Partitioning based on differential expression}\label{sec:mirna}
-The data set is from a deep sequencing analysis on microRNA \cite[]{neerincx_combination_2018}.
-The study was done in 88 treated colorectal cancer patients, with the aim of classifying treatment response, coded as either non-progressive/remission (70 patients) or progressive (18 patients).
-After pre-processing and normalisation, 2114 microRNAs remained.
-In addition to the 2114 microRNAs we incorporated 4 unpenalized clinical covariates into the analysis: prior use of adjuvant therapy (binary), the type of systemic treatment regimen (ternary), age, and primary tumor differentiation (binary).
+  \section{Application to MicroRNAs in colorectal cancer}\label{sec:colorectal}
+  \subsection{Additional comparisons}
+  We compared \texttt{gren} to two methods in addition to the ones in the MD
+  Section \ref{md-sec:colorectal}: 
+  the sparse group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and 
+  the group exponential lasso (\texttt{gel}) by \cite{breheny_group_2015}. For
+  all methods with a tuning parameter $\alpha$, we set 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ and present the AUC and Brier skill score 
+  on the test set in Figures 
+  \ref{fig:lines_micrornaseq_colorectal_cancer_res1_auc}--
+  \ref{fig:lines_micrornaseq_colorectal_cancer_res1_briers}.
+<<lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated AUC for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
+@
 
-In a preliminary experiment on different subjects, the microRNA expression levels of metastatic colorectal tumour tissue were compared to normal non-colorectal tissue and primary colorectal tumour tissue was compared to primary colorectal tumour tissue \cite[]{neerincx_mir_2015}.
-This yielded 221 microRNAs that were differentially expressed in both comparisons ($\text{FDR} \leq 0.05$), versus 1893 not differentially expressed microRNAs.
-We expect that incorporation of this partitioning enhances therapy response classification, because tumor-specific miRNAs might be more relevant than non-specific ones.
-In addition, we divided the differentially expressed microRNAs even further into 127 highly differentially expressed microRNAs ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs ($0.001 < \text{FDR} \leq 0.05$).
-We will refer to this second partitioning as the three-group setting, as opposed to the first, the two-group setting.
+<<lines_micrornaseq_colorectal_cancer_res1_briers, fig.cap="Estimated Brier skill score for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
+@
 
-We compared \texttt{gren} to ridge and elastic net regression, and \texttt{GRridge}.
-For the elastic net methods (including \texttt{gren}) we set $\alpha \in \{ 0.05, 0.5, 0.95\}$.
-The estimated penalty multipliers are according to expectation: in all group-regularized methods, the 221 differentially expressed microRNAs receive the smallest penalty (Figure \ref{fig:col_bar}a).
-In the three-group setting the pattern is again as expected (Figure \ref{fig:col_bar}b): the highly expressed group, receives the lowest penalty, followed by the medium expressed group, while the non-expressed group receives the strongest penalty.
-\texttt{GRridge} is not able to distinguish between the medium and non-differentially expressed groups of microRNAs.
+  \subsection{Random groups}\label{sec:randomgroups}
+  Considering that the features may be partitioned into many groups, with one 
+  parameter per group, we have to be aware of overfitting risks. We investigated
+  this using the data introduced in Section \ref{md-sec:colorectal} of the MD, 
+  randomly dividing the features into three groups. We fixed the group sizes to 
+  the group sizes used in the MD. Under this random 
+  partitioning of the features, we expect all penalty multipliers to be 
+  estimated as one if no overfitting occurs.
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.83\linewidth]{grEBEN_mirseq_Maarten_bar1.png}
-	\caption{Estimated penalty multipliers for (a) the two-group setting and (b) the three-group setting.}
-	\label{fig:col_bar}
-\end{figure}
-Predictive performance of the methods is measured by AUC and BSS, both estimated by leave-one-out cross-validation (LOOCV), and shown in Figure \ref{fig:col_per}.
-All group-regularized elastic net models consistently outperform their non-group-regularized counterparts in both settings, in terms of AUC and BSS.
-In both settings, \texttt{gren} with $\alpha=0.5$ and $\alpha=0.95$ outperforms the other methods for a large range of model sizes in terms of AUC.
-This is especially true for the larger models.
-\texttt{gren} with $\alpha=0.5$ outperforms the other methods with respect to BSS in the smaller model ranges, while \texttt{gren} with $\alpha=0.05$ becomes competitive for the larger models.
-A general pattern in both settings is that \texttt{GRridge} and \texttt{gren} with $\alpha=0.05$ perform similarly.
-This is not surprising, because an elastic net model with $\alpha=0.05$ is close to a ridge model, in terms of penalisation.
-In general, the classifiers with three group penalties are better than the ones with two group penalties when relatively few features are used.
-Comparing, for example, the models of size 20 for \texttt{gren} with $\alpha=0.5$, we have an AUC of 0.68 for the two-group setting, versus an AUC of 0.78 for the three-group setting.
-Note that models with few features are often desirable for clinical implementation.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=1\linewidth]{grEBEN_mirseq_Maarten_combined_performance.png}
-	\caption{Cross-validated performance as (a) AUC in the two-group setting, (b) Brier skill score in the two-group setting, (c) AUC in the three-group setting, and (d) Brier skill score in the three-group setting against the number of selected features for each method.}
-	\label{fig:col_per}
-\end{figure}
+<<>>=
+res <- read.table("results/micrornaseq_colorectal_cancer_res2.csv", 
+                  stringsAsFactors=FALSE)
+medians <- round(apply(res, 2, median), 2)
+@
+  We compared the estimated multipliers to the estimates by \texttt{GRridge}. 
+  Since these results depend on one specific randomisation of the groups, we 
+  repeated the procedure 100 times and present the results in Figure 
+  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res2_multipliers}. 
+  From this figure we see that the estimates for \texttt{gren} are relatively 
+  close to one.
+  The estimates by \texttt{GRridge} show more variation. We found no strong
+  indication of bias, with median estimated penalty parameters 
+  \Sexpr{medians[c(1:3)]} for
+  \texttt{GRridge} and \Sexpr{medians[c(4:6)]}; \Sexpr{medians[c(7:9)]},
+  and \Sexpr{medians[c(10:12)]} for \texttt{gren} with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
+<<boxplots_micrornaseq_colorectal_cancer_res2_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 3 random groups of features.", out.width="50%", fig.asp=1>>=
+@
 
-\subsection{Random groups}
-Considering that the features may be partitioned into many groups, with one parameter per group, we have to be aware of overfitting risks.
-We investigated this using the data introduced in Section \ref{sec:mirna}, randomly dividing the features into three groups.
-We fixed the group sizes to the group sizes used in the three-group setting in Section \ref{sec:mirna}.
-Under this random partitioning of the features, we expect all penalty multipliers to be estimated as one if overfitting does not occur.
-
-We compared the estimated multipliers to the estimates by \texttt{GRridge}.
-Since these results depend on one specific randomisation of the groups, we repeated the procedure 100 times and present the results in Figure \ref{fig:random1}.
-From this figure we see that the estimates for \texttt{gren} are close to one.
-The estimates by \texttt{GRridge} show much more variation.
-Additionally, \texttt{GRridge} gives slightly biased estimates: the median penalty parameter estimates are 1.16, 1, and 0.85 for the three groups.
-In contrast, the median estimates of \texttt{gren} are 1, 1, and 0.99.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.5\linewidth]{grEBEN_mirseq_Maarten_res6_boxplot.png}
-	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 3 random groups in colorectal miRNA data.}
-	\label{fig:random1}
-\end{figure}
-
-We repeated the simulation with ten evenly sized groups.
-Figure \ref{fig:random2} presents the estimated penalty multipliers for 100 random splits of the microRNA features into 10 evenly sized groups for \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
-From the Figure we see that \texttt{GRridge} shows more variation around one then the other models.
-In addition, the median penalty multipliers for the three \texttt{gren} models are all between 0.99 and 1.02, while the \texttt{GRridge} median estimates range from 1.04 to 1.09.
-This indicates that \texttt{GRridge} suffers from a slight bias in the estimation, while the \texttt{gren} estimates appear to be unbiased.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.9\linewidth]{grEBEN_mirseq_Maarten_res8_boxplot.png}
-	\caption{Estimated penalty multipliers of the group-regularized methods for 100 repeats of 10 random groups in colorectal miRNA data.}
-	\label{fig:random2}
-\end{figure}
+<<>>=
+res <- read.table("results/micrornaseq_colorectal_cancer_res3.csv", 
+                  stringsAsFactors=FALSE)
+medians <- round(sapply(c("grridge", paste0("gren", 1:3)), function(s) {
+  median(as.matrix(res[, grepl(s, colnames(res))]))}), 2)
+@
+  We repeated the simulation with ten evenly sized groups. Figure 
+  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers} presents 
+  the estimated penalty multipliers for 100 random 
+  splits of the microRNA features into 10 evenly sized groups for 
+  \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$. 
+  From the Figure we see that \texttt{GRridge} shows more variation around one 
+  then the other models. There is no indication of bias: the median penalty 
+  multiplier for 
+  \texttt{GRridge} is \Sexpr{medians[1]}, while the median multipliers in
+  \texttt{gren} are \Sexpr{medians[c(2:4)]} for 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
+<<boxplots_micrornaseq_colorectal_cancer_res3_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
+@
 
   \subsection{Stability of selection}
-One common criticism on the elastic net, and especially the lasso, is the instability of feature selection.
-That is, different data sets yield different sets of selected features.
-To investigate the stability of feature selection by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA data \cite[]{neerincx_combination_2018}, where we stratified by treatment response.
-For each of these bootstrap samples we estimated the regular elastic net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected 25 features (including the 5 unpenalized covariates).
-We calculated the size of all ${50}\choose{2}$ intersections of selected features and present the results in Figure \ref{fig:bootstrap}.
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.9\linewidth]{gren_mirseq_Maarten_overlap.png}
-	\caption{Overlap between stratified bootstrap samples for the elastic net models with (a) $\alpha=0.05$, (b) $\alpha=0.5$, and (c) $\alpha=0.95$.}
-	\label{fig:bootstrap}
-\end{figure}
-From this figure it can be seen that stability of feature selection in \texttt{gren} is higher than in the regular elastic net for all three $\alpha$'s.
+  One common criticism on the elastic net, and especially the lasso, is the 
+  instability of feature selection. That is, different data sets yield different
+  sets of selected features. To investigate the stability of feature selection 
+  by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA 
+  data \cite[]{neerincx_combination_2018}, where we stratified by the target of 
+  prediction, treatment response. 
+  For each of these bootstrap samples we estimated the regular elastic
+  net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected
+  25 features (including the 5 unpenalized covariates). We calculated the size 
+  of all $\genfrac(){0pt}{2}{50}{2}$ intersections of selected features and
+  present the results in Figure 
+  \ref{fig:hist_micrornaseq_colorectal_cancer_res4_overlap}.
+<<hist_micrornaseq_colorectal_cancer_res4_overlap, fig.cap="Stability of feature selection in colorectal cancer example for (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=2/3>>=
+@
 
-\section{Empirical-variational Bayes for the ridge and lasso}\label{sec:evbridgelasso}
+  \section{Application to RNAseq in lymph node cancer}
+  \label{sec:lymphnode}
+  In addition to the results presented in Section \ref{md-sec:lymphnode} of the 
+  MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 
+  \cite{simon_sparse-group_2013} and the group exponential lasso (\texttt{gel}) 
+  by \cite{breheny_group_2015}. The original group lasso methods are not 
+  designed to deal with overlapping groups, so in addition to using the versions
+  that allow for overlapping groups, we 
+  cross-tabulated the two co-data partitions to create one paritioning with 25 
+  groups for \texttt{cMCP} and \texttt{gel}.
+  We use $\alpha \in \{0.05, 0.5, 0.95 \}$ for the
+  methods that require this extra tuning parameter. The AUCs and Brier skill 
+  scores calculated on the
+  validation set are presented in Figures 
+  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_auc}--
+  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_briers}
+<<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated AUC for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_rnaseq_lymph_node_metastasis_res1_briers, fig.cap="Estimated Briers skill score for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
+@
+
+  \section{Application to metabolomics in alzheimer}\label{sec:alzheimer}
+	We applied our method to an Alzheimer's dataset from 
+	\cite{de_leeuw_blood-based_2017}. The
+	data consist of 87 diagnosed Alzheimer's patients with the $\epsilon$4
+	allele of the APOE gene and 87 subjects with subjective un-diagnosed cognitive
+	decline, without the $\epsilon$4 allele. The aim of the study was to diagnose 
+	subjects with Alzheimer's disease based on their metabolic profile. After 
+	pre-processing, 230 metabolite expression levels were obtained. To enhance
+	classification performance, we included two sources of co-data: (a) the 
+	relative standard devitation of the measurements (RSD), a commonly 
+	used quality score (smaller is better) for metabolites, binned into five 
+	groups, and (b) the node degree of the metabolites, based on a differential 
+	network analysis of the metabolites, binned into three categories: 0 degree, 
+	positive, lower than average degree, and postive, higher than average degree. 
+	We expect a lower RSD and a higher node degree to be indicative of an 
+	important metabolite and therefore receive less penalty.
+
+	To estimate performance measures AUC and Briers skill score, we split the data
+	into 120 training instances and 54 test instances. 
+	The training data is used to fit the models, while the test data
+	is used to estimate the performance measures on. This resulted in the 
+	estimated multipliers and performance measures presented in Figures 
+	\ref{fig:barplot_metabolomics_alzheimer_res1}--
+	\ref{fig:lines_metabolomics_alzheimer_res1_briers}.
+<<barplot_metabolomics_alzheimer_res1, fig.cap="Estimated penalty multiplies for the (a) quality score and the (b) network degree in the alzheimer example.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Estimated AUC for the alzheimer example.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_metabolomics_alzheimer_res1_briers, fig.cap="Estimated Brier skill score for the alzheimer example.", out.width="100%", fig.asp=1>>=
+@
+
+  In this example, \texttt{GRridge} and \texttt{gren} underperform due to a 
+  small number of features to learn the penalty parameters from. The other 
+  methods estimate one global penalty parameter, so do not suffer from the 
+  lower dimensionality of the problem.
+
+  \section{Application to microRNAs in cervical cancer}\label{sec:cervical}
+  A deep sequencing analysis on small non-coding ribonucleic acid (microRNAseq) 
+  was performed on 24 women with high-grade cervical intraepithelial 
+  neoplasia (CIN3) and 32 healthy women for the purpose of finding relevant 
+  screening markers for cervical cancer screening. The next generation 
+  sequencing analysis resulted in 2,576 transcripts. The data was normalized 
+  and pre-processed, rendering 752 transcripts. More details of the
+  data sets and the pre-processing steps are found in the supplementary   
+  material of \cite{novianti_better_2017}.
+
+  The 752 microRNAs were divided in three classes based on their conservation 
+  status. Conservation status refers to whether the microRNAs are only found in
+  humans or are conserved across different species. They come in three classes:
+  (a) non-conserved (535), (b) conserved in mammals (70), and (c) 
+  conserved across most vertebrates (147). We expect (c) to contain the most 
+  important microRNAs, followed by (b) and (a), because these microRNAs have not
+  been lost during evolution and might therefore regulate important processes.
+
+<<barplot_micrornaseq_cervical_cancer_res1, fig.cap="Estimated penalty multiplies for conservation status in cervical cancer example.", out.width="50%", fig.asp=1>>=
+@
+
+<<lines_micrornaseq_cervical_cancer_res1_auc, fig.cap="Cross-validated AUC in cervical cancer example", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_micrornaseq_cervical_cancer_res1_briers, fig.cap="Cross-validated Brier skill score in cervical cancer example", out.width="100%", fig.asp=1>>=
+@
+
+  \section{Empirical-variational Bayes for the ridge and lasso}
+  \label{sec:evbridgelasso}
   \subsection{Ridge regression}
-Ridge regression is a special case of the elastic net, that does not include an $L_1$-norm on the parameters. Its estimator is given by:
+  Ridge regression is a special case of the elastic net, that does not include
+  an $L_1$-norm on the parameters. Its estimator is given by:
   $$
-\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_2}{2} \norm{\bbeta}_2.
+  \hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+  \frac{\lambda_2}{2} \norm{\bbeta}_2.
   $$
-The Bayesian version of ridge regression is regression under a Gaussian prior. The generalised version of this prior is:
+  The Bayesian version of ridge regression is regression under a Gaussian prior.
+  The generalised version of this prior is:
   $$
-\bbeta \sim \prod_{g=1}^G \prod_{j \in \G(g)} \N(0,(\lambda_2 \lambda'_g)^{-1}).
+  \bbeta \sim \prod_{g=1}^G \prod_{j \in \G(g)} 
+  \N(0,(\lambda_2 \lambda'_g)^{-1}).
   $$
-Although the posterior is analytically calculated in the linear case, in the logistic case we must approximate it iteratively. The variational distributions in this case are:
+  Although the posterior is analytically calculated in the linear case, in the
+  logistic case we must approximate it iteratively. The variational 
+  distributions in this case are:
   \begin{align*}
@@
   \end{align*}
-The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net and are updated, together with $\bSigma$, until convergence:
+  The parameters $c_i$ and $\bmu$ remain unchanged compared to the elastic net
+  and are updated, together with $\bSigma$, until convergence:
   $$
-\bSigma^{t+1} = (\X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda')^{-1}\text{, with } \bOmega^{(t)} = \diag \[ \frac{m_i}{2 c^{(t)}_i} \tanh \( \frac{c^{(t)}_i}{2} \) \].
+  \bSigma^{t+1} = (\X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda')^{-1}
+  \text{, with } \bOmega^{(t)} = \diag \[ \frac{m_i}{2 c^{(t)}_i} 
+  \tanh \( \frac{c^{(t)}_i}{2} \) \].
   $$
 
-Updating the penalty parameters is done by solving the following (convex) constraint optimisation problem:
+  Updating the penalty parameters is done by solving the following (convex)
+  constraint optimisation problem:
   \begin{align*}
-\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+    \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G
+    |\G(g)| \log (\lambda'_g) - \frac{\lambda_2}{2} \sum_{g=1}^G 
+    \lambda'_g d^{(k)}_g \\
     & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
@@
   \subsection{Lasso regression}
-In lasso regression, another special case of the elastic net, we have the following estimator:
+  In lasso regression, another special case of the elastic net, we have the
+  following estimator:
   $$
-\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1.
+  \hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - 
+  \frac{\lambda_1}{2} \norm{\bbeta}_1.
   $$
-The Bayesian counterpart is regression under a Laplace prior on the model parameters. The generalised Bayesian prior may decomposed as:
+  The Bayesian counterpart is regression under a Laplace prior on the model
+  parameters. The generalised Bayesian prior may decomposed as:
   \begin{align*}
     \bbeta | \btau & \sim \prod_{j=1}^p \mathcal{N} \left(0,\tau_j \right), \\
-\btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \text{Exp} \left( \frac{\sqrt{\lambda'_g}\lambda_1^2}{8} \right).
+    \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \text{Exp} 
+    \left( \frac{\sqrt{\lambda'_g}\lambda_1^2}{8} \right).
   \end{align*}
 
-The variational distributions are similar as in the elastic net, except that we do not re-parametrise the $\tau_j$:
+  The variational distributions are similar as in the elastic net, except that
+  we do not re-parametrise the $\tau_j$:
   \begin{align*}
@@
     q^*_{\bomega} (\bomega) & \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i), \\
-q^*_{\btau}(\btau) & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{GIG} \(\frac{1}{2}, \phi_g, \chi_j\).
+    q^*_{\btau}(\btau) & \sim \prod_{g=1}^G \prod_{j \in \G(g)} 
+    \mathcal{GIG} \(\frac{1}{2}, \phi_g, \chi_j\).
   \end{align*}
-The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic net, while the $\phi_g$ are given by:
+  The parameters $c_i$ and $\bmu$ and remain unchanged compared to the elastic
+  net, while the $\phi_g$ are given by:
   $$
@@
   $$
-The following parameters are updated, together with the unchanged $c_i$ and $\bmu$, until convergence:
+  The following parameters are updated, together with the unchanged $c_i$ and
+  $\bmu$, until convergence:
   \begin{align*}
-\bSigma^{(t + 1)} &= ( \X \tr \bOmega^{(t)} \X + \Z^{(t)})^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\(\sqrt{\frac{\phi_g(j)}{\chi_j^{(t)}}}\), \\
-\chi_j^{(t + 1)} &= \bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \text{, for } j=1, \dots, p.
+    \bSigma^{(t + 1)} &= ( \X \tr \bOmega^{(t)} \X + \Z^{(t)})^{-1}, \\ 
+    & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) 
+    \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } 
+    \Z^{(t)}=\diag\(\sqrt{\frac{\phi_g(j)}{\chi_j^{(t)}}}\), \\
+    \chi_j^{(t + 1)} &= \bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 
+    \text{, for } j=1, \dots, p.
   \end{align*}
 
-Updating the penalty parameters is done by solving the following (convex) constraint problem:
+  Updating the penalty parameters is done by solving the following (convex)
+  constraint problem:
   \begin{align*}
-\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{\lambda_1^2}{8} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+    \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \sum_{g=1}^G |\G(g)|
+    \log (\lambda'_g) - \frac{\lambda_1^2}{8} \sum_{g=1}^G 
+    \lambda'_g d^{(k)}_g \\
     & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1,
@@
   $$
-d^{(k)}_g = \sum_{j \in \G(g)} \( \frac{1}{\phi_g^{(k)}} + \frac{\chi^{(k)}_j}{\phi_g^{(k)}} \).
+  d^{(k)}_g = \sum_{j \in \G(g)} \( \frac{1}{\phi_g^{(k)}} + 
+  \frac{\chi^{(k)}_j}{\phi_g^{(k)}} \).
   $$ 
-Note that the $\phi_g$ are updated after every penalty multiplier update iteration as well, in contrast to the elastic net.
+  Note that the $\phi_g$ are updated after every penalty multiplier update  
+  iteration as well, in contrast to the elastic net.
 
-\bibliographystyle{author_short3.bst} 
+	\bibliographystyle{biorefs}
 	\bibliography{refs}

### 2019-03-22 code/simulations.R

@@ -12,7 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-parallel <- FALSE
+parallel <- TRUE
 
 ### libraries
 library(gren)
@@ -206,6 +206,8 @@ if(parallel) {
 }
 
 res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
+  
+  print(paste("rep", k))
   set.seed(2019 + k)
   
   ztrain <- rmvnorm(n, rep(0, G), Sigma)
@@ -317,3 +319,411 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
 }
 if(parallel) {stopCluster(cluster)}
 save(res, file="results/simulations_res2.Rdata")
+
+
+
+################################# simulation 3 #################################
+n <- 100
+ntest <- 1000
+p <- 1000
+alpha <- 0.5
+lambda <- 100
+G <- 5
+q <- 0.5
+
+rho <- 0.7
+Sigma <- matrix(rho, nrow=p/G, ncol=p/G)
+diag(Sigma) <- 1
+
+part <- rep(c(1:G), each=p/G)
+csel <- 2^c(1:8)
+nreps <- 100
+
+### analysis splits in parallel
+ncores <- min(detectCores() - 1, nreps)
+cluster <- makeForkCluster(ncores)
+if(parallel) {
+  registerDoParallel(cluster)
+} else {
+  registerDoSEQ()
+}
+
+res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
+  set.seed(2019 + k)
+  
+  beta <- as.numeric(sapply(1:G, function(g) {
+    b <- renet(p/G, lambda*alpha, 0.5*(1 - alpha)*lambda);
+    b[abs(b)<=quantile(abs(b), q)] <- 0
+    return(b)}))
+  xtrain <- do.call(cbind, replicate(G, rmvnorm(n, mean=rep(0, p/G),
+                                                sigma=Sigma), simplify=FALSE))
+  ytrain <- rbinom(n, 1, as.numeric(1/(1 + exp(-xtrain %*% beta))))
+  
+  xtest <- do.call(cbind, replicate(G, rmvnorm(ntest, mean=rep(0, p/G),
+                                               sigma=Sigma), simplify=FALSE))
[…]

### 2019-03-22 code/simulations.R
commit before pull

@@ -12,7 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-parallel <- FALSE
+parallel <- TRUE
 
 ### libraries
 library(gren)
@@ -197,7 +197,8 @@ csel <- 2^c(1:8)
 nreps <- 100
 
 ### analysis splits in parallel
-ncores <- min(detectCores() - 1, nreps)
+# ncores <- min(detectCores() - 1, nreps)
+ncores <- 50
 cluster <- makeForkCluster(ncores)
 if(parallel) {
   registerDoParallel(cluster)
@@ -231,11 +232,14 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   
   # fitting models
   fit.gren1 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.05, 
-                    standardize=TRUE, trace=FALSE, psel=csel)
+                    standardize=TRUE, trace=FALSE, psel=csel,
+                    control=list(maxit.vb=10, maxit=200))
   fit.gren2 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.5, 
-                    standardize=TRUE, trace=FALSE, psel=csel)
+                    standardize=TRUE, trace=FALSE, psel=csel,
+                    control=list(maxit.vb=10, maxit=200))
   fit.gren3 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.95, 
-                    standardize=TRUE, trace=FALSE, psel=csel)
+                    standardize=TRUE, trace=FALSE, psel=csel,
+                    control=list(maxit.vb=10, maxit=200))
   
   fit.grridge <- grridge(t(xtrain), ytrain, list(part=split(1:p, part)))
   

### 2019-03-22 docs/manuscript_biostatistics.Rnw

--- 2019-02-05 docs/manuscipt.Rnw
@@
-\documentclass[a4paper,hidelinks]{article}
-\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}
+% set options
+<<settings, include=FALSE, echo=FALSE>>=
+opts_knit$set(root.dir="..", base.dir="../figs")
+opts_chunk$set(fig.align='center', fig.path="../figs/", 
+               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
+@
+
+% read figure code
+<<figures, include=FALSE>>=
+read_chunk("code/figures.R")
+@
+
+\documentclass[oupdraft]{bio}
+\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
+\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable}
 
@@
 
-% settings
-\graphicspath{{../figs/}}
-\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
-\setlength{\textwidth}{15cm}
+\begin{document}
+
+	% Title of paper
 	\title{Adaptive group-regularized logistic elastic net regression}
-\date{}
-\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}
 
-\begin{document}
+	% List of authors, with corresponding author marked by asterisk
+	\author{MAGNUS M. M\"UNCH$^{\ast 1,2}$, CAREL F.W. PEETERS$^{1}$, AAD W. VAN 
+	DER VAART$^{2}$, \\* MARK A. VAN DE WIEL$^{1,3}$\\[4pt]
+	% Author addresses
+	\textit{1. Department of Epidemiology \& Biostatistics, Amsterdam Public 
+	Health research institute, Amsterdam University medical centers, PO Box 7057,
+	1007 MB Amsterdam, The Netherlands \\*
+	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
+	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom}
+	\\[2pt]
+	% E-mail address for correspondence
+	{m.munch@vumc.nl}}
+
+	% Running headers of paper:
+	\markboth%
+	% First field is the short list of authors
+	{M. M. M\"unch and others}
+	% Second field is the short title of the paper
+	{Adaptive group-regularized logistic elastic net regression}
 
@@
 
-	\noindent
-	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
-	Amsterdam, The Netherlands\\
-	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
-	3. MRC Biostatistics Unit, Cambridge, United Kingdom \\
+	% Add a footnote for the corresponding author if one has been
+	% identified in the author list
+	\footnotetext{To whom correspondence should be addressed.}
 
 	\begin{abstract}
-		{In high-dimensional data settings, additional information on the features is often available.
-Examples of such external information in omics research are: (a) $p$-values from a previous study, (b) a summary of prior information, and (c) omics annotation.
-The inclusion of this information in the analysis may enhance classification performance and feature selection, but is not straightforward in the standard regression setting.
-As a solution to this problem, we propose a group-regularized (logistic) elastic net regression method, where each penalty parameter corresponds to a group of features based on the external information.
-The method, termed \texttt{gren}, makes use of the Bayesian formulation of logistic elastic net regression to estimate both the model and penalty parameters in an approximate empirical-variational Bayes framework.
-Simulations and an application to a colon cancer microRNA study show that, if the partitioning of the features is informative, classification performance and feature selection are indeed enhanced.}
+		{In high-dimensional data settings, additional information on the features 
+		is often available. Examples of such external information in omics research
+		are: (a) $p$-values from a previous study, (b) a summary of prior 
+		information, and (c) omics annotation. The inclusion of this information in
+		the analysis may enhance classification performance and feature selection,
+		but is not straightforward in the standard regression setting. As a solution
+		to this problem, we propose a group-regularized (logistic) elastic net
+		regression method, where each penalty parameter corresponds to a group of
+		features based on the external information. The method, termed 
+		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
+		regression to estimate both the model and penalty parameters in an 
+		approximate empirical-variational Bayes framework. Simulations and
+		applications in two different molecular marker and cancer type studies show
+		that, if the partitioning of the features is informative, classification 
+		performance and feature
+		selection are indeed enhanced.}
+		{Empirical Bayes; High-dimensional data; Prediction; Variational Bayes}
 	\end{abstract}
 
-	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes
+	\section{Introduction}\label{sec:introduction}
+	Prediction from high-dimensional data is a challenge common to many fields of
+	research. Examples of such high-dimensional prediction problems arise in
+	computer vision, stock market prediction and disease diagnosis from omics 
+	data. In this paper, we specifically focus on the latter setting. High
+	dimensionality of data introduces several issues in the estimation of 
+	prediction models, especially: (a) unidentifiable models, (b) highly 
+	correlated predictor variables and (c) non-trivial variable selection.
 
-	\noindent\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}
+	Attempts to tackle one or more of these issues have lead to the development of
+	many prediction methods \cite[]{fan_selective_2010}. Here we will focus on the
+	penalization approach. A well-known and popular penalization method is elastic
+	net regression \cite[]{zou_regularization_2005}, with its special cases ridge
+	regression \cite[]{hoerl_ridge_1970} and lasso regression 
+	\cite[]{tibshirani_regression_1996}. The elastic net approach takes issues 
+	(a)-(c) into account and has yielded many successful extensions. Recent work 
+	on elastic net regression (and its special cases) has focussed on increasing
+	prediction accuracy by the inclusion of prior knowledge on the variables
+	\cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.
 
-	\section{Introduction}
-	Prediction from high-dimensional data is a challenge common to many fields of research.
-Examples of such high-dimensional prediction problems arise in computer vision, stock market prediction and disease diagnosis from omics data.
-In this paper, we specifically focus on the latter setting.
-High dimensionality of data introduces several issues in the estimation of prediction models, especially: (a) unidentifiable models, (b) highly correlated predictor variables and (c) non-trivial variable selection.
+	In omics research, sources of such prior knowledge on the variables are often
+	available. The information can, for example, come in the form of (a) results 
+	on the same molecular features obtained in a previous study (e.g. $p$-values),
+	(b) information from a publicly available database that summarizes the prior
+	information on the molecular features involved (e.g., the Cancer Gene Census
+	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. a gene's 
+	chromosome location). Although this information can rarely be directly
+	included in the statistical analysis, it may still be useful and informative
+	for the study at hand.
 
-	Attempts to tackle one or more of these issues have lead to the development of many prediction methods \cite[]{fan_selective_2010}.
-Here we will focus on the penalization approach.
-A well-known and popular penalization method is elastic net regression \cite[]{zou_regularization_2005}, with its special cases ridge regression \cite[]{hoerl_ridge_1970} and lasso regression \cite[]{tibshirani_regression_1996}.
-The elastic net approach takes issues (a)-(c) into account and has yielded many successful extensions.
-Recent work on elastic net regression (and its special cases) has focussed on increasing prediction accuracy by the inclusion of prior knowledge on the variables \cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.
+	Two such studies are introduced in Sections \ref{sec:colorectal} and
+	\ref{sec:lymphnode}. The first study is concerned with treatment 
+	response prediction for colorectal cancer patients, based on sequenced
+	microRNAs. In the second study we attempt to predict lymph node metastases,
+	using RNAseq data. In both settings additional information on the 
+	features is available in the form of a partitioning. Furthermore, in both
+	examples data is sparse, so any extra included information may help to boost
+	predictive performance. Inclusion of such feature partitions in 
+	the analysis is not straightforward with standard methods.
 
-	In omics research, sources of such prior knowledge on the variables are often available.
-The information can, for example, come in the form of (a) results on the same molecular features obtained in a previous study (e.g.
-$p$-values), (b) information from a publicly available database that summarizes the prior information on the molecular features involved (e.g., the Cancer Gene Census \cite[]{futreal_census_2004}), and (c) omics annotation (e.g.
-the location of a gene on the chromosome).
-Although this information can rarely be directly included in the statistical analysis, it may still be useful and informative for the study at hand.
+	At first sight, such an \textit{a priori} grouping of the features suggests 
+	the group lasso
+	\cite[]{meier_group_2008, yuan_model_2006} or one of its extensions, because
+	the group lasso penalises and selects features at the group level. 
+	There are two limitations to this approach: the group lasso (a) selects groups
+	as a whole instead of single 
+	features and (b) does not penalise adaptively: all groups are penalised 
+	equally. Extensions such as the sparse group lasso 
+	\cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
 
-	An example of such a study with extra information available is investigated in Section \ref{sec:app}.
-INTRODUCE DATA.
-This problem is not straightforward to solve with the standard methods available in literature.
-At first sight, such an \textit{a priori} grouping of the features suggests the group lasso \cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which penalises and selects features at the group level.
-There are two limitations to the group lasso: (a) it selects groups as a whole instead of single features and (b) it is not adaptive: the amount of penalization a group receives depends on a `global' penalty parameter common to all groups.
-Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
+	A natural way to deal with (b) is through differential penalization: that is,
+	each group of variables receives its own penalty parameter. An apparent issue
+	with this differential penalization is the estimation of the penalty 
+	parameters. Naive estimation may be done by cross-validation. However, 
+	cross-validation requires re-estimation of the model over a grid of penalty
+	parameters. The size of this grid increases exponentially with the number of
+	penalty parameters. Consequently, it quickly becomes computationally 
+	infeasible. We therefore propose an efficient alternative: empirical Bayes
+	estimation of the penalty parameters, which corresponds to hyperparameter
+	estimation in the Bayesian prior framework. Because of the ubiquity of binary
+	outcome data in omics research, we focus on the logistic elastic net.
 
-	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter.
-An apparent issue with this differential penalization is the estimation of the penalty parameters.
-Naive estimation may be done by cross-validation.
-However, cross-validation requires re-estimation of the model over a grid of penalty parameters.
-The size of this grid increases exponentially with the number of penalty parameters.
-Consequently, it quickly becomes computationally infeasible.
-We therefore propose an efficient alternative: empirical Bayes estimation of the penalty parameters, which corresponds to hyperparameter estimation in the Bayesian prior framework.
-Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net.
-Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit.
-Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.
+	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
+	spirit. Nevertheless, one clear advantage of our method over VBVS is the
+	possibility of adaptive inclusion of external information on the features to
+	aid in prediction and feature selection. Bayesian versions of support vector 
+	machines have been used in classification problems as well 
+	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
+	adaptive inclusion of external information on the features.
 
-	In line with the above, our proposed method is (a) subjective: the few remaining tuning parameters may be cross-validated, (b) adaptive: prior information is automatically and subjectively weighed with respect to its informativeness, (c) fast compared to full Bayesian analysis or cross-validation, and (d) requires minimal input by the user: only the data and grouping of the features is required.
+	In line with the above, our proposed method is (a) subjective: the 
+	hyper-parameters are estimated from the data and not chosen on objective 
+	grounds, (b) adaptive: prior information is automatically and subjectively 
+	weighed with respect to its informativeness, (c) fast compared to full 
+	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
+	user: only the data and grouping of the features is required. We emphasize that 
+	our method is different from recent developments in grouped penalized 
+	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
+	features, the observations are partitioned.
 
-	We introduce the Bayesian logistic (generalised) elastic net model in Section \ref{sec:model}.
-In Section \ref{seq:estimation} we derive a variational approximation to this model and use this novel approximation in the empirical Bayes estimation of multiple, group-specific penalty parameters.
-In Sections \ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a simulation study and a real data example, respectively.
-In Section \ref{sec:discussion} we conclude by both discussing some differences and parallels between the proposed method, termed \texttt{gren}, and other methods in the literature.
+	The rest of the paper is structured as follows. We introduce the Bayesian 
+	logistic generalised elastic net model in Section
+	\ref{sec:model}. In Section \ref{sec:estimation} we derive a variational
+	approximation to this model and use this novel approximation in the empirical
+	Bayes estimation of multiple, group-specific penalty parameters. In Section
+	\ref{sec:simulations} we compare the method in a simulation study. In Sections
+	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
+	two data sets. In Section \ref{sec:discussion} 
+	we conclude by both discussing some differences and
+	parallels between the proposed method, termed \texttt{gren}, and other methods
+	in the literature.
 
 	\section{Model}\label{sec:model}
-	\subsection{Logistic elastic net regression}\label{sec:logitmodel}
-	In logistic regression the outcome variables are assumed to be binary or sums of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l \in \{0,1\}$ for $i=1, \dots, n$).
-The binomial logistic model relates the responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix} x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} \( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is the binomial distribution with number of trials $m$ and probability $\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + \exp(\x\tr_i \bbeta)]$.
-Note that if $m_i=1$ for $i=1, \dots, n$, the model reduces to a binary logistic regression model.
-Throughout the rest of the paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots & \x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.
+	In logistic regression the outcome variables are assumed to be binary or sums
+	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
+	\in \{0,1\}$ for $i=1, \dots, n$). The binomial logistic model relates the
+	responses to the $p$-dimensional covariate vectors $\x_i = \begin{bmatrix}
+	x_{i1} & \cdots & x_{ip}\end{bmatrix} \tr$ through $y_i \sim \mathcal{B} 
+	\( m_i, \expit ( \x_i\tr \bbeta ) \)$, where $\mathcal{B} (m,\upsilon)$ is 
+	the binomial distribution with number of trials $m$ and probability 
+	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
+	\exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model
+	reduces to a binary logistic regression model. Throughout the rest of the 
+	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
+	\x_n \end{bmatrix}\tr$ is standardized such that $\frac{1}{n}\sum_{i=1}^n
+	x_{ij}=0$ and $\frac{1}{n}\sum_{i=1}^n x_{ij}^2=1$ for $j=1, \dots, p$.
 
-	In elastic net regression, the penalised likelihood is maximised to yield parameter estimates:
-	$$
-	\hat{\bbeta} := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \norm{\bbeta}_1 - \frac{\lambda_2}{2} \norm{\bbeta}_2^2,
-	$$
-	with $\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ and $\y = \begin{bmatrix} y_1 & \cdots & y_n \end{bmatrix} \tr$.
-The likelihood term is thus complemented with a penalty term consisting of the $L_1$ and $L_2$-norm of the parameters, multiplied by respective penalty parameters $\lambda_1$ and $\lambda_2$.
-We have scaled the penalty parameters by $\frac{1}{2}$, to make them correspond with the Bayesian version in the following.
-The maximiser of the penalised likelihood corresponds to the posterior mode of a Bayesian elastic model with prior \cite[]{li_bayesian_2010,zou_regularization_2005}:
-	\begin{equation}\label{eq:elasticnetprior}
-	\beta_j \sim g(\lambda_1,\lambda_2) \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta_j| + \lambda_2 \beta_j^2)\].
-	\end{equation}
-	Here, $g(\lambda_1,\lambda_2)$ denotes a normalizing constant that is dependent on $\lambda_1$ and $\lambda_2$ (given in Section 2 of the Supplementary Material (SM), along with further details on the elastic net prior).
-The elastic net combines $L_1$ and $L_2$-norm penalisation such that the model parameters are shrunken towards zero.
-The $L_1$-norm may set some of the estimates exactly to zero, thus automatically selecting features.
-The $L_2$-norm ensures that correlated features behave similarly, as one would generally require.
-We extend the elastic net model to the generalised elastic net to allow for different penalty parameters per group of features.
+	Assume we have a partitioning of the features in $G$ groups, such that each
+	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
+	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
+	group-specific weight. In a generalised elastic net regression, the penalised 
+	likelihood is maximised to yield parameter estimates:
+	\begin{align}
+	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
+	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
+	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
+	  \label{eq:weightedlik}\\
+	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
+	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
+	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
+	  \nonumber
+	\end{align}
+	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
+	observed data $\y = \begin{bmatrix} 
+	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
+	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
+	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
+	group-specific weights may be interpreted as penalty multipliers. Note that
+	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.
 
-	\subsection{The generalised elastic net}\label{sec:genen}
-	Assume we have a partitioning of the features in $G$ groups, such that each feature belongs to one group.
-Let $\G(g)$ be the feature index set of group $g$ for $g=1, \dots, G$.
-To apply differential penalisation we include a group-specific weight $w_g \in \mathbb{R}_{>0}$.
-The model parameters are now estimated by:
-	\begin{align}
-	\hat{\bbeta} & := \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) - \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 \label{eq:weightedlik}\\
-	& = \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, \nonumber
-	\end{align}
-	where we wrote $w^2_g = \lambda'_g$ to emphasise that these group-specific weights may be interpreted as penalty multipliers.
-Throughout the following we assume that the geometric mean of the multipliers, weighted by their respective group sizes, is one, such that the average shrinkage of the model parameters is determined by the `global' $\lambda_1$ and $\lambda_2$.
-That is, we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G (\lambda'_g)^{|\G(g)|}=1$.
-The multiplier appears in square root form in the $L_1$-norm term to ensure that penalisation on the parameter level is the same for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in (\ref{eq:weightedlik}).
+  Throughout the following we assume that the geometric mean of the multipliers, 
+	weighted by their 
+	respective group sizes, is one, such that the average shrinkage of the model
+	parameters is determined by $\lambda_1$ and $\lambda_2$. That is,
+	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
+	(\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the
+	$L_1$-norm term to ensure that penalisation on the parameter level is the same
+	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
+	(\ref{eq:weightedlik}). The elastic net combines $L_1$ and $L_2$-norm 
+	penalisation such that the model parameters are shrunken towards zero. 
+	The $L_1$-norm may set some of the estimates exactly to zero, thus 
+	automatically selecting features. The $L_2$-norm ensures that correlated 
+	features behave similarly, as one would generally require.
 
-	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as a computationally more convenient scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$.
-Using this result, we write the generalised elastic net model in its Bayesian form as:
+	The maximiser of the penalised likelihood in the elastic net, corresponds 
+	to the posterior mode of a 
+	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
+	\cite{li_bayesian_2010} show that the elastic net prior 
+	(see Supplementary Material (SM) Section \ref{sm-sec:enetprior} for 
+	details on this prior) may be written as
+	a computationally more convenient scale mixture of normals, with mixing 
+	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
+	\tr$. Using this result, we write the generalised elastic net model in its 
+	Bayesian form as:
 	\begin{subequations}\label{eq:enmodel2}
 		\begin{align}
-		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i \bbeta ) \right),  \\
-		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
-		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right).
+		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
+		\bbeta ) \right),  \\
+		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
+		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
+		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
+		}{\lambda_1^2}, \left(1,\infty \right) \right).
 		\end{align}
 	\end{subequations}
-	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$.
-In this Bayesian formulation the penalty parameters in $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.
+	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
+	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
+	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
+	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
+	play the role of the hyperparameters in a Bayesian hierarchical model.
 
-	\section{Estimation}\label{seq:estimation}
-	\subsection{Empirical Bayes}\label{seq:empiricalbayes}
-	If the penalty parameters are known, estimation of the frequentist elastic net model parameters or finding the posterior of the generalized elastic net model is feasible with small adjustments of the available algorithms \cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers \cite[]{li_bayesian_2010}.
-Determining these penalty parameters, however, is not straightforward.
+	\section{Estimation}\label{sec:estimation}
+	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
+	If the penalty parameters are known, estimation of the frequentist elastic net
+	model parameters or finding the posterior of the generalized elastic net model
+	is feasible with small adjustments of the available algorithms 
+	\cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers
+	\cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is 
+	not straightforward.
 
-	In the frequentist elastic net without group-wise penalisation, two main strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by cross-validation over a two-dimensional grid of values or (ii) re-parametrise the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 \lambda_2 + \lambda_1}$ and $\lambda = 2 \lambda_2 + \lambda_1$, fix the proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global penalty parameter $\lambda$.
-Strategy (i) is advised by \cite{waldron_optimized_2011}, while (ii) is proposed by \cite{friedman_regularization_2010}.
-In the generalised elastic net setting, strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, respectively.
-$K$-fold cross validation over $D$ values then results in $K \cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate.
-Typically $K$ is set to 5, 10, or to the number of samples $n$, while $D$ is in the order of $100$, so that even for small $G$, the number of models to estimate is prohibitively large.
+	In the frequentist elastic net without group-wise penalisation, two main 
+	strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by 
+	cross-validation over a two-dimensional grid of values or (ii) re-parametrise 
+	the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 
+	\lambda_2 + \lambda_1}$ and $\lambda = 2 \lambda_2 + \lambda_1$, fix the 
+	proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global 
+	penalty parameter $\lambda$. Strategy (i) is advised by 
+	\cite{waldron_optimized_2011}, while (ii) is proposed by 
+	\cite{friedman_regularization_2010}. In the generalised elastic net setting, 
+	strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, 
+	respectively. $K$-fold cross validation over $D$ values then results in $K 
+	\cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate. Typically $K$ is 
+	set to 5, 10, or to the number of samples $n$, while $D$ is in the order of 
+	$100$, so that even for small $G$, the number of models to estimate is 
+	prohibitively large.
 
-	In the Bayesian framework, estimation of penalty parameters may be avoided by the addition of a hyperprior to the model hierarchy.
-The hyperprior takes the uncertainty in the penalty parameters into account by integrating over them.
-This approach introduces two issues.
-Firstly, the choice of hyperprior is not straightforward.
-Many authors suggest a hyperprior from the gamma family of distributions \cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, but the precise parametrisation of this gamma prior is not so obvious.
-A second issue is the loss of correspondence between the  Bayesian and frequentist elastic net.
-If the ultimate goal is feature selection, this correspondence may be exploited through the automatic feature selection property of the frequentist elastic net.
-Endowing the penalty parameters with a hyperprior obstructs their point estimation and, consequently, impedes automatic feature selection.
-Therefore, to circumvent the problem of hyperprior choice and allow for feature selection by the frequentist elastic net, we propose to estimate the penalty parameters by empirical Bayes.
+	In the Bayesian framework, estimation of penalty parameters may be avoided by 
+	the addition of a hyperprior to the model hierarchy. The hyperprior takes the 
+	uncertainty in the penalty parameters into account by integrating over them. 
+	This approach introduces two issues. Firstly, the choice of hyperprior is not 
+	straightforward. Many authors suggest a hyperprior from the gamma family of 
+	distributions 
+	\cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, 
+	but the precise parametrisation of this gamma prior is not so obvious. A 
+	second issue is the loss of correspondence between the  Bayesian and 
+	frequentist elastic net. If the ultimate goal is feature selection, this 
+	correspondence may be exploited through the automatic feature selection 
+	property of the frequentist elastic net. Endowing the penalty parameters with 
+	a hyperprior obstructs their point estimation and, consequently, impedes 
+	automatic feature selection. Therefore, to circumvent the problem of 
+	hyperprior choice and allow for feature selection by the frequentist elastic 
+	net, we propose to estimate the penalty parameters by empirical Bayes.
 
-	Many forms of empirical Bayes exist, the most formal one being maximisation of the marginal likelihood with respect to the hyperparameters.
-The resulting hyperparameter estimates are then plugged into the prior.
-The marginal likelihood is often introduced as a measure of model evidence given the observed data and is computed by integrating the product of likelihood and prior with respect to the model parameters.
-In the case of the elastic net introduced in (\ref{eq:enmodel2}) this results in the following marginal empirical Bayes posterior for $\bbeta$:
+	Many forms of empirical Bayes exist, the most formal one being maximisation of
+	the marginal likelihood with respect to the hyperparameters. The resulting 
+	hyperparameter estimates are then plugged into the prior. The marginal 
+	likelihood is often introduced as a measure of model evidence given the 
+	observed data and is computed by integrating the product of likelihood and 
+	prior with respect to the model parameters. In the case of the elastic net 
+	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
+	empirical Bayes posterior for $\bbeta$:
 	\begin{align}
-	p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = \frac{\int_{\btau} \L (\y ; \bbeta) \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \, d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
-	\hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
-	& = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. \label{eq:mmlestimate}
+	  p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) 
+	  \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = 
+	  \frac{\int_{\btau} \L (\y ; \bbeta) 
+	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
+	  d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
+	  \hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
+	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, 
+	  \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
+	  & = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta)
+	  \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. 
+	  \label{eq:mmlestimate}
 	\end{align}
-	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the elastic net.
-In the omics setting the integrals are also high-dimensional, in which case numerical and Monte Carlo approximation methods become tedious and computationally expensive.
-Moreover, Laplace approximation is known to suffer from low accuracy in many high-dimensional settings \cite[]{shun_laplace_1995}.
-In \cite{casella_empirical_2001} an EM algorithm is described that estimates the hyperparameters.
-This EM algorithm iteratively maximises the expected joint log likelihood, such that the sequence:
+	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
+	elastic net. In the omics setting the integrals are also high-dimensional, in 
+	which case numerical and Monte Carlo approximation methods become tedious and 
+	computationally expensive. Moreover, Laplace approximation is known to suffer 
+	from low accuracy in many high-dimensional settings 
+	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
+	is described that estimates the hyperparameters. This EM algorithm iteratively
+	maximises the expected joint log likelihood, such that the sequence:
 	\begin{equation}\label{eq:ebEM}
-	\blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
+	  \blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
+	  \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
 	\end{equation}
-	converges to a local maximum of the marginal likelihood.
-The difficulty herein is in the calculation of the expected joint log likelihood.
-\cite{casella_empirical_2001} suggests to approximate the expectation by its Monte Carlo expectation.
-Although elegant and simple, this method requires a converged MCMC sample from the posterior for every iteration: a computationally intensive procedure.
-\cite{roy_selection_2017} introduce generalised importance sampling for the Bayesian elastic net, such that we just need a limited, pre-specified number of MCMC chains.
-However, this still requires several converged chains, so is not feasible in many high-dimensional omics settings.
-We propose to tackle this problem by approximating the expectation in (\ref{eq:ebEM}) using variational Bayes.
+	converges to a local maximum of the marginal likelihood. The difficulty herein
+	is in the calculation of the expected joint log likelihood. 
+	\cite{casella_empirical_2001} suggests to approximate the expectation by its 
+	Monte Carlo expectation. Although elegant and simple, this method requires a 
+	converged MCMC sample from the posterior for every iteration: a 
+	computationally intensive procedure. \cite{roy_selection_2017} introduce
+	generalised importance sampling for the Bayesian elastic net, such that we
+	just need a limited, pre-specified number of MCMC chains. However, this still
+	requires several converged chains, so is not feasible in many high-dimensional
+	omics settings. We propose to tackle this problem by approximating the
+	expectation in (\ref{eq:ebEM}) using variational Bayes.
 
@@
 	Variational Bayes is a widely used method to approximate Bayesian posteriors. 
-It has successfully been applied in a wide range of applications, including genetic association studies \cite[]{carbonetto_scalable_2012} and gene network reconstruction \cite[]{leday_gene_2017}.
-In variational Bayes, the posterior is approximated by a tractable form and estimated by optimizing a lower bound on the marginal likelihood of this model (see Section 3 of the SM for the lower bound of the proposed model).
-For an extensive introduction and concise review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, respectively.
+	It has successfully been applied in a wide range of applications, including 
+	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
+	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
+	is approximated by a tractable form and estimated by optimizing a lower bound 
+	on the marginal likelihood of this model (see Section 
+	\ref{sm-sec:objectivefunction} of the SM for the 
+	lower bound of the proposed model). For an extensive introduction and concise 
+	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
+	respectively.
 
-	To simplify the computations of our variational approximation, we follow \cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for $i=1, \dots, n$.
-Conditional on $\bbeta$, the $\omega_i$ are independent of the $y_i$ and P\'{o}lya-Gamma distributed (see Section 4 of the SM for more details).
-We augment Model (\ref{eq:enmodel2}) with:
+	To simplify the computations of our variational approximation, we follow
+	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
+	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
+	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
+	\ref{sm-sec:polyagamma} of the SM for more 
+	details). We augment Model (\ref{eq:enmodel2}) with:
 	\begin{equation}\label{eq:logmodel}
@@
 	\end{equation}
-	Our variational Bayes approximation to the posterior distribution of (\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of parameters. We choose the blocks such that:
+	Our variational Bayes approximation to the posterior distribution of 
+	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
+	parameters. We choose the blocks such that:
 	\begin{equation}\label{eq:varbayesapproximation}
-	p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
+	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
+	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
 	\end{equation}
-	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, calculus of variations gives the optimal distributions $q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} [\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the Kullback-Leibler divergence of the posterior to the approximate distribution \cite[]{neal_view_1998}.
-The approximation in (\ref{eq:varbayesapproximation}) renders both the posterior parameter calculations and the expected joint log likelihood as introduced in (\ref{eq:ebEM}) tractable.
+	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
+	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
+	calculus of variations gives the optimal distributions 
+	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
+	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
+	Kullback-Leibler divergence of the posterior to the approximate distribution 
+	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
+	renders both the posterior parameter calculations and the expected joint log 
+	likelihood as introduced in (\ref{eq:ebEM}) tractable.
 
-	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal distributions in our variational Bayes implementation for the model parameters as:
+	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
+	distributions in our variational Bayes implementation for the model parameters
+	as:
 	\begin{equation}\label{eq:varbayesdistr}
-	q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)\text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
+	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
+	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
+	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
+	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
 	\end{equation}
-	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian distribution (See SM Section 5 for the derivations).
-The so-called variational parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we update them by:
+	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
+	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
+	for the derivations). The so-called variational
+	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
+	update them by:
 	\begin{subequations}\label{eq:VBupdateequations}
 		\begin{align}
-		\bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
+		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
+		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
+		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
+		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
+		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
 		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
-		c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
-		\chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
+		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
+		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
+		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
+		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
 		\end{align}
 	\end{subequations}
-	until convergence.
-Here, $\bLambda'$ is a diagonal matrix with entries $\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & \cdots & m_n \end{bmatrix} \tr$.
-Furthermore, $\mathbf{A}_{jj}$ and $\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th element of a column vector, respectively.
-Naive calculation of the variational parameters is computationally expensive.
-In Section 6 of the SM we show that informed calculation of the parameters results in a significant reduction of computational complexity.
+	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
+	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
+	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
+	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
+	element of a column vector, respectively. Naive calculation of the variational
+	parameters is computationally expensive. In Section 
+	\ref{sm-sec:computationalcomplexity} of the SM we show that 
+	informed calculation of the parameters results in a significant reduction of 
+	computational complexity.
 
-	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
-	Variational Bayes was shown to underestimate the posterior variance of the parameters, both numerically and theoretically, in several settings \cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,wang_inadequacy_2005}.
-This coincides with our experience that the global penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated.
-To prevent overestimation we use the parametrisation of the elastic net in \cite{friedman_regularization_2010} as introduced in Section \ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by cross-validation of the regular elastic net model, such that the overall penalisation is determined by cross-validation of only $\lambda$.
-By combining cross-validation of the global penalty parameter $\lambda$ with empirical Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more robust to underestimation of the variational posterior variances.
-The remaining issue is the choice of $\alpha$.
-\cite{hastie_glmnet_2016} recommend to either fix $\alpha$ \textit{a priori}, or compare the results for several choices of $\alpha$.
-We recommend the latter.
-Otherwise we note that, in our experience, the choice $\alpha=0.5$ often gives good results.
+	\subsection{Empirical-variational Bayes}\label{sec:empvarbayes}
+	Variational Bayes was shown to underestimate the posterior variance of the 
+	parameters, both numerically and theoretically, in several settings 
+	\cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,
+	wang_inadequacy_2005}. This coincides with our experience that the global 
+	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To 
+	prevent overestimation we use the parametrisation of the elastic net in 
+	\cite{friedman_regularization_2010} as introduced in Section 
+	\ref{sec:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
+	cross-validation of the regular elastic net model, such that the overall 
+	penalisation is determined by cross-validation of only $\lambda$. By combining
+	cross-validation of the global penalty parameter $\lambda$ with empirical 
+	Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} 
+	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more 
+	robust to underestimation of the variational posterior variances. The 
+	remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend
+	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
+	choices of $\alpha$. We recommend the latter.
 
-	For estimating the penalty multipliers, the intractable posterior expectation in (\ref{eq:ebEM}) is approximated by the variational posterior:
-	$$
-	\E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} \] = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
-	$$
-	where $C$ is constant in $\blambda'$ (see SM Section 7 for the full derivation).
-The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j \in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$.
-An estimate of the new penalty multipliers is now given by:
+	To estimate the penalty multipliers, the intractable posterior expectation
+	in (\ref{eq:ebEM}) is approximated using the variational posterior:
+	\begin{equation}\label{eq:approximatedloglikelihood}
+	  \E_{Q} \[ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | 
+	  \blambda'^{(k)} \]
+	  = \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) 
+	  \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g + C,
+	\end{equation}
+	where $C$ is constant in $\blambda'$ (see SM Section 
+	\ref{sm-sec:expectedloglikelihood} for the full 
+	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
+	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
+	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
+	of the new penalty multipliers is now given by:
 	\begin{subequations}\label{eq:mmlupdateequation}
 		\begin{align}
-		\blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
+		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
+		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
 		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
@@
 	\end{subequations}
-	Although the solution to (\ref{eq:mmlupdateequation}) is not available in closed form, this convex problem is easily solved by a numerical optimisation routine.
-The full procedure is summarized in Section 8 of the SM.
+	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
+	closed form, this convex problem is easily solved by a numerical optimisation 
+	routine. The full procedure is summarized in Section 
+	\ref{sm-sec:fullestimationprocedure} of the SM.
 
 	\subsection{Feature selection}
-	Feature selection is often desirable in high-dimensional prediction problems and omics data problems are no exception.
-A selection of biomarkers may lead to a large cost reduction by supporting targeted assays.
-Bayesian feature selection is often done by inspection of posterior credible intervals.
-Similarly as in frequentist hypothesis testing, we may select a feature if zero is not contained in the credible interval.
-However, the Bayesian lasso's credible intervals (a special case of the elastic net) are known to suffer from low frequentist coverage in sparse settings \cite[]{castillo_bayesian_2015}.
-We often assume sparsity of the features in omics research, so selection by credible intervals is generally inconsistent.
+	Feature selection is often desirable in high-dimensional prediction problems 
+	and omics data problems are no exception. For example, biomarker selection 
+	may lead 
+	to a large reduction in costs by supporting targeted assays. Bayesian feature
+	selection is often done by inspection of posterior credible intervals. 
+	Similarly as in frequentist hypothesis testing, we may select a feature if 
+	zero is not in the credible interval. However, the Bayesian lasso's 
+	(a special case of the elastic net) credible intervals  are known to suffer 
+	from low frequentist coverage in sparse settings 
+	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
+	omics research, so selection by credible intervals is generally inconsistent. 
 	We therefore propose to select features in the frequentist paradigm.
 
-	As shortly touched upon in Section \ref{seq:empiricalbayes}, frequentist feature selection is trivial after estimation of the penalty multipliers.
-We simply plug the estimated penalty parameters into some frequentist elastic net algorithm that allows for differential penalization.
-In our own package \texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} \cite[]{friedman_regularization_2010}.
-In the frequentist elastic net, feature selection is then done automatically.
-Furthermore, to select a specific number of features, we simply adjust the global $\lambda$ until we select the desired number of features.
+	As shortly touched upon in Section \ref{sec:empiricalbayes}, frequentist 
+	feature selection is trivial after estimation of the penalty multipliers. We 
+	simply plug the estimated penalty parameters into some frequentist elastic net
+	algorithm that allows for differential penalization. In our own package 
+	\texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} 
+	\cite[]{friedman_regularization_2010}. In the frequentist elastic net, feature
+	selection is then done automatically. Furthermore, to select a specific number
+	of features, we simply adjust the global $\lambda$ until we select the desired
+	number of features.
 
 	\section{Simulations}\label{sec:simulations}
-	\subsection{Setup}\label{sec:setup}
-	To assess variable selection and predictive performance we conduct a simulation study in which we compare \texttt{gren} to the regular elastic net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, group mimimax concave penalty (\texttt{grMCP}) \cite[]{breheny_penalized_2009}, and the group exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}.
-\texttt{GRridge} is similar to \texttt{gren} in the sense that it estimates group-specific penalty multipliers.
-The two main differences with \texttt{gren} are (i) the absence of an $L_1$-norm penalty and (ii) the estimation procedure.
-The other methods are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} and not adaptive on the group level.
-In contrast to the original group lasso, all of them are able to select single features, instead of just on the group level.
+	To assess classification and variable selection performance we conducted a 
+	simulation study in which we compared \texttt{gren} to the regular elastic 
+	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
+	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
+	composite mimimax concave penalty (\texttt{cMCP}) 
+	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
+	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
+	\texttt{gren} in the sense that it estimates group-specific penalty 
+	multipliers. The two main differences with \texttt{gren} are (i) the absence 
+	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
+	are extensions of the group lasso and not adaptive on the group level. 
+	However, in contrast to the original group lasso,
+	they select single features, instead of just whole groups.
 
-	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, we fix the proportion of $L_1$-norm penalty to three different values: $\alpha \in \{0.05, 0.5, 0.95 \}$.
-The first setting closely resembles the ridge setting, where $\alpha=0$, while the third setting is similar to the lasso with $\alpha=1$.
-For the methods that do automatic variable selection (all methods except the regular ridge and \texttt{GRridge}), we estimate the models for a range of model sizes.
-
-	Predictive performance is measured by area under the receiver operator curve (AUC) and Brier skill score (BSS).
-AUC gives the area under the curve of sensitivity versus $1-$ specificity and equals the probability that the classifier ranks a randomly chosen case higher than a randomly chosen non-case.
-The BSS is a normalised version of the Brier score: $\text{BSS} = 1 - \frac{\sum_{i=1}^n (y_i - \hat{y}_i)^2}{\sum_{i=1}^n (y_i - \bar{y})^2}$,
-	where $\hat{y}_i=\expit(\x_i \tr \hat{\bbeta})$ is the prediction for observation $i$ and $\bar{y}=n^{-1} \sum_{i=1}^n y_i$.
-The BSS measures the fraction of Brier score improvement compared to a non-informative model.
-Quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960} and parameter estimation by mean squared error (MSE).
-Both are presented in SM Section 10.
+	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{cMCP}, 
+	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
+	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
+	such that the first setting closely resembles the ridge setting, 
+	where $\alpha=0$, while the third setting is similar to the lasso with 
+	$\alpha=1$. In \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}, $\alpha$
+	determines the ratio of group-level penalization ($\alpha=0$) to 
+	feature level ($\alpha=1$) penalization. For the methods that do 
+	automatic feature selection, we estimate models for a range of model sizes:
+	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.
 
@@
 	\begin{enumerate}
-		\item differential signal between the groups and uniformly distributed model parameters,
+		\item differential signal between the groups and uniformly distributed model
+		parameters,
 		\item a large number of small groups of features,
-		\item no differential signal between the groups, but strong correlations within groups of features, 
-		\item differential signal between the groups and heavy-tailed distributed model parameters, and
+		\item no differential signal between the groups, but strong correlations 
+		within groups of features,
+		\item differential signal between the groups and heavy-tailed distributed 
+		model parameters, and
 		\item a very sparse setting with no signal in some of the groups.
 	\end{enumerate}
-	Full descriptions of the scenarios are given in Section 10 of the SM.
-In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform.
-In the second scenario, we expect \texttt{gren} and \textt{GRridge} to overfit due to the large number of hyperparameters to estimate.
-The rest of the methods are conjectured to perform better, due to a lack of overfitting.
-In scenario three, we expect all methods to perform badly due to the high correlation between the features.
-In scenario three
+	Full descriptions of the scenarios are given in Section 
+	\ref{sm-sec:simulations} of 
+	the SM. In the first scenario, the distribution of model parameters is far
+	from the assumed ridge/elastic net/lasso distributions, so we expect
+	all methods to underperform. In the second scenario, which is similar to
+	the simulations presented in the original group lasso paper 
+	\cite[]{yuan_model_2006}, we expect \texttt{gren}
+	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
+	estimate. The group lasso extensions are expected to perform better: they 
+	estimate just one penalty parameter, so are less sensitive to overfitting.
+	In scenario three, we expect all methods to underperform
+	due to the high correlations between the features. The fourth 
+	scenario is expected to yield a better \texttt{gren} performance, since
+	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
+	detect differential signal between
+	the groups. Scenario five is expected to yield a bad performance for all 
+	methods, because of the sparse truth and lack of differential 
+	signal.
 
-	In all scenarios we created a training set of $n=100$ samples to estimate the models and a test set of $n_{\text{test}}=1000$ samples on which we compute the performance measures.
-The features $\x_i$, $i=1, \dots, n$, are sampled from a $p$-dimensional Gaussian distribution centred at zero, where we set $p=1000$.
-We let the covariance matrix of the features vary over the scenarios.
-Likewise for the model parameters $\beta_j$, $j=1, \dots, p$.
-We simulate the outcome data $y_i$, $i=1, \dots, n$, from a binary logistic model: $y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$.
-To mitigate the influence of random variation in the simulations, we repeated every scenario 100 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.
+	In all scenarios we created a training set of $n=100$ samples to estimate the 
+	models and a test set of $n_{\text{test}}=1000$ samples to compute 
+	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
+	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
+	$p=900$ in scenario 2 and $p=1000$ in the other scenarios. We varied the 
+	covariance matrix of the features and simulation of 
+	the model parameters $\beta_j$, $j=1, \dots, p$ between the
+	scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
+	parametrisations). We 
+	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
+	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
+	of random variation, we repeated each scenario 100 times 
+	and report predictive performance measures 
+	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
+	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
+	Additional results and performance measures are presented in Section 
+	\ref{sm-sec:simulations} of the SM.
 
-	 The results of this simulation are presented in Figures \ref{fig:sim1_performance1_set1}-\ref{fig:sim1_performance1_set3}.
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set1.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set1}
-	\end{figure}
+	- description of results
 
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set2.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set2}
-	\end{figure}
+	\section{Application to microRNAs in colorectal cancer}
+	\label{sec:colorectal}
+  We investigated the performance of \texttt{gren} on data from a 
+  deep sequencing study in microRNAs 
+  \cite[]{neerincx_combination_2018}. The aim of the study was to predict 
+  treatment response in 88 treated colorectal cancer patients, coded as either
+  non-progressive/remission (70 patients) or progressive 
+  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
+  remained. In addition to the microRNAs, four unpenalized clinical
+  covariates were included in the analysis: prior use of adjuvant therapy 
+  (binary), the type of systemic treatment regimen (ternary), age, and primary 
+  tumor differentiation (binary).
 
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set3.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 1.}
-		\label{fig:sim1_performance1_set3}
-	\end{figure}
+  In a preliminary experiment on different subjects, the microRNA expression 
+  levels of metastatic colorectal tumour tissue were compared to normal 
+  non-colorectal tissue and primary colorectal tumour tissue was compared to
+  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 127
+  microRNAs that were differentially expressed in both comparisons with
+  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 
+  microRNAs that were not differentially expressed. We expect that incorporation
+  of this partitioning enhances therapy response 
+  classification, because tumor-specific microRNAs are likely to be more 
+  relevant than non-specific microRNAs.
 
-	The results of this simulation 1 are presented in Figures \ref{fig:sim4_performance1_set1}-\ref{fig:sim4_performance1_set3}.
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set1.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set1}
-	\end{figure}
+  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
+  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
+  three methods, we only present \texttt{cMCP} here; the results for 
+  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
+  [! section ref here]. For the methods with a
+  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
+  performance and refer
+  the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
+  different $\alpha$'s.
 
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set2.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set2}
-	\end{figure}
+  To estimate performance, we split the data into 61 training and 27 test 
+  instances, stratified by treatment response. We estimated the models on the 
+  training data and calculated AUC on the test data. We present the AUC for a
+  range of model sizes, together with the estimated 
+  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
+  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
+  performance measures are presented in Section \ref{sm-sec:colorectal}
+  of the SM. In addition, we investigated the sensitivity of
+  the multiplier estimation on this dataset in SM Section 
+  \ref{sm-sec:colorectal}. We experienced numerical issues with 
+  \texttt{GRridge} when cross-validating 
+  the global $\lambda$ on the training data. We therefore estimated this
+  global parameter on the full data for \texttt{GRridge}.
 
-	\begin{figure}[!h]
-		\centering
-		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set3.png}
-		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 4.}
-		\label{fig:sim4_performance1_set3}
-	\end{figure}
+  The penalty multipliers estimated by 
+  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
+  expressed group receives the lowest penalty, followed by the medium expressed
+  group, while the non-expressed group receives the strongest penalty 
+  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
+  \texttt{gren} seems to outperform the other methods for a range of model 
+  sizes in terms of AUC 
+  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
+  This is especially pronounced for the larger models. The performance of
+  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
+  all group lasso extensions that we investigated.
+  (SM Section \ref{sm-sec:colorectal}). Overall, it seems as though the 
+  inclusion of the extra information on the 
+  features benefited predictive performance: both \texttt{GRridge} and 
+  \texttt{gren} outperform their respective non-group-regularized versions, 
+  albeit it only slightly for \texttt{GRridge}.
 
-	\subsection{Results}\label{sec:results}
+  \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
+  The aim of our second example study was to predict lymph node metastasis in
+  oral cancer patients using sequenced RNAs from TCGA 
+  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
+  3096 transformed and normalised TCGA RNASeqv2 profiles of head-and-neck 
+  squamous cell carcinomas on 133 patients with HPV-negative tumours in the oral
+  cavity. Of these patients, 76 suffered from lymph node metastasis, while 57 
+  did not. For a thorough introduction of these data, see 
+  \cite{te_beest_improved_2017}.
 
-	\section{Application}\label{sec:app}
+  We considered two sources of external feature information: (a) the 
+  cis-correlations between the RNASeqv2 data and TCGA DNA copy numbers on 
+  the same patients, quantified by Kendall's $\tau$ and binned into five groups.
+  In addition, we used (b) $p$-values from a microarray data set
+  described in \cite{mes_prognostic_2017}, again binned into five groups. 
+  We expected features 
+  with a large positive Kendall's $\tau$ to be more important in 
+  metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
+  features with low external $p$-values to be more important.
+
+  We compared \texttt{gren} to the same methods as in Section 
+  \ref{sec:colorectal}. However, since the feature information
+  consists of two partitions with overlapping groups, we used an extension of
+  \texttt{cMCP} that allows for
+  overlapping groups \cite[]{zeng_overlapping_2016}.
+  We estimated AUC for the methods on an independent validation set of 97 
+  patients \cite[]{mes_prognostic_2017}. The features 
+  in this validation set were
+  microarray expression levels, normalised to account
+  for a difference in scale compared to the RNAseq training data. We present the
+  results on this validation set, together with the estimated penalty 
+  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
+  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
+  comparisons and performance measures are presented in the SM Section 
+  \ref{sm-sec:lymphnode}.
+
+  The estimated penalty multipliers follow the expected pattern 
+  (Figure \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}a): 
+  larger cis-correlations receive smaller penalties. 
+  Likewise, smaller external $p$-values are penalized less. This pattern is
+  slightly less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the
+  other methods. The predictive performance of
+  \texttt{gren} is slighty better than the other methods for a range of model
+  sizes. In this example, ridge and \texttt{GRridge} perform almost 
+  identical, while \texttt{gren} outperforms the regular elastic net.
 
 	\section{Discussion}\label{sec:discussion}
-	In a taxonomy of Bayesian methods, the proposed method may be considered a local shrinkage model, as opposed to the global-local shrinkage priors that \cite{polson_local_2012,bernardo_shrink_2011} discuss.
-They characterise certain desirable properties of these global-local shrinkage priors in high dimensions, which, for example, the horseshoe possesses \cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}.
-In our case, global shrinkage would imply adding another hyperprior for the global $\lambda_1$ and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters.
-We argue however, that if the groups are informative, the empirical Bayes estimation of the (semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than full Bayes shrinkage of the global penalty parameters, because the latter does not use any known structure to model the variability in the hyperparameters.
-Nonetheless, an interesting direction of future research is the extension of the group-regularized elastic net to a group-regularized horseshoe model, since the horseshoe has been shown to handle sparsity well and render better coverage of credibility intervals than lasso-type priors \cite[]{van_der_pas_horseshoe_2014}.
+	In a taxonomy of Bayesian methods, \texttt{gren} may be considered a 
+	local shrinkage model, as opposed to the global-local shrinkage priors that 
+	\cite{polson_local_2012,bernardo_shrink_2011} discuss. They characterise 
+	certain desirable properties of these global-local shrinkage priors in high 
+	dimensions, which, for example, the horseshoe possesses 
+	\cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}. In our case, global 
+	shrinkage would imply adding another hyperprior for the global $\lambda_1$ 
+	and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters. We argue however,
+	that if the groups are informative, the empirical Bayes estimation of the 
+	(semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than 
+	full Bayes shrinkage of the global penalty parameters, because the latter does
+	not use any known structure to model the variability in the hyperparameters. 
+	Nonetheless, an interesting direction of future research is the extension of 
+	the group-regularized elastic net to a group-regularized horseshoe model, 
+	since the horseshoe has been shown to handle sparsity well and render better
+	coverage of credibility intervals than lasso-type priors 
+	\cite[]{van_der_pas_horseshoe_2014}.
 
-	Although our method can be considered weakly adaptive, it is different from the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net \cite[]{zou_adaptive_2009} in the sense that it adapts to external information rather than to the data.
-It also differs in the scale of adaptation: in the adaptive lasso and elastic net the adaptive weights are feature specific, while in our case they are estimated on the group level, rendering the adaptation more robust.
-As can be seen from the simulations and real data example in Sections \ref{sec:simulations} and \ref{sec:app}, adaptation to external data may be beneficial for prediction and feature selection.
-We believe that this is due to the `borrowing of information' effect: estimates that are believed to be similar are shrunken in the same way, yielding overall, better estimates.
+	Although our method is essentially a reweighted elastic net and can be
+	considered weakly adaptive, it is different from 
+	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
+	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
+	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
+	rather than to the primary data. It also differs in the scale of adaptation: 
+	in the adaptive lasso and elastic net the adaptive weights are feature 
+	specific, while in our case they are estimated on the group level, rendering
+	the adaptation more robust. As can be seen from both the simulations in 
+	Section \ref{sec:simulations} and the applications to different outcomes 
+	(prognostic or diagnostic) and marker types 
+	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
+	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
+	\ref{sm-sec:cervical}, adaptation to 
+	external data may be beneficial for prediction and feature selection. 
+	We believe that this is due to the `borrowing of information' effect: 
+	estimates of feature effects that behave similarly are shrunken similarly, 
+	yielding overall, better estimates.
 
-	Another obvious comparison is to the group lasso \cite[]{meier_group_2008,yuan_model_2006}.
-Although it is similar in the sense that it shrinks on the group level, the group lasso is built upon an entirely different philosophy: it sets whole groups of features to zero.
-The intended application of such a group-wise penalty is to small interpretable groups of features, like, for example, dummies of a categorical variable.
-Another difference between \texttt{gren} and the group lasso is the number of penalty parameters.
-\texttt{gren} estimates one parameter per group, while the group lasso estimates one overall penalty parameter; it is thereby less flexible in differential shrinkage of the parameters.
-In addition, in the simulations and data applications of \cite{van_de_wiel_better_2016}, group lasso prediction performed inferior to \texttt{GRridge} prediction.
+	As touched upon in Section \ref{sec:introduction}, an obvious 
+	comparison is to the group lasso \cite[]{meier_group_2008,
+	yuan_model_2006} and its extensions. Although the group lasso is similar in 
+	the sense that it shrinks on the 
+	group level, it is built upon an entirely different philosophy: 
+	its intended application is to small interpretable groups of features, like, 
+	for example, dummies of a categorical variable. Another difference between 
+	\texttt{gren} and the group lasso is the number of penalty parameters. 
+	\texttt{gren} estimates one parameter per group, while the group lasso
+	estimates one overall penalty parameter; it is thereby less flexible in 
+	differential shrinkage of the parameters. In addition, we show in our 
+	simulations and data applications that \texttt{gren} is competitive and often
+	superior to (extensions of) the group lasso.
 
-	A possible weak point of the proposed method is the double EM loop.
-The double loop increases the chance of ending up in a local optimum.
-In the application discussed above we investigated the occurrence of multiple optima, but never encountered them.
-This does not guarantee that local optima do not occur, but it provides some evidence that local optima are not ubiquitous.
-Local optima can often be avoided by an informed choice of starting values.
-In our experience, reasonable starting values are obtained by running a group-regularized ridge regression and using the estimated penalty multipliers as starting values for \texttt{gren}.
-In addition, we note that even a local optimum might give `good enough' predictions in many practical settings.
+	A common criticism of the lasso (and elastic net) is its instability of 
+	feature 
+	selection: different data instances lead to different sets of selected
+	features. We investigated the stability of selection on the data from Section
+	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
+	the original data and calculated the sizes of all intersections of selected
+	features. We compared \texttt{gren} to the 
+	regular elastic net and found that the inclusion of extra information
+	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
+	In addition, we investigated 
+	the bias and variance of the penalty multipliers by examination of the
+	distribution of estimated penalty parameters from 100 random partitionings 
+	of the features
+  (see SM Section \ref{sm-sec:colorectal}). We found that the penalty parameters
+  tend to cluster around one, as expected for an unbiased estimation method, and
+  that the variance of the penalty multipliers is less than for the similar
+  \texttt{GRridge} method.
+
+	A possible weak point of the proposed method is the double EM loop. The double
+	loop increases the chance of ending up in a local optimum. In the applications
+	discussed in Sections \ref{sec:colorectal}--
+	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
+	\ref{sm-sec:cervical}, we investigated the occurrence of multiple optima, 
+	but never encountered them. This does not guarantee that local optima do not
+	occur, but 
+	it provides some evidence that local optima are not ubiquitous. In any case,
+	bad local optima can often be avoided by an informed choice of starting 
+	values. In our 
+	experience, reasonable starting values are obtained by running a 
+	group-regularized ridge regression and using the estimated penalty multipliers
+	as starting values for \texttt{gren}. Alternatively, a skeptical user may set 
+	the starting values for all penalty multipliers to one, thereby starting from
+	the regular elastic net and adaptively adjusting the penalty parameters to a
+	group-regularized model. The latter is the default approach in our \texttt{R}
+	package
+
+<<>>=
+bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench1.csv", 
+                     stringsAsFactors=FALSE)
+bench2 <- read.table("results/rnaseq_lymph_node_metastasis_bench1.csv", 
+                     stringsAsFactors=FALSE)
+times1 <- round(bench1$time*10^(-9), 2)
+times2 <- round(bench2$time*10^(-9), 2)
+session <- gsub("_", ".", devtools::session_info()$platform$system)
+@
+  The double EM loop runs the danger of excessive computation time. We have   
+  compared computation times to the
+  other methods introduced in Sections \ref{sec:colorectal}--
+  \ref{sec:lymphnode}. All timings were obtained on a Macbook 
+  Pro 2016 running \Sexpr{session}. 
+  The timings of \texttt{gren} are competitive. On the colorectal cancer example
+  we achieved the following times:
+  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
+  respectively. \texttt{GRridge} is 
+  slightly faster with \Sexpr{times1[4]} seconds, while \texttt{SGL} was slightly
+  slower: \Sexpr{times1[c(5:7)]} seconds, with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
+  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
+  The results for the lymph node cancer example are similar. \texttt{gren} 
+  achieved 
+  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
+  respectively.
+  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
+  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
+  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}. 
+  We note that a considerable speedup may be achieved by iterating between 
+  single EB and single VB updates, instead of running the VB EM 
+  until convergence. This works in practice, because an EM tends to take a few 
+  initial large steps towards the optimum, followed by a series of smaller 
+  steps. We argue that the first few iterations of the VB EM render the 
+  approximate expectation in \ref{eq:approximatedloglikelihood}
+  accurate enough for a 
+  reliable EB step.
 
 	\section*{Software}
-	The method is available as an \texttt{R} package from \url{https://CRAN.R-project.org/package=gren}.
+	A (stable) \texttt{R} package is available from 
+	\url{https://CRAN.R-project.org/package=gren}.
 
 	\section*{Supplementary Material}
-	Supplementary Material is available online from \url{http://pub.math.leidenuniv.nl/~munchmm/}.
+	Supplementary Material is available online from 
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All results may be recreated
+	from \url{https://github.com/magnusmunch/gren}.
 
 	\section*{Acknowledgements}
-	\textit{Conflict of Interest}: None declared.
+	\textit{Conflict of Interest}: None declared. We thank the editors and 
+	referees for their useful suggestions and comments.
 
 	\section*{Funding}
-	This research has received funding from the European Research Council under ERC Grant Agreement 320637.
+	This research has received funding from the European Research Council under 
+	ERC Grant Agreement 320637.
 
-	\bibliographystyle{author_short3.bst}
+	\bibliographystyle{biorefs}
 	\bibliography{refs}
 
+<<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res3_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 3.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res4_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res5_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the lymph node metastasis example.", out.width="100%", fig.asp=1>>=
+@
+
 \end{document}

### 2019-03-22 docs/manuscript.Rnw

@@
 	feature level ($\alpha=1$) penalization. For the methods that do 
-	automatic feature selection, we estimate models for a range of model sizes.
+	automatic feature selection, we estimate models for a range of model sizes:
+	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.
 
@@
 <<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res3_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 3.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res4_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res5_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5.", out.width="100%", fig.asp=1/2>>=
 @

### 2019-03-26 code/figures.R
simulation 2 running

@@ -75,6 +75,10 @@ abline(h=plot.data[[1]][, 3], col=col[1], lty=lty[1])
 abline(h=plot.data[[2]][, 3], col=col[2], lty=lty[1])
 par(opar)
 
+# ---- lines_simulations_res2_auc_briers ----
+library(sp)
+load(file="results/simulations_res2.Rdata")
+
 # ---- lines_simulations_res3_auc_briers ----
 library(sp)
 load(file="results/simulations_res3.Rdata")

### 2019-03-26 docs/supplement.Rnw
simulation 2 running

@@
 <<settings, include=FALSE, echo=FALSE>>=
-opts_knit$set(root.dir="..", base.dir="../figs")
-opts_chunk$set(fig.align='center', fig.path="../figs/", 
+knitr::opts_knit$set(root.dir="..", base.dir="../figs")
+knitr::opts_chunk$set(fig.align='center', fig.path="../figs/", 
                       echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
@@
 <<figures, include=FALSE>>=
-read_chunk("code/figures.R")
+knitr::read_chunk("code/figures.R")
 @
@@
 	\end{align*}
-	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. The
+	where $\mathcal{U} (a, b)$ denotes the uniform distribution on $(a,b)$. We set
+	$b_1 \approx 0.18$ such that the mean expected signal 
+	$\overline{\E(\beta_j^g)}=0.07$. Furthermore, we
+	introduce correlation between the features by setting the covariances
+	$\bSigma_{ij} = 0.5^{|i-j|}$.The
 	model parameters clearly do not follow the relatively heavy-tailed elastic net
 	distribution. Because of this mismatch between the assumed and true 
-	distributions, we expect poor performance by \texttt{gren}. Furthermore, we
-	introduce correlation between the features by having covariance elements
-	$\bSigma_{ij} = 0.5^{|i-j|}$.
+	distributions, we expect poor performance by \texttt{gren}.
 
 	\subsection{Scenario 2}
+	For this scenario we loosely follow the simulation setup in 
+	\cite{meier_group_2008}. We create $G=300$ groups of features and $p=900$
+	features in total, such that we
+	have 3 features per group. There are 50 active groups, with 2 active 
+	parameters with the same value and one inactive parameter per group, such 
+	that in total we have 100 non-zero model parameters. The active model 
+	parameters are fixed and decrease linearly from $30/51$ to $3/510$, such that
+	the mean parameter $\overline{\beta}_j=1/30$. In this example, we included
+	an intercept; $\beta_0=10$.
+
+	The predictors are sampled in such a way that they represent dummy variables
+	for categorical variables. We start with sampling 300 latent variables:
+	$\mathbf{z} \sim \mathcal{N} (\mathbf{0}, \bSigma)$, with 
+	$\bSigma_{ij} = 0.5^{|i-j|}$, $i,j=1, \dots, G$. These latent latent 
+	variables are then transformed to dummies by:
+	$$
+	x_{ij} = 
+	\begin{cases}
+	  1 & \text{if } z_g \in (q_i, q_{i+1}), \\
+	  0 & \text{otherwise},
+	\end{cases}
+	$$
+	where the $q_i$ are the $0$, $1/3$, $2/3$, and $1$th quantiles of the normal
+	distribution.
+
 	The second scenario contains many groups of features and is geared towards the
-	non-adaptive group-regularized methods. These methods are able to incorporate
-	the grouping of the features, but were originally designed for small 
+	group lasso and its extensions. These methods are able to incorporate
+	the grouping of the features, and were originally designed for small 
 	interpretable groups, such as dummies of a categorical variable. \texttt{gren}
@@
 
-	For this scenario we follow the second simulation example in 
-	\cite{meier_group_2008}. They set $n=100$, $G=250$, and $p=1000$, such that we
-	have 4 features per group. The first 10 groups are active and those 40 model
-	parameters are set to 0.2. The predictors are simulated from a $p$-dimensional
-	Gaussian distribution with covariance matrix $\bSigma_{ij} = 0.5^{|i-j|}$.
-
 	\subsection{Scenario 3}
-	A third scenario is modelled by considering a situation in which there is no
-	differential signal between the groups, but the correlation within groups is
-	very high. We conjecture that this leads to overestimation of effect within
-	the groups by \texttt{gren} and consequently, performance will suffer.
-
-	To that end we sample the predictors in $G=5$ separate blocks of 200 
-	variables, corresponding to the partitioning. The variances of the features
-	are set to one, while all covariances are 0.7. The 1000 model parameters are
-	simulated in two steps. First, we draw 1000 parameters from the elastic net
-	distribution, parametrised as in \cite{friedman_regularization_2010}, where
-	$\alpha=0.5$ and $\lambda=100$. The second step consists of setting the 
+	In this scenario we sample the 1000 model parameters in $G=5$ groups of 200.
+	For each group we generate the parameters in two steps. First we draw 200
+	parameters from the elastic net distribution, parametrised as in 
+	\cite{friedman_regularization_2010}, where $\alpha=0.5$ and $\lambda=100$.
+	The second step consists of setting the 
 	(in absolute value) smallest 100 parameters in a group to zero. This results
 	in a total of 500 zero and 500 non-zero parameters, evenly distributed over
-	the five groups.
+	the five groups. We sample the features from a multivariate normal 
+	distribution with all variances 1 and all covariances 0.7.
+
+	This third scenario is modelled by considering a situation in which there is no
+	differential signal between the groups, but the correlation within groups is
+	very high. We conjecture that this leads to unstable estimation of the 
+	penalty multipliers by \texttt{gren} and consequently, performance will suffer.
 
 	\subsection{Scenario 4}
-	This scenario is similar to Scenario 3, except each groups penalty parameter
-	is multiplied by one of : $\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$ and the
-	covariance matrix of the predictors is block diagonal, with block sizes $25
-	\times 25$, off-diagonals set to $\rho=0.7$, and diagonals to $\sigma^2 = 1$.
+	This scenario is similar to Scenario 3, except that we apply differential 
+	shrinkage over the groups with penalty multipliers: 
+	$\lambda' \in \{ 0.14, 0.51, 1.95, 7.39 \}$. 
+	Furthermore, we change the covariance matrix of the predictors to a block
+	diagonal, with block sizes $25
+	\times 25$, such that each group of 200 features consists of 8 blocks of 
+	correlated features. The covariance off-diagonals and diagonal in each block 
+	are set to $\rho=0.7$ and $\sigma^2 = 1$, respectively.
 	We expect \texttt{gren} to pick up this differential signal and outperform the
@@
 	\subsection{Scenario 5}
-	In the last scenario we create a sparse true model that contains some 
-	differential signal. We impl
+	In this last scenario, we create $G=10$ groups of 100 features.
+	Of the 10 groups 8 do not contain any signal. In the two active groups, we set 
+	85 model parameters to zero and 15 to 0.2 and 0.5, respectively, such that the
+	mean signal $\overline{\beta}_j=21/2000$. The features are simulated from the
+	same distribution as in Scenario 1.
+
+	This sparse truth with differential signal between the groups is expected to
+	yield a somewhat better performance by \texttt{gren} than the regular elastic
+	net and \texttt{GRridge}. The group lasso extensions are expected to give a 
+	similar performance.
 

### 2019-03-26 code/simulations.R

@@ -43,7 +43,7 @@ for(i in 1:p) {
     Sigma[i, j] <- rho^abs(i - j)
   }
 }
-b <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
+b1 <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
 beta <- numeric(p)
 
 part <- rep(c(1:G), each=p/G)
@@ -66,7 +66,7 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
     idz <- c(((g-1)*p/G + 1):(g*p/G - rg))
     ida <- c(((g*p/G - rg) + 1):(g*p/G))
     beta[idz] <- 0
-    beta[ida] <- runif(rg, 0, b*f^(g-1))
+    beta[ida] <- runif(rg, 0, b1*f^(g-1))
   }
   xtrain <- rmvnorm(n, rep(0, p), Sigma)
   ytrain <- rbinom(n, 1, as.numeric(1/(1 + exp(-xtrain %*% beta))))
@@ -607,6 +607,14 @@ for(i in 1:p) {
 beta.group <- c(rep(0, 8), 0.2, 0.5)
 q.zero <- 0.85
 
+beta <- numeric(p)
+for(g in 1:G) {
+  beta[((g - 1)*p/G + 1):(g*p/G)] <- rep(beta.group[g], p/G)
+  if(beta.group[g]!=0) {
+    beta[((g - 1)*p/G + 1):((g - 1)*p/G + q.zero*p/G)] <- 0
+  }
+}
+
 part <- rep(c(1:G), each=p/G)
 csel <- 2^c(1:8)
 nreps <- 100
@@ -624,14 +632,6 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   
   print(paste("rep", k))
   set.seed(2019 + k)
-  
-  beta <- numeric(p)
-  for(g in 1:G) {
-    beta[((g - 1)*p/G + 1):(g*p/G)] <- rep(beta.group[g], p/G)
-    if(beta.group[g]!=0) {
-      beta[((g - 1)*p/G + 1):((g - 1)*p/G + q.zero*p/G)] <- 0
-    }
-  }
   
   xtrain <- rmvnorm(n, mean=rep(0, p), sigma=Sigma)
   ytrain <- rbinom(n, 1, as.numeric(1/(1 + exp(-xtrain %*% beta))))

### 2019-03-27 docs/manuscript.Rnw
incorporated suggestions by Mark

@@
 	Two such studies are introduced in Sections \ref{sec:colorectal} and
-	\ref{sec:lymphnode}. The first study is concerned with treatment 
+	\ref{sec:oral}. The first study is concerned with treatment 
 	response prediction for colorectal cancer patients, based on sequenced
@@
 	the group lasso penalises and selects features at the group level. 
-	There are two limitations to this approach: the group lasso (a) selects groups
-	as a whole instead of single 
+	There are two limitations to this approach: the group lasso (a) selects entire
+	groups instead of single 
 	features and (b) does not penalise adaptively: all groups are penalised 
 	equally. Extensions such as the sparse group lasso 
-	\cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
+	\cite[]{simon_sparse-group_2013} deal with (a), but do not address (b).
 
@@
 
-	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
+	Recently, \cite{zhang_novel_2019} introduced VBVS, a method similar to ours in
 	spirit. Nevertheless, one clear advantage of our method over VBVS is the
@@
 	our method is different from recent developments in grouped penalized 
-	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
+	regression such as \cite{dondelinger_joint_2018}, where instead of the 
 	features, the observations are partitioned.
@@
 	\ref{sec:simulations} we compare the method in a simulation study. In Sections
-	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
+	\ref{sec:colorectal} and \ref{sec:oral}, we demonstrate the approach in 
 	two data sets. In Section \ref{sec:discussion} 
@@
 	distributions 
-	\cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, 
+	\cite[]{alhamzawi_bayesian_2018,nengjun_yi_bayesian_2013,
+	kyung_penalized_2010}, 
 	but the precise parametrisation of this gamma prior is not so obvious. A 
@@
 	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
-	choices of $\alpha$. We recommend the latter.
+	choices of $\alpha$. We use the latter.
 
@@
   [! section ref here]. For the methods with a
-  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
+  tuning parameter $\alpha$, we picked the $\alpha$ 
+  (two $\alpha$'s for \texttt{gren}) that gave the best
   performance and refer
@@
   all group lasso extensions that we investigated.
-  (SM Section \ref{sm-sec:colorectal}). Overall, it seems as though the 
+  (SM Section \ref{sm-sec:colorectal}). Overall, the 
   inclusion of the extra information on the 
@@
 
-  \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
-  The aim of our second example study was to predict lymph node metastasis in
+  \section{Application to RNAseq in oral cancer}\label{sec:oral}
+  The aim of our second example study is to predict lymph node metastasis in
   oral cancer patients using sequenced RNAs from TCGA 
   \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
-  3096 transformed and normalised TCGA RNASeqv2 profiles of head-and-neck 
-  squamous cell carcinomas on 133 patients with HPV-negative tumours in the oral
-  cavity. Of these patients, 76 suffered from lymph node metastasis, while 57 
+  3096 transformed and normalised TCGA RNASeqv2 expression values of 3096 genes 
+  for 133 HPV-negative oral tumours. Of the corresponding patients, 76 suffered 
+  from lymph node metastasis, while 57 
   did not. For a thorough introduction of these data, see 
@@
   the same patients, quantified by Kendall's $\tau$ and binned into five groups.
-  In addition, we used (b) $p$-values from a microarray data set
+  In addition, we used (b) $p$-values from an independent microarray data set
   described in \cite{mes_prognostic_2017}, again binned into five groups. 
@@
   comparisons and performance measures are presented in the SM Section 
-  \ref{sm-sec:lymphnode}.
+  \ref{sm-sec:oral}.
 
@@
 @
+
+  \section{Additional applications}\label{sec:applications}
+  Two more applications are presented in SM Sections \ref{sm-sec:alzheimer} and
+  \ref{sm-sec:cervical}. The first one is concerned with diagnosis of 
+  Alzheimer's from 230 metabolites' expression levels in 174 subjects. We 
+  included several sources of extra information on the metabolites. In this 
+  example, \texttt{gren} performs worse than the group lasso methods. This is 
+  due to the smaller number of features to learn the penalty parameters from. 
+  Additionally, the metabolites are highly collinear, which further impairs 
+  estimation. The group lasso extensions depend on just one penalty parameter,
+  so suffer less from the small number of features.
+
+  In the second extra example, we attempt to diagnose 56 women with cervical 
+  lesions using 2576 sequenced microRNAs with the purpose of 
+  selecting relevent markers for cervical cancer screening programmes. We use 
+  include one partitioning of the features to enhance predictive performance and 
+  feature selection. In this example, \texttt{gren} outperforms the 
+  regular elastic net with respect to predictive performance for a range of 
+  model sizes.
 
@@
 	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
-	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
+	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}--
 	\ref{sm-sec:cervical}, adaptation to 
@@
 	discussed in Sections \ref{sec:colorectal}--
-	\ref{sec:lymphnode}, and SM Sections \ref{sm-sec:colorectal}--
+	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}--
 	\ref{sm-sec:cervical}, we investigated the occurrence of multiple optima, 
@@
                      stringsAsFactors=FALSE)
+
 bench2 <- read.table("results/rnaseq_lymph_node_metastasis_bench1.csv", 
@@
 @
-  The double EM loop runs the danger of excessive computation time. We have   
-  compared computation times to the
+  The double EM loop runs the danger of excessive computation time. In our 
+  implementation we have reduced computational time considerably by applying
+  some computational tricks (see SM Section 
+  \ref{sm-sec:computationalcomplexity}). In addition, the computationally heavy
+  parts were implemented in \texttt{C++} to further reduce the computation time. 
+  To assess this reduction, we have compared computation times to the
   other methods introduced in Sections \ref{sec:colorectal}--
-  \ref{sec:lymphnode}. All timings were obtained on a Macbook 
+  \ref{sec:oral}. All timings were obtained on a Macbook 
   Pro 2016 running \Sexpr{session}. 
@@
   \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
-  respectively. \texttt{GRridge} is 
-  slightly faster with \Sexpr{times1[4]} seconds, while \texttt{SGL} was slightly
-  slower: \Sexpr{times1[c(5:7)]} seconds, with 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+  respectively. \texttt{GRridge} and \texttt{SGL} with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
+  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
+  respectively.
   \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
@@
   (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}. 
-  We note that a considerable speedup may be achieved by iterating between 
-  single EB and single VB updates, instead of running the VB EM 
-  until convergence. This works in practice, because an EM tends to take a few 
-  initial large steps towards the optimum, followed by a series of smaller 
-  steps. We argue that the first few iterations of the VB EM render the 
-  approximate expectation in \ref{eq:approximatedloglikelihood}
-  accurate enough for a 
-  reliable EB step.
+  % We note that a considerable speedup may be achieved by iterating between 
+  % single EB and single VB updates, instead of running the VB EM 
+  % until convergence. This works in practice, because an EM tends to take a few 
+  % initial large steps towards the optimum, followed by a series of smaller 
+  % steps. We argue that the first few iterations of the VB EM render the 
+  % approximate expectation in \ref{eq:approximatedloglikelihood}
+  % accurate enough for a 
+  % reliable EB step.
 
@@
 	Supplementary Material is available online from 
-	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All results may be recreated
-	from \url{https://github.com/magnusmunch/gren}.
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}. The results and documents may 
+	be recreated from \url{https://github.com/magnusmunch/gren}.
 
 	\section*{Acknowledgements}
-	\textit{Conflict of Interest}: None declared. We thank the editors and 
-	referees for their useful suggestions and comments.
+	We thank the editor and referees for their useful suggestions and comments. 
+	\textit{Conflict of Interest}: None declared.
 

### 2019-03-27 docs/supplement.Rnw
incorporated suggestions by Mark

@@
   Additional results to the applications in the MD Sections 
-  \ref{md-sec:colorectal}--\ref{md-sec:lymphnode}. We
+  \ref{md-sec:colorectal}--\ref{md-sec:oral}. We
   apply our method to a study on metabolomics in Alzheimer's in Section
@@
 
-  \section{Application to RNAseq in lymph node cancer}
-  \label{sec:lymphnode}
-  In addition to the results presented in Section \ref{md-sec:lymphnode} of the 
+  \section{Application to RNAseq in oral cancer}
+  \label{sec:oral}
+  In addition to the results presented in Section \ref{md-sec:oral} of the 
   MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 
@@
   screening markers for cervical cancer screening. The next generation 
-  sequencing analysis resulted in 2,576 transcripts. The data was normalized 
+  sequencing analysis resulted in 2576 transcripts. The data was normalized 
   and pre-processed, rendering 752 transcripts. More details of the
@@
 
+  \section{Gibbs sampler}
+
+  \begin{align*}
+    \bbeta | \bomega, \bpsi & \sim \mathcal{N} (\bSigma, \bmu) \\
+      & \text{with } \bSigma = [\X \tr \diag (\bomega_i) \X + 
+      \lambda_2 \bLambda' + \lambda_2 \bLambda' \diag (\psi_j^{-1})]^{-1} 
+      \text{ and } \bmu = \bSigma \X \tr (\y - \frac{1}{2} \mathbf{m}), \\
+    \omega_i | \bbeta & \sim \mathcal{PG} (m_i, |\x_i \tr \bbeta|), \\
+    \psi_j^{-1} | \beta_j & \sim \mathcal{IG} (\lambda_1/ (2 \lambda_2 
+    \sqrt{\lambda'_{g(j)}} |\beta_j|), \lambda_1^2/(4 \lambda_2)).
+  \end{align*}
+
   \bibliographystyle{author_short3.bst} 

### 2019-03-27 code/figures.R

@@ -646,22 +646,24 @@ methods <- c("ridge", "grridge", paste0("gren", c(1:3)), paste0("enet", c(1:3)),
              paste0("sgl", c(1:3)), paste0("cmcp", c(1:3)), 
              paste0("gel", c(1:3)), paste0("ocmcp", c(1:3)), 
              paste0("ogel", c(1:3)))
-col <- bpy.colors(length(methods), cutoff.tail=0.1)
+col <- bpy.colors(9, cutoff.tail=0.1)
 lty <- 1:length(methods)
-labels <- expression("ridge", "GRridge", "gren,"~alpha==0.05, 
-                     "gren,"~alpha==0.5, "gren,"~alpha==0.95, 
-                     "enet,"~alpha==0.05, "enet,"~alpha==0.5, 
-                     "enet,"~alpha==0.95, "SGL,"~alpha==0.05, "SGL,"~alpha==0.5, 
-                     "SGL,"~alpha==0.95, "cMCP,"~alpha==0.05, 
-                     "cMCP,"~alpha==0.5, "cMCP,"~alpha==0.95, 
-                     "gel,"~alpha==0.05, "gel,"~alpha==0.5, "gel,"~alpha==0.95,
-                     "OcMCP,"~alpha==0.05, "OcMCP,"~alpha==0.5, 
-                     "OcMCP,"~alpha==0.95, "Ogel,"~alpha==0.05, 
-                     "Ogel,"~alpha==0.5, "Ogel,"~alpha==0.95)
+labels <- list(expression("ridge", "GRridge", "gren,"~alpha==0.05, 
+                          "enet,"~alpha==0.05, "SGL,"~alpha==0.05, 
+                          "cMCP,"~alpha==0.05, "gel,"~alpha==0.05, 
+                          "OcMCP,"~alpha==0.05, "Ogel,"~alpha==0.05),
+               expression("ridge", "GRridge", "gren,"~alpha==0.5, 
+                          "enet,"~alpha==0.5, "SGL,"~alpha==0.5, 
+                          "cMCP,"~alpha==0.5, "gel,"~alpha==0.5, 
+                          "OcMCP,"~alpha==0.5, "Ogel,"~alpha==0.5),
+               expression("ridge", "GRridge", "gren,"~alpha==0.95, 
+                          "enet,"~alpha==0.95, "SGL,"~alpha==0.95, 
[…]

### 2019-03-27 code/functions.R

@@ -61,24 +61,6 @@ sel.grpreg <- function(X, y, group=1:ncol(X),
   
 }
 
-n <- 100
-p <- 20
-x <- matrix(rnorm(n*p), ncol=p, nrow=n)
-beta <- -9.5:9.5
-y <- rbinom(n, 1, as.numeric(1/(1 + exp(-x %*% beta))))
-m <- rep(1, n)
-alpha <- 0.5
-lambda <- 1
-lambdamult <- rep(1, p)
-intercept <- FALSE
-control <- list(nsamples=1000)
-
-fit.mcmc <- mcmc.gren(x, y, m=rep(1, nrow(x)), 
-                      alpha=0.5, lambda=1, lambdamult=rep(1, p), 
-                      intercept=FALSE, control=list(nsamples=1000))
-plot(c(beta), apply(fit.mcmc$samples$beta, 1, mean))
-
-
 # mcmc version of VB steps
 # devtools::install_version("BayesLogit", "0.2-0")
 mcmc.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL, 
@@ -142,5 +124,123 @@ mcmc.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL,
   return(out)
   
 }
+
+n <- 100
+p <- 20
+x <- matrix(rnorm(n*p), ncol=p, nrow=n)
+beta <- -9.5:9.5
+y <- rbinom(n, 1, as.numeric(1/(1 + exp(-x %*% beta))))
+m <- rep(1, n)
+alpha <- 0.5
+lambda <- 1
+lambdamult <- rep(1, p)
+
+
+fit.vb <- vb.gren(x=x, y=y, m=m, alpha=alpha, lambda=lambda, 
+                  lambdamult=lambdamult, intercept=TRUE, 
+                  control=list(maxit=100, epsilon=0.001))
+fit.mcmc <- mcmc.gren(x=x, y=y, m=m, alpha=alpha, lambda=lambda, 
+                      lambdamult=lambdamult, intercept=TRUE, 
+                      control=list(nsamples=10000))
+test1 <- diag(fit.vb$posterior$sigma)[-1]
+test2 <- fit.vb$posterior$LRdsigma
+test3 <- apply(fit.mcmc$samples$beta, 1, var)[-1]
+plot(test1, test3)
+plot(test2, test3)
+plot(test1, test2)
+
[…]

### 2019-04-01 code/simulations.R
finished simulation 2

@@ -12,7 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-parallel <- TRUE
+parallel <- FALSE
 
 ### libraries
 library(gren)

### 2019-04-01 docs/manuscript.Rnw
added all suggestions

--- 2019-03-27 docs/manuscript.Rnw
@@
 titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
-xr-hyper,hyperref}
+xr-hyper,hyperref,enumerate}
 \externaldocument[sm-]{supplement}
@@
 
-	Recently, \cite{zhang_novel_2019} introduced VBVS, a method similar to ours in
+	In line with the above, our proposed method is (a) objective: the 
+	hyper-parameters are estimated from the data and not chosen on subjective 
+	grounds, (b) adaptive: prior information is automatically and subjectively 
+	weighed with respect to its informativeness, (c) fast compared to full 
+	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
+	user: only the data and grouping of the features is required. We emphasize 
+	that our method is different from recent developments in grouped penalized 
+	regression such as \cite{dondelinger_joint_2018}, where instead of the 
+	features, the observations are partitioned.
+
+	Recently, \cite{zhang_novel_2019} introduced Variational Bayes Variable
+	Selection (VBVS), a method similar to ours in
 	spirit. Nevertheless, one clear advantage of our method over VBVS is the
@@
 	adaptive inclusion of external information on the features.
-
-	In line with the above, our proposed method is (a) subjective: the 
-	hyper-parameters are estimated from the data and not chosen on objective 
-	grounds, (b) adaptive: prior information is automatically and subjectively 
-	weighed with respect to its informativeness, (c) fast compared to full 
-	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
-	user: only the data and grouping of the features is required. We emphasize that 
-	our method is different from recent developments in grouped penalized 
-	regression such as \cite{dondelinger_joint_2018}, where instead of the 
-	features, the observations are partitioned.
 
@@
 	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016},     
-	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
 	composite mimimax concave penalty (\texttt{cMCP}) 
@@
 	However, in contrast to the original group lasso,
-	they select single features, instead of just whole groups.
+	they select single features, instead of complete groups.
 
-	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{cMCP}, 
+	For the regular elastic net, \texttt{gren}, \texttt{cMCP}, 
 	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
@@
 	where $\alpha=0$, while the third setting is similar to the lasso with 
-	$\alpha=1$. In \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}, $\alpha$
+	$\alpha=1$. In \texttt{cMCP} and \texttt{gel}, $\alpha$
 	determines the ratio of group-level penalization ($\alpha=0$) to 
@@
 	We simulated data according to five different scenarios:
-	\begin{enumerate}
-		\item differential signal between the groups and uniformly distributed model
+	\begin{enumerate}[(i)]
+		\item \label{it:scenario1} differential signal between the groups and uniformly
+		distributed model
 		parameters,
-		\item a large number of small groups of features,
-		\item no differential signal between the groups, but strong correlations 
+		\item \label{it:scenario2} a large number of small groups of features,
+		\item \label{it:scenario3} no differential signal between the groups, but 
+		strong correlations 
 		within groups of features,
-		\item differential signal between the groups and heavy-tailed distributed 
+		\item \label{it:scenario4} differential signal between the groups and 
+		heavy-tailed distributed 
 		model parameters, and
-		\item a very sparse setting with no signal in some of the groups.
+		\item \label{it:scenario5}a very sparse setting with no signal in some of 
+		the groups.
 	\end{enumerate}
@@
 	\ref{sm-sec:simulations} of 
-	the SM. In the first scenario, the distribution of model parameters is far
+	the SM. In Scenario (\ref{it:scenario1}), the distribution of model 
+	parameters is far
 	from the assumed ridge/elastic net/lasso distributions, so we expect
-	all methods to underperform. In the second scenario, which is similar to
+	all methods to underperform. In Scenario (\ref{it:scenario2}), 
+	which is similar to
 	the simulations presented in the original group lasso paper 
@@
 	estimate just one penalty parameter, so are less sensitive to overfitting.
-	In scenario three, we expect all methods to underperform
-	due to the high correlations between the features. The fourth 
-	scenario is expected to yield a better \texttt{gren} performance, since
+	In Scenario (\ref{it:scenario3}), we expect all methods to underperform
+	due to the high correlations between the features. The
+	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
+	performance, since
 	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
 	detect differential signal between
-	the groups. Scenario five is expected to yield a bad performance for all 
+	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
+	performance for all 
 	methods, because of the sparse truth and lack of differential 
@@
 	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
-	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
-	$p=900$ in scenario 2 and $p=1000$ in the other scenarios. We varied the 
+	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
+	from $p=900$ to $p=1000$ over the Scenarios. We varied the 
 	covariance matrix of the features and simulation of 
 	the model parameters $\beta_j$, $j=1, \dots, p$ between the
-	scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
+  Scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
 	parametrisations). We 
@@
 	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
-	Additional results and performance measures are presented in Section 
+	For the methods with an $\alpha$ tuning parameter, we show the one that gives
+	the best performance. Results for the other $\alpha$'s and additional
+	performance measures are presented in Section 
 	\ref{sm-sec:simulations} of the SM.
 
-	- description of results
+	In Scenario (\ref{it:scenario1})
+	\texttt{gren} and \texttt{GRridge} don't suffer much from the wrong 
+	distributional assumptions as they outperform the other methods.
+	Scenario (\ref{it:scenario2}) follows our expectation \texttt{gren}, and to
+	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
+	the regular elastic net outperforms the group lasso methods, which are 
+	designed to handle situations such as Scenario (\ref{it:scenario2}).
+	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
+	regular elastic net suffer more from the higher correlations. Here,
+	the regular elastic net outperforms \texttt{gren}, which is an indication
+	that the estimation of the penalty parameters is impaired by high
+	correlation between features. Scenario (\ref{it:scenario4}) follows the 
+	expected pattern, with the group-regularized methods (\texttt{GRridge} and
+	\texttt{gren}) outperforming their respective non group-regularized 
+	counterparts, as well as the sparse group lasso extensions.
+	Surprisingly, in Scenario (\ref{it:scenario5}), \texttt{gren} outperforms all
+	other methods, which is an indication that \texttt{gren} is able to pick up 
+	very sparse differential signals between the groups.
 
 <<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res2_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 2.", out.width="100%", fig.asp=1/2>>=
 @
@@
   three methods, we only present \texttt{cMCP} here; the results for 
-  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
-  [! section ref here]. For the methods with a
+  another extension of the group lasso, the sparse
+  group lasso by \cite{simon_sparse-group_2013}
+  (\texttt{SGL}), and \texttt{gel} are presented in the SM Section 
+  \ref{sm-sec:colorectal}. For the methods with a
   tuning parameter $\alpha$, we picked the $\alpha$ 
-  (two $\alpha$'s for \texttt{gren}) that gave the best
-  performance and refer
+  that gave the best performance and refer
   the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
@@
   (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
-  \texttt{gren} seems to outperform the other methods for a range of model 
-  sizes in terms of AUC 
+  \texttt{gren} seems to outperform the other methods for a range of smaller 
+  model sizes in terms of AUC 
   (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
-  This is especially pronounced for the larger models. The performance of
+  Selection of larger models is difficult with large values of $\alpha$, an 
+  inherent property of the strong $L_1$-norm penalization. With lower values of
+  $\alpha$, larger models are possible (see SM Section \ref{sm-sec:colorectal})
+  for comparisons with smaller values of $\alpha$). The performance of
   \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
@@
   \texttt{gren} outperform their respective non-group-regularized versions, 
-  albeit it only slightly for \texttt{GRridge}.
+  albeit only slightly for \texttt{GRridge}.
 <<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@@
   multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
-  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
+  \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}. Additional
   comparisons and performance measures are presented in the SM Section 
@@
   The estimated penalty multipliers follow the expected pattern 
-  (Figure \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}a): 
+  (Figure \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}a): 
   larger cis-correlations receive smaller penalties. 
@@
   identical, while \texttt{gren} outperforms the regular elastic net.
-<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the lymph node metastasis example.", out.width="100%", fig.asp=1>>=
+<<barplot_lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
 @
@@
   due to the smaller number of features to learn the penalty parameters from. 
-  Additionally, the metabolites are highly collinear, which further impairs 
-  estimation. The group lasso extensions depend on just one penalty parameter,
-  so suffer less from the small number of features.
+  Additionally, many metabolites are strongly negatively correlated, which
+  further impairs estimation. The group lasso extensions depend on just one
+  penalty parameter, so suffer less from the small number of features and
+  negative correlations.
 
@@
 	the adaptation more robust. As can be seen from both the simulations in 
-	Section \ref{sec:simulations} and the applications to different outcomes 
-	(prognostic or diagnostic) and marker types 
-	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
-	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}--
-	\ref{sm-sec:cervical}, adaptation to 
+	Section \ref{sec:simulations} and the applications in Sections 
+	\ref{sec:colorectal}-\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}-
+	\ref{sm-sec:cervical} adaptation to 
 	external data may be beneficial for prediction and feature selection. 
+	We show that to be true for both prognostic and diagnostic targets,
+	as well as a range of marker types (RNAseq, microRNAs, or metabolomics).
 	We believe that this is due to the `borrowing of information' effect: 
@@
 
-bench2 <- read.table("results/rnaseq_lymph_node_metastasis_bench1.csv", 
+bench2 <- read.table("results/rnaseq_oral_cancer_bench1.csv", 
                      stringsAsFactors=FALSE)
@@
   (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
-  The results for the lymph node cancer example are similar. \texttt{gren} 
+  The results for the oral cancer example are similar. \texttt{gren} 
   achieved 
@@
 	Supplementary Material is available online from 
-	\url{http://pub.math.leidenuniv.nl/~munchmm/}. The results and documents may 
-	be recreated from \url{https://github.com/magnusmunch/gren}.
+	\url{http://biostatistics.oxfordjournals.org}.
+
+	\section*{Reproducible Research}
+	All results and documents may be recreated from 
+	\url{https://github.com/magnusmunch/gren}.
+
+	\section*{Funding}
+	This research has received funding from the European Research Council under 
+	ERC Grant Agreement 320637.
 
@@
 	\textit{Conflict of Interest}: None declared.
-
-	\section*{Funding}
-	This research has received funding from the European Research Council under 
-	ERC Grant Agreement 320637.
 

### 2019-04-01 docs/manuscript_biostatistics.Rnw
added all suggestions

@@
 \documentclass[oupdraft]{bio}
-\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
-\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable}
+\usepackage{xr-hyper}
+% \usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, 
+% citecolor=citecolor]{hyperref}
+\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,
+enumerate}
+
+\externaldocument[sm-]{supplement}
 
@@
 
-	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
+	In line with the above, our proposed method is (a) subjective: the 
+	hyper-parameters are estimated from the data and not chosen on objective 
+	grounds, (b) adaptive: prior information is automatically and subjectively 
+	weighed with respect to its informativeness, (c) fast compared to full 
+	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
+	user: only the data and grouping of the features is required. We emphasize 
+	that our method is different from recent developments in grouped penalized 
+	regression such as \cite{dondelinger_joint_2018}, where instead of the 
+	features, the observations are partitioned.
+
+	Recently, \cite{zhang_novel_2019} introduced VBVS, a method similar to ours in
 	spirit. Nevertheless, one clear advantage of our method over VBVS is the
@@
 
-	In line with the above, our proposed method is (a) subjective: the 
-	hyper-parameters are estimated from the data and not chosen on objective 
-	grounds, (b) adaptive: prior information is automatically and subjectively 
-	weighed with respect to its informativeness, (c) fast compared to full 
-	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
-	user: only the data and grouping of the features is required. We emphasize that 
-	our method is different from recent developments in grouped penalized 
-	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
-	features, the observations are partitioned.
-
 	The rest of the paper is structured as follows. We introduce the Bayesian 
@@
 
-	- description of results
+	In Scenario (\ref{it:scenario1})
+	\texttt{gren} and \texttt{GRridge} don't suffer much from the wrong 
+	distributional assumptions as they outperform the other methods.
+
+	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
+	regular elastic net suffer more from the higher correlations. Here,
+	the regular elastic net outperforms \texttt{gren}, which is an indication
+	that the estimation of the penalty parameters is impaired by high
+	correlation between features. Scenario (\ref{it:scenario4}) follows the 
+	expected pattern, with the group-regularized methods (\texttt{GRridge} and
+	\texttt{gren}) outperforming their respective non group-regularized 
+	counterparts, as well as the sparse group lasso extensions.
+	Surprisingly, in Scenario (\ref{it:scenario5}), \texttt{gren} outperforms all
+	other methods, which is an indication that \texttt{gren} is able to pick up 
+	very sparse differential signals between the groups.
 
@@
   three methods, we only present \texttt{cMCP} here; the results for 
-  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
+  another extension of the group lasso, the sparse
+  group lasso by \cite{simon_sparse-group_2013}
+  (\texttt{SGL}), and \texttt{gel} are presented in the SM Section 
   [! section ref here]. For the methods with a
-  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
+  tuning parameter $\alpha$, we picked the $\alpha$ 
+  (two $\alpha$'s for \texttt{gren}) that gave the best
   performance and refer
@@
   identical, while \texttt{gren} outperforms the regular elastic net.
+
+  \section{Additional applications}\label{sec:applications}
+  Two more applications are presented in SM Sections \ref{sm-sec:alzheimer} and
+  \ref{sm-sec:cervical}. The first one is concerned with diagnosis of 
+  Alzheimer's from 230 metabolites' expression levels in 174 subjects. We 
+  included several sources of extra information on the metabolites. In this 
+  example, \texttt{gren} performs worse than the group lasso methods. This is 
+  due to the smaller number of features to learn the penalty parameters from. 
+  Additionally, the metabolites are highly collinear, which further impairs 
+  estimation. The group lasso extensions depend on just one penalty parameter,
+  so suffer less from the small number of features.
+
+  In the second extra example, we attempt to diagnose 56 women with cervical 
+  lesions using 2576 sequenced microRNAs with the purpose of 
+  selecting relevent markers for cervical cancer screening programmes. We use 
+  include one partitioning of the features to enhance predictive performance and 
+  feature selection. In this example, \texttt{gren} outperforms the 
+  regular elastic net with respect to predictive performance for a range of 
+  model sizes.
 
@@
 	Supplementary Material is available online from
-	\url{http://pub.math.leidenuniv.nl/~munchmm/}. All results may be recreated
-	from \url{https://github.com/magnusmunch/gren}.
+	\url{http://pub.math.leidenuniv.nl/~munchmm/}. The results and documents may
+	be recreated from \url{https://github.com/magnusmunch/gren}.
 
 	\section*{Acknowledgements}
-	\textit{Conflict of Interest}: None declared. We thank the editors and 
-	referees for their useful suggestions and comments.
+	We thank the editor and referees for their useful suggestions and comments.
+	\textit{Conflict of Interest}: None declared.
 

### 2019-04-01 docs/supplement.Rnw
added all suggestions

@@
   proposed method is shown in Section \ref{sec:fullestimationprocedure}. 
+  Extensions of our method are
+  presented in Section \ref{sec:extensions}. The simulations in MD Section
+  \ref{md-sec:simulations} are elaborated in Section \ref{sec:simulations}. 
   Additional results to the applications in the MD Sections 
-  \ref{md-sec:colorectal}--\ref{md-sec:oral}. We
+  \ref{md-sec:colorectal}--\ref{md-sec:oral} are presented in Sections 
+  \ref{sec:colorectal}--\ref{sec:oral}. We
   apply our method to a study on metabolomics in Alzheimer's in Section
@@
 
-  Inspection of the updating equations in the MD, learns us that the 
+  Inspection of the updating equations (\ref{md-eq:VBupdateequations}) in the 
+  MD, learns us that the 
   last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
@@
 
+	\subsection{Additional results}
+	Additional results to MD Section \ref{md-sec:simulations} are given in 
+	Figures \ref{fig:lines2_simulations_res1_auc_briers}-
+	\ref{fig:lines2_simulations_res5_auc_briers}.
+
+<<lines2_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res2_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 2.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res3_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 3.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res4_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res5_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5.", out.width="100%", fig.asp=1/2>>=
+@
+
   \section{Application to MicroRNAs in colorectal cancer}\label{sec:colorectal}
@@
   \ref{fig:lines_micrornaseq_colorectal_cancer_res1_briers}.
-<<lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated AUC for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
-@
-
-<<lines_micrornaseq_colorectal_cancer_res1_briers, fig.cap="Estimated Brier skill score for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
+<<lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated AUC for the colorectal cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", fig.asp=1>>=
+@
+
+<<lines_micrornaseq_colorectal_cancer_res1_briers, fig.cap="Estimated Brier skill score for the colorectal cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
 @
@@
   validation set are presented in Figures 
-  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_auc}--
-  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_briers}
-<<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated AUC for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
-@
-
-<<lines_rnaseq_lymph_node_metastasis_res1_briers, fig.cap="Estimated Briers skill score for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
+  \ref{fig:lines_rnaseq_oral_cancer_res1_auc}--
+  \ref{fig:lines_rnaseq_oral_cancer_res1_briers}
+<<lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated AUC for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_rnaseq_oral_cancer_res1_briers, fig.cap="Estimated Briers skill score for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
 @
@@
 
-  In this example, \texttt{GRridge} and \texttt{gren} underperform due to a 
-  small number of features to learn the penalty parameters from. The other 
+  In this example, \texttt{GRridge} and \texttt{gren} underperform due to:
+  (i) a small number of features to learn the penalty parameters from and (ii)
+  Large negative correlations between the metabolites (see Figure 
+  \ref{fig:heatmap_metabolomics_alzheimer_res1_cor}). Both (i) and (ii) make 
+  estimation of the penalty multipliers difficult. The other 
   methods estimate one global penalty parameter, so do not suffer from the 
-  lower dimensionality of the problem.
+  lower dimensionality of the problem and the large negative correlation.
+
+<<heatmap_metabolomics_alzheimer_res1_cor, fig.cap="Correlations between the metabolites.", out.width="50%", fig.asp=1>>=
+library(Biobase)
+library(lattice)
+load("data/ESetMbolCSFPR2.Rdata")
+pheno <- pData(ESetMbolCSFPR2)
+metabol <- t(exprs(ESetMbolCSFPR2))
+metabol.apoe <- metabol[(pheno$D_diag_name=="Probable AD" & 
+                           pheno$APOE=="E4YES") |
+                          (pheno$D_diag_name=="Subjectieve klachten" & 
+                             pheno$APOE=="E4NO"), ]
+cor.mat <- cor(scale(metabol.apoe))
+levelplot(cor.mat, xlab=NULL, ylab=NULL, scales=list(draw=FALSE))
+@
 
@@
 
+  The estimated multipliers, AUC, and Brier skill score
+  are in Figures \ref{fig:barplot_micrornaseq_cervical_cancer_res1}--
+  \ref{fig:lines_micrornaseq_cervical_cancer_res1_briers}, respectively. In this
+  example the group regularized methods outperform the regular methods with
+  respect to discrimination (AUC) and calibration (Brier skill score).
+
 <<barplot_micrornaseq_cervical_cancer_res1, fig.cap="Estimated penalty multiplies for conservation status in cervical cancer example.", out.width="50%", fig.asp=1>>=
@@
 
-  \section{Gibbs sampler}
-
-  \begin{align*}
-    \bbeta | \bomega, \bpsi & \sim \mathcal{N} (\bSigma, \bmu) \\
-      & \text{with } \bSigma = [\X \tr \diag (\bomega_i) \X + 
-      \lambda_2 \bLambda' + \lambda_2 \bLambda' \diag (\psi_j^{-1})]^{-1} 
-      \text{ and } \bmu = \bSigma \X \tr (\y - \frac{1}{2} \mathbf{m}), \\
-    \omega_i | \bbeta & \sim \mathcal{PG} (m_i, |\x_i \tr \bbeta|), \\
-    \psi_j^{-1} | \beta_j & \sim \mathcal{IG} (\lambda_1/ (2 \lambda_2 
-    \sqrt{\lambda'_{g(j)}} |\beta_j|), \lambda_1^2/(4 \lambda_2)).
-  \end{align*}
+  % \section{Gibbs sampler}
+  %   
+  % \begin{align*}
+  %   \bbeta | \bomega, \bpsi & \sim \mathcal{N} (\bSigma, \bmu) \\
+  %     & \text{with } \bSigma = [\X \tr \diag (\bomega_i) \X + 
+  %     \lambda_2 \bLambda' + \lambda_2 \bLambda' \diag (\psi_j^{-1})]^{-1} 
+  %     \text{ and } \bmu = \bSigma \X \tr (\y - \frac{1}{2} \mathbf{m}), \\
+  %   \omega_i | \bbeta & \sim \mathcal{PG} (m_i, |\x_i \tr \bbeta|), \\
+  %   \psi_j^{-1} | \beta_j & \sim \mathcal{IG} (\lambda_1/ (2 \lambda_2 
+  %   \sqrt{\lambda'_{g(j)}} |\beta_j|), \lambda_1^2/(4 \lambda_2)).
+  % \end{align*}
 

### 2019-04-01 code/figures.R

@@ -4,35 +4,39 @@ library(sp)
 load(file="results/simulations_res1.Rdata")
 plot.data <- c(sapply(c("grridge", "ridge"), function(m) {
   unname(cbind(NA, mean(sapply(res, function(s) {
-    s$auc[grep(m, names(s$auc))[1]]}), na.rm=TRUE),
+    s$auc[substr(names(s$auc), 1, nchar(m))==m]}), na.rm=TRUE),
     mean(sapply(res, function(s) {
-      s$auc[grep(m, names(s$briers))[1]]}), na.rm=TRUE)))}, simplify=FALSE),
+      s$auc[substr(names(s$briers), 1, nchar(m))==m]}), na.rm=TRUE)))}, 
+  simplify=FALSE),
   sapply(c(paste0("gren", 1:3), paste0("enet", 1:3),
            paste0("cmcp", 1:3), paste0("gel", 1:3)), function(m) {
              cbind(rowMeans(sapply(res, function(s) {
-               vec <- s$psel[grep(m, names(s$psel))]
+               vec <- s$psel[substr(names(s$psel), 1, nchar(m))==m]
                replace(rep(NA, 8), 1:length(vec), vec)}), 
                na.rm=TRUE),
                rowMeans(sapply(res, function(s) {
-                 vec <- s$auc[grep(m, names(s$auc))]
+                 vec <- s$auc[substr(names(s$auc), 1, nchar(m))==m]
                  replace(rep(NA, 8), 1:length(vec), vec)}), 
                  na.rm=TRUE),
                rowMeans(sapply(res, function(s) {
-                 vec <- s$briers[grep(m, names(s$briers))]
+                 vec <- s$briers[substr(names(s$briers), 1, nchar(m))==m]
                  replace(rep(NA, 8), 1:length(vec), vec)}), 
                  na.rm=TRUE))}, simplify=FALSE))
 
 methods <- c("GRridge", "ridge", "gren", "enet", "cmcp", "gel")
-labels <- c(methods, expression(alpha==0.05), expression(alpha==0.5), 
-            expression(alpha==0.95))
[…]

### 2019-04-01 code/metabolomics_alzheimer.R

@@ -13,6 +13,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 
 ### libraries
 library(Biobase)
+library(lattice)
 library(gren)
 library(GRridge)
 library(grpregOverlap)

### 2019-04-02 docs/manuscript.Rnw
small changes to text of manuscript.Rnw and supplement.Rnw

@@
 % objects
-\newcommand{\x}{\mathbf{x}}
 \newcommand{\bbeta}{\bm{\beta}}
-\newcommand{\X}{\mathbf{X}}
-\newcommand{\y}{\mathbf{y}}
-\newcommand{\G}{\mathcal{G}}
 \newcommand{\btau}{\bm{\tau}}
-\newcommand{\blambda}{\bm{\lambda}}
+\newcommand{\balpha}{\bm{\alpha}}
 \newcommand{\bomega}{\bm{\omega}}
+\newcommand{\bOmega}{\bm{\Omega}}
+\newcommand{\bSigma}{\bm{\Sigma}}
+\newcommand{\bkappa}{\bm{\kappa}}
 \newcommand{\btheta}{\bm{\theta}}
 \newcommand{\bmu}{\bm{\mu}}
-\newcommand{\bSigma}{\bm{\Sigma}}
 \newcommand{\bpsi}{\bm{\psi}}
-\newcommand{\bOmega}{\bm{\Omega}}
+\newcommand{\blambda}{\bm{\lambda}}
+\newcommand{\bLambda}{\bm{\Lambda}}
+\newcommand{\bphi}{\bm{\phi}}
+\newcommand{\Y}{\mathbf{Y}}
+\newcommand{\y}{\mathbf{y}}
+\newcommand{\X}{\mathbf{X}}
+\newcommand{\x}{\mathbf{x}}
+\newcommand{\I}{\mathbf{I}}
+\newcommand{\T}{\mathbf{T}}
 \newcommand{\Z}{\mathbf{Z}}
-\newcommand{\bLambda}{\bm{\Lambda}}
+\newcommand{\W}{\mathbf{W}}
+\renewcommand{\O}{\mathbf{O}}
+\newcommand{\B}{\mathbf{B}}
+\renewcommand{\H}{\mathbf{H}}
+\newcommand{\U}{\mathbf{U}}
+\newcommand{\D}{\mathbf{D}}
+\newcommand{\Vm}{\mathbf{V}}
+\newcommand{\rv}{\mathbf{r}}
+\newcommand{\A}{\mathbf{A}}
+\newcommand{\Q}{\mathbf{Q}}
+\newcommand{\Sm}{\mathbf{S}}
+\newcommand{\0}{\bm{0}}
+\newcommand{\tx}{\tilde{\mathbf{x}}}
+\newcommand{\hy}{\hat{y}}
+\newcommand{\tm}{\tilde{m}}
+\newcommand{\tkappa}{\tilde{\kappa}}
 \newcommand{\m}{\mathbf{m}}
+\newcommand{\C}{\mathbf{C}}
+\renewcommand{\v}{\mathbf{v}}
+\newcommand{\g}{\mathbf{g}}
+\newcommand{\s}{\mathbf{s}}
+\newcommand{\ones}{\mathbf{1}}
 
 % functions and operators
+\renewcommand{\L}{\mathcal{L}}
+\newcommand{\G}{\mathcal{G}}
+\renewcommand{\P}{\mathcal{P}}
+\newcommand{\argmax}{\text{argmax} \,}
+\newcommand{\expit}{\text{expit}}
+\newcommand{\erfc}{\text{erfc}}
+\newcommand{\E}{\mathbb{E}}
+\newcommand{\V}{\mathbb{V}}
 \newcommand{\tr}{^{\text{T}}}
-\newcommand{\expit}{\text{expit}}
-\newcommand{\argmax}{\text{argmax} \,}
-\renewcommand{\L}{\mathcal{L}}
-\newcommand{\E}{\mathbb{E}}
 \newcommand{\diag}{\text{diag}}
+\newcommand{\KL}{D_{\text{KL}}}
 \newcommand{\trace}{\text{tr}}
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+
+% distributions
+\newcommand{\N}{\text{N}}
+\newcommand{\TG}{\text{TG}}
+\newcommand{\Bi}{\text{Binom}}
+\newcommand{\PG}{\text{PG}}
+\newcommand{\Mu}{\text{Multi}}
+\newcommand{\GIG}{\text{GIG}}
+\newcommand{\IGauss}{\text{IGauss}}
+\newcommand{\Un}{\text{U}}
 
@@
 \renewcommand{\]}{\right]}
-\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 
@@
 		approximate empirical-variational Bayes framework. Simulations and
-		applications in two different molecular marker and cancer type studies show
+		applications in two different cancer studies show
 		that, if the partitioning of the features is informative, classification 
@@
 	\ref{sec:oral}. The first study is concerned with treatment 
-	response prediction for colorectal cancer patients, based on sequenced
-	microRNAs. In the second study we attempt to predict lymph node metastases,
-	using RNAseq data. In both settings additional information on the 
+	response prediction for colorectal cancer patients, using sequenced
+	microRNAs. In the second study we predict lymph node metastases in oral cancer
+	patients, using RNAseq data. In both settings additional information on the 
 	features is available in the form of a partitioning. Furthermore, in both
-	examples data is sparse, so any extra included information may help to boost
-	predictive performance. Inclusion of such feature partitions in 
+	studies data is sparse, so any extra included information may help to boost
+	predictive performance. Inclusion of such extra information in 
 	the analysis is not straightforward with standard methods.
@@
 	the group lasso
-	\cite[]{meier_group_2008, yuan_model_2006} or one of its extensions, because
+	\cite[]{meier_group_2008, yuan_model_2006}, because
 	the group lasso penalises and selects features at the group level. 
@@
 	each group of variables receives its own penalty parameter. An apparent issue
-	with this differential penalization is the estimation of the penalty 
+	with differential penalization is the estimation of the penalty 
 	parameters. Naive estimation may be done by cross-validation. However, 
@@
 	spirit. Nevertheless, one clear advantage of our method over VBVS is the
-	possibility of adaptive inclusion of external information on the features to
-	aid in prediction and feature selection. Bayesian versions of support vector 
+	possibility of adaptive inclusion of external information on the features. 
+	Bayesian versions of support vector 
 	machines have been used in classification problems as well 
-	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
+	\cite[]{chakraborty_bayesian_2011}, but these methods lack 
 	adaptive inclusion of external information on the features.
@@
 	\begin{align}
-	  p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) 
-	  \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = 
+	  p_{\hat{\blambda}}(\bbeta | \y) & =
+	  \frac{\int_{\btau} \L_{\hat{\blambda}}(\y, 
+	  \bbeta, \btau) \, d\btau}{p_{\hat{\blambda}}(\y)} = 
 	  \frac{\int_{\btau} \L (\y ; \bbeta) 
 	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
-	  d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
-	  \hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
-	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, 
-	  \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
-	  & = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta)
-	  \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. 
-	  \label{eq:mmlestimate}
+	  d\btau}{p_{\hat{\blambda}}(\y)} \nonumber \\
+	  \hat{\blambda} & := 
+	  \underset{\blambda}{\argmax} p_{\blambda}(\y) = 
+	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} 
+	  \L_{\blambda}(\y, \bbeta, \btau) \, d\bbeta d\btau. \label{eq:mmlestimate}
 	\end{align}
@@
 	\end{equation}
-	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
-	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
+	Writing $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 
+	\end{bmatrix} = \begin{bmatrix} \bomega & \bbeta & \btau \end{bmatrix}$, 
 	calculus of variations gives the optimal distributions 
@@
 	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
-	distributions in our variational Bayes implementation for the model parameters
-	as:
+	distributions as:
 	\begin{equation}\label{eq:varbayesdistr}
@@
 	\ref{sm-sec:computationalcomplexity} of the SM we show that 
-	informed calculation of the parameters results in a significant reduction of 
+	informed calculation results in a significant reduction of 
 	computational complexity.
@@
 	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
-	estimate. The group lasso extensions are expected to perform better: they 
-	estimate just one penalty parameter, so are less sensitive to overfitting.
-	In Scenario (\ref{it:scenario3}), we expect all methods to underperform
-	due to the high correlations between the features. The
+	estimate. In Scenario (\ref{it:scenario3}), we expect all methods to 
+	underperform due to the high correlations between the features. The
 	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
@@
 	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
-	performance for all 
-	methods, because of the sparse truth and lack of differential 
-	signal.
+	performance for all methods, because of the sparse truth.
 
@@
 	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
-	from $p=900$ to $p=1000$ over the Scenarios. We varied the 
-	covariance matrix of the features and simulation of 
-	the model parameters $\beta_j$, $j=1, \dots, p$ between the
-  Scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
-	parametrisations). We 
+	from $p=900$ to $p=1000$. The covariance matrix of the features and 
+	simulation of the model parameters $\beta_j$, $j=1, \dots, p$ varies between
+  Scenarios. We 
 	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
@@
 	of random variation, we repeated each scenario 100 times 
-	and report predictive performance measures 
-	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
-	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
-	For the methods with an $\alpha$ tuning parameter, we show the one that gives
-	the best performance. Results for the other $\alpha$'s and additional
-	performance measures are presented in Section 
-	\ref{sm-sec:simulations} of the SM.
+	and report average area under the receiver operator curve (AUC) and 
+	Brier skill score in Figures \ref{fig:lines_simulations_res1_auc_briers}-
+	\ref{fig:lines_simulations_res5_auc_briers}. AUC and Brier skill score are 
+	measures of discrimination and calibration, respectively, and for both a 
+	higher value indicates better predictive performance.
+	For the methods with an $\alpha$ parameter, we display the best-performing 
+	$\alpha$. Results for the other $\alpha$'s are presented in 
+	Section \ref{sm-sec:simulations} of the SM.
 
+	Overall, the ridge methods outperform the elastic net methods, which in turn
+	outperform the group lasso methods, in terms of Brier skill score.
 	In Scenario (\ref{it:scenario1})
-	\texttt{gren} and \texttt{GRridge} don't suffer much from the wrong 
-	distributional assumptions as they outperform the other methods.
-	Scenario (\ref{it:scenario2}) follows our expectation \texttt{gren}, and to
+	\texttt{gren} and \texttt{GRridge} outperform the other methods in terms
+	of AUC.
+	Scenario (\ref{it:scenario2}) the AUC follows our expectation: \texttt{gren}, 
+	and to
 	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
-	the regular elastic net outperforms the group lasso methods, which are 
-	designed to handle situations such as Scenario (\ref{it:scenario2}).
+	the regular elastic net outperforms the group lasso methods.
 	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
-	regular elastic net suffer more from the higher correlations. Here,
-	the regular elastic net outperforms \texttt{gren}, which is an indication
-	that the estimation of the penalty parameters is impaired by high
-	correlation between features. Scenario (\ref{it:scenario4}) follows the 
-	expected pattern, with the group-regularized methods (\texttt{GRridge} and
-	\texttt{gren}) outperforming their respective non group-regularized 
+	regular elastic net suffer from the high correlations. The regular elastic net 
+	outperforms \texttt{gren}, which is an indication that high correlations
+	impair penalty parameter estimation. Scenario (\ref{it:scenario4}) follows the 
+	expected pattern, with \texttt{GRridge} and
+	\texttt{gren} outperforming their respective non group-regularized 
 	counterparts, as well as the sparse group lasso extensions.
@@
 	other methods, which is an indication that \texttt{gren} is able to pick up 
-	very sparse differential signals between the groups.
+	the sparse differential signal.
 
-<<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
+<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario1}).", out.width="100%", fig.asp=1/2>>=
 @
 
-<<lines_simulations_res2_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 2.", out.width="100%", fig.asp=1/2>>=
+<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario2}).", out.width="100%", fig.asp=1/2>>=
 @
 
-<<lines_simulations_res3_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 3.", out.width="100%", fig.asp=1/2>>=
+<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario3}).", out.width="100%", fig.asp=1/2>>=
 @
 
-<<lines_simulations_res4_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4.", out.width="100%", fig.asp=1/2>>=
+<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario4}).", out.width="100%", fig.asp=1/2>>=
 @
 
-<<lines_simulations_res5_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5.", out.width="100%", fig.asp=1/2>>=
+<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario5}).", out.width="100%", fig.asp=1/2>>=
 @
@@
   We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
-  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
-  three methods, we only present \texttt{cMCP} here; the results for 
-  another extension of the group lasso, the sparse
-  group lasso by \cite{simon_sparse-group_2013}
-  (\texttt{SGL}), and \texttt{gel} are presented in the SM Section 
+  elastic net, sparse group lasso (\texttt{SGL}) by 
+  \cite{simon_sparse-group_2013}, \texttt{cMCP}, and
+  \texttt{gel}. Of the latter 
+  three methods, we only present the best performing one, \texttt{cMCP} here.
+  The results for \texttt{SGL} and \texttt{gel} are presented in SM Section 
   \ref{sm-sec:colorectal}. For the methods with a
-  tuning parameter $\alpha$, we picked the $\alpha$ 
-  that gave the best performance and refer
-  the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
-  different $\alpha$'s.
+  tuning parameter $\alpha$, we show the best performing $\alpha$ here and refer
+  the reader to SM Section \ref{sm-sec:colorectal} for the remaining $\alpha$'s.
 
@@
   instances, stratified by treatment response. We estimated the models on the 
-  training data and calculated AUC on the test data. We present the AUC for a
+  training data and calculated AUC on the test data. We present AUC for a
   range of model sizes, together with the estimated 
   penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
-  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
-  performance measures are presented in Section \ref{sm-sec:colorectal}
-  of the SM. In addition, we investigated the sensitivity of
-  the multiplier estimation on this dataset in SM Section 
-  \ref{sm-sec:colorectal}. We experienced numerical issues with 
-  \texttt{GRridge} when cross-validating 
-  the global $\lambda$ on the training data. We therefore estimated this
+  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. Brier skill 
+  scores are presented in SM Section \ref{sm-sec:colorectal}. 
+  In addition, we investigated the sensitivity of
+  the multiplier estimation in SM Section 
+  \ref{sm-sec:colorectal}. Cross-validation of the global $\lambda$ in 
+  \texttt{GRridge} resulted in numerical issues. We therefore estimated this
   global parameter on the full data for \texttt{GRridge}.
 
-  The penalty multipliers estimated by 
-  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
+  The estimated penalty multipliers are according to expectation: the highly
   expressed group receives the lowest penalty, followed by the medium expressed
-  group, while the non-expressed group receives the strongest penalty 
-  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
+  group and the non-expressed group. 
   \texttt{gren} seems to outperform the other methods for a range of smaller 
-  model sizes in terms of AUC 
-  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
-  Selection of larger models is difficult with large values of $\alpha$, an 
-  inherent property of the strong $L_1$-norm penalization. With lower values of
-  $\alpha$, larger models are possible (see SM Section \ref{sm-sec:colorectal})
-  for comparisons with smaller values of $\alpha$). The performance of
-  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
-  all group lasso extensions that we investigated.
-  (SM Section \ref{sm-sec:colorectal}). Overall, the 
-  inclusion of the extra information on the 
-  features benefited predictive performance: both \texttt{GRridge} and 
+  model sizes. Selection of larger models is impaired by the large
+  $\alpha$, a property inherent to $L_1$-norm penalization. With a smaller
+  $\alpha$ larger models are possible (see SM Section \ref{sm-sec:colorectal}). 
+  The performance of \texttt{cMCP} is somewhat unstable. This unstable 
+  estimation is an issue in all investigated group lasso extensions. 
+  Overall, the inclusion of the extra information on the 
+  features benefits predictive performance: both \texttt{GRridge} and 
   \texttt{gren} outperform their respective non-group-regularized versions, 
@@
   \section{Application to RNAseq in oral cancer}\label{sec:oral}
-  The aim of our second example study is to predict lymph node metastasis in
+  The aim of this second study is to predict lymph node metastasis in
   oral cancer patients using sequenced RNAs from TCGA 
   \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
-  3096 transformed and normalised TCGA RNASeqv2 expression values of 3096 genes 
+  3096 transformed and normalised TCGA RNASeqv2 gene expression values 
   for 133 HPV-negative oral tumours. Of the corresponding patients, 76 suffered 
@@
   described in \cite{mes_prognostic_2017}, again binned into five groups. 
-  We expected features 
+  We expect features 
   with a large positive Kendall's $\tau$ to be more important in 
@@
   \ref{sec:colorectal}. However, since the feature information
-  consists of two partitions with overlapping groups, we used an extension of
-  \texttt{cMCP} that allows for
-  overlapping groups \cite[]{zeng_overlapping_2016}.
-  We estimated AUC for the methods on an independent validation set of 97 
-  patients \cite[]{mes_prognostic_2017}. The features 
-  in this validation set were
+  consists of two partitions with overlapping groups, we used extensions of
+  \texttt{cMCP} and \texttt{gel} that allow for
+  overlapping groups \cite[]{zeng_overlapping_2016}. In \texttt{SGL} we
+  cross-tabulated the two partitions to create one grouping of the features.
+  We estimated AUC on an independent validation set of 97 
+  patients \cite[]{mes_prognostic_2017}, containing features with
   microarray expression levels, normalised to account
-  for a difference in scale compared to the RNAseq training data. We present the
-  results on this validation set, together with the estimated penalty 
-  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
-  \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}. Additional
-  comparisons and performance measures are presented in the SM Section 
-  \ref{sm-sec:oral}.
+  for a scale difference with the RNAseq training data. We present estimated
+  AUC on this validation set for \texttt{GRridge}, ridge, \texttt{gren}, 
+  elastic net, and the best performing group lasso extension,
+  \texttt{cMCP}, together with the estimated penalty multipliers in Figure
+  \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}. For the methods with
+  an $\alpha$ parameter, we pick the best performing one.
+  Results for \texttt{SGL}, \texttt{gel}, other values of $\alpha$, and the 
+  Brier skill scores are presented in SM Section \ref{sm-sec:oral}.
 
-  The estimated penalty multipliers follow the expected pattern 
-  (Figure \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}a): 
-  larger cis-correlations receive smaller penalties. 
-  Likewise, smaller external $p$-values are penalized less. This pattern is
-  slightly less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the
-  other methods. The predictive performance of
+  The estimated penalty multipliers follow the expected pattern: 
+  larger cis-correlations receive smaller penalties and smaller $p$-values are
+  penalized less. The AUC of
   \texttt{gren} is slighty better than the other methods for a range of model
@@
   \section{Additional applications}\label{sec:applications}
-  Two more applications are presented in SM Sections \ref{sm-sec:alzheimer} and
-  \ref{sm-sec:cervical}. The first one is concerned with diagnosis of 
+  Two additional applications are presented in SM Sections 
+  \ref{sm-sec:alzheimer} and \ref{sm-sec:cervical}. The first one is concerned
+  with diagnosis of 
   Alzheimer's from 230 metabolites' expression levels in 174 subjects. We 
@@
   Additionally, many metabolites are strongly negatively correlated, which
-  further impairs estimation. The group lasso extensions depend on just one
-  penalty parameter, so suffer less from the small number of features and
-  negative correlations.
+  further impairs penalty parameter estimation.
 
-  In the second extra example, we attempt to diagnose 56 women with cervical 
-  lesions using 2576 sequenced microRNAs with the purpose of 
-  selecting relevent markers for cervical cancer screening programmes. We use 
-  include one partitioning of the features to enhance predictive performance and 
+  In the second application, the aim is to diagnose 56 women with cervical 
+  lesions using 2576 sequenced microRNAs. We include a grouping of the features 
+  to enhance predictive performance and 
   feature selection. In this example, \texttt{gren} outperforms the 
@@
 	full Bayes shrinkage of the global penalty parameters, because the latter does
-	not use any known structure to model the variability in the hyperparameters. 
+	not use any known structure to model variability in the hyperparameters. 
 	Nonetheless, an interesting direction of future research is the extension of 
@@
 	We show that to be true for both prognostic and diagnostic targets,
-	as well as a range of marker types (RNAseq, microRNAs, or metabolomics).
+	as well as a range of marker types (RNAseqs, microRNAs, and metabolites).
 	We believe that this is due to the `borrowing of information' effect: 
@@
 	for example, dummies of a categorical variable. Another difference between 
-	\texttt{gren} and the group lasso is the number of penalty parameters. 
-	\texttt{gren} estimates one parameter per group, while the group lasso
-	estimates one overall penalty parameter; it is thereby less flexible in 
-	differential shrinkage of the parameters. In addition, we show in our 
-	simulations and data applications that \texttt{gren} is competitive and often
-	superior to (extensions of) the group lasso.
+	\texttt{gren} and the group lasso is the form of the penalty. 
+	\texttt{gren} estimates one parameter per group. Once these are estimated
+	\texttt{gren} fits a reweigthed elastic net. The group lasso uses a 
+	qualitatively different penalty with one overall penalty parameter; 
+	it is thereby less flexible in differential shrinkage of the parameters. Our 
+	simulations and data applications show that \texttt{gren} is competitive and
+	often superior to (extensions of) the group lasso.
 
@@
 	selection: different data instances lead to different sets of selected
-	features. We investigated the stability of selection on the data from Section
-	\ref{sec:colorectal}. To that end we created several bootstrap samples from 
-	the original data and calculated the sizes of all intersections of selected
-	features. We compared \texttt{gren} to the 
-	regular elastic net and found that the inclusion of extra information
+	features. To investigate the stability of selection on the data from Section
+	\ref{sec:colorectal} we created bootstrap samples from 
+	the original data and calculated the intersections sizes of selected
+	features. Compared to the regular elastic net, the inclusion of extra 
+	information
 	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
-	In addition, we investigated 
-	the bias and variance of the penalty multipliers by examination of the
-	distribution of estimated penalty parameters from 100 random partitionings 
-	of the features
-  (see SM Section \ref{sm-sec:colorectal}). We found that the penalty parameters
-  tend to cluster around one, as expected for an unbiased estimation method, and
-  that the variance of the penalty multipliers is less than for the similar
-  \texttt{GRridge} method.
+	In addition, we investigated the performance of penalty multiplier estimation 
+	by creating 100 random partitionings of the features
+  (see SM Section \ref{sm-sec:colorectal}). We found that the estimated penalty 
+  parameters tend to cluster around one, as expected with a random grouping.
 
-	A possible weak point of the proposed method is the double EM loop. The double
-	loop increases the chance of ending up in a local optimum. In the applications
-	discussed in Sections \ref{sec:colorectal}--
-	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}--
+	A possible weak point of the proposed method is the double EM loop, which has
+	positive probability of ending up in a local optimum. In the applications
+	discussed in Sections \ref{sec:colorectal}-
+	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}-
 	\ref{sm-sec:cervical}, we investigated the occurrence of multiple optima, 
@@
 	values. In our 
-	experience, reasonable starting values are obtained by running a 
-	group-regularized ridge regression and using the estimated penalty multipliers
-	as starting values for \texttt{gren}. Alternatively, a skeptical user may set 
+	experience, reasonable starting values are obtained from an initial 
+	\texttt{GRridge} fit. Alternatively, a skeptical user may set 
 	the starting values for all penalty multipliers to one, thereby starting from
@@
 @
-  The double EM loop runs the danger of excessive computation time. In our 
+  An EM algorithm runs the danger of excessive computation time. In our 
   implementation we have reduced computational time considerably by applying
   some computational tricks (see SM Section 
-  \ref{sm-sec:computationalcomplexity}). In addition, the computationally heavy
-  parts were implemented in \texttt{C++} to further reduce the computation time. 
-  To assess this reduction, we have compared computation times to the
-  other methods introduced in Sections \ref{sec:colorectal}--
+  \ref{sm-sec:computationalcomplexity}) and implementing computationally heavy
+  parts in \texttt{C++}. To assess computation times we compared \texttt{gren}
+  other methods introduced in Sections \ref{sec:simulations}-
   \ref{sec:oral}. All timings were obtained on a Macbook 
@@
   (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}.
-  % We note that a considerable speedup may be achieved by iterating between 
-  % single EB and single VB updates, instead of running the VB EM 
-  % until convergence. This works in practice, because an EM tends to take a few 
-  % initial large steps towards the optimum, followed by a series of smaller 
-  % steps. We argue that the first few iterations of the VB EM render the 
-  % approximate expectation in \ref{eq:approximatedloglikelihood}
-  % accurate enough for a 
-  % reliable EB step.
 

### 2019-04-02 docs/supplement.Rnw
small changes to text of manuscript.Rnw and supplement.Rnw

@@
 
-% vectors and matrices
+% objects
 \newcommand{\bbeta}{\bm{\beta}}
@@
 \newcommand{\s}{\mathbf{s}}
-\renewcommand{\c}{\mathbf{c}}
 \newcommand{\ones}{\mathbf{1}}
@@
 \setlength{\textwidth}{15cm}
-\title{Adaptive group-regularized logistic elastic net regression Supplementary 
-material}
+\title{Supplementary material to 
+`Adaptive group-regularized logistic elastic net regression'}
 \date{\today}
@@
   $$
-  Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, 
+  Let $\mathbf{c} = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, 
   in the following, all expectations and variances are with respect to 
@@
       \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
-      \E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - 
-      \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} 
+      \E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\mathbf{c})] - n \log 2 - 
+      \frac{1}{2} \m \tr \mathbf{c} - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} 
       \sum_{i=1}^n c_i^2 \E [\omega_i] \\
@@
 	    & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) 
-	    \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - 
-	    \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
+	    \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \mathbf{c}^{(t)} - 
+	    \log [1 + \exp(\mathbf{c}^{(t)})] \right\} \label{eq:elboomega4}\\
 	    & \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} 
@@
     \text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + 
-    \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2}
+    \m \tr \{ \frac{1}{2} \mathbf{c}^{(t)} - \log [ \exp(\mathbf{c}^{(t)}) + 
+    1 ]\} +
+    \frac{1}{2}
     \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
@@
 
-  Now we rewrite the calculation of $\c = \begin{bmatrix} c_1 & \cdots & c_n
+  Now we rewrite the calculation of $\mathbf{c} = \begin{bmatrix} c_1 & 
+  \cdots & c_n
   \end{bmatrix} \tr$ as:
   \begin{align*}
-    \c & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + (\x_1 \tr \bmu)^2} & 
+    \mathbf{c} & = \begin{bmatrix} \sqrt{\x_1 \bSigma \x_1 + 
+    (\x_1 \tr \bmu)^2} & 
     \cdots & \sqrt{\x_n \bSigma \x_n + (\x_n \tr \bmu)^2} \end{bmatrix} \tr \\
@@
   \section{Simulations}\label{sec:simulations}
-  The quality of feature selection is assessed by Cohen's kappa 
-  \cite[]{cohen_coefficient_1960}. This kappa is measures similarity between two
-  sets of categorical values. One set contains the feature selection indicators:
-  $s(j) := \mathbbm{1}\{\hat{\beta}_j \neq 0\}$, while the other set contains
-  the true non-zero feature indicators: $t(j) := \mathbbm{1}\{\beta_j \neq 0\}$.
-  We calculate Cohen's $\kappa$ as:
-	\begin{align*}
-	  \kappa & = \frac{f_o - f_e}{1 - f_e}, \text{ where } f_o = p^{-1}
-	  \sum_{j=1}^p \mathbbm{1}\{s(j)=t(j)\} \\ & \text{ and } f_e = p^{-2} 
-	  \left\{ \(\sum_{j=1}^p s(j)\) \cdot \(\sum_{j=1}^p t(j)\) + 
-	  \[\sum_{j=1}^p (1 - s(j)) \] \cdot \[\sum_{j=1}^p (1 - t(j)) \]\right\}.
-	\end{align*}
-	Here, $f_o$ and $f_e$ denote the frequencies of correctly identified and
-	expected correctly identified features, respectively. A positive or negative
-	kappa indicate better and worse feature selection than expected by chance,
-	respectively. In addition, we evaluate estimation accuracy by mean squared
-	error: $\text{MSE} = p^{-1} \sum_{j=1}^p (\beta_j - \hat{\beta}_j)^2$.
-
   \subsection{Scenario 1}
@@
 	\subsection{Additional results}
-	Additional results to MD Section \ref{md-sec:simulations} are given in 
+	Results for different values of the tuning parameter $\alpha$ than presented
+	in MD Section \ref{md-sec:simulations} are given in 
 	Figures \ref{fig:lines2_simulations_res1_auc_briers}-
-	\ref{fig:lines2_simulations_res5_auc_briers}.
+	\ref{fig:lines2_simulations_res5_auc_briers}. Overall, these results follow 
+	the same pattern as the ones presented in MD Section \ref{md-sec:simulations}.
 
@@
 <<heatmap_metabolomics_alzheimer_res1_cor, fig.cap="Correlations between the metabolites.", out.width="50%", fig.asp=1>>=
-library(Biobase)
 library(lattice)
-load("data/ESetMbolCSFPR2.Rdata")
-pheno <- pData(ESetMbolCSFPR2)
-metabol <- t(exprs(ESetMbolCSFPR2))
-metabol.apoe <- metabol[(pheno$D_diag_name=="Probable AD" & 
-                           pheno$APOE=="E4YES") |
-                          (pheno$D_diag_name=="Subjectieve klachten" & 
-                             pheno$APOE=="E4NO"), ]
+load("data/metabolomics_alzheimer_data.Rdata")
 cor.mat <- cor(scale(metabol.apoe))

### 2019-06-28 docs/biostatistics_main.Rnw
commit before second revision

--- 2019-04-02 docs/manuscript.Rnw
@@
 
-\documentclass[a4paper,hidelinks]{article}
-\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
-amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
-titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
-xr-hyper,hyperref,enumerate}
-\externaldocument[sm-]{supplement}
+\documentclass[oupdraft]{bio}
+\usepackage{xr-hyper}
+% \usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, 
+% citecolor=citecolor]{hyperref}
+\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,
+enumerate}
+
+\externaldocument[sm-]{biostatistics_supp}
+
+% Add history information for the article if required
+%\history{Received August 1, 2010;
+%	revised October 1, 2010;
+%	accepted for publication November 1, 2010}
 
 % objects
+\newcommand{\x}{\mathbf{x}}
 \newcommand{\bbeta}{\bm{\beta}}
+\newcommand{\X}{\mathbf{X}}
+\newcommand{\y}{\mathbf{y}}
+\newcommand{\G}{\mathcal{G}}
 \newcommand{\btau}{\bm{\tau}}
-\newcommand{\balpha}{\bm{\alpha}}
+\newcommand{\blambda}{\bm{\lambda}}
 \newcommand{\bomega}{\bm{\omega}}
-\newcommand{\bOmega}{\bm{\Omega}}
-\newcommand{\bSigma}{\bm{\Sigma}}
-\newcommand{\bkappa}{\bm{\kappa}}
 \newcommand{\btheta}{\bm{\theta}}
 \newcommand{\bmu}{\bm{\mu}}
+\newcommand{\bSigma}{\bm{\Sigma}}
 \newcommand{\bpsi}{\bm{\psi}}
-\newcommand{\blambda}{\bm{\lambda}}
+\newcommand{\bOmega}{\bm{\Omega}}
+\newcommand{\Z}{\mathbf{Z}}
 \newcommand{\bLambda}{\bm{\Lambda}}
-\newcommand{\bphi}{\bm{\phi}}
-\newcommand{\Y}{\mathbf{Y}}
-\newcommand{\y}{\mathbf{y}}
-\newcommand{\X}{\mathbf{X}}
-\newcommand{\x}{\mathbf{x}}
-\newcommand{\I}{\mathbf{I}}
-\newcommand{\T}{\mathbf{T}}
-\newcommand{\Z}{\mathbf{Z}}
-\newcommand{\W}{\mathbf{W}}
-\renewcommand{\O}{\mathbf{O}}
-\newcommand{\B}{\mathbf{B}}
-\renewcommand{\H}{\mathbf{H}}
-\newcommand{\U}{\mathbf{U}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\Vm}{\mathbf{V}}
-\newcommand{\rv}{\mathbf{r}}
-\newcommand{\A}{\mathbf{A}}
-\newcommand{\Q}{\mathbf{Q}}
-\newcommand{\Sm}{\mathbf{S}}
-\newcommand{\0}{\bm{0}}
-\newcommand{\tx}{\tilde{\mathbf{x}}}
-\newcommand{\hy}{\hat{y}}
-\newcommand{\tm}{\tilde{m}}
-\newcommand{\tkappa}{\tilde{\kappa}}
 \newcommand{\m}{\mathbf{m}}
-\newcommand{\C}{\mathbf{C}}
-\renewcommand{\v}{\mathbf{v}}
-\newcommand{\g}{\mathbf{g}}
-\newcommand{\s}{\mathbf{s}}
-\newcommand{\ones}{\mathbf{1}}
 
 % functions and operators
+\newcommand{\tr}{^{\text{T}}}
+\newcommand{\expit}{\text{expit}}
+\newcommand{\argmax}{\text{argmax} \,}
 \renewcommand{\L}{\mathcal{L}}
-\newcommand{\G}{\mathcal{G}}
-\renewcommand{\P}{\mathcal{P}}
-\newcommand{\argmax}{\text{argmax} \,}
-\newcommand{\expit}{\text{expit}}
-\newcommand{\erfc}{\text{erfc}}
 \newcommand{\E}{\mathbb{E}}
-\newcommand{\V}{\mathbb{V}}
-\newcommand{\tr}{^{\text{T}}}
 \newcommand{\diag}{\text{diag}}
-\newcommand{\KL}{D_{\text{KL}}}
 \newcommand{\trace}{\text{tr}}
-\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
-
-% distributions
-\newcommand{\N}{\text{N}}
-\newcommand{\TG}{\text{TG}}
-\newcommand{\Bi}{\text{Binom}}
-\newcommand{\PG}{\text{PG}}
-\newcommand{\Mu}{\text{Multi}}
-\newcommand{\GIG}{\text{GIG}}
-\newcommand{\IGauss}{\text{IGauss}}
-\newcommand{\Un}{\text{U}}
 
@@
 \renewcommand{\]}{\right]}
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-
-% settings
-\graphicspath{{../figs/}}
-\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
-\setlength{\textwidth}{15cm}
-\title{Adaptive group-regularized logistic elastic net regression}
-\date{\today}
-\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
-\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
-Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
 
@@
 
+	% Title of paper
+	\title{Adaptive group-regularized logistic elastic net regression}
+
+	% List of authors, with corresponding author marked by asterisk
+	\author{MAGNUS M. M\"UNCH$^{\ast 1,2}$, CAREL F.W. PEETERS$^{1}$, AAD W. VAN
+	DER VAART$^{2}$, \\* MARK A. VAN DE WIEL$^{1,3}$\\[4pt]
+	% Author addresses
+	\textit{1. Department of Epidemiology \& Biostatistics, Amsterdam Public
+	Health research institute, Amsterdam University medical centers, PO Box 7057,
+	1007 MB Amsterdam, The Netherlands \\*
+	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
+	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom}
+	\\[2pt]
+	% E-mail address for correspondence
+	{m.munch@vumc.nl}}
+
+	% Running headers of paper:
+	\markboth%
+	% First field is the short list of authors
+	{M. M. M\"unch and others}
+	% Second field is the short title of the paper
+	{Adaptive group-regularized logistic elastic net regression}
+
 	\maketitle
 
-	\noindent
-	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
-	research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
-	Amsterdam, The Netherlands \\*
-	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
-	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom
+	% Add a footnote for the corresponding author if one has been
+	% identified in the author list
+	\footnotetext{To whom correspondence should be addressed.}
 
@@
 		approximate empirical-variational Bayes framework. Simulations and
-		applications in two different cancer studies show
+		applications in two different molecular marker and cancer type studies show
 		that, if the partitioning of the features is informative, classification 
@@
 		selection are indeed enhanced.}
+		{Empirical Bayes; High-dimensional data; Prediction; Variational Bayes}
 	\end{abstract}
-
-	\noindent\textbf{Keywords}: Empirical Bayes; High-dimensional data; 
-	Prediction; Variational Bayes
-
-	\noindent\textbf{Software available from}: 
-	\url{https://CRAN.R-project.org/package=gren}
 
@@
 	\end{align}
-	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
+	where $\L(\y ; \bbeta)$ denotes the likelihood function of the 
 	observed data $\y = \begin{bmatrix} 
@@
 
-<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario1}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario2}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario3}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario4}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario5}).", out.width="100%", fig.asp=1/2>>=
-@
-
 	\section{Application to microRNAs in colorectal cancer}
@@
   albeit only slightly for \texttt{GRridge}.
-<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
-@
 
@@
   identical, while \texttt{gren} outperforms the regular elastic net.
-<<barplot_lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
-@
 
@@
 
-	\bibliographystyle{author_short3.bst}
+	\bibliographystyle{biorefs}
 	\bibliography{refs}
 
-	\section*{Session info}
+<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario1}).", out.width="100%", fig.asp=1/2>>=
+@
 
-<<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=
-devtools::session_info()
+<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario2}).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario3}).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario4}).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario5}).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<barplot_lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
 @

### 2019-06-28 docs/supplement.Rnw
commit before second revision

@@
 
-<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Estimated AUC for the alzheimer example.", out.width="100%", fig.asp=1>>=
-@
-
-<<lines_metabolomics_alzheimer_res1_briers, fig.cap="Estimated Brier skill score for the alzheimer example.", out.width="100%", fig.asp=1>>=
+<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Estimated AUC for the alzheimer example with with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_metabolomics_alzheimer_res1_briers, fig.cap="Estimated Brier skill score for the alzheimer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
 @

### 2019-06-28 code/figures.R

@@ -1048,10 +1048,10 @@ methods <- c("ridge", "grridge", paste0("gren", c(1:3)), paste0("enet", c(1:3)),
              paste0("sgl", c(1:3)), paste0("cmcp", c(1:3)), 
              paste0("gel", c(1:3)), paste0("ocmcp", c(1:3)), 
              paste0("ogel", c(1:3)))
-col <- bpy.colors(9, cutoff.tail=0.1)
-lty <- 1:length(methods)
 labels <- list("ridge", "GRridge", "gren", "enet", "SGL", "cMCP", 
                "gel", "OcMCP", "Ogel")
+col <- bpy.colors(length(labels), cutoff.tail=0.1)
+lty <- 1:length(labels)
 
 plot.data <- lapply(methods, function(m) {
   aggregate(auc[, substr(colnames(auc), 1, nchar(m))==m], 
@@ -1065,12 +1065,12 @@ layout(matrix(c(rep(c(1, 1, 2, 2), 2), rep(c(0, 3, 3, 0), 2)),
 plot(plot.data[[3]], type="l", xlab="Number of selected features", 
      ylab="AUC", main="(a)", ylim=range(auc), xlim=c(0, 500), 
      col=col[3], lty=lty[3])
-lines(plot.data[[6]], col=col[4], lty=lty[6])
-lines(plot.data[[9]], col=col[5], lty=lty[9])
-lines(plot.data[[12]], col=col[6], lty=lty[12])
-lines(plot.data[[15]], col=col[7], lty=lty[15])
-lines(plot.data[[18]], col=col[8], lty=lty[18])
-lines(plot.data[[21]], col=col[9], lty=lty[21])
+lines(plot.data[[6]], col=col[4], lty=lty[4])
+lines(plot.data[[9]], col=col[5], lty=lty[5])
+lines(plot.data[[12]], col=col[6], lty=lty[6])
+lines(plot.data[[15]], col=col[7], lty=lty[7])
+lines(plot.data[[18]], col=col[8], lty=lty[8])
+lines(plot.data[[21]], col=col[9], lty=lty[9])
 abline(h=plot.data[[1]][, 2], col=col[1], lty=lty[1])
 abline(h=plot.data[[2]][, 2], col=col[2], lty=lty[2])
 legend("bottomright", legend=labels, col=col, 
@@ -1079,24 +1079,24 @@ legend("bottomright", legend=labels, col=col,
[…]

### 2019-07-29 docs/biostatistics_main.Rnw
just before resubmit

@@
 
-\externaldocument[sm-]{biostatistics_supp}
+\externaldocument[sm-]{supplement}
 
@@
 		is often available. Examples of such external information in omics research
-		are: (a) $p$-values from a previous study, (b) a summary of prior 
-		information, and (c) omics annotation. The inclusion of this information in
+		are: (a) $p$-values from a previous study, and (b) omics annotation. The
+		inclusion of this information in
 		the analysis may enhance classification performance and feature selection,
-		but is not straightforward in the standard regression setting. As a solution
-		to this problem, we propose a group-regularized (logistic) elastic net
+		but is not straightforward. We propose a
+		group-regularized (logistic) elastic net
 		regression method, where each penalty parameter corresponds to a group of
@@
 		approximate empirical-variational Bayes framework. Simulations and
-		applications in two different molecular marker and cancer type studies show
-		that, if the partitioning of the features is informative, classification 
-		performance and feature
-		selection are indeed enhanced.}
+		applications to three cancer genomics studies and one Alzheimer metabolomics
+		study show that, if the partitioning of the features is informative, 
+		classification performance and feature selection are indeed enhanced.}
 		{Empirical Bayes; High-dimensional data; Prediction; Variational Bayes}
@@
 	\section{Introduction}\label{sec:introduction}
-	Prediction from high-dimensional data is a challenge common to many fields of
-	research. Examples of such high-dimensional prediction problems arise in
-	computer vision, stock market prediction and disease diagnosis from omics 
-	data. In this paper, we specifically focus on the latter setting. High
-	dimensionality of data introduces several issues in the estimation of 
-	prediction models, especially: (a) unidentifiable models, (b) highly 
-	correlated predictor variables and (c) non-trivial variable selection.
+	In cancer genomics studies one is often faced with relatively small sample 
+	sizes as compared to the number of features. Data pooling may alleviate the 
+	curse of dimensionality, but does not apply to research settings with a unique
+	set up, and cannot integrate other sources of information. However, external
+	information on the features (e.g. genes) is often ubiquitously available in 
+	the public domain. We aim to use this information to improve sparse 
+	prediction. The information may come in as feature groups: e.g. the chromosome
+	on which a gene is located (24 groups), relation to a CpG island for 
+	methylation probes (around 6 groups), or membership of a known gene signature 
+	(2 groups). Alternatively, it may be continuous, such as $p$-values from a an
+	external study. We introduce a method that allows to systematically use
+	multiple sources of such external information to improve high-dimensional
+	prediction.
 
-	Attempts to tackle one or more of these issues have lead to the development of
-	many prediction methods \cite[]{fan_selective_2010}. Here we will focus on the
-	penalization approach. A well-known and popular penalization method is elastic
-	net regression \cite[]{zou_regularization_2005}, with its special cases ridge
-	regression \cite[]{hoerl_ridge_1970} and lasso regression 
-	\cite[]{tibshirani_regression_1996}. The elastic net approach takes issues 
-	(a)-(c) into account and has yielded many successful extensions. Recent work 
-	on elastic net regression (and its special cases) has focussed on increasing
-	prediction accuracy by the inclusion of prior knowledge on the variables
-	\cite[]{lee_variable_2017,van_de_wiel_better_2016,tai_incorporating_2007}.
+  Our methodology is motivated by four recent, small $n$ clinical omics 
+  studies, discussed in detail in Section 
+  \ref{sec:simandapp}. The studies concern 
+  treatment response prediction for colorectal cancer patients, based on 
+  sequenced microRNAs ($n$ = 88); Lymph node metastasis prediction for oral 
+  cancer, using RNAseq data ($n$=133); Cervival cancer diagnostics based on 
+  microRNAs ($n$=56) and Alzheimer diagnosis based on metabolomics ($n$=87).
+  Several
+  sources of information on the features were used including: external 
+  $p$-values, correlation with DNA markers, the conservation status of microRNAs 
+  and node degree in an estimated molecular network.
 
-	In omics research, sources of such prior knowledge on the variables are often
-	available. The information can, for example, come in the form of (a) results 
-	on the same molecular features obtained in a previous study (e.g. $p$-values),
-	(b) information from a publicly available database that summarizes the prior
-	information on the molecular features involved (e.g., the Cancer Gene Census
-	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. a gene's 
-	chromosome location). Although this information can rarely be directly
-	included in the statistical analysis, it may still be useful and informative
-	for the study at hand.
+  As basic prediction model we use the (logistic) elastic net 
+  regression \cite[]{zou_regularization_2005}, which combines the 
+  desirable properties of its special cases ridge
+  \cite[]{hoerl_ridge_1970} and lasso regression 
+  \cite[]{tibshirani_regression_1996}: de-correlation and feature selection. 
+  It has been demonstrated that the prediction accuracy of penalised regression
+  can improve by the inclusion of prior knowledge on the variables. Available
+  methods, however, either handle one source of external information only
+  \cite[]{lee_variable_2017,tai_incorporating_2007}, or do not aim for sparsity
+  \cite[]{van_de_wiel_better_2016}.
 
-	Two such studies are introduced in Sections \ref{sec:colorectal} and
-	\ref{sec:oral}. The first study is concerned with treatment 
-	response prediction for colorectal cancer patients, using sequenced
-	microRNAs. In the second study we predict lymph node metastases in oral cancer
-	patients, using RNAseq data. In both settings additional information on the 
-	features is available in the form of a partitioning. Furthermore, in both
-	studies data is sparse, so any extra included information may help to boost
-	predictive performance. Inclusion of such extra information in 
-	the analysis is not straightforward with standard methods.
+  Like others, we assume the external information to be available as feature 
+  groups; for continuous information like $p$-values we propose a simple, 
+  data-based discretisation (see Section \ref{sec:external}).  At first sight, 
+  such \textit{a priori} grouping of the features suggests the group lasso 
+  \cite[]{meier_group_2008} 
+  or one of its extensions such as the group smoothly clipped absolute 
+  deviations (\texttt{grSCAD}) and the group minimax concave penalty 
+  (\texttt{grMCP}) 
+  \cite[]{huang_selective_2012}. These methods penalise and select 
+  features at the group level. This comes with two limitations: the group lasso 
+  (a) selects entire groups instead of single features and (b) does not
+  penalise adaptively: all groups are penalised
+  equally. Extensions such as the sparse group lasso 
+  \cite[]{simon_sparse-group_2013} partly deal with (a), but do not address (b).
+  Our way to deal with (b) is through differential penalization. That is, each
+  group of features receives its own penalty parameter: the group-regularized 
+  elastic net (\texttt{gren}). An apparent issue with differential 
+  penalization 
+  is the estimation of the penalty parameters. Naive estimation may be done by
+  cross-validation (CV). However, CV requires re-estimation of the 
+  model over a grid, which grows
+  exponentially with the number of penalty parameters. Consequently, it quickly
+  becomes computationally infeasible. We therefore propose an efficient 
+  alternative: empirical-variational Bayes estimation of the penalty parameters,
+  which corresponds to hyperparameter estimation in the Bayesian prior 
+  framework. Because of the ubiquity of binary outcome data in clinical omics
+  research, we focus on the logistic elastic net.
 
-	At first sight, such an \textit{a priori} grouping of the features suggests 
-	the group lasso
-	\cite[]{meier_group_2008, yuan_model_2006}, because
-	the group lasso penalises and selects features at the group level. 
-	There are two limitations to this approach: the group lasso (a) selects entire
-	groups instead of single 
-	features and (b) does not penalise adaptively: all groups are penalised 
-	equally. Extensions such as the sparse group lasso 
-	\cite[]{simon_sparse-group_2013} deal with (a), but do not address (b).
+  Recently, \cite{zhang_novel_2019} introduced a method similar to ours as
+  it also applies variational Bayes (VB) for feature selection in logistic
+  regression. An advantage of our method, however, is the adaptive inclusion of
+  external information on the features to aid in prediction and feature 
+  selection, as \cite{zhang_novel_2019} do not estimate feature- or 
+  group-specific penalty 
+  weights. Bayesian versions of support vector machines have been used in
+  classification problems as well \cite[]{chakraborty_bayesian_2011},
+  but these methods also lack adaptive inclusion of external information on the
+  features. In line with the
+  above, our proposed method is (a) data-driven: the
+  hyperparameters are estimated from the data, (b) adaptive: prior information
+  is automatically
+  weighted with respect to its informativeness, (c) fast compared to full 
+  Bayesian analysis or CV, and (d) easy to use: only the data and
+  grouping of the features are required as input.
 
-	A natural way to deal with (b) is through differential penalization: that is,
-	each group of variables receives its own penalty parameter. An apparent issue
-	with differential penalization is the estimation of the penalty 
-	parameters. Naive estimation may be done by cross-validation. However, 
-	cross-validation requires re-estimation of the model over a grid of penalty
-	parameters. The size of this grid increases exponentially with the number of
-	penalty parameters. Consequently, it quickly becomes computationally 
-	infeasible. We therefore propose an efficient alternative: empirical Bayes
-	estimation of the penalty parameters, which corresponds to hyperparameter
-	estimation in the Bayesian prior framework. Because of the ubiquity of binary
-	outcome data in omics research, we focus on the logistic elastic net.
-
-	In line with the above, our proposed method is (a) objective: the 
-	hyper-parameters are estimated from the data and not chosen on subjective 
-	grounds, (b) adaptive: prior information is automatically and subjectively 
-	weighed with respect to its informativeness, (c) fast compared to full 
-	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
-	user: only the data and grouping of the features is required. We emphasize 
-	that our method is different from recent developments in grouped penalized 
-	regression such as \cite{dondelinger_joint_2018}, where instead of the 
-	features, the observations are partitioned.
-
-	Recently, \cite{zhang_novel_2019} introduced Variational Bayes Variable
-	Selection (VBVS), a method similar to ours in
-	spirit. Nevertheless, one clear advantage of our method over VBVS is the
-	possibility of adaptive inclusion of external information on the features. 
-	Bayesian versions of support vector 
-	machines have been used in classification problems as well 
-	\cite[]{chakraborty_bayesian_2011}, but these methods lack 
-	adaptive inclusion of external information on the features.
-
-	The rest of the paper is structured as follows. We introduce the Bayesian 
-	logistic generalised elastic net model in Section
-	\ref{sec:model}. In Section \ref{sec:estimation} we derive a variational
-	approximation to this model and use this novel approximation in the empirical
-	Bayes estimation of multiple, group-specific penalty parameters. In Section
-	\ref{sec:simulations} we compare the method in a simulation study. In Sections
-	\ref{sec:colorectal} and \ref{sec:oral}, we demonstrate the approach in 
-	two data sets. In Section \ref{sec:discussion} 
-	we conclude by both discussing some differences and
-	parallels between the proposed method, termed \texttt{gren}, and other methods
-	in the literature.
+  The paper is structured as follows. We introduce the model in Section 
+  \ref{sec:model}. In Section 
+  \ref{sec:external} we shortly discuss possible sources of co-data.
+  In Section \ref{sec:estimation} we derive a VB approximation to the
+  model introduced in Section \ref{sec:model}
+  and use this novel approximation in the empirical Bayes (EB) estimation of 
+  multiple, group-specific penalty parameters. In Section \ref{sec:simandapp}
+  we compare the method in a simulation study, demonstrate the benefit of the 
+  approach for two 
+  data sets, and summarise results for two additional data sets. 
+  We conclude with a 
+  discussion.
 
 	\section{Model}\label{sec:model}
-	In logistic regression the outcome variables are assumed to be binary or sums
+	The outcome variables are assumed to be binary or sums
 	of $m_i$ disjoint binary Bernoulli trials ($y_i = \sum_{l=1}^{m_i} k_l, k_l 
@@
 	$\upsilon$, and $\expit \( \x_i\tr \bbeta \) = \exp(\x\tr_i \bbeta)/[1 + 
-	\exp(\x\tr_i \bbeta)]$. Note that if $m_i=1$ for $i=1, \dots, n$, the model
-	reduces to a binary logistic regression model. Throughout the rest of the 
+	\exp(\x\tr_i \bbeta)]$. Throughout the rest of the 
 	paper we assume that the model matrix $\X = \begin{bmatrix} \x_1 & \cdots &
@@
 	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
-	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
-	group-specific weight. In a generalised elastic net regression, the penalised 
-	likelihood is maximised to yield parameter estimates:
-	\begin{align}
-	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
-	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
-	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
-	  \label{eq:weightedlik}\\
-	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
+	$g$ for $g=1, \dots, G$ and let $\lambda'_g \in \mathbb{R}_{>0}$ denote a 
+	group-specific penalty weight. In a generalised elastic net regression, the
+	penalised likelihood is maximised to yield parameter estimates:
+	\begin{equation}\label{eq:weightedlik}
+	  \hat{\bbeta} := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  
+	  \frac{\lambda_1}{2} 
 	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
 	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
-	  \nonumber
-	\end{align}
+	\end{equation}
 	where $\L(\y ; \bbeta)$ denotes the likelihood function of the 
@@
 	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
-	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
-	group-specific weights may be interpreted as penalty multipliers. Note that
+	parameters. From (\ref{eq:weightedlik}) we see that the
+	$\lambda'_g$'s may be interpreted as penalty multipliers. Note that
 	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.
@@
 	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
-	(\lambda'_g)^{|\G(g)|}=1$. The multiplier appears in square root form in the
-	$L_1$-norm term to ensure that penalisation on the parameter level is the same
-	for the $L_1$ and $L_2$-norm terms, as can be seen from the parametrisation in
-	(\ref{eq:weightedlik}). The elastic net combines $L_1$ and $L_2$-norm 
-	penalisation such that the model parameters are shrunken towards zero. 
-	The $L_1$-norm may set some of the estimates exactly to zero, thus 
-	automatically selecting features. The $L_2$-norm ensures that correlated 
-	features behave similarly, as one would generally require.
+	(\lambda'_g)^{|\G(g)|}=1$. The multipliers appear in square root form in the
+	$L_1$-norm term to ensure that penalisation on the parameter level scales
+	with the norm. The $L_1$-norm sets some of the estimates exactly to zero, thus 
+	automatically selecting features. The $L_2$-norm ensures collinearity is 
+	well-handled.
 
@@
 	to the posterior mode of a 
-	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
+	Bayesian elastic model \cite[]{zou_regularization_2005}.
 	\cite{li_bayesian_2010} show that the elastic net prior 
-	(see Supplementary Material (SM) Section \ref{sm-sec:enetprior} for 
-	details on this prior) may be written as
+	(see Supplementary Material (SM) Section 2 for 
+	details) may be written as
 	a computationally more convenient scale mixture of normals, with mixing 
@@
 	\tr$. Using this result, we write the generalised elastic net model in its 
-	Bayesian form as:
+	Bayesian form:
 	\begin{subequations}\label{eq:enmodel2}
@@
 
+  \section{External information sources}\label{sec:external}
+  We describe possible sources of external information in omics
+  studies that may provide the feature groups $\G(g)$. Firstly, there are 
+  biologically motivated partitioning of features that are
+  easily retrieved from online repositories. These are already 
+  discretised and may be included in the analysis as-is. Examples are:
+  (i) pathway memberships of genes,
+  (ii) classes of metabolites (SM Section 12),
+  and (iii) conservation status of microRNAs (SM Section 
+  13).
+  A second type of external information comes in the form of continuous data.
+  Examples are:
+  (iv) $p$-values or false discovery rates (FDRs) from a 
+  different, related study (Section \ref{sec:colorectal}), and
+  (v) quality scores of the features (SM Section 12), 
+  and (vii) the node degrees of a network estimated on the feature 
+  data (SM Section 12).
+
+  \texttt{gren} requires discretised external information, so in the case of 
+  continuous external data, some form of discretisation is required. In some
+  cases discretisation comes naturally. For example, with external data 
+  type (iv) one might consider the `standard' cut-offs 0.05, 0.01,
+  and 0.001. If the choice of cut-offs is not straightforward, we propose a 
+  a data-driven heuristic, which renders a relatively finer discretisation grid 
+  for data dense than for data sparse domains. Then,
+  adaptation to external information is more pronounced in high density areas.
+  We propose to fit a piecewise linear spline to the empirical cumulative
+  distribution function of the external data and automatically select knot
+  locations as cut-offs \cite[]{spiriti_knot_2013}. They also provide a 
+  data-driven method to choose the
+  number of knots, for a given maximum number of knots. The maximum number
+  of knots should be chosen such that each group contains enough features for 
+  stable estimation of the penalty weights. As a rule of thumb, we advise
+  at least 20 features per group.
+
+  In many practical settings, the external information will be incomplete. 
+  We suggest to use a separate
+  group of features with missing external information. We prefer this solution
+  to setting the penalty multipliers for this group to one, because the absence 
+  of external information might be informative from the perspective of 
+  prediction.
+
 	\section{Estimation}\label{sec:estimation}
 	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
-	If the penalty parameters are known, estimation of the frequentist elastic net
-	model parameters or finding the posterior of the generalized elastic net model
+	If the penalty parameters are known, estimation of the elastic net
+	model parameters
 	is feasible with small adjustments of the available algorithms 
-	\cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers
-	\cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is 
+	\cite[]{friedman_regularization_2010,zou_regularization_2005}. Determining 
+	these penalty parameters, however, is 
 	not straightforward.
@@
 	strategies are used: (i) estimate both $\lambda_1$ and $\lambda_2$ by 
-	cross-validation over a two-dimensional grid of values or (ii) re-parametrise 
+	CV over a two-dimensional grid of values \cite[]{waldron_optimized_2011} or 
+	(ii) re-parametrise 
 	the problem in terms of penalty parameters $\alpha= \frac{\lambda_1}{2 
@@
 	proportion of $L_1$-norm penalty $\alpha$ and cross-validate the global 
-	penalty parameter $\lambda$. Strategy (i) is advised by 
-	\cite{waldron_optimized_2011}, while (ii) is proposed by 
-	\cite{friedman_regularization_2010}. In the generalised elastic net setting, 
+	penalty parameter $\lambda$ \cite[]{friedman_regularization_2010}. In the
+	generalised elastic net setting, 
 	strategies (i) and (ii) imply $2 + G$ and $1 + G$ penalty parameters, 
-	respectively. $K$-fold cross validation over $D$ values then results in $K 
+	respectively. $K$-fold CV over $D$ values then results in $K 
 	\cdot D^{2 + G}$ and $K \cdot D^{1 + G}$ models to estimate. Typically $K$ is 
@@
 	$100$, so that even for small $G$, the number of models to estimate is 
-	prohibitively large.
+	very large.
 
@@
 	distributions 
-	\cite[]{alhamzawi_bayesian_2018,nengjun_yi_bayesian_2013,
+	\cite[]{alhamzawi_bayesian_2018,
 	kyung_penalized_2010}, 
-	but the precise parametrisation of this gamma prior is not so obvious. A 
-	second issue is the loss of correspondence between the  Bayesian and 
-	frequentist elastic net. If the ultimate goal is feature selection, this 
+	but the precise parametrisation of this gamma prior is not so obvious. 
+	Secondly, the correspondence between the  Bayesian and 
+	frequentist elastic net is lost. This 
 	correspondence may be exploited through the automatic feature selection 
@@
 	hyperprior choice and allow for feature selection by the frequentist elastic 
-	net, we propose to estimate the penalty parameters by empirical Bayes.
+	net, we propose to estimate the penalty parameters by EB.
 
-	Many forms of empirical Bayes exist, the most formal one being maximisation of
+	The most formal form of EB is maximisation of
 	the marginal likelihood with respect to the hyperparameters. The resulting 
@@
 	prior with respect to the model parameters. In the case of the elastic net 
-	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
-	empirical Bayes posterior for $\bbeta$:
-	\begin{align}
-	  p_{\hat{\blambda}}(\bbeta | \y) & =
-	  \frac{\int_{\btau} \L_{\hat{\blambda}}(\y, 
-	  \bbeta, \btau) \, d\btau}{p_{\hat{\blambda}}(\y)} = 
-	  \frac{\int_{\btau} \L (\y ; \bbeta) 
-	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
-	  d\btau}{p_{\hat{\blambda}}(\y)} \nonumber \\
-	  \hat{\blambda} & := 
+	introduced in (\ref{eq:enmodel2}) EB comes down to finding:
+	\begin{equation}\label{eq:mmlestimate}
+	  \hat{\blambda} := 
 	  \underset{\blambda}{\argmax} p_{\blambda}(\y) = 
 	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} 
-	  \L_{\blambda}(\y, \bbeta, \btau) \, d\bbeta d\btau. \label{eq:mmlestimate}
-	\end{align}
+	  \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, 
+	  d\bbeta d\btau. 
+	\end{equation}
 	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
@@
 	  \blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
-	  \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
+	  \log \[ \L (\y ; \bbeta) \pi_{\blambda}(\bbeta | \btau) 
+	  \pi_{\blambda}(\btau) \] | \blambda^{(k)} \]
 	\end{equation}
@@
 	omics settings. We propose to tackle this problem by approximating the
-	expectation in (\ref{eq:ebEM}) using variational Bayes.
+	expectation in (\ref{eq:ebEM}) using VB.
 
@@
 	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
-	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
+	reconstruction \cite[]{leday_gene_2017}. In VB, the posterior 
 	is approximated by a tractable form and estimated by optimizing a lower bound 
-	on the marginal likelihood of this model (see Section 
-	\ref{sm-sec:objectivefunction} of the SM for the 
+	on the marginal likelihood of this model (see SM Section 
+	3 for the 
 	lower bound of the proposed model). For an extensive introduction and concise 
-	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
-	respectively.
+	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}.
 
-	To simplify the computations of our variational approximation, we follow
+	To simplify the computations of our VB approximation, we follow
 	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
 	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
-	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
-	\ref{sm-sec:polyagamma} of the SM for more 
+	$y_i$ and P\'{o}lya-Gamma distributed (see SM Section 
+	4 for more 
 	details). We augment Model (\ref{eq:enmodel2}) with:
@@
 	\end{equation}
-	Our variational Bayes approximation to the posterior distribution of 
+	Our VB approximation to the posterior distribution of 
 	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
@@
 	\end{bmatrix} = \begin{bmatrix} \bomega & \bbeta & \btau \end{bmatrix}$, 
-	calculus of variations gives the optimal distributions 
+	calculus of variations renders the optimal distributions 
 	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
@@
 	Kullback-Leibler divergence of the posterior to the approximate distribution 
-	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
+	\cite[]{beal_variational_2003}. The approximation in 
+	(\ref{eq:varbayesapproximation})
 	renders both the posterior parameter calculations and the expected joint log 
-	likelihood as introduced in (\ref{eq:ebEM}) tractable.
+	likelihood in (\ref{eq:ebEM}) tractable.
 
@@
 	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
-	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
+	distribution (See SM Section 5 
 	for the derivations). The so-called variational
 	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
-	update them by:
-	\begin{subequations}\label{eq:VBupdateequations}
-		\begin{align}
-		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
-		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
-		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
-		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
-		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
-		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
-		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
-		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
-		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
-		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
-		\end{align}
-	\end{subequations}
-	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
-	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
-	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
-	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
-	element of a column vector, respectively. Naive calculation of the variational
-	parameters is computationally expensive. In Section 
-	\ref{sm-sec:computationalcomplexity} of the SM we show that 
+	update them by (13) in SM Section 
+	6 until convergence. 
+	Naive calculation of the variational
+	parameters is computationally expensive. In SM Section 
+	7 we show that 
 	informed calculation results in a significant reduction of 
@@
 	\subsection{Empirical-variational Bayes}\label{sec:empvarbayes}
-	Variational Bayes was shown to underestimate the posterior variance of the 
+	VB was shown to underestimate the posterior variance of the 
 	parameters, both numerically and theoretically, in several settings 
-	\cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,
+	\cite[]{rue_approximate_2009,
 	wang_inadequacy_2005}. This coincides with our experience that the global 
-	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To 
-	prevent overestimation we use the parametrisation of the elastic net in 
-	\cite{friedman_regularization_2010} as introduced in Section 
+	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated,
+	because they are inversely related to the posterior variances of the 
+	$\beta_j$. To 
+	prevent overestimation we use the parametrisation of
+	\cite{friedman_regularization_2010} as discussed in Section 
 	\ref{sec:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
-	cross-validation of the regular elastic net model, such that the overall 
-	penalisation is determined by cross-validation of only $\lambda$. By combining
-	cross-validation of the global penalty parameter $\lambda$ with empirical 
-	Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} 
-	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more 
-	robust to underestimation of the variational posterior variances. The 
-	remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend
-	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
+	CV of the regular elastic net model, such that the overall 
+	penalisation is determined by CV of only $\lambda$. By combining
+	CV of the global penalty parameter $\lambda$ with EB estimation 
+	of the penalty multipliers $\blambda' = \begin{bmatrix} 
+	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is 
+	robust to underestimation of the VB posterior variances. 
+	For $\alpha$, \cite{hastie_glmnet_2016} recommend
+	to either fix it \textit{a priori}, or compare results for several 
 	choices of $\alpha$. We use the latter.
@@
 	To estimate the penalty multipliers, the intractable posterior expectation
-	in (\ref{eq:ebEM}) is approximated using the variational posterior:
+	in (\ref{eq:ebEM}) is approximated using the VB posterior:
 	\begin{equation}\label{eq:approximatedloglikelihood}
@@
 	where $C$ is constant in $\blambda'$ (see SM Section 
-	\ref{sm-sec:expectedloglikelihood} for the full 
-	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
-	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
-	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
-	of the new penalty multipliers is now given by:
-	\begin{subequations}\label{eq:mmlupdateequation}
-		\begin{align}
-		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
-		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
-		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
-		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
-		\end{align}
-	\end{subequations}
-	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
+	6 for the full 
+	derivation and the $d^{(k)}_g$ terms). An estimate
+	of the new penalty multipliers is now given by 
+	(14) in SM Section 6. 
+	Although the solution to
+	(14) in the SM is not available in
 	closed form, this convex problem is easily solved by a numerical optimisation 
-	routine. The full procedure is summarized in Section 
-	\ref{sm-sec:fullestimationprocedure} of the SM.
+	routine. The full estimation procedure is summarised in Algorithm 
+	\ref{alg:summaryalg}.
+	\begin{algorithm}[h!]
+    \caption{Group-regularized empirical Bayes elastic net}
+    \label{alg:summaryalg}
+	  \begin{algorithmic}[h!]
+		  \Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
+		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
+		  \State{Estimate $\lambda$ by CV of the regular elastic net
+		  model}
+		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
+		  \epsilon_1$} 
+		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
+		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
+		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
+		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
+		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
+		  for $j=1, \dots, p$ using (13) in the SM}
+		  \EndWhile 
+		  \State{Update $\blambda'$ by (14) in the SM}
+		  \EndWhile
+	  \end{algorithmic}
+  \end{algorithm}
 
-	\subsection{Feature selection}
-	Feature selection is often desirable in high-dimensional prediction problems 
-	and omics data problems are no exception. For example, biomarker selection 
-	may lead 
+	\subsection{Feature selection}\label{sec:featureselection}
+	Feature selection is often desirable in high-dimensional prediction problems. 
+	For example, biomarker selection may lead 
 	to a large reduction in costs by supporting targeted assays. Bayesian feature
 	selection is often done by inspection of posterior credible intervals. 
-	Similarly as in frequentist hypothesis testing, we may select a feature if 
-	zero is not in the credible interval. However, the Bayesian lasso's 
+	However, the Bayesian lasso's 
 	(a special case of the elastic net) credible intervals  are known to suffer 
-	from low frequentist coverage in sparse settings 
-	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
-	omics research, so selection by credible intervals is generally inconsistent. 
+	from low frequentist coverage \cite[]{castillo_bayesian_2015}. 
 	We therefore propose to select features in the frequentist paradigm.
 
-	As shortly touched upon in Section \ref{sec:empiricalbayes}, frequentist 
+	Frequentist 
 	feature selection is trivial after estimation of the penalty multipliers. We 
-	simply plug the estimated penalty parameters into some frequentist elastic net
-	algorithm that allows for differential penalization. In our own package 
-	\texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} 
-	\cite[]{friedman_regularization_2010}. In the frequentist elastic net, feature
-	selection is then done automatically. Furthermore, to select a specific number
-	of features, we simply adjust the global $\lambda$ until we select the desired
-	number of features.
+	therefore simply plug the estimated penalty parameters into some frequentist
+	elastic net
+	algorithm that allows for differential penalization. In our package 
+	\texttt{gren}, we involve the \texttt{R}-package \texttt{glmnet} 
+	\cite[]{friedman_regularization_2010}, which automatically selects features.
+	Furthermore, to select a specific number
+	of features, we simply adjust the global $\lambda$ to render the desired
+	number.
 
-	\section{Simulations}\label{sec:simulations}
-	To assess classification and variable selection performance we conducted a 
+  \section{Simulations and applications}\label{sec:simandapp}
+	\subsection{Simulations}\label{sec:simulations}
+	We conducted a 
 	simulation study in which we compared \texttt{gren} to the regular elastic 
@@
 
-	For the regular elastic net, \texttt{gren}, \texttt{cMCP}, 
-	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
-	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
-	such that the first setting closely resembles the ridge setting, 
-	where $\alpha=0$, while the third setting is similar to the lasso with 
-	$\alpha=1$. In \texttt{cMCP} and \texttt{gel}, $\alpha$
-	determines the ratio of group-level penalization ($\alpha=0$) to 
-	feature level ($\alpha=1$) penalization. For the methods that do 
-	automatic feature selection, we estimate models for a range of model sizes:
-	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.
+	We simulated data according to five different scenarios:
+	(i) differential signal between the groups and uniformly distributed model
+	parameters; (ii) a large number of small groups of features; (iii) no 
+	differential signal between the groups, but strong correlations within groups
+	of features; (iv) differential signal between the groups and 
+	heavy-tailed distributed model parameters; and (v) a very sparse setting with 
+	no signal in some of the groups.
 
-	We simulated data according to five different scenarios:
-	\begin{enumerate}[(i)]
-		\item \label{it:scenario1} differential signal between the groups and uniformly
-		distributed model
-		parameters,
-		\item \label{it:scenario2} a large number of small groups of features,
-		\item \label{it:scenario3} no differential signal between the groups, but 
-		strong correlations 
-		within groups of features,
-		\item \label{it:scenario4} differential signal between the groups and 
-		heavy-tailed distributed 
-		model parameters, and
-		\item \label{it:scenario5}a very sparse setting with no signal in some of 
-		the groups.
-	\end{enumerate}
-	Full descriptions of the scenarios are given in Section 
-	\ref{sm-sec:simulations} of 
-	the SM. In Scenario (\ref{it:scenario1}), the distribution of model 
-	parameters is far
-	from the assumed ridge/elastic net/lasso distributions, so we expect
-	all methods to underperform. In Scenario (\ref{it:scenario2}), 
-	which is similar to
-	the simulations presented in the original group lasso paper 
-	\cite[]{yuan_model_2006}, we expect \texttt{gren}
-	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
-	estimate. In Scenario (\ref{it:scenario3}), we expect all methods to 
-	underperform due to the high correlations between the features. The
-	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
-	performance, since
-	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
-	detect differential signal between
-	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
-	performance for all methods, because of the sparse truth.
+	In all scenarios the $y_i$ are sampled from the logistic model introduced in
+	Section \ref{sec:model}, where the $\x_i \tr$ are multivariate Gaussian and
+	$\bbeta$ is scenario-dependent. Area under the receiver operator curve (AUC)
+	and Brier skill score, averaged over 100 repeats, were used to evaluate 
+	performance for models trained on $n=100$ samples and $p \approx 1000$ 
+	features. Full descriptions of the scenarios and 
+	corresponding results are given in SM Section 
+	9. Here, we summarise the results.
 
-	In all scenarios we created a training set of $n=100$ samples to estimate the 
-	models and a test set of $n_{\text{test}}=1000$ samples to compute 
-	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
-	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
-	from $p=900$ to $p=1000$. The covariance matrix of the features and 
-	simulation of the model parameters $\beta_j$, $j=1, \dots, p$ varies between
-  Scenarios. We 
-	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
-	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
-	of random variation, we repeated each scenario 100 times 
-	and report average area under the receiver operator curve (AUC) and 
-	Brier skill score in Figures \ref{fig:lines_simulations_res1_auc_briers}-
-	\ref{fig:lines_simulations_res5_auc_briers}. AUC and Brier skill score are 
-	measures of discrimination and calibration, respectively, and for both a 
-	higher value indicates better predictive performance.
-	For the methods with an $\alpha$ parameter, we display the best-performing 
-	$\alpha$. Results for the other $\alpha$'s are presented in 
-	Section \ref{sm-sec:simulations} of the SM.
-
-	Overall, the ridge methods outperform the elastic net methods, which in turn
-	outperform the group lasso methods, in terms of Brier skill score.
-	In Scenario (\ref{it:scenario1})
+	In terms of Brier skill score, the ridge methods generally outperform the 
+	elastic net methods, which in turn outperform the group lasso methods.
+	In Scenario (i)
 	\texttt{gren} and \texttt{GRridge} outperform the other methods in terms
 	of AUC.
-	Scenario (\ref{it:scenario2}) the AUC follows our expectation: \texttt{gren}, 
+	In scenario (ii) the AUC follows our expectation: \texttt{gren}, 
 	and to
-	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
-	the regular elastic net outperforms the group lasso methods.
-	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
+	a lesser extent \texttt{GRridge} underperform due to overfitting. 
+	The regular elastic net outperforms the group lasso methods.
+	In Scenario (iii), \texttt{gren} and to a lesser extent the
 	regular elastic net suffer from the high correlations. The regular elastic net 
 	outperforms \texttt{gren}, which is an indication that high correlations
-	impair penalty parameter estimation. Scenario (\ref{it:scenario4}) follows the 
+	impair penalty parameter estimation. Scenario (iv) follows the 
 	expected pattern, with \texttt{GRridge} and
@@
 	counterparts, as well as the sparse group lasso extensions.
-	Surprisingly, in Scenario (\ref{it:scenario5}), \texttt{gren} outperforms all
+	In Scenario (v), \texttt{gren} outperforms all
 	other methods, which is an indication that \texttt{gren} is able to pick up 
@@
 
-	\section{Application to microRNAs in colorectal cancer}
+	\subsection{Application to microRNAs in colorectal cancer}
 	\label{sec:colorectal}
   We investigated the performance of \texttt{gren} on data from a 
-  deep sequencing study in microRNAs 
+  microRNA sequencing study
   \cite[]{neerincx_combination_2018}. The aim of the study was to predict 
-  treatment response in 88 treated colorectal cancer patients, coded as either
+  treatment response in 88 colorectal cancer patients, coded as either
   non-progressive/remission (70 patients) or progressive 
   (18 patients). After pre-processing and normalisation, 2114 microRNAs 
-  remained. In addition to the microRNAs, four unpenalized clinical
+  remained. Four unpenalized clinical
   covariates were included in the analysis: prior use of adjuvant therapy 
-  (binary), the type of systemic treatment regimen (ternary), age, and primary 
+  (binary), type of systemic treatment regimen (ternary), age, and primary 
   tumor differentiation (binary).
@@
   In a preliminary experiment on different subjects, the microRNA expression 
-  levels of metastatic colorectal tumour tissue were compared to normal 
-  non-colorectal tissue and primary colorectal tumour tissue was compared to
-  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 127
-  microRNAs that were differentially expressed in both comparisons with
-  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 
-  microRNAs that were not differentially expressed. We expect that incorporation
-  of this partitioning enhances therapy response 
+  levels of primary and metastatic colorectal tumour tissues were compared to 
+  their normal tissue counterparts \cite[]{neerincx_mir_2015}. The two resulting
+  FDRs were combined through the harmonic mean \cite[]{wilson_harmonic_2019} 
+  and discretised using the method
+  described in Section \ref{sec:external}. This yielded four groups of features:
+  (i) $\text{FDR} \leq 0.0001$, (ii) $0.0001 < \text{FDR} \leq 0.0186$,
+  (iii) $0.0186 < \text{FDR}$, and (iv) missing FDR. We expect that 
+  incorporation of this partitioning enhances therapy response 
   classification, because tumor-specific microRNAs are likely to be more 
-  relevant than non-specific microRNAs.
+  relevant than non-specific ones.
 
   We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
+  random forest \cite[]{breiman_random_2001},
   elastic net, sparse group lasso (\texttt{SGL}) by 
@@
   \texttt{gel}. Of the latter 
-  three methods, we only present the best performing one, \texttt{cMCP} here.
+  three methods, we only present the best performing one, \texttt{cMCP}, here.
   The results for \texttt{SGL} and \texttt{gel} are presented in SM Section 
-  \ref{sm-sec:colorectal}. For the methods with a
+  10. For the methods with a
   tuning parameter $\alpha$, we show the best performing $\alpha$ here and refer
-  the reader to SM Section \ref{sm-sec:colorectal} for the remaining $\alpha$'s.
+  the reader to SM Section 10 for the remaining $\alpha$'s.
 
   To estimate performance, we split the data into 61 training and 27 test 
-  instances, stratified by treatment response. We estimated the models on the 
+  samples, stratified by treatment response. We estimated the models on the 
   training data and calculated AUC on the test data. We present AUC for a
@@
   \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. Brier skill 
-  scores are presented in SM Section \ref{sm-sec:colorectal}. 
+  scores are presented in SM Section 10. 
   In addition, we investigated the sensitivity of
   the multiplier estimation in SM Section 
-  \ref{sm-sec:colorectal}. Cross-validation of the global $\lambda$ in 
-  \texttt{GRridge} resulted in numerical issues. We therefore estimated this
-  global parameter on the full data for \texttt{GRridge}.
+  10.
 
-  The estimated penalty multipliers are according to expectation: the highly
-  expressed group receives the lowest penalty, followed by the medium expressed
-  group and the non-expressed group. 
-  \texttt{gren} seems to outperform the other methods for a range of smaller 
-  model sizes. Selection of larger models is impaired by the large
+  The estimated penalty multipliers are according to expectation: the small
+  FDR group receives the lowest penalty, followed by the medium and large
+  FDR groups. The missing FDR group receives the strongest penalty, thereby
+  confirming that absence of 
+  information is informative itself. We observe that
+  \texttt{gren} outperforms the other methods for smaller 
+  models. Selection of larger models is impaired by the large
   $\alpha$, a property inherent to $L_1$-norm penalization. With a smaller
-  $\alpha$ larger models are possible (see SM Section \ref{sm-sec:colorectal}). 
+  $\alpha$ larger models are possible (see SM Section 10). 
   The performance of \texttt{cMCP} is somewhat unstable. This unstable 
@@
   \texttt{gren} outperform their respective non-group-regularized versions, 
-  albeit only slightly for \texttt{GRridge}.
+  albeit only slightly for \texttt{GRridge}. The random forest performs worse 
+  here.
 
-  \section{Application to RNAseq in oral cancer}\label{sec:oral}
+  \subsection{Application to RNAseq in oral cancer}\label{sec:oral}
   The aim of this second study is to predict lymph node metastasis in
@@
   cis-correlations between the RNASeqv2 data and TCGA DNA copy numbers on 
-  the same patients, quantified by Kendall's $\tau$ and binned into five groups.
+  the same patients, quantified by Kendall's $\tau$ and binned into five groups
+  using the rule from Section \ref{sec:external}.
   In addition, we used (b) $p$-values from an independent microarray data set
@@
   cross-tabulated the two partitions to create one grouping of the features.
-  We estimated AUC on an independent validation set of 97 
-  patients \cite[]{mes_prognostic_2017}, containing features with
-  microarray expression levels, normalised to account
-  for a scale difference with the RNAseq training data. We present estimated
-  AUC on this validation set for \texttt{GRridge}, ridge, \texttt{gren}, 
+  We estimated AUC on another independent validation set of 97 
+  patients \cite[]{mes_prognostic_2017}, containing microarray features, 
+  normalised to account
+  for a scale difference with the RNAseq data. We present estimated
+  AUC on this validation set for \texttt{GRridge}, ridge, random forest,
+  \texttt{gren}, 
   elastic net, and the best performing group lasso extension,
   \texttt{cMCP}, together with the estimated penalty multipliers in Figure
-  \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}. For the methods with
+  \ref{fig:barplot_lines_rnaseq_oral_cancer_metastasis_res1_auc}. For the 
+  methods with
   an $\alpha$ parameter, we pick the best performing one.
   Results for \texttt{SGL}, \texttt{gel}, other values of $\alpha$, and the 
-  Brier skill scores are presented in SM Section \ref{sm-sec:oral}.
+  Brier skill scores are presented in SM Section 11.
 
@@
   \texttt{gren} is slighty better than the other methods for a range of model
-  sizes. In this example, ridge and \texttt{GRridge} perform almost 
+  sizes. In this example, ridge, \texttt{GRridge}, and random forest
+  perform almost 
   identical, while \texttt{gren} outperforms the regular elastic net.
 
-  \section{Additional applications}\label{sec:applications}
+  \subsection{Additional applications}\label{sec:applications}
   Two additional applications are presented in SM Sections 
-  \ref{sm-sec:alzheimer} and \ref{sm-sec:cervical}. The first one is concerned
+  12 and 13. The first one is concerned
   with diagnosis of 
@@
   included several sources of extra information on the metabolites. In this 
-  example, \texttt{gren} performs worse than the group lasso methods. This is 
+  example, \texttt{gren} performs worse than the group lasso methods for larger
+  models. This is 
   due to the smaller number of features to learn the penalty parameters from. 
@@
   regular elastic net with respect to predictive performance for a range of 
-  model sizes.
+  model sizes. Random forest is competitive to \texttt{gren}, but requires all
+  features.
 
@@
 	local shrinkage model, as opposed to the global-local shrinkage priors that 
-	\cite{polson_local_2012,bernardo_shrink_2011} discuss. They characterise 
+	\cite{polson_local_2012} discuss. They characterise 
 	certain desirable properties of these global-local shrinkage priors in high 
 	dimensions, which, for example, the horseshoe possesses 
-	\cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}. In our case, global 
+	\cite[]{carvalho_horseshoe_2010}. In our case, global 
 	shrinkage would imply adding another hyperprior for the global $\lambda_1$ 
 	and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters. We argue however,
-	that if the groups are informative, the empirical Bayes estimation of the 
-	(semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than 
+	that if the groups are informative, the EB estimation of the 
+	(semi-) global shrinkage parameters $\lambda'_g$ may be more beneficial than 
 	full Bayes shrinkage of the global penalty parameters, because the latter does
@@
 	considered weakly adaptive, it is different from 
-	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
-	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
-	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
+	the adaptive lasso \cite[]{huang_adaptive_2008,zou_adaptive_2006} and adaptive
+	elastic net 
+	\cite[]{zou_adaptive_2009} because it adapts to external information
 	rather than to the primary data. It also differs in the scale of adaptation: 
@@
 	specific, while in our case they are estimated on the group level, rendering
-	the adaptation more robust. As can be seen from both the simulations in 
-	Section \ref{sec:simulations} and the applications in Sections 
-	\ref{sec:colorectal}-\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}-
-	\ref{sm-sec:cervical} adaptation to 
-	external data may be beneficial for prediction and feature selection. 
-	We show that to be true for both prognostic and diagnostic targets,
-	as well as a range of marker types (RNAseqs, microRNAs, and metabolites).
-	We believe that this is due to the `borrowing of information' effect: 
+	the adaptation more robust. Both the simulations and the applications
+	illustrate that adaptation to 
+	external data may be beneficial for prediction and feature selection for a 
+	range of marker types (RNAseqs, microRNAs, and metabolites) due to the 
+	`borrowing of information' effect: 
 	estimates of feature effects that behave similarly are shrunken similarly, 
-	yielding overall, better estimates.
+	yielding overall, better predictions.
 
 	As touched upon in Section \ref{sec:introduction}, an obvious 
-	comparison is to the group lasso \cite[]{meier_group_2008,
-	yuan_model_2006} and its extensions. Although the group lasso is similar in 
+	comparison is to the group lasso \cite[]{meier_group_2008} and its extensions.
+	Although the group lasso is similar in 
 	the sense that it shrinks on the 
@@
 	its intended application is to small interpretable groups of features, like, 
-	for example, dummies of a categorical variable. Another difference between 
-	\texttt{gren} and the group lasso is the form of the penalty. 
+	for example, gene pathways. Another difference between 
+	\texttt{gren} and the group lasso is the form of the penalty: 
 	\texttt{gren} estimates one parameter per group. Once these are estimated
-	\texttt{gren} fits a reweigthed elastic net. The group lasso uses a 
-	qualitatively different penalty with one overall penalty parameter; 
+	\texttt{gren} fits a reweigthed elastic net. The group lasso uses one overall
+	penalty parameter; 
 	it is thereby less flexible in differential shrinkage of the parameters. Our 
@@
 	often superior to (extensions of) the group lasso.
+
+	A recent development is to group samples
+	rather than features \cite[]{dondelinger_joint_2018} to allow different levels 
+	of sparsity across sample groups. This approach does not incorporate feature
+	information, and uses CV to estimate the penalty parameters. An
+	interesting future line of research would be to combine this approach with 
+	\texttt{gren}.
 
@@
 	\ref{sec:colorectal} we created bootstrap samples from 
-	the original data and calculated the intersections sizes of selected
-	features. Compared to the regular elastic net, the inclusion of extra 
+	the original data and calculated the sizes of the selected feature set 
+	intersections. Compared to the regular elastic net, the inclusion of extra 
 	information
-	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
-	In addition, we investigated the performance of penalty multiplier estimation 
-	by creating 100 random partitionings of the features
-  (see SM Section \ref{sm-sec:colorectal}). We found that the estimated penalty 
-  parameters tend to cluster around one, as expected with a random grouping.
+	increases the selection stability (see SM Section 10). 
+	In addition, we investigated penalty multiplier estimation 
+	on 100 random partitionings of the features
+  (SM Section 10). We found that the estimated penalty 
+  parameters tend to cluster around one, as desired with random groups.
 
-	A possible weak point of the proposed method is the double EM loop, which has
-	positive probability of ending up in a local optimum. In the applications
-	discussed in Sections \ref{sec:colorectal}-
-	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}-
-	\ref{sm-sec:cervical}, we investigated the occurrence of multiple optima, 
-	but never encountered them. This does not guarantee that local optima do not
-	occur, but 
-	it provides some evidence that local optima are not ubiquitous. In any case,
-	bad local optima can often be avoided by an informed choice of starting 
-	values. In our 
-	experience, reasonable starting values are obtained from an initial 
-	\texttt{GRridge} fit. Alternatively, a skeptical user may set 
-	the starting values for all penalty multipliers to one, thereby starting from
-	the regular elastic net and adaptively adjusting the penalty parameters to a
-	group-regularized model. The latter is the default approach in our \texttt{R}
-	package
+	A possible weak point of the proposed method is the double EM loop, which may
+	end up in a local optimum, depending on the starting values of the algorithm.
+	Reasonable starting values, e.g. obtained by applying \texttt{GRridge}, could
+	alleviate this issue.
+	The default of \texttt{gren}, however, is to simply start from the regular
+	elstic net, i.e., penalty multipliers equal to one, and than adapt these. In 
+	the applications, we investigated the occurrence of multiple optima, but never
+	encountered them. This does not guarantee that local optima do not occur, but 
+	it provides some evidence that local optima are not ubiquitous.
 
-<<>>=
-bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench1.csv", 
-                     stringsAsFactors=FALSE)
+	Proper uncertainty quantification in the elastic net is an open problem.
+	If uncertainy quantification is required, Gibbs samples from 
+  the posterior (with the estimated hyperparameters) could be drawn. However,
+  we do not recommend Bayesian uncertainty quantification
+  for the Bayesian elastic net due to bad frequentist properties of lasso-like
+  posteriors \cite[]{castillo_bayesian_2015} in sparse (omics) settings. Hence,
+  selected features should be interpreted with caution, but are nonetheless
+  deemed useful in prediction.
 
-bench2 <- read.table("results/rnaseq_oral_cancer_bench1.csv", 
-                     stringsAsFactors=FALSE)
-times1 <- round(bench1$time*10^(-9), 2)
-times2 <- round(bench2$time*10^(-9), 2)
-session <- gsub("_", ".", devtools::session_info()$platform$system)
-@
   An EM algorithm runs the danger of excessive computation time. In our 
-  implementation we have reduced computational time considerably by applying
-  some computational tricks (see SM Section 
-  \ref{sm-sec:computationalcomplexity}) and implementing computationally heavy
+  implementation we have reduced computational time considerably by
+  some computational shortcuts (see SM Section 
+  7) and implementing some
   parts in \texttt{C++}. To assess computation times we compared \texttt{gren}
   other methods introduced in Sections \ref{sec:simulations}-
-  \ref{sec:oral}. All timings were obtained on a Macbook 
-  Pro 2016 running \Sexpr{session}. 
-  The timings of \texttt{gren} are competitive. On the colorectal cancer example
-  we achieved the following times:
-  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
-  respectively. \texttt{GRridge} and \texttt{SGL} with 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
-  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
-  respectively.
-  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
-  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
-  The results for the oral cancer example are similar. \texttt{gren} 
-  achieved 
-  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
-  respectively.
-  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
-  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
-  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
-  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}.
+  \ref{sec:oral} on a Macbook Pro 2016 running 
+  \Sexpr{gsub("_", ".", devtools::session_info()$platform$system)} and present the
+  results in SM Section 14. In general we found that 
+  \texttt{gren} is similar in times as \texttt{GRridge}, faster than 
+  \texttt{SGL}, and slower than \texttt{cMCP}, and \texttt{gel}.
 
@@
 
-<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario1}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario2}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario3}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario4}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario5}).", out.width="100%", fig.asp=1/2>>=
-@
-
 <<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@@
 
-<<barplot_lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
+<<barplot_lines_rnaseq_oral_cancer_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
 @

### 2019-07-29 docs/manuscript.Rnw
just before resubmit

@@
 \newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\renewcommand{\algorithmicrequire}{\textbf{input:}}
+\renewcommand{\algorithmicensure}{\textbf{output:}}
 
@@
 		approximate empirical-variational Bayes framework. Simulations and
-		applications in two different cancer studies show
-		that, if the partitioning of the features is informative, classification 
-		performance and feature
-		selection are indeed enhanced.}
+		applications to three cancer genomics studies and one Alzheimer metabolomics
+		study show that, if the partitioning of the features is informative, 
+		classification performance and feature selection are indeed enhanced.}
 	\end{abstract}
@@
 	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
-	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
-	group-specific weight. In a generalised elastic net regression, the penalised 
-	likelihood is maximised to yield parameter estimates:
-	\begin{align}
-	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
-	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
-	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
-	  \label{eq:weightedlik}\\
-	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
+	$g$ for $g=1, \dots, G$ and let $\lambda'_g \in \mathbb{R}_{>0}$ denote a 
+	group-specific penalty weight. In a generalised elastic net regression, the
+	penalised likelihood is maximised to yield parameter estimates:
+	\begin{equation}\label{eq:weightedlik}
+	  \hat{\bbeta} := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  
+	  \frac{\lambda_1}{2} 
 	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
 	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
-	  \nonumber
-	\end{align}
-	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
+	\end{equation}
+	where $\L(\y ; \bbeta)$ denotes the likelihood function of the 
 	observed data $\y = \begin{bmatrix} 
@@
 	to the posterior mode of a 
-	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
+	Bayesian elastic model \cite[]{zou_regularization_2005}.
 	\cite{li_bayesian_2010} show that the elastic net prior 
 	(see Supplementary Material (SM) Section \ref{sm-sec:enetprior} for 
-	details on this prior) may be written as
+	details) may be written as
 	a computationally more convenient scale mixture of normals, with mixing 
@@
 
+  \section{External information sources}\label{sec:external}
+  We describe possible sources of external information in omics
+  studies that may provide the feature groups $\G(g)$. Firstly, there are 
+  biologically motivated partitioning of features that are
+  easily retrieved from online repositories. These are already 
+  discretised and may be included in the analysis as-is. Examples are:
+  (i) pathway memberships of genes,
+  (ii) classes of metabolites (see SM Section \ref{sm-sec:alzheimer}),
+  and (iii) conservation status of microRNAs (see SM Section 
+  \ref{sm-sec:cervical}).
+  A second type of external information comes in the form of continuous data.
+  Examples are:
+  (iv) $p$-values or false discovery rates (FDRs) from a 
+  different, related study (see Section \ref{sec:colorectal}), and
+  (v) quality scores of the features (see SM Section \ref{sm-sec:alzheimer}), 
+  and (vii) the node degrees of a network estimated on the feature 
+  data (see SM Section \ref{sm-sec:alzheimer}).
+
+  \texttt{gren} requires discretised external information, so in the case of 
+  continuous external data, some form of discretisation is required. In some
+  cases discretisation comes naturally. For example, with external data 
+  type (iv) one might consider the `standard' cut-offs 0.05, 0.01,
+  and 0.001. If the choice of cut-offs is not straightforward, we propose a 
+  a data-driven heuristic, which renders a relatively finer discretisation grid 
+  for data dense than for data sparse domains. Then,
+  adaptation to external information is more pronounced in high density areas.
+  We propose to fit a piecewise linear spline to the empirical cumulative
+  distribution function of the external data and automatically select knot
+  locations as cut-offs \cite[]{spiriti_knot_2013}. They also provide a 
+  data-driven method to choose the
+  number of knots, for a given maximum number of knots. The maximum number
+  of knots should be chosen such that each group contains enough features for 
+  stable estimation of the penalty weights. As a rule of thumb, we advise
+  at least 20 features per group.
+
+  In many practical settings, the external information will be incomplete. 
+  We suggest to use a separate
+  group of features with missing external information. We prefer this solution
+  to setting the penalty multipliers for this group to one, because the absence 
+  of external information might be informative from the perspective of 
+  prediction.
+
 	\section{Estimation}\label{sec:estimation}
 	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
-	If the penalty parameters are known, estimation of the frequentist elastic net
-	model parameters or finding the posterior of the generalized elastic net model
+	If the penalty parameters are known, estimation of the elastic net
+	model parameters
 	is feasible with small adjustments of the available algorithms 
-	\cite[]{friedman_regularization_2010,zou_regularization_2005} or MCMC samplers
-	\cite[]{li_bayesian_2010}. Determining these penalty parameters, however, is 
+	\cite[]{friedman_regularization_2010,zou_regularization_2005}. Determining 
+	these penalty parameters, however, is 
 	not straightforward.
@@
 	genetic association studies \cite[]{carbonetto_scalable_2012} and gene network
-	reconstruction \cite[]{leday_gene_2017}. In variational Bayes, the posterior 
+	reconstruction \cite[]{leday_gene_2017}. In VB, the posterior 
 	is approximated by a tractable form and estimated by optimizing a lower bound 
-	on the marginal likelihood of this model (see Section 
-	\ref{sm-sec:objectivefunction} of the SM for the 
+	on the marginal likelihood of this model (see SM Section 
+	\ref{sm-sec:objectivefunction} for the 
 	lower bound of the proposed model). For an extensive introduction and concise 
-	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}, 
-	respectively.
+	review, see \cite{beal_variational_2003} and \cite{blei_variational_2017}.
 
-	To simplify the computations of our variational approximation, we follow
+	To simplify the computations of our VB approximation, we follow
 	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
 	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
-	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
-	\ref{sm-sec:polyagamma} of the SM for more 
+	$y_i$ and P\'{o}lya-Gamma distributed (see SM Section 
+	\ref{sm-sec:polyagamma} for more 
 	details). We augment Model (\ref{eq:enmodel2}) with:
@@
 	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
-	update them by:
-	\begin{subequations}\label{eq:VBupdateequations}
-		\begin{align}
-		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
-		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
-		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
-		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
-		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
-		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
-		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
-		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
-		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
-		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
-		\end{align}
-	\end{subequations}
-	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
-	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
-	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
-	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
-	element of a column vector, respectively. Naive calculation of the variational
-	parameters is computationally expensive. In Section 
-	\ref{sm-sec:computationalcomplexity} of the SM we show that 
+	update them by (\ref{sm-eq:VBupdateequations}) in SM Section 
+	\ref{sm-sec:updatingequations} until convergence. 
+	Naive calculation of the variational
+	parameters is computationally expensive. In SM Section 
+	\ref{sm-sec:computationalcomplexity} we show that 
 	informed calculation results in a significant reduction of 
@@
 	where $C$ is constant in $\blambda'$ (see SM Section 
-	\ref{sm-sec:expectedloglikelihood} for the full 
-	derivation). The $d_g^{(k)}$ terms are calculated as: $d^{(k)}_g = \sum_{j 
-	\in \mathcal{G} (g)} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + 
-	\alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. An estimate
-	of the new penalty multipliers is now given by:
-	\begin{subequations}\label{eq:mmlupdateequation}
-		\begin{align}
-		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
-		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
-		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
-		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
-		\end{align}
-	\end{subequations}
-	Although the solution to (\ref{eq:mmlupdateequation}) is not available in
+	\ref{sm-sec:updatingequations} for the full 
+	derivation and the $d^{(k)}_g$ terms). An estimate
+	of the new penalty multipliers is now given by 
+	(\ref{sm-eq:mmlupdateequation}) in SM Section \ref{sm-sec:updatingequations}. 
+	Although the solution to
+	(\ref{sm-eq:mmlupdateequation}) in the SM is not available in
 	closed form, this convex problem is easily solved by a numerical optimisation 
-	routine. The full procedure is summarized in Section 
-	\ref{sm-sec:fullestimationprocedure} of the SM.
+	routine. The full estimation procedure is summarised in Algorithm 
+	\ref{alg:summaryalg}.
+	\begin{algorithm}[h!]
+    \caption{Group-regularized empirical Bayes elastic net}
+    \label{alg:summaryalg}
+	  \begin{algorithmic}[h!]
+		  \Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
+		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
+		  \State{Estimate $\lambda$ by CV of the regular elastic net
+		  model}
+		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
+		  \epsilon_1$} 
+		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
+		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
+		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
+		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
+		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
+		  for $j=1, \dots, p$ using (\ref{sm-eq:VBupdateequations}) in the SM}
+		  \EndWhile 
+		  \State{Update $\blambda'$ by (\ref{sm-eq:mmlupdateequation}) in the SM}
+		  \EndWhile
+	  \end{algorithmic}
+  \end{algorithm}
 
-	\subsection{Feature selection}
-	Feature selection is often desirable in high-dimensional prediction problems 
-	and omics data problems are no exception. For example, biomarker selection 
-	may lead 
+	\subsection{Feature selection}\label{sec:featureselection}
+	Feature selection is often desirable in high-dimensional prediction problems. 
+	For example, biomarker selection may lead 
 	to a large reduction in costs by supporting targeted assays. Bayesian feature
 	selection is often done by inspection of posterior credible intervals. 
-	Similarly as in frequentist hypothesis testing, we may select a feature if 
-	zero is not in the credible interval. However, the Bayesian lasso's 
+	However, the Bayesian lasso's 
 	(a special case of the elastic net) credible intervals  are known to suffer 
-	from low frequentist coverage in sparse settings 
-	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
-	omics research, so selection by credible intervals is generally inconsistent. 
+	from low frequentist coverage \cite[]{castillo_bayesian_2015}. 
 	We therefore propose to select features in the frequentist paradigm.
 
-	As shortly touched upon in Section \ref{sec:empiricalbayes}, frequentist 
+	Frequentist 
 	feature selection is trivial after estimation of the penalty multipliers. We 
-	simply plug the estimated penalty parameters into some frequentist elastic net
-	algorithm that allows for differential penalization. In our own package 
-	\texttt{gren}, we use the \texttt{R}-package \texttt{glmnet} 
-	\cite[]{friedman_regularization_2010}. In the frequentist elastic net, feature
-	selection is then done automatically. Furthermore, to select a specific number
-	of features, we simply adjust the global $\lambda$ until we select the desired
-	number of features.
+	therefore simply plug the estimated penalty parameters into some frequentist
+	elastic net
+	algorithm that allows for differential penalization. In our package 
+	\texttt{gren}, we involve the \texttt{R}-package \texttt{glmnet} 
+	\cite[]{friedman_regularization_2010}, which automatically selects features.
+	Furthermore, to select a specific number
+	of features, we simply adjust the global $\lambda$ to render the desired
+	number.
 
-	\section{Simulations}\label{sec:simulations}
-	To assess classification and variable selection performance we conducted a 
+  \section{Simulations and applications}\label{sec:simandapp}
+	\subsection{Simulations}\label{sec:simulations}
+	We conducted a 
 	simulation study in which we compared \texttt{gren} to the regular elastic 
@@
 
-	For the regular elastic net, \texttt{gren}, \texttt{cMCP}, 
-	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
-	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
-	such that the first setting closely resembles the ridge setting, 
-	where $\alpha=0$, while the third setting is similar to the lasso with 
-	$\alpha=1$. In \texttt{cMCP} and \texttt{gel}, $\alpha$
-	determines the ratio of group-level penalization ($\alpha=0$) to 
-	feature level ($\alpha=1$) penalization. For the methods that do 
-	automatic feature selection, we estimate models for a range of model sizes:
-	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.
+	We simulated data according to five different scenarios:
+	(i) differential signal between the groups and uniformly distributed model
+	parameters; (ii) a large number of small groups of features; (iii) no 
+	differential signal between the groups, but strong correlations within groups
+	of features; (iv) differential signal between the groups and 
+	heavy-tailed distributed model parameters; and (v) a very sparse setting with 
+	no signal in some of the groups.
 
-	We simulated data according to five different scenarios:
-	\begin{enumerate}[(i)]
-		\item \label{it:scenario1} differential signal between the groups and uniformly
-		distributed model
-		parameters,
-		\item \label{it:scenario2} a large number of small groups of features,
-		\item \label{it:scenario3} no differential signal between the groups, but 
-		strong correlations 
-		within groups of features,
-		\item \label{it:scenario4} differential signal between the groups and 
-		heavy-tailed distributed 
-		model parameters, and
-		\item \label{it:scenario5}a very sparse setting with no signal in some of 
-		the groups.
-	\end{enumerate}
-	Full descriptions of the scenarios are given in Section 
-	\ref{sm-sec:simulations} of 
-	the SM. In Scenario (\ref{it:scenario1}), the distribution of model 
-	parameters is far
-	from the assumed ridge/elastic net/lasso distributions, so we expect
-	all methods to underperform. In Scenario (\ref{it:scenario2}), 
-	which is similar to
-	the simulations presented in the original group lasso paper 
-	\cite[]{yuan_model_2006}, we expect \texttt{gren}
-	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
-	estimate. In Scenario (\ref{it:scenario3}), we expect all methods to 
-	underperform due to the high correlations between the features. The
-	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
-	performance, since
-	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
-	detect differential signal between
-	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
-	performance for all methods, because of the sparse truth.
+	In all scenarios the $y_i$ are sampled from the logistic model introduced in
+	Section \ref{sec:model}, where the $\x_i \tr$ are multivariate Gaussian and
+	$\bbeta$ is scenario-dependent. Area under the receiver operator curve (AUC)
+	and Brier skill score, averaged over 100 repeats, were used to evaluate 
+	performance for models trained on $n=100$ samples and $p \approx 1000$ 
+	features. Full descriptions of the scenarios and 
+	corresponding results are given in SM Section 
+	\ref{sm-sec:simulations}. Here, we summarise the results.
 
-	In all scenarios we created a training set of $n=100$ samples to estimate the 
-	models and a test set of $n_{\text{test}}=1000$ samples to compute 
-	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
-	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
-	from $p=900$ to $p=1000$. The covariance matrix of the features and 
-	simulation of the model parameters $\beta_j$, $j=1, \dots, p$ varies between
-  Scenarios. We 
-	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
-	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
-	of random variation, we repeated each scenario 100 times 
-	and report average area under the receiver operator curve (AUC) and 
-	Brier skill score in Figures \ref{fig:lines_simulations_res1_auc_briers}-
-	\ref{fig:lines_simulations_res5_auc_briers}. AUC and Brier skill score are 
-	measures of discrimination and calibration, respectively, and for both a 
-	higher value indicates better predictive performance.
-	For the methods with an $\alpha$ parameter, we display the best-performing 
-	$\alpha$. Results for the other $\alpha$'s are presented in 
-	Section \ref{sm-sec:simulations} of the SM.
-
-	Overall, the ridge methods outperform the elastic net methods, which in turn
-	outperform the group lasso methods, in terms of Brier skill score.
-	In Scenario (\ref{it:scenario1})
+	In terms of Brier skill score, the ridge methods generally outperform the 
+	elastic net methods, which in turn outperform the group lasso methods.
+	In Scenario (i)
 	\texttt{gren} and \texttt{GRridge} outperform the other methods in terms
 	of AUC.
-	Scenario (\ref{it:scenario2}) the AUC follows our expectation: \texttt{gren}, 
+	In scenario (ii) the AUC follows our expectation: \texttt{gren}, 
 	and to
-	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
-	the regular elastic net outperforms the group lasso methods.
-	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
+	a lesser extent \texttt{GRridge} underperform due to overfitting. 
+	The regular elastic net outperforms the group lasso methods.
+	In Scenario (iii), \texttt{gren} and to a lesser extent the
 	regular elastic net suffer from the high correlations. The regular elastic net 
 	outperforms \texttt{gren}, which is an indication that high correlations
-	impair penalty parameter estimation. Scenario (\ref{it:scenario4}) follows the 
+	impair penalty parameter estimation. Scenario (iv) follows the 
 	expected pattern, with \texttt{GRridge} and
@@
 
-<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario1}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario2}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario3}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario4}).", out.width="100%", fig.asp=1/2>>=
-@
-
-<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario5}).", out.width="100%", fig.asp=1/2>>=
-@
-
-	\section{Application to microRNAs in colorectal cancer}
+	\subsection{Application to microRNAs in colorectal cancer}
 	\label{sec:colorectal}
   We investigated the performance of \texttt{gren} on data from a 
-  deep sequencing study in microRNAs 
+  microRNA sequencing study
   \cite[]{neerincx_combination_2018}. The aim of the study was to predict 
-  treatment response in 88 treated colorectal cancer patients, coded as either
+  treatment response in 88 colorectal cancer patients, coded as either
   non-progressive/remission (70 patients) or progressive 
   (18 patients). After pre-processing and normalisation, 2114 microRNAs 
-  remained. In addition to the microRNAs, four unpenalized clinical
+  remained. Four unpenalized clinical
   covariates were included in the analysis: prior use of adjuvant therapy 
-  (binary), the type of systemic treatment regimen (ternary), age, and primary 
+  (binary), type of systemic treatment regimen (ternary), age, and primary 
   tumor differentiation (binary).
@@
   \texttt{gel}. Of the latter 
-  three methods, we only present the best performing one, \texttt{cMCP} here.
+  three methods, we only present the best performing one, \texttt{cMCP}, here.
   The results for \texttt{SGL} and \texttt{gel} are presented in SM Section 
@@
   To estimate performance, we split the data into 61 training and 27 test 
-  instances, stratified by treatment response. We estimated the models on the 
+  samples, stratified by treatment response. We estimated the models on the 
   training data and calculated AUC on the test data. We present AUC for a
@@
   the multiplier estimation in SM Section 
-  \ref{sm-sec:colorectal}. Cross-validation of the global $\lambda$ in 
-  \texttt{GRridge} resulted in numerical issues. We therefore estimated this
-  global parameter on the full data for \texttt{GRridge}.
+  \ref{sm-sec:colorectal}.
 
-  The estimated penalty multipliers are according to expectation: the highly
-  expressed group receives the lowest penalty, followed by the medium expressed
-  group and the non-expressed group. 
-  \texttt{gren} seems to outperform the other methods for a range of smaller 
-  model sizes. Selection of larger models is impaired by the large
+  The estimated penalty multipliers are according to expectation: the small
+  FDR group receives the lowest penalty, followed by the medium and large
+  FDR groups. The missing FDR group receives the strongest penalty, thereby
+  confirming that absence of 
+  information is informative itself. We observe that
+  \texttt{gren} outperforms the other methods for smaller 
+  models. Selection of larger models is impaired by the large
   $\alpha$, a property inherent to $L_1$-norm penalization. With a smaller
@@
   \texttt{gren} outperform their respective non-group-regularized versions, 
-  albeit only slightly for \texttt{GRridge}.
+  albeit only slightly for \texttt{GRridge}. The random forest performs worse 
+  here.
 <<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@@
 
-  \section{Application to RNAseq in oral cancer}\label{sec:oral}
+  \subsection{Application to RNAseq in oral cancer}\label{sec:oral}
   The aim of this second study is to predict lymph node metastasis in
@@
   cross-tabulated the two partitions to create one grouping of the features.
-  We estimated AUC on an independent validation set of 97 
-  patients \cite[]{mes_prognostic_2017}, containing features with
-  microarray expression levels, normalised to account
-  for a scale difference with the RNAseq training data. We present estimated
-  AUC on this validation set for \texttt{GRridge}, ridge, \texttt{gren}, 
+  We estimated AUC on another independent validation set of 97 
+  patients \cite[]{mes_prognostic_2017}, containing microarray features, 
+  normalised to account
+  for a scale difference with the RNAseq data. We present estimated
+  AUC on this validation set for \texttt{GRridge}, ridge, random forest,
+  \texttt{gren}, 
   elastic net, and the best performing group lasso extension,
   \texttt{cMCP}, together with the estimated penalty multipliers in Figure
-  \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}. For the methods with
+  \ref{fig:barplot_lines_rnaseq_oral_cancer_metastasis_res1_auc}. For the 
+  methods with
   an $\alpha$ parameter, we pick the best performing one.
@@
   \texttt{gren} is slighty better than the other methods for a range of model
-  sizes. In this example, ridge and \texttt{GRridge} perform almost 
+  sizes. In this example, ridge, \texttt{GRridge}, and random forest
+  perform almost 
   identical, while \texttt{gren} outperforms the regular elastic net.
-<<barplot_lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
+<<barplot_lines_rnaseq_oral_cancer_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
 @
 
-  \section{Additional applications}\label{sec:applications}
+  \subsection{Additional applications}\label{sec:applications}
   Two additional applications are presented in SM Sections 
@@
 	\ref{sec:colorectal} we created bootstrap samples from 
-	the original data and calculated the intersections sizes of selected
-	features. Compared to the regular elastic net, the inclusion of extra 
+	the original data and calculated the sizes of the selected feature set 
+	intersections. Compared to the regular elastic net, the inclusion of extra 
 	information
-	increased the selection stability (see SM Section \ref{sm-sec:colorectal}). 
-	In addition, we investigated the performance of penalty multiplier estimation 
-	by creating 100 random partitionings of the features
+	increases the selection stability (see SM Section \ref{sm-sec:colorectal}). 
+	In addition, we investigated penalty multiplier estimation 
+	on 100 random partitionings of the features
   (see SM Section \ref{sm-sec:colorectal}). We found that the estimated penalty 
-  parameters tend to cluster around one, as expected with a random grouping.
+  parameters tend to cluster around one, as desired with random groups.
 
-	A possible weak point of the proposed method is the double EM loop, which has
-	positive probability of ending up in a local optimum. In the applications
-	discussed in Sections \ref{sec:colorectal}-
-	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}-
-	\ref{sm-sec:cervical}, we investigated the occurrence of multiple optima, 
-	but never encountered them. This does not guarantee that local optima do not
-	occur, but 
-	it provides some evidence that local optima are not ubiquitous. In any case,
-	bad local optima can often be avoided by an informed choice of starting 
-	values. In our 
-	experience, reasonable starting values are obtained from an initial 
-	\texttt{GRridge} fit. Alternatively, a skeptical user may set 
-	the starting values for all penalty multipliers to one, thereby starting from
-	the regular elastic net and adaptively adjusting the penalty parameters to a
-	group-regularized model. The latter is the default approach in our \texttt{R}
-	package
+	A possible weak point of the proposed method is the double EM loop, which may
+	end up in a local optimum, depending on the starting values of the algorithm.
+	Reasonable starting values, e.g. obtained by applying \texttt{GRridge}, could
+	alleviate this issue.
+	The default of \texttt{gren}, however, is to simply start from the regular
+	elstic net, i.e., penalty multipliers equal to one, and than adapt these. In 
+	the applications, we investigated the occurrence of multiple optima, but never
+	encountered them. This does not guarantee that local optima do not occur, but 
+	it provides some evidence that local optima are not ubiquitous.
 
-<<>>=
-bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench1.csv", 
-                     stringsAsFactors=FALSE)
+	Proper uncertainty quantification in the elastic net is an open problem.
+	If uncertainy quantification is required, Gibbs samples from 
+  the posterior (with the estimated hyperparameters) could be drawn. However,
+  we do not recommend Bayesian uncertainty quantification
+  for the Bayesian elastic net due to bad frequentist properties of lasso-like
+  posteriors \cite[]{castillo_bayesian_2015} in sparse (omics) settings. Hence,
+  selected features should be interpreted with caution, but are nonetheless
+  deemed useful in prediction.
 
-bench2 <- read.table("results/rnaseq_oral_cancer_bench1.csv", 
-                     stringsAsFactors=FALSE)
-times1 <- round(bench1$time*10^(-9), 2)
-times2 <- round(bench2$time*10^(-9), 2)
-session <- gsub("_", ".", devtools::session_info()$platform$system)
-@
   An EM algorithm runs the danger of excessive computation time. In our 
-  implementation we have reduced computational time considerably by applying
-  some computational tricks (see SM Section 
-  \ref{sm-sec:computationalcomplexity}) and implementing computationally heavy
+  implementation we have reduced computational time considerably by
+  some computational shortcuts (see SM Section 
+  \ref{sm-sec:computationalcomplexity}) and implementing some
   parts in \texttt{C++}. To assess computation times we compared \texttt{gren}
   other methods introduced in Sections \ref{sec:simulations}-
-  \ref{sec:oral}. All timings were obtained on a Macbook 
-  Pro 2016 running \Sexpr{session}. 
-  The timings of \texttt{gren} are competitive. On the colorectal cancer example
-  we achieved the following times:
-  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
-  respectively. \texttt{GRridge} and \texttt{SGL} with 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
-  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
-  respectively.
-  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
-  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
-  The results for the oral cancer example are similar. \texttt{gren} 
-  achieved 
-  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
-  respectively.
-  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
-  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
-  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
-  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
-  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}.
+  \ref{sec:oral} on a Macbook Pro 2016 running 
+  \Sexpr{gsub("_", ".", devtools::session_info()$platform$system)} and present the
+  results in SM Section \ref{sm-sec:timings}. In general we found that 
+  \texttt{gren} is similar in times as \texttt{GRridge}, faster than 
+  \texttt{SGL}, and slower than \texttt{cMCP}, and \texttt{gel}.
 

### 2019-07-29 docs/supplement.Rnw
just before resubmit

@@
   derivations of the variational posterior used in the MD are shown in Section
-  \ref{sec:variationaldistributions}. Section
+  \ref{sec:variationaldistributions}. In Section \ref{sec:updatingequations}
+  the variational and empirical Bayes updating equations for the algorithm are
+  given. Section
   \ref{sec:computationalcomplexity} shows how we may reduce the computational
-  complexity of our variational Bayes procedure. Section 
-  \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected
-  joint log likelihood that is maximised to estimate the penalty multipliers as
-  explained in Section \ref{md-sec:estimation} of the MD. The full procedure of
-  the
-  proposed method is shown in Section \ref{sec:fullestimationprocedure}. 
-  Extensions of our method are
+  complexity of our variational Bayes procedure. Extensions of our
+  method are
   presented in Section \ref{sec:extensions}. The simulations in MD Section
@@
   \ref{sec:alzheimer} and to study in microRNAs in cervical cancer data 
-  in Section \ref{sec:cervical}. Finally, 
+  in Section \ref{sec:cervical}. The computational times
+  of the method are explored in Section \ref{sec:timings}. Finally, 
   we introduce the ridge and lasso versions of the proposed method in Section
@@
 
-  Inspection of the updating equations (\ref{md-eq:VBupdateequations}) in the 
-  MD, learns us that the 
+  Inspection of the updating equations (\ref{eq:VBupdateequations}), 
+  learns us that the 
   last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that
@@
 
+  \section{Updating equations}\label{sec:updatingequations}
+  If we fill in the expectations and variances in the variational posterior
+  from Section \ref{sec:variationaldistributions} and maximise with respect
+  to the variational parameters, we arrive at the following updating equations:
+  \begin{subequations}\label{eq:VBupdateequations}
+		\begin{align}
+		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
+		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
+		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
+		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
+		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
+		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
+		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
+		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
+		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
+		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
+		\end{align}
+	\end{subequations}
+	which we iterate until convergence. Here, $\bLambda'$ is a diagonal matrix 
+	with entries 
+	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
+	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
+	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
+	element of a column vector, respectively.
+
+	To obtain the empirical Bayes updating equations, we first derive the 
+	expected log likelihood under the variational posterior as:
+  \begin{align*}
+    \E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ]
+    &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}]
+    \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}
+    (\beta_j | \tau_j) | \blambda'^{(k)}] \\
+    & \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  
+    \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G 
+    \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1}
+    \beta_j^2 | \blambda'^{(k)} \) \\ 
+    & \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - 
+    \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
+  \end{align*}
+  with the coefficients $d_g^{(k)}$ calculated as:
+  \begin{align*}
+    d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1}
+    \beta_j^2 | \blambda'^{(k)} \) \\
+    & = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} 
+    ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} 
+    ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} 
+    ( \beta_j | \blambda'^{(k)} ) \] \\
+    & = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + 
+    (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} 
+    \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
+  \end{align*}
+  Here, $\propto$ denotes proportionality with respect to $\blambda'$. The 
+  resulting updating equations for the hyperparameters are now:
+  \begin{subequations}\label{eq:mmlupdateequation}
+		\begin{align}
+		  \blambda'^{(k+1)} & = \underset{\blambda'}{\argmax} \frac{1}{2} 
+		  \sum_{g=1}^G |\G(g)| \log (\lambda'_g) - \frac{(1 - \alpha) \lambda}{4} 
+		  \sum_{g=1}^G \lambda'_g d^{(k)}_g \\
+		  & \text{subject to } \prod_{g=1}^G (\lambda'_g)^{|\G(g)|} = 1.
+		\end{align}
+	\end{subequations}
+
   \section{Computational complexity}\label{sec:computationalcomplexity}
@@
   of the algorithm, especially with possibly many iterations. However, 
-  inspection of the updating equations in (\ref{md-eq:VBupdateequations})
-  in the MD, reveals
+  inspection of the updating equations in (\ref{eq:VBupdateequations}), reveals
   that we only require the inner products $\x_i \tr \bSigma \x_i$, $\x_i 
@@
   of complexity of $\mathcal{O}(n^2p)$.
-
-  \section{Variational Bayes expected joint log likelihood}
-  \label{sec:expectedloglikelihood}
-  Here, let $\propto$ denote proportionality with respect to $\blambda'$. Then
-  the variational Bayes expected joint log likelihood is derived as follows:
-  \begin{align*}
-    \E_{Q} [ \log \L_{\blambda'}(\y, \bomega, \bbeta, \btau) | \blambda'^{(k)} ]
-    &\propto \E_Q [\log \pi_{\blambda'}(\bbeta | \btau) | \blambda'^{(k)}]
-    \propto \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)} \E_Q [\log \pi_{\blambda'}
-    (\beta_j | \tau_j) | \blambda'^{(k)}] \\
-    & \propto \frac{1}{2} \sum_{g=1}^G \sum_{j \in \mathcal{G}(g)}  
-    \log \lambda'_g - \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G 
-    \sum_{j \in \mathcal{G}(g)} \lambda'_g \E_Q \( \frac{\tau_j}{\tau_j - 1}
-    \beta_j^2 | \blambda'^{(k)} \) \\ 
-    & \propto \frac{1}{2} \sum_{g=1}^G |\G(g)|  \log \lambda'_g - 
-    \frac{(1 - \alpha) \lambda}{4} \sum_{g=1}^G \lambda'_g d_g^{(k)},
-  \end{align*}
-  with the coefficients $d_g^{(k)}$ calculated as:
-  \begin{align*}
-    d_g^{(k)} & = \sum_{j \in \mathcal{G}(g)} \E_Q \( \frac{\tau_j}{\tau_j - 1}
-    \beta_j^2 | \blambda'^{(k)} \) \\
-    & = \sum_{j \in \mathcal{G}(g)} \[1 + \E_{q_{\psi_j}} 
-    ( \psi_j^{-1} | \blambda'^{(k)} ) \]  \[ \E_{q_{\beta_j}} 
-    ( \beta_j | \blambda'^{(k)} )^2 + \V_{q_{\beta_j}} 
-    ( \beta_j | \blambda'^{(k)} ) \] \\
-    & = \sum_{j \in \mathcal{G}(g)} \[\bSigma^{(k)}_{jj} + 
-    (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} 
-    \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\).
-  \end{align*}
-
-  \section{Full estimation procedure}\label{sec:fullestimationprocedure}
-  The expectation and maximisation steps introduced in the Main Document Section
-  \ref{md-sec:estimation} result
-  in a procedure as outlined in Algorithm \ref{alg:summaryalg}:
-  \begin{algorithm}
-    \caption{Group-regularized empirical Bayes elastic net}
-    \label{alg:summaryalg}
-	  \begin{algorithmic}[h]
-		  \Require $\X, \y, \G, \alpha, \epsilon_1, \epsilon_2$
-		  \Ensure $\lambda, \blambda', \bSigma, \bmu$
-		  \State{Estimate $\lambda$ by cross-validation of the regular elastic net
-		  model}
-		  \While{$|\frac{\blambda'^{(k + 1)} - \blambda'^{(k)}}{\blambda'^{(k)}}| >
-		  \epsilon_1$}
-		  \While{$\underset{ij}{\max} \, |\frac{\bSigma_{ij}^{(k+1)} - 
-		  \bSigma_{ij}^{(k)}}{\bSigma_{ij}^{(k)}}| > \epsilon_2$ or 
-		  $\underset{i}{\max} \, |\frac{\bmu_{i}^{(k+1)} - 
-		  \bmu_{i}^{(k)}}{\bmu_{i}^{(k)}}| > \epsilon_2$}
-		  \State{Update $\bSigma$, $\bmu$, $c_i$ for $i=1, \dots, n$ and $\chi_j$ 
-		  for $j=1, \dots, p$ using (\ref{md-eq:VBupdateequations}) in the MD}
-		  \EndWhile
-		  \State{Update $\blambda'$ by (\ref{md-eq:mmlupdateequation}) in the MD}
-		  \EndWhile
-	  \end{algorithmic}
-  \end{algorithm}
 
@@
   \section{Simulations}\label{sec:simulations}
+  We simulated data according to five different scenarios that are explained
+  in more detail below. For the regular elastic net, \texttt{gren},
+  \texttt{cMCP}, 
+	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
+	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
+	such that the first setting closely resembles the ridge setting, 
+	where $\alpha=0$, while the third setting is similar to the lasso with 
+	$\alpha=1$. In \texttt{cMCP} and \texttt{gel}, $\alpha$
+	determines the ratio of group-level penalization ($\alpha=0$) to 
+	feature level ($\alpha=1$) penalization. For the methods that do 
+	automatic feature selection, we estimate models for a range of model sizes:
+	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.
+
+	In all scenarios we created a training set of $n=100$ samples to estimate the 
+	models and a test set of $n_{\text{test}}=1000$ samples to compute 
+	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
+	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
+	from $p=900$ to $p=1000$. The covariance matrix of the features and 
+	simulation of the model parameters $\beta_j$, $j=1, \dots, p$ varies between
+  Scenarios. We 
+	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
+	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
+	of random variation, we repeated each scenario 100 times 
+	and report average area under the receiver operator curve (AUC) and 
+	Brier skill score in Figures \ref{fig:lines_simulations_res1_auc_briers}-
+	\ref{fig:lines_simulations_res5_auc_briers}. AUC and Brier skill score are 
+	measures of discrimination and calibration, respectively, and for both a 
+	higher value indicates better predictive performance.
+	For the methods with an $\alpha$ parameter, we display the best-performing 
+	$\alpha$. Results for the other $\alpha$'s are presented in 
+	Figures \ref{fig:lines2_simulations_res1_auc_briers}-
+	\ref{fig:lines2_simulations_res5_auc_briers}-
+
+<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (i).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (ii).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (iii).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (iv).", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines_simulations_res5_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (v).", out.width="100%", fig.asp=1/2>>=
+@
+
   \subsection{Scenario 1}
@@
   validation set are presented in Figures 
-  \ref{fig:lines_rnaseq_oral_cancer_res1_auc}--
-  \ref{fig:lines_rnaseq_oral_cancer_res1_briers}
-<<lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated AUC for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
-@
-
-<<lines_rnaseq_oral_cancer_res1_briers, fig.cap="Estimated Briers skill score for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
+  \ref{fig:lines_rnaseq_oral_cancer_metastasis_res1_auc}--
+  \ref{fig:lines_rnaseq_oral_cancer_metastasis_res1_briers}
+<<lines_rnaseq_oral_cancer_metastasis_res1_auc, fig.cap="Estimated AUC for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
+@
+
+<<lines_rnaseq_oral_cancer_metastasis_res1_briers, fig.cap="Estimated Briers skill score for the oral cancer example with (a) $\\alpha=0.05$, (b) $\\alpha=0.5$, and (c) $\\alpha=0.95$.", out.width="100%", fig.asp=1>>=
 @
@@
 
+  \section{Computational times}\label{sec:timings}
+<<>>=
+bench1 <- read.table("results/micrornaseq_colorectal_cancer_bench1.csv", 
+                     stringsAsFactors=FALSE)
+
+bench2 <- read.table("results/rnaseq_oral_cancer_metastasis_bench1.csv", 
+                     stringsAsFactors=FALSE)
+times1 <- round(bench1$time*10^(-9), 2)
+times2 <- round(bench2$time*10^(-9), 2)
+@
+  On the colorectal cancer example in MD Section \ref{md-sec:colorectal}
+  we achieved the following times:
+  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
+  respectively. \texttt{GRridge} and \texttt{SGL} with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
+  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
+  respectively.
+  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
+  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
+  The results for the oral cancer example in MD Section \ref{md-sec:oral} 
+  are similar. \texttt{gren} 
+  achieved 
+  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
+  respectively.
+  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
+  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
+  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
+  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
+  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}.
+
   \section{Empirical-variational Bayes for the ridge and lasso}

### 2019-07-29 code/figures.R

@@ -319,25 +319,29 @@ pred <- as.matrix(res[substr(rownames(res), 1, 4)=="pred", ])
 psel <- as.matrix(res[substr(rownames(res), 1, 4)=="psel", ])
 auc <- as.matrix(res[substr(rownames(res), 1, 3)=="auc", ])
 
-methods1 <- c("grridge", expression("gren, "~alpha==0.05, "gren, "~alpha==0.5,
+methods1 <- c("GRridge", expression("gren, "~alpha==0.05, "gren, "~alpha==0.5,
                                     "gren, "~alpha==0.95), "no co-data")
-methods2 <- c("grridge", "ridge", "gren2", "enet1", "cmcp1")
+methods2 <- c("grridge2", "ridge", "rf", "gren5", "enet1", "cmcp1")
 col1 <- grey.colors(length(methods1) - 1, start=0.3, end=0.9, gamma=2.2, 
                     alpha=NULL)
-col2 <- bpy.colors(length(methods2), cutoff.tail=0.1)
+col2 <- bpy.colors(length(methods2) + 1, cutoff.tail=0.1)[
+  -(length(methods2) + 1)]
 lty1 <- 2
 lty2 <- c(1, 2)
-labels1 <- list(diff.expr=expression("FDR">0.05,
-                                     atop(NA, atop(textstyle(0.05>=phantom(0)),
-                                                   textstyle("FDR" > 0.001))),
-                                     "FDR"<=0.001))
-labels2 <- expression("GRridge", "ridge",
+labels1 <- list(diff.expr=expression(atop(NA, atop(textstyle("FDR"), 
+                                                   textstyle(""<=0.0001))),
+                                     atop(NA, atop(textstyle(0.0001<phantom(0)),
+                                                   textstyle("FDR"<=0.0186))),
+                                     atop(NA, atop(textstyle("FDR"), 
+                                                   textstyle("">0.0186))), 
+                                     "no FDR"))
+labels2 <- expression("GRridge", "ridge", "random forest",
                       "gren,"~alpha==0.5, "enet"~alpha==0.05, 
                       "cMCP"~alpha==0.05)
 
-plot.data1 <- lapply(1:length(fit.gren1$lambdag), function(s) {
[…]

### 2019-07-29 code/metabolomics_alzheimer.R

@@ -17,6 +17,9 @@ library(gren)
 library(GRridge)
 library(grpregOverlap)
 library(SGL)
+library(randomForestSRC)
+library(freeknotsplines)
+library(microbenchmark)
 
 ### load data
 load("data/metabolomics_alzheimer_data.Rdata")
@@ -34,9 +37,13 @@ id.train <- c(sample(which(pheno.apoe$D_diag_name=="Probable AD"),
 platformcode <- feat$PlatformCode2
 
 # quality gives groups based on a quality measure of the assayed features
-part.RSDqc <- CreatePartition(feat$RSDqc, ngroup=3)
+# part.RSDqc <- CreatePartition(feat$RSDqc, ngroup=3)
 quality <- rep(c(1:3), times=sapply(part.RSDqc, length))[
   order(unlist(part.RSDqc))]
+
+kn <- knots(ecdf(feat$RSDqc))
+spl <- freelsgen(kn, 1:length(kn), degree=1, numknot=2, seed=2019, stream=0)
+quality2 <- (feat$RSDqc <= spl@optknot[1]) + (feat$RSDqc <= spl@optknot[2]) + 1
 
 # sds gives groups based on the standard deviations of the features
 part.sds.apoe.train <- CreatePartition(apply(metabol.apoe[id.train, ], 1, sd), 
@@ -58,58 +65,77 @@ ytest <- alzheim.apoe[-id.train]
 xtest <- scale(metabol.apoe[-id.train, ])
 p <- ncol(xtrain)
 part <- list(quality=quality, degree=degree)
+part2 <- list(quality=quality2, degree=degree)
 part.gl <- as.numeric(as.factor(paste(quality, degree)))
 part.ogl <- unlist(lapply(list(quality=split(1:p, part$quality), 
                                degree=split(1:p, part$degree)), function(part) {
                                  lapply(part, function(s) {
                                    colnames(xtrain)[s]})}), recursive=FALSE)
 
-fit.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE)
-fit.gren2 <- gren(xtrain, ytrain, partitions=part, alpha=0.5, trace=FALSE)
-fit.gren3 <- gren(xtrain, ytrain, partitions=part, alpha=0.95, trace=FALSE)
+bench <- microbenchmark(
+fit.gren1 <- gren(xtrain, ytrain, partitions=part, alpha=0.05, trace=FALSE),
[…]

### 2019-07-29 code/micrornaseq_cervical_cancer.R

@@ -16,6 +16,8 @@ library(gren)
 library(GRridge)
 library(grpreg)
 library(SGL)
+library(randomForestSRC)
+library(microbenchmark)
 
 ### load data
 load("data/mirsData.RData")
@@ -47,30 +49,39 @@ n <- nrow(x)
 
 csel <- c(seq(1, 5, 1), seq(7, 15, 2), seq(18, 30, 3), seq(34, 50, 4))
 
+bench <- microbenchmark(
 fit.gren1 <- gren(x, y, partitions=list(part=part), alpha=0.05, 
-                  standardize=TRUE, trace=FALSE, psel=csel)
+                  standardize=TRUE, trace=FALSE, psel=csel),
 fit.gren2 <- gren(x, y, partitions=list(part=part), alpha=0.5, 
-                  standardize=TRUE, trace=FALSE, psel=csel)
+                  standardize=TRUE, trace=FALSE, psel=csel),
 fit.gren3 <- gren(x, y, partitions=list(part=part), alpha=0.95, 
-                  standardize=TRUE, trace=FALSE, psel=csel)
+                  standardize=TRUE, trace=FALSE, psel=csel),
 
-fit.grridge <- grridge(t(x), y, list(part=split(1:p, part)))
+fit.grridge <- grridge(t(x), y, list(part=split(1:p, part))),
 
-save(fit.grridge, fit.gren1, fit.gren2, fit.gren3, 
+fit.rf <- rfsrc(y ~ ., data=data.frame(y=y, x=x), 
+                var.used="all.trees", ntree=5000, importance="none"),
+times=1, control=list(order="inorder"))
+
+rownames(bench) <- c(paste0("gren", 1:3), "grridge", "rf")
+write.table(bench, file="results/micrornaseq_cervical_cancer_bench1.csv")
+
+save(fit.grridge, fit.gren1, fit.gren2, fit.gren3, fit.rf,
      file="results/micrornaseq_cervical_cancer_fit1.Rdata")
 
+### cross-validating predictions
 nfolds <- 5
 foldid <- sample(rep(1:nfolds, times=round(c(rep(
   n %/% nfolds + as.numeric((n %% nfolds)!=0), times=n %% nfolds),
   rep(n %/% nfolds, times=nfolds - n %% nfolds)))))
 
-pred <- as.data.frame(matrix(NA, nrow=n, ncol=6*length(csel) + 2))
[…]

### 2019-07-29 code/micrornaseq_colorectal_cancer.R

@@ -16,6 +16,9 @@ library(gren)
 library(GRridge)
 library(grpreg)
 library(SGL)
+library(randomForestSRC)
+library(harmonicmeanp)
+library(freeknotsplines)
 library(foreach)
 library(doParallel)
 library(microbenchmark)
@@ -25,23 +28,48 @@ parallel <- TRUE
 
 ### load data
 load("data/forMagnusN88.Rdata")
+codata <- read.table("data/results_all.txt", header=TRUE)
 
 ### create model matrix for unpenalized covariates
 unpenal <- model.matrix(~ adjth + thscheme + age + pcrcdiff, data=datfr)[, -1]
-
-### create partitioning based on FDR <= 0.05 and FDR <= 0.001
-miRNA.BFDR <- as.character(TumMirs$miRNA)[TumMirs$BFDR_PNminP < 0.001]
-miRNA.TumMirs <- as.character(TumMirs$miRNA)
-miRNA <- as.character(sapply(rownames(mirnormcen_resp), function(s) {
-  strsplit(s, split=" ")[[1]][1]}))
-diff.expr <- miRNA %in% miRNA.BFDR + miRNA %in% miRNA.TumMirs + 1
 
 ### target vector
 benefit <- as.numeric(resp) - 1
 
 ### mirna data
 micrornas <- t(mirnormcen_resp)
-colnames(micrornas) <- miRNA
+colnames(micrornas) <- sapply(strsplit(colnames(micrornas), " "), "[[", 1)
+
+
+### create partitioning based on FDR <= 0.05 and FDR <= 0.001
+# miRNA.BFDR <- as.character(TumMirs$miRNA)[TumMirs$BFDR_PNminP < 0.001]
+# miRNA.TumMirs <- as.character(TumMirs$miRNA)
+# miRNA <- as.character(sapply(rownames(mirnormcen_resp), function(s) {
+#   strsplit(s, split=" ")[[1]][1]}))
+# diff.expr <- miRNA %in% miRNA.BFDR + miRNA %in% miRNA.TumMirs + 1
+# 
+# sum(codata$BFDR_MNminM <= 0.001 & codata$BFDR_PNminP <= 0.001 & 
+#       codata$miRNA %in% colnames(micrornas))
+# 
+# sum(codata$BFDR_MNminM > 0.001 & codata$BFDR_MNminM <= 0.05 &
[…]

### 2019-07-29 code/rnaseq_oral_cancer_metastasis.R

@@ -17,6 +17,8 @@ library(gren)
 library(GRridge)
 library(grpregOverlap)
 library(SGL)
+library(randomForestSRC)
+library(freeknotsplines)
 library(microbenchmark)
 
 ### load data
@@ -42,6 +44,16 @@ pvpart0 <- grridge2gren(CreatePartition(CDT$pvalsVUmc, ngroup=5,
 corpart0 <- grridge2gren(CreatePartition(CDT$Corrs, ngroup=5, decreasing=TRUE))
 partLNM <- list(corr=corpart0, pv=pvpart0)
 
+kn1 <- knots(ecdf(CDT$Corrs))
+kn2 <- knots(ecdf(CDT$pvalsVUmc))
+spl1 <- freelsgen(kn1, 1:length(kn1), degree=1, numknot=4, seed=2019, stream=0)
+spl2 <- freelsgen(kn2, 1:length(kn2), degree=1, numknot=4, seed=2019, stream=0)
+
+corr2 <- (CDT$Corrs <= spl1@optknot[1]) + (CDT$Corrs <= spl1@optknot[2]) +
+  (CDT$Corrs <= spl1@optknot[3]) + (CDT$Corrs <= spl1@optknot[4]) + 1
+pv2 <- (CDT$pvalsVUmc <= spl2@optknot[1]) + (CDT$pvalsVUmc <= spl2@optknot[2]) +
+  (CDT$pvalsVUmc <= spl2@optknot[3]) + (CDT$pvalsVUmc <= spl2@optknot[4]) + 1
+
 ################################### model 1 ####################################
 ### fitting the models
 set.seed(2019)
@@ -50,6 +62,7 @@ xtrain <- Tr
 ytest <- as.numeric(levels(RespValidation))[RespValidation]
 xtest <- ValidationData[, selectgenes]
 part <- partLNM
+part2 <- list(corr=corr2, pv=pv2)
 part.ogl <- unlist(lapply(list(corr=split(1:p, part$corr), 
                                pv=split(1:p, part$pv)), function(part) {
                                  lapply(part, function(s) {
@@ -105,17 +118,28 @@ fit.ogel3 <- grpregOverlap(xtrain, ytrain, part.ogl, penalty="gel",
                             family="binomial", alpha=0.95)
 fit.ogel3 <- list(cve=NA, cvse=NA, lambda=fit.ogel3$lambda, fit=fit.ogel3, 
                    min=1, lambda.min=fit.ogel3$lambda, null.dev=NA, pe=NA)
[…]

### 2019-11-29 docs/biostatistics_main.Rnw
added extra simulations

@@
 opts_chunk$set(fig.align='center', fig.path="../figs/", 
-               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
+               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht',
+               dev="postscript")
 @
@@
 \documentclass[oupdraft]{bio}
-\usepackage{xr-hyper}
 % \usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, 
@@
 
-\externaldocument[sm-]{supplement}
+% \externaldocument[sm-]{supplement}
 
@@
   (i) pathway memberships of genes,
-  (ii) classes of metabolites (SM Section 12),
-  and (iii) conservation status of microRNAs (SM Section 
+  (ii) classes of metabolites (see SM Section 12),
+  and (iii) conservation status of microRNAs (see SM Section 
   13).
@@
   (iv) $p$-values or false discovery rates (FDRs) from a 
-  different, related study (Section \ref{sec:colorectal}), and
-  (v) quality scores of the features (SM Section 12), 
+  different, related study (see Section \ref{sec:colorectal}), and
+  (v) quality scores of the features (see SM Section 12), 
   and (vii) the node degrees of a network estimated on the feature 
-  data (SM Section 12).
+  data (see SM Section 12).
 
@@
   a data-driven heuristic, which renders a relatively finer discretisation grid 
-  for data dense than for data sparse domains. Then,
+  for data-dense than for data-sparse domains. Then,
   adaptation to external information is more pronounced in high density areas.

### 2019-11-29 docs/manuscript.Rnw
added extra simulations

@@
   data sets, and summarise results for two additional data sets. 
-  We conclude with a 
-  discussion.
+  We conclude with a discussion of some of the benefits and drawbacks of the
+  proposed \texttt{gren}.
 
@@
 	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
-	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
+	$\lambda_1 > 0$ and $\lambda_2 > 0$ are the `global' penalty 
 	parameters. From (\ref{eq:weightedlik}) we see that the
@@
 	we calibrate the $\lambda'_g$ such that $\prod_{g=1}^G 
-	(\lambda'_g)^{|\G(g)|}=1$. The multipliers appear in square root form in the
+	(\lambda'_g)^{|\G(g)|}=1$, with $|\G(g)|$ the number of features in group $g$. 
+	The multipliers appear in square root form in the
 	$L_1$-norm term to ensure that penalisation on the parameter level scales
@@
 	automatically selecting features. The $L_2$-norm ensures collinearity is 
-	well-handled.
+	well-handled. Addition of the penalty terms also prevents quasi-complete
+	separation in logistic regression, a common phenomenon in small $n$ studies.
 
@@
 
+	In addition to simulations from the correct model, i.e., model
+
 	\subsection{Application to microRNAs in colorectal cancer}

### 2019-11-29 docs/supplement.Rnw
added extra simulations

@@
 	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
-	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
+	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. 
+	To mitigate the influence
 	of random variation, we repeated each scenario 100 times 
@@
 <<lines2_simulations_res5_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5.", out.width="100%", fig.asp=1/2>>=
+@
+
+  \subsection{Model misspecification}
+  To assess the effects of model misspecification we conducted a series of 
+  simulations under different modes of misspecification and simulation settings.
+  We investigate model misspecification by comparing \texttt{gren} to the
+  regular elastic net.
+  To investigate the effect of misspecifying the link function, we simulated
+  according to Scenarios 4 and 5, where we changed the response simulation:
+	$y_i \sim \mathcal{B}(1, [\exp (\x_i \tr \bbeta) + \kappa]/
+	[\exp (\x_i \tr \bbeta) + 1 + \kappa])$ as in \cite{komori_asymmetric_2016}, 
+	where the link function is unbalanced for $\kappa \neq 0$. We use 
+	$\kappa \in \{ 0.1, 0.5, 1 \}$, which leads to 6 different simulations 
+	from Scenarios 4 and 5 in total. We present the results in Figures
+	\ref{fig:lines2_simulations_res6_auc_briers}-
+	\ref{fig:lines2_simulations_res11_auc_briers}. From the Figures we see that 
+	\texttt{gren} outperforms the regular elastic net even with a misspecified 
+	link function in terms of AUC. For the Brier skill score there seems to be 
+	some detrimental effect of the misspecification.
+
+<<lines2_simulations_res6_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4 for a misspecified link function with $\\kappa=0.1$.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res7_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4 for a misspecified link function with $\\kappa=0.5$.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res8_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4 for a misspecified link function with $\\kappa=1$.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res9_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5 for a misspecified link function with $\\kappa=0.1$.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res10_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5 for a misspecified link function with $\\kappa=0.5$.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res11_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 5 for a misspecified link function with $\\kappa=1$.", out.width="100%", fig.asp=1/2>>=
+@
+
+  In addition to misspecification of the response relation, model 
+  misspecification may also occur in the feature relation. To investigate this
+  we simulated data that includes non-linear relations between linear predictor
+  and features. To do so, we considered the following link function for part of 
+  the features: $\expit [\max(0,x_{ij})\beta_j]$. We apply this non-linear 
+  relation in simulation Scenarios 4 and 5, where we consider three fractions
+  of features for which we apply the non-linear relation: 0.2, 0.4, and 0.6, 
+  evenly divided over the groups of features. 
+  Meaning that 0.8, 0.6, and 0.4 of the features are linearly related to the
+  link function, respectively. This leads again to 6 different simulation 
+  settings: three in Scenario 4 and three in Scenario 5. We present the results
+  in Figures \ref{fig:lines2_simulations_res12_auc_briers}-
+  \ref{fig:lines2_simulations_res14_auc_briers}.
+  From the Figures we see that \texttt{gren} still outperforms the regular 
+  elastic net, albeit by a slightly smaller margin then with a correctly
+  specified feature relation.
+
+<<lines2_simulations_res12_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4 for a non-linear feature relation with the fraction of non-linear relations 0.2.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res13_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4 for a non-linear feature relation with the fraction of non-linear relations 0.4.", out.width="100%", fig.asp=1/2>>=
+@
+
+<<lines2_simulations_res14_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 4 for a non-linear feature relation with the fraction of non-linear relations 0.6.", out.width="100%", fig.asp=1/2>>=
 @

### 2019-11-29 code/figures.R

@@ -835,6 +835,679 @@ abline(h=plot.data[[1]][, 3], col=col[1], lty=lty[1])
 abline(h=plot.data[[2]][, 3], col=col[2], lty=lty[1])
 par(opar)
 
+# ---- lines2_simulations_res6_auc_briers ----
+library(sp)
+load(file="results/simulations_res6.Rdata")
+plot.data <- 
+  sapply(c(paste0("gren", 1:3), paste0("enet", 1:3)), function(m) {
+             cbind(rowMeans(sapply(res, function(s) {
+               vec <- s$psel[substr(names(s$psel), 1, nchar(m))==m]
+               replace(rep(NA, 8), 1:length(vec), vec)}), 
+               na.rm=TRUE),
+               rowMeans(sapply(res, function(s) {
+                 vec <- s$auc[substr(names(s$auc), 1, nchar(m))==m]
+                 replace(rep(NA, 8), 1:length(vec), vec)}), 
+                 na.rm=TRUE),
+               rowMeans(sapply(res, function(s) {
+                 vec <- s$briers[substr(names(s$briers), 1, nchar(m))==m]
+                 replace(rep(NA, 8), 1:length(vec), vec)}), 
+                 na.rm=TRUE))}, simplify=FALSE)
+
+methods <- c("gren", "enet")
+labels <- c(methods, expression(alpha==0.05), expression(alpha==0.5), 
+            expression(alpha==0.95))
+
+col <- bpy.colors(length(methods), cutoff.tail=0.2)
+lty <- c(1:3)
+
+xlim <- range(sapply(plot.data, function(s) {s[, 1]}), na.rm=TRUE)
+ylim1 <- range(sapply(plot.data, function(s) {s[, 2]}), na.rm=TRUE)
+ylim2 <- range(sapply(plot.data, function(s) {s[, 3]}), 
+               na.rm=TRUE)
+
+opar <- par(no.readonly=TRUE)
+par(mar=opar$mar*c(1, 1.3, 1, 1))
+layout(matrix(rep(c(1, 1, 2, 2), 2), 2, 4, byrow=TRUE))
+plot(plot.data[[1]][, c(1, 2)], type="l", main="(a)", ylab="AUC", 
[…]

### 2019-11-29 code/functions.R

@@ -125,30 +125,6 @@ mcmc.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL,
   
 }
 
-n <- 100
-p <- 20
-x <- matrix(rnorm(n*p), ncol=p, nrow=n)
-beta <- -9.5:9.5
-y <- rbinom(n, 1, as.numeric(1/(1 + exp(-x %*% beta))))
-m <- rep(1, n)
-alpha <- 0.5
-lambda <- 1
-lambdamult <- rep(1, p)
-
-
-fit.vb <- vb.gren(x=x, y=y, m=m, alpha=alpha, lambda=lambda, 
-                  lambdamult=lambdamult, intercept=TRUE, 
-                  control=list(maxit=100, epsilon=0.001))
-fit.mcmc <- mcmc.gren(x=x, y=y, m=m, alpha=alpha, lambda=lambda, 
-                      lambdamult=lambdamult, intercept=TRUE, 
-                      control=list(nsamples=10000))
-test1 <- diag(fit.vb$posterior$sigma)[-1]
-test2 <- fit.vb$posterior$LRdsigma
-test3 <- apply(fit.mcmc$samples$beta, 1, var)[-1]
-plot(test1, test3)
-plot(test2, test3)
-plot(test1, test2)
-
 vb.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL, 
                     lambdamult, intercept=TRUE, 
                     control=list(maxit=100, epsilon=0.001)) {

### 2019-11-29 code/simulations.R

@@ -27,574 +27,1844 @@ library(mvtnorm)
 ### loading functions
 source("code/functions.R")
 
-################################# simulation 1 #################################
-n <- 100
-ntest <- 1000
-p <- 1000
-G <- 5
-rho <- 0.5
-rg <- 50
-beta.mean <- 0.07
-f <- 1.6
-
-Sigma <- diag(p)
-for(i in 1:p) {
-  for(j in 1:p) {
-    Sigma[i, j] <- rho^abs(i - j)
-  }
-}
-b1 <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
-beta <- numeric(p)
-
-part <- rep(c(1:G), each=p/G)
-csel <- 2^c(1:8)
-nreps <- 100
-
-### analysis splits in parallel
-ncores <- min(detectCores() - 1, nreps)
-cluster <- makeForkCluster(ncores)
-if(parallel) {
-  registerDoParallel(cluster)
-} else {
-  registerDoSEQ()
-}
-
-res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
-  set.seed(2019 + k)
-  
-  for(g in 1:G) {
-    idz <- c(((g-1)*p/G + 1):(g*p/G - rg))
-    ida <- c(((g*p/G - rg) + 1):(g*p/G))
-    beta[idz] <- 0
-    beta[ida] <- runif(rg, 0, b1*f^(g-1))
-  }
-  xtrain <- rmvnorm(n, rep(0, p), Sigma)
-  ytrain <- rbinom(n, 1, as.numeric(1/(1 + exp(-xtrain %*% beta))))
-  
-  xtest <- rmvnorm(ntest, rep(0, p), Sigma)
-  ytest <- rbinom(ntest, 1, as.numeric(1/(1 + exp(-xtest %*% beta))))
-  
-  # fitting models
-  fit.gren1 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.05, 
-                    standardize=TRUE, trace=FALSE, psel=csel)
-  fit.gren2 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.5, 
-                    standardize=TRUE, trace=FALSE, psel=csel)
-  fit.gren3 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.95, 
-                    standardize=TRUE, trace=FALSE, psel=csel)
-  
[…]

### 2019-11-29 rpackage/src/est_param.cpp

@@ -1,3 +1,4 @@
+// [[Rcpp::depends(RcppArmadillo)]]
 #include <RcppArmadillo.h>
 #include <math.h>
 

### 2019-11-29 code/simulations.R
added extra simulations

@@ -12,7 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-parallel <- FALSE
+parallel <- TRUE
 
 ### libraries
 library(gren)
@@ -1895,14 +1895,14 @@ csel <- 2^c(1:8)
 nreps <- 100
 
 ### analysis splits in parallel
-ncores <- min(detectCores() - 1, nreps)
+# ncores <- min(detectCores() - 1, nreps)
+ncores <- nreps
 cluster <- makeForkCluster(ncores)
 if(parallel) {
   registerDoParallel(cluster)
 } else {
   registerDoSEQ()
 }
-
 res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   
   print(paste("rep", k))

### 2019-11-29 docs/manuscript_biostatistics.Rnw

--- 2019-03-22 docs/manuscript_biostatistics.Rnw
@@
 \documentclass[oupdraft]{bio}
-\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
+\usepackage{xr-hyper}
+\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, 
+citecolor=citecolor]{hyperref}
 \usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable}
+
+\externaldocument[sm-]{supplement}
 
@@
 	our method is different from recent developments in grouped penalized
-	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
+	regression such as \cite{dondelinger_joint_2018}, where instead of the
 	features, the observations are partitioned.
@@
 	distributions
-	\cite[]{alhamzawi_bayesian_2017,mallick_bayesian_2013,kyung_penalized_2010}, 
+	\cite[]{alhamzawi_bayesian_2018,nengjun_yi_bayesian_2013,
+	kyung_penalized_2010},
 	but the precise parametrisation of this gamma prior is not so obvious. A

### 2019-12-02 docs/manuscript.Rnw
merged old versions from onedrive

@@
 	In addition to simulations from the correct model, i.e., model 
+	(\ref{eq:enmodel2}), we simulated from an incorrect model to investigate 
+	the performance of \texttt{gren} under model misspecification. We investigated
+	both link misspecification and non-linear feature effects misspecification. 
+	The details of the simulations and detailed results are given in SM Section 
+	\ref{sm-sec:modelmisspecification}. In all investigated Scenarios 
+	\texttt{gren} outperformed the regular elastic net in terms of predictive 
+	performance; an indication that \texttt{gren} is somewhat robust to model
+	misspecification.
 

### 2019-12-02 docs/supplement.Rnw
merged old versions from onedrive

@@
   $$
-  Let $\mathbf{c} = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, 
+  Let $\mathbf{c} = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. 
+  Furthermore, 
   in the following, all expectations and variances are with respect to 
@@
 
-  \subsection{Model misspecification}
+  \subsection{Model misspecification}\label{sec:modelmisspecification}
   To assess the effects of model misspecification we conducted a series of 

### 2019-12-02 code/figures.R

@@ -4,23 +4,6 @@ library(sp)
 load(file="results/simulations_res1.Rdata")
 plot.data <- c(sapply(c("grridge", "ridge"), function(m) {
   unname(cbind(NA, mean(sapply(res, function(s) {
-<<<<<<< HEAD
-    s$auc[grep(m, names(s$auc))[1]]}), na.rm=TRUE),
-    mean(sapply(res, function(s) {
-      s$auc[grep(m, names(s$briers))[1]]}), na.rm=TRUE)))}, simplify=FALSE),
-  sapply(c(paste0("gren", 1:3), paste0("enet", 1:3),
-           paste0("cmcp", 1:3), paste0("gel", 1:3)), function(m) {
-             cbind(rowMeans(sapply(res, function(s) {
-               vec <- s$psel[grep(m, names(s$psel))]
-               replace(rep(NA, 8), 1:length(vec), vec)}), 
-               na.rm=TRUE),
-               rowMeans(sapply(res, function(s) {
-                 vec <- s$auc[grep(m, names(s$auc))]
-                 replace(rep(NA, 8), 1:length(vec), vec)}), 
-                 na.rm=TRUE),
-               rowMeans(sapply(res, function(s) {
-                 vec <- s$briers[grep(m, names(s$briers))]
-=======
     s$auc[substr(names(s$auc), 1, nchar(m))==m]}), na.rm=TRUE),
     mean(sapply(res, function(s) {
       s$auc[substr(names(s$briers), 1, nchar(m))==m]}), na.rm=TRUE)))}, 
@@ -565,7 +548,6 @@ plot.data <- c(sapply(c("grridge", "ridge"), function(m) {
                  na.rm=TRUE),
                rowMeans(sapply(res, function(s) {
                  vec <- s$briers[substr(names(s$briers), 1, nchar(m))==m]
->>>>>>> 31db1da7fc892df40e05ce4287d82f38e942befc
                  replace(rep(NA, 8), 1:length(vec), vec)}), 
                  na.rm=TRUE))}, simplify=FALSE))
 
@@ -622,36 +604,11 @@ abline(h=plot.data[[1]][, 3], col=col[1], lty=lty[1])
 abline(h=plot.data[[2]][, 3], col=col[2], lty=lty[1])
 par(opar)
 
-<<<<<<< HEAD
-# ---- lines_simulations_res2_auc_briers ----
-library(sp)
-load(file="results/simulations_res2.Rdata")
-
[…]

### 2019-12-02 code/functions.R

@@ -61,27 +61,6 @@ sel.grpreg <- function(X, y, group=1:ncol(X),
   
 }
 
-<<<<<<< HEAD
-n <- 100
-p <- 20
-x <- matrix(rnorm(n*p), ncol=p, nrow=n)
-beta <- -9.5:9.5
-y <- rbinom(n, 1, as.numeric(1/(1 + exp(-x %*% beta))))
-m <- rep(1, n)
-alpha <- 0.5
-lambda <- 1
-lambdamult <- rep(1, p)
-intercept <- FALSE
-control <- list(nsamples=1000)
-
-fit.mcmc <- mcmc.gren(x, y, m=rep(1, nrow(x)), 
-                      alpha=0.5, lambda=1, lambdamult=rep(1, p), 
-                      intercept=FALSE, control=list(nsamples=1000))
-plot(c(beta), apply(fit.mcmc$samples$beta, 1, mean))
-
-
-=======
->>>>>>> 31db1da7fc892df40e05ce4287d82f38e942befc
 # mcmc version of VB steps
 # devtools::install_version("BayesLogit", "0.2-0")
 mcmc.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL, 
@@ -146,8 +125,6 @@ mcmc.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL,
   
 }
 
-<<<<<<< HEAD
-=======
 vb.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL, 
                     lambdamult, intercept=TRUE, 
                     control=list(maxit=100, epsilon=0.001)) {
@@ -242,5 +219,3 @@ vb.gren <- function(x, y, m=rep(1, nrow(x)), alpha=0.5, lambda=NULL,
 
 
 
->>>>>>> 31db1da7fc892df40e05ce4287d82f38e942befc
-                

### 2019-12-02 code/simulations.R

@@ -12,11 +12,7 @@ if(substr(system('git log -n 1 --format="%h %aN %s %ad"', intern=TRUE), 1, 7)!=
 }
 
 ### parallelisation
-<<<<<<< HEAD
-parallel <- FALSE
-=======
 parallel <- TRUE
->>>>>>> 31db1da7fc892df40e05ce4287d82f38e942befc
 
 ### libraries
 library(gren)
@@ -31,576 +27,6 @@ library(mvtnorm)
 ### loading functions
 source("code/functions.R")
 
-<<<<<<< HEAD
-################################# simulation 1 #################################
-n <- 100
-ntest <- 1000
-p <- 1000
-G <- 5
-rho <- 0.5
-rg <- 50
-beta.mean <- 0.07
-f <- 1.6
-
-Sigma <- diag(p)
-for(i in 1:p) {
-  for(j in 1:p) {
-    Sigma[i, j] <- rho^abs(i - j)
-  }
-}
-b1 <- 2*p*beta.mean/sum(rg*f^(c(1:G) - 1))
-beta <- numeric(p)
-
-part <- rep(c(1:G), each=p/G)
-csel <- 2^c(1:8)
-nreps <- 100
-
-### analysis splits in parallel
-ncores <- min(detectCores() - 1, nreps)
-cluster <- makeForkCluster(ncores)
-if(parallel) {
-  registerDoParallel(cluster)
-} else {
-  registerDoSEQ()
-}
-
-res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
-  set.seed(2019 + k)
-  
-  for(g in 1:G) {
-    idz <- c(((g-1)*p/G + 1):(g*p/G - rg))
-    ida <- c(((g*p/G - rg) + 1):(g*p/G))
-    beta[idz] <- 0
-    beta[ida] <- runif(rg, 0, b1*f^(g-1))
-  }
-  xtrain <- rmvnorm(n, rep(0, p), Sigma)
-  ytrain <- rbinom(n, 1, as.numeric(1/(1 + exp(-xtrain %*% beta))))
-  
-  xtest <- rmvnorm(ntest, rep(0, p), Sigma)
-  ytest <- rbinom(ntest, 1, as.numeric(1/(1 + exp(-xtest %*% beta))))
-  
-  # fitting models
-  fit.gren1 <- gren(xtrain, ytrain, partitions=list(part=part), alpha=0.05, 
[…]

### 2019-12-02 code/simulations.R
fixed error in simulations.R

@@ -1695,6 +1695,7 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   print(paste("rep", k))
   set.seed(2019 + k)
 
+  xtrain <- rmvnorm(n, mean=rep(0, p), sigma=Sigma)
   ind.mat <- matrix(rep(rep(c(TRUE, FALSE),
                             times=c(frac.relu*p/G, (1 - frac.relu)*p/G)), G),
                     nrow=n, ncol=p, byrow=TRUE)
@@ -1803,6 +1804,7 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   print(paste("rep", k))
   set.seed(2019 + k)
 
+  xtrain <- rmvnorm(n, mean=rep(0, p), sigma=Sigma)
   ind.mat <- matrix(rep(rep(c(TRUE, FALSE),
                             times=c(frac.relu*p/G, (1 - frac.relu)*p/G)), G),
                     nrow=n, ncol=p, byrow=TRUE)
@@ -1910,6 +1912,7 @@ res <- foreach(k=c(1:nreps), .errorhandling="pass") %dopar% {
   print(paste("rep", k))
   set.seed(2019 + k)
   
+  xtrain <- rmvnorm(n, mean=rep(0, p), sigma=Sigma)
   ind.mat <- matrix(rep(rep(c(TRUE, FALSE), 
                             times=c(frac.relu*p/G, (1 - frac.relu)*p/G)), G), 
                     nrow=n, ncol=p, byrow=TRUE)
